---
output:
  github_document:
    toc: false
    html_preview: false
    md_extensions: -smart
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
options(width = 90)
options(cli.unicode = FALSE)
default_output_hook <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(
  output = function(x, options) {
    default_output_hook(gsub("[ \t]+(?=\n)", "", x, perl = TRUE), options)
  }
)
demo_data_path <- if (file.exists("data/ledgr_demo_bars.rda")) {
  "data/ledgr_demo_bars.rda"
} else {
  file.path("..", "data", "ledgr_demo_bars.rda")
}
if (!exists("ledgr_demo_bars") && file.exists(demo_data_path)) {
  load(demo_data_path)
}
```

# ledgr

ledgr is an event-sourced systematic trading research framework for R.

In v0.1.x, ledgr focuses on deterministic research: sealed market-data
snapshots, experiment-first backtests, durable run metadata, strategy
provenance, comparison tables, and low-code TTR indicators. Paper trading and
live trading adapters are planned for later releases and are not available in
the current package.

Most backtesting tools compute results directly from price arrays. ledgr records
each decision and state change as an immutable event, then derives trades,
equity, and metrics from that ledger.

```text
sealed snapshot -> experiment -> run -> event ledger -> results
```

For the longer design arc, see the
[`research-to-production`](https://blechturm.github.io/ledgr/articles/research-to-production.html)
article on the pkgdown site.

Not sure whether ledgr fits your workflow? Start with
[Who ledgr is for](https://blechturm.github.io/ledgr/articles/who-ledgr-is-for.html).

## Install

```{r install, eval=FALSE}
if (!requireNamespace("pak", quietly = TRUE)) install.packages("pak")
pak::pak("blechturm/ledgr")
```

The README uses `dplyr` and `tibble` for compact example output. They are
suggested packages for documentation and examples; ledgr strategies themselves
use the pulse context shown below.

```{r attach, message=FALSE}
library(ledgr)
library(dplyr)
library(tibble)
data("ledgr_demo_bars", package = "ledgr")
```

## First Experiment

Use the bundled demo bars for a first run. They are deterministic and require no
network access.

```{r bars}
bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% c("DEMO_01", "DEMO_02"),
    between(ts_utc, ledgr_utc("2019-01-01"), ledgr_utc("2019-06-30"))
  )

bars |>
  slice_head(n = 4)
```

Create a sealed snapshot. A snapshot is the immutable data artifact every run
uses.

```{r snapshot}
snapshot <- ledgr_snapshot_from_df(bars)
```

Strategies receive a pulse context `ctx` and a parameter list `params`. They
return target holdings: a named numeric vector with one desired quantity per
instrument.

```{r strategy}
strategy <- function(ctx, params) {
  targets <- ctx$flat()

  for (id in ctx$universe) {
    sma <- ctx$feature(id, "sma_20")
    if (is.finite(sma) && ctx$close(id) > sma) {
      targets[id] <- params$qty
    }
  }

  targets
}
```

Bundle the snapshot, strategy, indicators, starting state, and execution options
into an experiment. Construction validates the object; it does not run the
strategy or write run artifacts.

```{r experiment}
features <- list(ledgr_ind_sma(20))

exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = strategy,
  features = features,
  opening = ledgr_opening(cash = 10000)
)

exp
```

Run the experiment with explicit parameters.

```{r run}
bt <- exp |>
  ledgr_run(params = list(qty = 10), run_id = "readme_sma_20")

bt
```

Inspect result views. These are derived from the recorded event ledger.

```{r inspect}
summary(bt)
ledgr_results(bt, what = "trades")
```

## Compare Runs

Run another parameter set against the same experiment and compare stored
results. Comparison reads existing artifacts; it does not recompute strategies.

```{r compare}
bt_qty_20 <- exp |>
  ledgr_run(params = list(qty = 20), run_id = "readme_sma_20_qty_20")

ledgr_compare_runs(snapshot, run_ids = c("readme_sma_20", "readme_sma_20_qty_20"))
```

## Durable Research

For durable research, create the snapshot in a project DuckDB file:

```r
snapshot <- ledgr_snapshot_from_df(bars, db_path = "research.duckdb")
```

In a later R session, reopen the sealed snapshot and continue from the snapshot
handle:

```r
snapshot <- ledgr_snapshot_load("research.duckdb", snapshot_id = "my_snapshot")
ledgr_run_list(snapshot)
ledgr_run_info(snapshot, "readme_sma_20")
```

After snapshot creation or loading, normal experiment-store operations take the
snapshot handle rather than a raw database path.

## Scope

v0.1.7 is the experiment-first research API. It does not ship parameter sweep
execution, broker adapters, paper trading, live trading, or short-selling
semantics. Those are separate roadmap items with different state and safety
requirements.

`ledgr_run()` returns a live handle. The run artifacts are already durable when
the run finishes. Most result inspection opens and closes its own read
connection; explicit `close(bt)` is resource cleanup for long sessions, explicit
opens, and lazy result cursors.

```{r cleanup}
close(bt)
close(bt_qty_20)
ledgr_snapshot_close(snapshot)
```

## Documentation

```{r docs, eval=FALSE}
help(package = "ledgr")
utils::packageDescription("ledgr")[c("Package", "Version", "Title")]
vignette(package = "ledgr")
system.file("doc", package = "ledgr")
system.file("doc", "strategy-development.html", package = "ledgr")
```

The `system.file()` calls are useful in noninteractive `Rscript` and agent
workflows where opening pkgdown in a browser is not the first step. The
pkgdown-only background articles are on the website, while installed vignettes
focus on package workflows.

Design packets are in `inst/design/`, including the current v0.1.7.3 packet.
