---
output:
  github_document:
    toc: false
    html_preview: false
    md_extensions: -smart
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
options(width = 90)
options(cli.unicode = FALSE)
if (file.exists("DESCRIPTION") && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(".", quiet = TRUE)
}
default_output_hook <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(
  output = function(x, options) {
    x <- gsub("Database:(\\s+)[^\\r\\n]+[.]duckdb", "Database:\\1<temporary DuckDB path>", x, perl = TRUE)
    x <- gsub("[ \t]+(?=\n)", "", x, perl = TRUE)
    default_output_hook(x, options)
  }
)
```

# ledgr

<img src="man/figures/logo.svg" align="right" alt="ledgr logo" width="160" class="ledgr-readme-logo" />

ledgr is an event-sourced systematic trading research framework for R.

Use it when you want a backtest result to be more than a temporary object in an
R session. ledgr starts from sealed market-data snapshots, runs strategies
through an experiment boundary, records event-sourced results, and lets you
reopen the evidence later.

```text
sealed snapshot -> experiment -> run -> event ledger -> results
```

The setup is not overhead. The setup is the audit trail.

## Install

```{r install, eval=FALSE}
if (!requireNamespace("pak", quietly = TRUE)) install.packages("pak")
pak::pak("blechturm/ledgr")
```

```{r attach, message=FALSE}
library(ledgr)
library(dplyr)

data("ledgr_demo_bars", package = "ledgr")
```

## Run A Small Backtest

Start with the package-owned demo bars. Real research should seal your own
market data, but the demo data keeps this first run local and deterministic.

```{r bars}
bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% c("DEMO_01", "DEMO_02"),
    between(ts_utc, ledgr_utc("2019-01-01"), ledgr_utc("2019-06-30"))
  )

bars |>
  slice_head(n = 4)
```

Seal the bars, declare the strategy boundary, and run one parameter set.

```{r quick-run}
snapshot <- ledgr_snapshot_from_df(
  bars,
  snapshot_id = "readme_demo"
)

features <- ledgr_feature_map(
  fast = ledgr_ind_sma(ledgr_param("fast_n")),
  slow = ledgr_ind_sma(ledgr_param("slow_n"))
)

exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = ledgr_demo_sma_crossover_strategy(),
  features = features,
  opening = ledgr_opening(cash = 10000),
  cost_model = ledgr_cost_zero()
)

bt <- ledgr_run(
  exp,
  feature_params = list(fast_n = 10L, slow_n = 40L),
  params = list(qty = 10, threshold = 0),
  run_id = "readme_sma_crossover"
)

summary(bt)
```

## Inspect The Evidence

The result views are derived from recorded events. The ledger is the source of
truth; trades, equity, and metrics are views over that evidence.

```{r inspect}
ledgr_results(bt, what = "trades")
head(ledgr_results(bt, what = "equity"), 3)
```

Stored strategy provenance is inspectable without rerunning or evaluating the
strategy source. Use `trust = FALSE` for source and metadata inspection.

```{r extract-strategy}
stored_strategy <- ledgr_run_strategy(snapshot, "readme_sma_crossover", trust = FALSE)
list(
  reproducibility_level = stored_strategy$reproducibility_level,
  hash_verified = stored_strategy$hash_verified,
  strategy_params = stored_strategy$strategy_params
)
```

Hash verification proves stored-text identity, not code safety. Use
`trust = TRUE` only when you already trust the store and intentionally want to
recover a function object.

## Where To Go Next

| Question | Article |
| --- | --- |
| I want the full research loop: snapshot, sweep, promotion, reopen. | [Research Workflow](https://blechturm.github.io/ledgr/articles/research-workflow.html) |
| I want to write strategies correctly. | [Strategy Development](https://blechturm.github.io/ledgr/articles/strategy-development.html) |
| I want feature maps, indicators, and active aliases. | [Indicators](https://blechturm.github.io/ledgr/articles/indicators.html) |
| I want exploratory sweeps and candidate promotion. | [Sweeps](https://blechturm.github.io/ledgr/articles/sweeps.html) |
| I want sealed snapshots, durable stores, backup, and reopen. | [Experiment Store](https://blechturm.github.io/ledgr/articles/experiment-store.html) |
| I want hashes, provenance tiers, and limits of recovery. | [Reproducibility](https://blechturm.github.io/ledgr/articles/reproducibility.html) |
| I want fills, trades, equity, metrics, and metric context. | [Metrics And Accounting](https://blechturm.github.io/ledgr/articles/metrics-and-accounting.html) |

Start with the pkgdown site for the full article set:
<https://blechturm.github.io/ledgr/>.

Installed package help remains available from R:

```{r docs, eval=FALSE}
help(package = "ledgr")
vignette(package = "ledgr")
```

## Ecosystem

ledgr connects to the R finance ecosystem through adapters. The core is narrow
by design: `data -> pulse -> decision -> fill -> ledger event -> portfolio
state`. Everything outside that sequence, such as data vendors, indicators,
charting, and analytics, can be provided by packages that already do those
things well.

| ledgr owns | Other packages can own |
|---|---|
| sealed snapshots and hashes | market-data acquisition |
| pulse construction and no-lookahead contexts | indicator calculations through adapters |
| target validation, fills, and ledger events | charting and visualization |
| run identity, provenance, and result reconstruction | downstream analytics and reporting |

This posture is deliberate. If you want an all-in-one charting or
array-backtesting package, ledgr may not be the shortest path. Choose ledgr
when you want the audit trail and adapter boundary to be explicit.

## Scope

The current ledgr research API is experiment-first and includes sequential
exploratory sweep support. It does not ship automatic ranking, `ledgr_tune()`,
parallel sweep, walk-forward/PBO/CSCV helpers, full sweep artifact persistence,
broker adapters, paper trading, live trading, or short-selling semantics. Those
are separate roadmap items with different state and safety requirements.

`ledgr_run()` returns a live handle. The run artifacts are already durable when
the run finishes. Most result inspection opens and closes its own read
connection; explicit `close(bt)` is resource cleanup for long sessions,
explicit opens, and lazy result cursors.

```{r cleanup, include=FALSE}
if (exists("bt")) close(bt)
if (exists("snapshot")) ledgr_snapshot_close(snapshot)
```

## Pre-CRAN Compatibility

ledgr is not yet on CRAN. Until the first CRAN release, stored artifacts,
database schemas, config hashes, provenance formats, and experimental APIs may
change without backward compatibility or a deprecation cycle. Treat pre-CRAN
ledgr as a research/development package and expect to rerun experiments after
upgrading. Once ledgr is released on CRAN, the project will define an explicit
compatibility and deprecation policy.
