---
output:
  github_document:
    toc: false
    html_preview: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
options(width = 90)
options(cli.unicode = FALSE)
default_output_hook <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(
  output = function(x, options) {
    default_output_hook(gsub("[ \t]+(?=\n)", "", x, perl = TRUE), options)
  }
)
```

# ledgr

ledgr is an event-sourced backtesting engine for R.

It is designed for one thing: making backtests reproducible and auditable.

Most backtesting tools compute results from full price arrays. ledgr records
decisions and state changes as events, then derives trades, equity, and metrics
from that ledger.

```text
data -> sealed snapshot -> pulses -> event ledger -> results
```

That means results come from recorded history, not from a hidden intermediate
calculation. ledgr is not a live-trading system, broker adapter, or optimization
framework.

## Install

```{r install, eval=FALSE}
if (!requireNamespace("pak", quietly = TRUE)) install.packages("pak")
pak::pak("blechturm/ledgr")
```

Then attach ledgr and tibble. ledgr returns tidy tibbles for inspection, so it
fits naturally into tidyverse-style workflows without requiring the full
tidyverse in the first run.

```{r attach}
library(ledgr)
library(tibble)
```

## First Backtest

This first run creates two synthetic instruments, defines a target-position
strategy, and runs one backtest. The data is deliberately generated in the
example so the code works in a fresh R session without local files or network
data.

First create a small OHLCV data set:

```{r create-bars}
set.seed(20260425)
calendar <- seq.Date(as.Date("2020-01-01"), as.Date("2020-02-14"), by = "day")
dates <- calendar[!(weekdays(calendar) %in% c("Saturday", "Sunday"))]

make_bars <- function(instrument_id, start_price, drift) {
  n <- length(dates)
  close <- start_price + cumsum(drift + stats::rnorm(n, mean = 0, sd = 0.35))
  open <- c(start_price, close[-n])

  data.frame(
    ts_utc = as.POSIXct(dates, tz = "UTC"),
    instrument_id = instrument_id,
    open = round(open, 2),
    high = round(pmax(open, close) + 0.45, 2),
    low = round(pmin(open, close) - 0.45, 2),
    close = round(close, 2),
    volume = seq.int(1000L, 1000L + n - 1L),
    stringsAsFactors = FALSE
  )
}

bars <- rbind(
  make_bars("AAA", start_price = 100, drift = 0.18),
  make_bars("BBB", start_price = 80, drift = -0.04)
)

bars |> as_tibble() |> head(4)
```

Every ledgr strategy ultimately returns target holding amounts. This vector:

```{r target-example}
c(AAA = 10, BBB = 0)
```

means: hold 10 units of `AAA` and 0 units of `BBB`. Names matter because ledgr
has to match each target to `ctx$universe`; values are numeric position
quantities, not labels such as `"LONG"` or `"FLAT"`.

Now define a strategy that reads the close price at the current pulse and
returns target holdings for the full universe. `ctx$targets()` creates a flat
target vector over `ctx$universe`; the strategy then changes only the holdings
it wants to own:

```{r define-strategy}
strategy <- function(ctx) {
  targets <- ctx$targets()

  if (ctx$close("AAA") > 100.4) {
    targets["AAA"] <- 10
  }

  if (ctx$close("BBB") > 80.0) {
    targets["BBB"] <- 5
  }

  targets
}
```

Run the backtest:

```{r run-backtest}
bt <- ledgr_backtest(
  data = bars,
  strategy = strategy,
  start = min(bars$ts_utc),
  end = max(bars$ts_utc),
  initial_cash = 10000,
  run_id = "readme-demo"
)

bt
```

Inspect the derived results:

```{r inspect-results}
summary(bt)
bt |> as_tibble(what = "trades")
```

The numbers come from generated toy data. They validate the API path, not the
strategy. Exact printed values can change after intentional engine changes. The
important invariant is not the exact toy numbers; it is that the same data and
same strategy produce the same ledger and equity curve.

## Why ledgr?

Most backtesting tools assume reproducibility. ledgr makes it testable.

- Same data and the same strategy produce identical normalized ledger and
  equity outputs.
- All reported results come from the event ledger.
- Sealed snapshot hashes let you detect when the input data changed.

This matters because research often gets revisited months later. ledgr gives
you a way to prove which data and strategy produced the result you are looking
at.

Many tools compute results from full price arrays. ledgr follows the sequential
process more closely:

- new data arrives;
- the strategy sees one decision point in time, called a pulse;
- positions and cash can change;
- the system moves forward.

At each pulse, the strategy only sees data available at that timestamp. Every
state change is recorded as an event. The ledger is the source of truth; trades,
equity, and metrics are derived views of that ledger.

The first call to `ledgr_backtest(data = bars, ...)` created a sealed snapshot,
then called the same canonical engine used by explicit snapshot workflows. We
seal this data so you cannot accidentally lie to your future self.

The quick ledger/equity count below confirms that the visible result is backed
by recorded events and derived equity rows.

```{r result-shape}
c(
  ledger_rows = nrow(bt |> as_tibble(what = "ledger")),
  equity_rows = nrow(bt |> as_tibble(what = "equity"))
)
```

## Trust Check: Determinism (Optional But Important)

```{r determinism-check}
normalize_result <- function(x) {
  if ("run_id" %in% names(x)) x$run_id <- NULL
  if ("event_id" %in% names(x)) x$event_id <- NULL

  x[] <- lapply(x, function(col) {
    if (inherits(col, "POSIXt")) {
      format(as.POSIXct(col, tz = "UTC"), "%Y-%m-%dT%H:%M:%SZ", tz = "UTC")
    } else if (is.numeric(col)) {
      round(col, 10)
    } else {
      col
    }
  })

  row.names(x) <- NULL
  x
}

run_once <- function(run_id) {
  bt <- ledgr_backtest(
    data = bars,
    strategy = strategy,
    start = min(bars$ts_utc),
    end = max(bars$ts_utc),
    initial_cash = 10000,
    run_id = run_id
  )

  list(
    ledger = normalize_result(bt |> as_tibble(what = "ledger")),
    equity = normalize_result(bt |> as_tibble(what = "equity"))
  )
}

run_a <- run_once("readme-a")
run_b <- run_once("readme-b")

c(
  same_ledger = identical(run_a$ledger, run_b$ledger),
  same_equity = identical(run_a$equity, run_b$equity)
)
```

The two runs use different run identifiers and different temporary databases.
After identity columns are removed, the ledger and equity curve match exactly.
That is the ledgr difference: replay is testable instead of assumed. If the
sealed input data changes, the snapshot hash changes with it.

## What To Try Next

Good next edits are small and observable:

- change the `AAA` or `BBB` target quantities in `strategy()`;
- change the threshold values that trigger a position;
- add an indicator such as `ledgr_ind_sma(5)`;
- inspect a single decision point with `ledgr_pulse_snapshot()`.

## Documentation

```{r docs, eval=FALSE}
help(package = "ledgr")
utils::packageDescription("ledgr")[c("Package", "Version", "Title")]
```

The v0.1.3 design packet is in `inst/design/ledgr_v0_1_3_spec_packet/`.
The v0.1.2 packet records the engine and UX foundation this release builds on.
