---
output:
  github_document:
    toc: false
    html_preview: false
    md_extensions: -smart
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
options(width = 90)
options(cli.unicode = FALSE)
default_output_hook <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(
  output = function(x, options) {
    x <- gsub("Database:(\\s+)[^\\r\\n]+[.]duckdb", "Database:\\1<temporary DuckDB path>", x, perl = TRUE)
    x <- gsub("[ \t]+(?=\n)", "", x, perl = TRUE)
    default_output_hook(x, options)
  }
)
demo_data_path <- if (file.exists("data/ledgr_demo_bars.rda")) {
  "data/ledgr_demo_bars.rda"
} else {
  file.path("..", "data", "ledgr_demo_bars.rda")
}
if (!exists("ledgr_demo_bars") && file.exists(demo_data_path)) {
  load(demo_data_path)
}
```

# ledgr

<img src="man/figures/logo.svg" align="right" alt="ledgr logo" width="160" class="ledgr-readme-logo" />

ledgr is an event-sourced systematic trading research framework for R.

In v0.1.x, ledgr focuses on deterministic research: sealed market-data
snapshots, experiment-first backtests, durable run metadata, strategy
provenance, comparison tables, and low-code TTR indicators. Paper trading and
live trading adapters are planned for later releases and are not available in
the current package.

Most backtesting tools compute results directly from price arrays. ledgr records
each decision and state change as an immutable event, then derives trades,
equity, and metrics from that ledger.

```text
sealed snapshot -> experiment -> run -> event ledger -> results
```

ledgr connects to the R finance ecosystem through adapters.

For the longer design arc, see the
[`research-to-production`](https://blechturm.github.io/ledgr/articles/research-to-production.html)
article on the pkgdown site.

Not sure whether ledgr fits your workflow? Start with
[Who ledgr is for](https://blechturm.github.io/ledgr/articles/who-ledgr-is-for.html).

## Install

```{r install, eval=FALSE}
if (!requireNamespace("pak", quietly = TRUE)) install.packages("pak")
pak::pak("blechturm/ledgr")
```

The README uses `dplyr` and `tibble` for compact example output. They are
suggested packages for documentation and examples; ledgr strategies themselves
use the pulse context shown below.

```{r attach, message=FALSE}
library(ledgr)
library(dplyr)
library(tibble)
data("ledgr_demo_bars", package = "ledgr")
```

## First Experiment

Use the bundled demo bars for a first run. They are deterministic and require no
network access.

```{r bars}
bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% c("DEMO_01", "DEMO_02"),
    between(ts_utc, ledgr::ledgr_utc("2019-01-01"), ledgr::ledgr_utc("2019-06-30"))
  )

bars |>
  slice_head(n = 4)
```

Create a sealed snapshot. A snapshot is the immutable data artifact every run
uses.
The setup is not overhead. The setup is the audit trail.

```{r snapshot}
snapshot <- ledgr_snapshot_from_df(bars)
```

First-contact examples use the demo SMA-crossover strategy fixture. It is a
small teaching strategy, not an investment recommendation. The feature map stays
explicit so the two namespaces are visible: `feature_params` materialize
indicators, while `params` are the values the strategy reads at each pulse.

```{r strategy}
features <- ledgr_feature_map(
  fast = ledgr_ind_sma(ledgr_param("fast_n")),
  slow = ledgr_ind_sma(ledgr_param("slow_n"))
)

strategy <- ledgr_demo_sma_crossover_strategy()
```

Bundle the snapshot, strategy, indicators, starting state, and execution options
into an experiment. Construction validates the object; it does not run the
strategy or write run artifacts.

```{r experiment}
exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = strategy,
  features = features,
  opening = ledgr_opening(cash = 10000)
)

exp
```

Run the experiment with explicit parameters.

```{r run}
bt <- exp |>
  ledgr_run(
    feature_params = list(fast_n = 10L, slow_n = 40L),
    params = list(qty = 10, threshold = 0),
    run_id = "readme_sma_crossover"
  )

bt
```

Inspect result views. These are derived from the recorded event ledger.

```{r inspect}
summary(bt)
ledgr_results(bt, what = "trades")
```

## Compare Runs

Run another parameter set against the same experiment and compare stored
results. Comparison reads existing artifacts; it does not recompute strategies.

```{r compare}
bt_qty_20 <- exp |>
  ledgr_run(
    feature_params = list(fast_n = 10L, slow_n = 40L),
    params = list(qty = 20, threshold = 0),
    run_id = "readme_sma_crossover_qty_20"
  )

ledgr_compare_runs(snapshot, run_ids = c("readme_sma_crossover", "readme_sma_crossover_qty_20"))
```

## Explore A Sweep

Use `ledgr_sweep()` for lightweight exploration. Active-alias sweeps keep
feature parameters and strategy parameters separate, then compose them into an
executable grid. Sweep results are candidate summaries, not committed run
artifacts, and ledgr does not rank candidates automatically.

```{r sweep, eval=FALSE}
feature_grid <- ledgr_feature_grid(
  fast_n = c(10L, 20L),
  slow_n = c(40L, 80L),
  .filter = fast_n < slow_n
)

strategy_grid <- ledgr_strategy_grid(
  threshold = c(0, 0.01),
  qty = c(10, 20)
)

grid <- ledgr_grid_cross(features = feature_grid, strategy = strategy_grid)
precomputed <- ledgr_precompute_features(exp, grid)
results <- ledgr_sweep(exp, grid, precomputed_features = precomputed, seed = 2026)

candidate <- results |>
  filter(status == "DONE") |>
  arrange(desc(sharpe_ratio)) |>
  ledgr_candidate(1)

bt_promoted <- ledgr_promote(
  exp,
  candidate,
  run_id = "readme_promoted_sweep_candidate",
  note = "Same-snapshot replay of an exploratory candidate."
)
```

Same-snapshot replay is useful for audit and debugging, but it remains
in-sample. For evaluation, sweep on a manually created train snapshot, lock the
selected params with `ledgr_candidate()`, and promote against a held-out test
snapshot with `require_same_snapshot = FALSE`. The
[`sweeps`](https://blechturm.github.io/ledgr/articles/sweeps.html) article
shows that train/sweep/test discipline and explains `execution_seed`,
row-level provenance, promotion context, failure rows, and deferred sweep
artifact persistence.

Stored strategy provenance is inspectable without rerunning or evaluating the
strategy source. Use the default `trust = FALSE` path for safe source and
metadata inspection:

```{r extract-strategy}
stored_strategy <- ledgr_extract_strategy(snapshot, "readme_sma_crossover", trust = FALSE)
source_lines <- strsplit(stored_strategy$strategy_source_text, "\n", fixed = TRUE)[[1]]
cat(paste(c(head(source_lines, 8), "..."), collapse = "\n"))
```

Hash verification proves stored-text identity, not code safety. Use
`trust = TRUE` only when you already trust the store and intentionally want to
recover a function object.

## Durable Research

For durable research, create the snapshot in a project DuckDB file:

```r
snapshot <- ledgr_snapshot_from_df(bars, db_path = "research.duckdb")
```

In a later R session, reopen the sealed snapshot and continue from the snapshot
handle:

```r
snapshot <- ledgr_snapshot_load("research.duckdb", snapshot_id = "my_snapshot")
ledgr_run_list(snapshot)
ledgr_run_info(snapshot, "readme_sma_crossover")
```

After snapshot creation or loading, normal experiment-store operations take the
snapshot handle rather than a raw database path.

## Ecosystem

ledgr connects to the R finance ecosystem through adapters. The core is narrow
by design: `data -> pulse -> decision -> fill -> ledger event -> portfolio
state`. Everything outside that sequence, such as data vendors, indicators,
charting, and analytics, can be provided by packages that already do those
things well.

| ledgr owns | Other packages can own |
|---|---|
| sealed snapshots and hashes | market-data acquisition |
| pulse construction and no-lookahead contexts | indicator calculations through adapters |
| target validation, fills, and ledger events | charting and visualization |
| run identity, provenance, and result reconstruction | downstream analytics and reporting |

This posture is deliberate. If you want an all-in-one charting or
array-backtesting package, ledgr may not be the shortest path. Choose ledgr
when you want the audit trail and adapter boundary to be explicit.

## Scope

The current ledgr research API is experiment-first and includes sequential
exploratory sweep support. It does not ship automatic ranking, `ledgr_tune()`,
parallel sweep, walk-forward/PBO/CSCV helpers, full sweep artifact persistence,
broker adapters, paper trading, live trading, or short-selling semantics. Those
are separate roadmap items with different state and safety requirements.

`ledgr_run()` returns a live handle. The run artifacts are already durable when
the run finishes. Most result inspection opens and closes its own read
connection; explicit `close(bt)` is resource cleanup for long sessions, explicit
opens, and lazy result cursors.

```{r cleanup}
close(bt)
close(bt_qty_20)
ledgr_snapshot_close(snapshot)
```

## Documentation

Start with the pkgdown site for the full article set:
<https://blechturm.github.io/ledgr/>.

Installed package help remains available from R:

```{r docs, eval=FALSE}
help(package = "ledgr")
vignette(package = "ledgr")
```

## Pre-CRAN Compatibility

ledgr is not yet on CRAN. Until the first CRAN release, stored artifacts,
database schemas, config hashes, provenance formats, and experimental APIs may
change without backward compatibility or a deprecation cycle. Treat pre-CRAN
ledgr as a research/development package and expect to rerun experiments after
upgrading. Once ledgr is released on CRAN, the project will define an explicit
compatibility and deprecation policy.
