---
title: "Maintainer Review: Feature Value Path"
subtitle: "From bars to snapshot to experiment to ctx$feature()"
format:
  html:
    toc: true
    toc-depth: 3
execute:
  eval: false
  warning: false
  message: false
---

## Purpose

This is an internal maintainer review notebook. It is not installed
documentation and it is not a contract.

The review question:

> How does ledgr take ordinary bars, seal them into a snapshot, register a
> feature on an experiment, compute that feature, attach it to a pulse context,
> and make it visible through `ctx$feature()` inside strategy execution?

This path is relevant to v0.1.8.1 because:

- LDG-2202 needs a feature lifecycle guide.
- LDG-2210 needs multi-output indicator bundles to flatten at the right
  boundary without creating a second runtime feature system.

## How To Use This Notebook

Run the setup chunk first. After that, each executable review chunk creates its
own fresh fixture. That means you can run sections independently without
depending on stale objects from earlier chunks.

Each chunk follows the same pattern:

1. `fx <- make_feature_path_fixture(...)`
2. inspect one part of `fx`
3. `close_feature_path_fixture(fx)`

If a chunk errors before cleanup, run the cleanup chunk at the end.

## Visual Orientation

These diagrams are here to make the code reading less abstract. They are not a
formal architecture spec. They are a visual checklist for the objects you will
see in the runnable chunks.

The diagram chunks use `DiagrammeR::grViz()` so they can be run interactively in
the notebook. If `DiagrammeR` is not installed in the active R library, run:

```r
install.packages("DiagrammeR", repos = "https://cloud.r-project.org")
```

### Data And Artifact Flow

This is the package-level path you are tracing. The feature path is the lower
branch that joins execution through the pulse context.

```{r diagram-data-artifact-flow}
if (!requireNamespace("DiagrammeR", quietly = TRUE)) {
  message("Install DiagrammeR to run this diagram: install.packages(\"DiagrammeR\")")
} else {
  DiagrammeR::grViz('
digraph data_artifact_flow {
  graph [rankdir = LR, bgcolor = "white"]
  node [shape = box, style = "rounded,filled", fillcolor = "#f7f7f7", color = "#555555", fontname = "Arial"]
  edge [color = "#555555", fontname = "Arial"]

  bars [label = "bars data.frame\\nts_utc, instrument_id, OHLCV"]
  snapshot_fn [label = "ledgr_snapshot_from_df()"]
  snapshot [label = "sealed ledgr_snapshot\\nsnapshot_id + snapshot_hash", fillcolor = "#e8f1ff"]
  feature_ctor [label = "ledgr_ind_sma(20)"]
  feature_map [label = "ledgr_feature_map()\\nalias -> indicator"]
  experiment [label = "ledgr_experiment()\\nsnapshot + strategy + features", fillcolor = "#fff6df"]
  run [label = "ledgr_run(exp, params)"]
  runner [label = "persistent run orchestration"]
  fold [label = "ledgr_execute_fold()", fillcolor = "#edf7ed"]
  ctx [label = "pulse ctx\\nOHLCV + portfolio + feature_table", fillcolor = "#edf7ed"]
  accessor [label = "ctx$feature(id, feature_id)"]
  strategy [label = "strategy(ctx, params)\\nreturns target vector"]
  events [label = "ledger events"]
  results [label = "equity / ledger / metrics"]

  bars -> snapshot_fn -> snapshot
  feature_ctor -> feature_map -> experiment
  snapshot -> experiment -> run -> runner -> fold -> ctx
  ctx -> accessor -> strategy -> events -> results
  strategy -> fold
}
')
}
```

Review prompt:

```{markdown}
Which node is the first durable artifact? Which node is the first strategy-time
object?
```

### Function Composition View

This view is about ownership. Each box names the function family that owns a
boundary. When reviewing code, try to decide whether each function is doing
orchestration, validation, materialization, execution, or inspection.

```{r diagram-function-composition}
if (!requireNamespace("DiagrammeR", quietly = TRUE)) {
  message("Install DiagrammeR to run this diagram: install.packages(\"DiagrammeR\")")
} else {
  DiagrammeR::grViz('
digraph function_composition {
  graph [rankdir = TB, compound = true, bgcolor = "white"]
  node [shape = box, style = "rounded,filled", fillcolor = "#f7f7f7", color = "#555555", fontname = "Arial"]
  edge [color = "#555555", fontname = "Arial"]

  subgraph cluster_authoring {
    label = "Authoring / declarations"
    color = "#d6d6d6"
    ind [label = "ledgr_ind_sma()\\nR/indicators_builtin.R"]
    fmap [label = "ledgr_feature_map()\\nR/feature-map.R"]
    expfn [label = "ledgr_experiment()\\nR/experiment.R"]
  }

  subgraph cluster_snapshot {
    label = "Snapshot layer"
    color = "#d6d6d6"
    snapdf [label = "ledgr_snapshot_from_df()\\nR/snapshot_adapters.R"]
    seal [label = "validation + seal\\nR/snapshot.R"]
  }

  subgraph cluster_materialization {
    label = "Feature materialization"
    color = "#d6d6d6"
    mat [label = "ledgr_experiment_materialize_features()\\nR/experiment.R"]
    ids [label = "ledgr_feature_id()\\nledgr_indicator_fingerprint()"]
    compute [label = "feature computation / cache"]
  }

  subgraph cluster_execution {
    label = "Execution"
    color = "#d6d6d6"
    runfn [label = "ledgr_run()\\nR/backtest.R"]
    runfold [label = "ledgr_run_fold()\\nR/backtest-runner.R"]
    foldcore [label = "ledgr_execute_fold()\\nR/fold-core.R"]
    attach [label = "ledgr_attach_feature_helpers()\\nR/pulse-context.R"]
    lookup [label = "ledgr_feature_accessor()\\nR/pulse-context.R"]
  }

  subgraph cluster_inspection {
    label = "Inspection"
    color = "#d6d6d6"
    pulse [label = "ledgr_pulse_snapshot()\\nR/indicator_dev.R"]
    long [label = "ledgr_pulse_features()\\nR/feature-inspection.R"]
    wide [label = "ledgr_pulse_wide()\\nR/feature-inspection.R"]
  }

  ind -> fmap -> expfn
  snapdf -> seal -> expfn
  expfn -> mat -> ids -> compute
  expfn -> runfn -> runfold -> foldcore -> attach -> lookup
  compute -> attach
  seal -> pulse -> long
  pulse -> wide
  compute -> pulse
}
')
}
```

Review prompt:

```{markdown}
Which functions are public user-facing helpers? Which functions are private
engine boundaries?
```

### Adapter And Boundary View

This view separates adapters from runtime engine objects. The important
maintainer question is whether a helper adapts user input into ledgr's existing
contracts, or whether it creates a new runtime path.

```{r diagram-adapter-boundaries}
if (!requireNamespace("DiagrammeR", quietly = TRUE)) {
  message("Install DiagrammeR to run this diagram: install.packages(\"DiagrammeR\")")
} else {
  DiagrammeR::grViz('
digraph adapter_boundaries {
  graph [rankdir = LR, compound = true, bgcolor = "white"]
  node [shape = box, style = "rounded,filled", fillcolor = "#f7f7f7", color = "#555555", fontname = "Arial"]
  edge [color = "#555555", fontname = "Arial"]

  subgraph cluster_snapshot_adapters {
    label = "Snapshot adapters"
    color = "#d6d6d6"
    df [label = "data.frame bars"]
    csv [label = "CSV bars"]
    yahoo [label = "Yahoo / external bars"]
    snapshot [label = "ledgr_snapshot", fillcolor = "#e8f1ff"]
  }

  subgraph cluster_feature_adapters {
    label = "Feature authoring adapters"
    color = "#d6d6d6"
    builtin [label = "built-in indicator\\nledgr_ind_sma()"]
    ttr [label = "TTR adapter\\nledgr_ind_ttr()"]
    custom [label = "custom indicator\\nledgr_indicator()"]
    fmap [label = "ledgr_feature_map"]
    indicators [label = "ordinary ledgr_indicator list"]
  }

  subgraph cluster_runtime_contract {
    label = "Runtime contract"
    color = "#d6d6d6"
    feature_table [label = "feature_table\\nts + instrument + feature_id + value", fillcolor = "#edf7ed"]
    ctx [label = "pulse ctx", fillcolor = "#edf7ed"]
    targets [label = "named numeric targets"]
    events [label = "events"]
  }

  df -> snapshot
  csv -> snapshot
  yahoo -> snapshot
  builtin -> fmap
  ttr -> fmap
  custom -> fmap
  fmap -> indicators
  indicators -> feature_table -> ctx -> targets -> events
}
')
}
```

Review prompt:

```{markdown}
Where is ledgr intentionally flexible for users? Where does it narrow back to a
strict engine contract?
```

### LDG-2210 Bundle Flattening Point

The multi-output bundle UX should live on the authoring side. It should flatten
before feature computation and before the fold core sees feature definitions.

```{r diagram-bundle-flattening}
if (!requireNamespace("DiagrammeR", quietly = TRUE)) {
  message("Install DiagrammeR to run this diagram: install.packages(\"DiagrammeR\")")
} else {
  DiagrammeR::grViz('
digraph bundle_flattening {
  graph [rankdir = LR, bgcolor = "white"]
  node [shape = box, style = "rounded,filled", fillcolor = "#f7f7f7", color = "#555555", fontname = "Arial"]
  edge [color = "#555555", fontname = "Arial"]

  bundle_helper [label = "ledgr_ind_ttr_outputs()\\nauthoring helper"]
  bundle [label = "ledgr_indicator_bundle\\nnot a runtime feature", fillcolor = "#fff6df"]
  flatten [label = "flatten at feature declaration boundary", fillcolor = "#fff6df"]
  indicators [label = "ordinary ledgr_indicator objects\\none per selected output"]
  ids [label = "output-specific feature IDs\\nand fingerprints"]
  engine [label = "existing feature computation\\nand ctx$feature() path", fillcolor = "#edf7ed"]
  bad [label = "forbidden:\\nruntime multi-output feature object", fillcolor = "#ffecec", color = "#aa5555"]

  bundle_helper -> bundle -> flatten -> indicators -> ids -> engine
  bundle -> bad [style = dashed, label = "must not become", color = "#aa5555"]
}
')
}
```

Review prompt:

```{markdown}
After reviewing the code, which function boundary should own bundle flattening?
Why would flattening later be risky?
```

## Setup

This chunk loads the working tree when `pkgload` is available, then defines a
small fixture builder used by every later chunk.

```{r setup}
if (requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(".", quiet = TRUE)
} else {
  library(ledgr)
}

options(width = 100)

make_feature_path_fixture <- function(run = FALSE) {
  data("ledgr_demo_bars", package = "ledgr")

  bars <- ledgr_demo_bars[
    ledgr_demo_bars$instrument_id %in% c("DEMO_01", "DEMO_02") &
      ledgr_demo_bars$ts_utc >= ledgr_utc("2019-01-01") &
      ledgr_demo_bars$ts_utc <= ledgr_utc("2019-03-31"),
  ]

  snapshot <- ledgr_snapshot_from_df(
    bars,
    snapshot_id = paste0("feature-path-review-", Sys.getpid(), "-", sample.int(1e6, 1L))
  )

  sma_feature <- ledgr_ind_sma(20)
  feature_map <- ledgr_feature_map(
    sma = sma_feature
  )

  pulse <- ledgr_pulse_snapshot(
    snapshot = snapshot,
    universe = c("DEMO_01", "DEMO_02"),
    ts_utc = ledgr_utc("2019-02-15"),
    features = feature_map
  )

  observed <- new.env(parent = emptyenv())
  observed$rows <- list()

  strategy <- function(ctx, params) {
    targets <- ctx$flat()

    for (id in ctx$universe) {
      sma <- ctx$feature(id, "sma_20")

      if (length(observed$rows) < 8L) {
        observed$rows[[length(observed$rows) + 1L]] <- data.frame(
          ts_utc = ctx$ts_utc,
          instrument_id = id,
          close = ctx$close(id),
          sma_20 = sma,
          stringsAsFactors = FALSE
        )
      }

      if (is.finite(sma) && ctx$close(id) > sma) {
        targets[id] <- params$qty
      }
    }

    targets
  }

  exp <- ledgr_experiment(
    snapshot = snapshot,
    strategy = strategy,
    features = feature_map,
    opening = ledgr_opening(cash = 10000)
  )

  bt <- NULL
  if (isTRUE(run)) {
    bt <- ledgr_run(
      exp,
      params = list(qty = 10),
      run_id = paste0("feature_path_review_", Sys.getpid(), "_", sample.int(1e6, 1L))
    )
  }

  list(
    bars = bars,
    snapshot = snapshot,
    sma_feature = sma_feature,
    feature_map = feature_map,
    pulse = pulse,
    observed = observed,
    strategy = strategy,
    exp = exp,
    bt = bt
  )
}

close_feature_path_fixture <- function(fx) {
  if (is.list(fx) && inherits(fx$bt, "ledgr_backtest")) {
    close(fx$bt)
  }
  if (is.list(fx) && inherits(fx$pulse, "ledgr_pulse_context")) {
    close(fx$pulse)
  }
  if (is.list(fx) && inherits(fx$snapshot, "ledgr_snapshot")) {
    ledgr_snapshot_close(fx$snapshot)
  }
  invisible(TRUE)
}
```

## Full Workflow Smoke

Run this once first. It proves the review fixture works end to end.

```{r full-workflow-smoke}
fx <- make_feature_path_fixture(run = TRUE)

head(fx$bars)
ledgr_snapshot_info(fx$snapshot)
ledgr_feature_id(fx$feature_map)
fx$pulse$feature("DEMO_01", "sma_20")
do.call(rbind, fx$observed$rows)
summary(fx$bt)
tail(ledgr_results(fx$bt, what = "equity"), 4)

close_feature_path_fixture(fx)
```

Maintainer note:

```{markdown}
What surprised me in the full workflow:
```

## Checkpoint 1: Bars Become A Sealed Snapshot

Run the chunk, then read the listed source files.

```{r snapshot-checkpoint}
fx <- make_feature_path_fixture()

head(fx$bars)
str(fx$bars)
fx$snapshot
ledgr_snapshot_info(fx$snapshot)

close_feature_path_fixture(fx)
```

Read:

- `R/snapshot.R`
- `R/db-schema-create.R`
- any helper opened from `ledgr_snapshot_from_df()`

Review questions:

- Which function accepts the user data frame?
- Where are required bar columns checked?
- Where is the snapshot ID assigned?
- Where is the snapshot sealed?
- What metadata proves this snapshot is the immutable input artifact for later
  execution?

Maintainer notes:

```{markdown}
Write the bars-to-sealed-snapshot path here.
```

Boundary table:

| Boundary | Owner | Input | Output | Invariant |
| --- | --- | --- | --- | --- |
| Bars to snapshot | TODO | TODO | TODO | TODO |
| Snapshot seal | TODO | TODO | TODO | TODO |

## Checkpoint 2: Indicator Becomes A Feature Definition

Run the chunk, then read the listed source files.

```{r feature-definition-checkpoint}
fx <- make_feature_path_fixture()

fx$sma_feature
fx$feature_map
ledgr_feature_id(fx$feature_map)
ledgr_feature_contracts(fx$feature_map)

close_feature_path_fixture(fx)
```

Read:

- `R/indicators.R`
- `R/indicator-ttr.R` if the path opens it
- `R/feature-inspection.R`
- `R/feature-cache.R`

Review questions:

- What does `ledgr_ind_sma(20)` return?
- What is the difference between the raw indicator and `ledgr_feature_map()`?
- Which helper gives the feature its engine ID?
- Which fields define warmup?
- Where is feature fingerprinting implemented?

Maintainer notes:

```{markdown}
Write what an indicator object is in ledgr terms.
```

Boundary table:

| Boundary | Owner | Input | Output | Invariant |
| --- | --- | --- | --- | --- |
| Indicator constructor | TODO | TODO | TODO | TODO |
| Feature map alias | TODO | TODO | TODO | TODO |
| Feature ID/fingerprint | TODO | TODO | TODO | TODO |

## Checkpoint 3: Pulse Inspection Shows Feature Values

Run the chunk, then read the listed source files.

```{r pulse-checkpoint}
fx <- make_feature_path_fixture()

fx$pulse$universe
fx$pulse$close("DEMO_01")
fx$pulse$feature("DEMO_01", "sma_20")
ledgr_pulse_features(fx$pulse, fx$feature_map)
ledgr_pulse_wide(fx$pulse, fx$feature_map)

close_feature_path_fixture(fx)
```

Read:

- `R/pulse-context.R`
- `R/feature-inspection.R`

Review questions:

- What data exists on the pulse before helper functions are called?
- What does the scalar accessor show?
- What does the long feature view show?
- What does the wide feature view show?
- Why does `ledgr_pulse_features(pulse, feature_map)` require a feature map
  when a second argument is supplied?

Maintainer notes:

```{markdown}
Write how pulse inspection relates to strategy execution.
```

Boundary table:

| Boundary | Owner | Input | Output | Invariant |
| --- | --- | --- | --- | --- |
| Pulse snapshot | TODO | TODO | TODO | TODO |
| Scalar feature accessor | TODO | TODO | TODO | TODO |
| Feature inspection views | TODO | TODO | TODO | TODO |

## Checkpoint 4: Experiment Stores The Feature Declaration

Run the chunk, then read the listed source files.

```{r experiment-checkpoint}
fx <- make_feature_path_fixture()

fx$exp
materialized <- ledgr:::ledgr_experiment_materialize_features(
  fx$exp,
  params = list(qty = 10)
)

materialized
ledgr_feature_id(materialized)
ledgr_feature_contracts(materialized)

close_feature_path_fixture(fx)
```

Read:

- `R/experiment.R`
- `ledgr_experiment()`
- `ledgr_experiment_materialize_features()`

Review questions:

- Where does the experiment store the original feature declaration?
- What does materialization do for static features?
- What would materialization do for a feature factory?
- Why does sweep need the materialization boundary per candidate?
- Where should a multi-output bundle flatten so the rest of the engine sees
  ordinary single-output features?

Maintainer notes:

```{markdown}
Write the experiment-to-materialized-features path here.
```

Boundary table:

| Boundary | Owner | Input | Output | Invariant |
| --- | --- | --- | --- | --- |
| Experiment construction | TODO | TODO | TODO | TODO |
| Feature materialization | TODO | TODO | TODO | TODO |
| Bundle flattening target | TODO | TODO | TODO | TODO |

## Checkpoint 5: The Strategy Reads `ctx$feature()`

Run the chunk, then read the listed source files.

```{r strategy-run-checkpoint}
fx <- make_feature_path_fixture(run = TRUE)

do.call(rbind, fx$observed$rows)
tail(ledgr_results(fx$bt, what = "equity"), 4)
head(ledgr_results(fx$bt, what = "ledger"), 6)

close_feature_path_fixture(fx)
```

Read:

- `R/backtest.R`
- `R/backtest-runner.R`
- `R/fold-core.R`
- `R/pulse-context.R`

Review questions:

- Which public function starts the run?
- Where does persistent run orchestration end?
- What is passed into `ledgr_execute_fold()`?
- Where are pulse contexts built?
- Where are feature helpers attached?
- Where are events emitted?

Maintainer notes:

```{markdown}
Write the run path from ledgr_run() to strategy(ctx, params).
```

Boundary table:

| Boundary | Owner | Input | Output | Invariant |
| --- | --- | --- | --- | --- |
| Public run entry | TODO | TODO | TODO | TODO |
| Fold core entry | TODO | TODO | TODO | TODO |
| Strategy call | TODO | TODO | TODO | TODO |

## Checkpoint 6: `ctx$feature()` Internals

Run the chunk to remind yourself of the visible behavior, then read the source.

```{r accessor-checkpoint}
fx <- make_feature_path_fixture()

fx$pulse$feature("DEMO_01", "sma_20")
fx$pulse$feature("DEMO_02", "sma_20")

close_feature_path_fixture(fx)
```

Read:

- `R/pulse-context.R`
- `ledgr_feature_accessor()`
- `ledgr_features_wide()`
- `ledgr_attach_feature_helpers()`

Review questions:

- What input does `ledgr_feature_accessor()` close over?
- How does `ctx$feature(id, feature_id)` find one scalar value?
- What happens when the feature ID is unknown?
- What happens when feature data is absent?
- Does helper construction allocate per pulse?
- Why does that matter for the v0.1.8.3 optimization roadmap?

Maintainer notes:

```{markdown}
Write the accessor implementation in plain language.
```

Boundary table:

| Boundary | Owner | Input | Output | Invariant |
| --- | --- | --- | --- | --- |
| Feature helper attachment | TODO | TODO | TODO | TODO |
| Scalar feature lookup | TODO | TODO | TODO | TODO |
| Missing feature behavior | TODO | TODO | TODO | TODO |

## Checkpoint 7: Tests That Protect This Path

This is a reading checkpoint rather than a workflow chunk.

Read:

- `tests/testthat/test-precompute-features.R`
- `tests/testthat/test-sweep.R`
- `tests/testthat/test-indicator-tools.R`
- `tests/testthat/test-feature-inspection.R`
- any test file that mentions `ctx$feature`

Review questions:

- Which test proves a feature-consuming strategy sees the same value in sweep
  and committed run paths?
- Which test proves feature factories resolve per candidate?
- Which test pins feature IDs or fingerprints?
- Which test proves precomputed features are consumed rather than recomputed?
- Which invariant is not clearly tested?

Maintainer notes:

```{markdown}
Write the test map here.
```

Test map:

| Invariant | Test file | Test name / rough locator |
| --- | --- | --- |
| Feature value reaches strategy | TODO | TODO |
| Feature IDs stable | TODO | TODO |
| Fingerprints stable | TODO | TODO |
| Sweep factory per candidate | TODO | TODO |
| Precomputed feature validation | TODO | TODO |

## Final Trace: Explain It From Memory

Close the source files, then answer this without copying from the source:

> A user writes a strategy that calls `ctx$feature("DEMO_01", "sma_20")`.
> Starting from `ledgr_snapshot_from_df()` and
> `ledgr_experiment(features = ledgr_feature_map(sma = ledgr_ind_sma(20)))`,
> what has to happen before that value is available inside the strategy?

Your rough explanation:

```{markdown}
Write your explanation here.
```

Then reopen the code and correct your explanation:

```{markdown}
Corrections after rereading:
```

## Cleanup

If you interrupted a chunk before it reached `close_feature_path_fixture(fx)`,
run this.

```{r cleanup}
if (exists("fx")) {
  close_feature_path_fixture(fx)
}
```

## Questions / Findings

Use this section for things to ask an agent, turn into a ticket note, or verify
with tests. Do not fix them inside this notebook.

| Question or finding | File/function | Severity | Follow-up |
| --- | --- | --- | --- |
| TODO | TODO | TODO | TODO |

## Review Checklist Before LDG-2202 / LDG-2210

- [ ] I can explain how bars become a sealed snapshot.
- [ ] I can explain where feature declarations are stored.
- [ ] I can explain how params materialize feature factories.
- [ ] I can explain where bundle flattening should happen.
- [ ] I can explain how feature IDs and fingerprints are generated.
- [ ] I can explain what feature table shape reaches the fold core.
- [ ] I can explain how `ctx$feature()` finds a value.
- [ ] I can name the tests that protect feature value lookup.
- [ ] I can name the tests that protect sweep feature factory resolution.

## Optional Appendix: Search Commands

Use these only when direct file reading gets stuck.

```{bash}
rg -n "ledgr_experiment_materialize_features|ledgr_attach_feature|ledgr_features_wide|ledgr_feature_accessor|ctx\\$feature|feature_defs|materialize_features|ledgr_execute_fold" R
```

```{bash}
rg -n "feature_table|features_wide|feature\\(|ctx\\$features|ctx\\$feature" R tests/testthat
```

```{bash}
rg -n "ledgr_resolve_candidate_features|ledgr_resolve_feature_candidates|ledgr_precompute_features|candidate_features|feature_set_hash" R/precompute-features.R R/sweep.R
```
