---
title: "Cash Distributions"
format:
  html:
    toc: true
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{Cash Distributions}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 90)
options(cli.unicode = FALSE)
pkg_root <- if (file.exists("DESCRIPTION")) "." else if (file.exists("../DESCRIPTION")) ".." else NA_character_
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
```

ledgr can post a fully evidenced ordinary cash distribution without enabling
the broader availability model. The research preset uses the vendor-normalized
gross amount and credits it at the ex-dividend session close.

That timing is a model assumption. It prevents a missing-asset dip on the
ex-date, but makes cash spendable before a broker might actually pay it.

## A Minimal Evidenced Dividend

```{r}
#| label: dividend-data

times <- as.POSIXct("2020-01-01 16:00:00", tz = "UTC") + 86400 * 0:3
bars <- data.frame(
  instrument_id = "AAA",
  ts_utc = times,
  open = c(100, 90, 45, 46),
  high = c(100, 90, 45, 46),
  low = c(100, 90, 45, 46),
  close = c(100, 90, 45, 46),
  volume = 1000
)
terms <- data.frame(
  fact_id = "example-dividend",
  subtype = "ordinary_cash_dividend",
  parent_instrument_id = "AAA",
  entitlement_time = times[[2]],
  effective_time = times[[4]],
  knowledge_time = times[[1]],
  complete = TRUE,
  provenance_tier = "snapshot_bound",
  gross_cash_per_parent_unit = 1.25,
  gross_cash_validated = TRUE,
  recipient_identity_validated = FALSE,
  recipient_quantity_validated = FALSE,
  source = "vignette-example"
)

store <- tempfile(fileext = ".duckdb")
snapshot <- ledgr_snapshot_from_df(
  bars,
  facts = ledgr_facts(ledgr_facts_equity_corporate_actions(terms)),
  price_basis = "split_adjusted",
  db_path = store
)
opening <- ledgr_opening(
  cash = 1000,
  positions = c(AAA = 2),
  cost_basis = c(AAA = 100)
)
hold <- function(ctx, params) ctx$positions
```

The later halving of the price is deliberately present. The canonical cash
term is already normalized to the sealed snapshot's unit basis; ledgr does not
reapply a vendor adjustment.

```{r}
#| label: research-preset

research <- ledgr_experiment(
  snapshot,
  hold,
  opening = opening,
  corporate_action_policy = ledgr_corporate_actions_research(),
  cost_model = ledgr_cost_zero()
)
research_run <- ledgr_run(research, run_id = "dividend-research")
summary(research_run)
```

The modeled result reports one `cash_amount.gross` choice, one
`cash_posting.effective_close` choice and gross cash posted of 2.5. It does
not claim a payment date, withholding, investor tax or broker-net amount.

This release does not claim corporate-action completeness, broker-exact
settlement, net cash, tax correctness, or exact recipient exposure.

## Strict Refusal

The strict preset uses the same facts and refuses to model the cash effect.

```{r}
#| label: strict-preset

strict <- ledgr_experiment(
  snapshot,
  hold,
  opening = opening,
  corporate_action_policy = ledgr_corporate_actions_strict(),
  cost_model = ledgr_cost_zero()
)
strict_result <- tryCatch(
  ledgr_run(strict, run_id = "dividend-strict"),
  ledgr_corporate_action_unsupported = function(error) {
    paste("Strict policy refused:", conditionMessage(error))
  }
)
strict_result
```

```{r}
#| label: cleanup
#| include: false

close(research_run)
ledgr_snapshot_close(snapshot)
unlink(store)
```
