---
title: "Cash Distributions"
format:
  html:
    toc: true
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{Cash Distributions}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 90)
options(cli.unicode = FALSE)
pkg_root <- if (file.exists("DESCRIPTION")) "." else if (file.exists("../DESCRIPTION")) ".." else NA_character_
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
library(ledgr)
```

Your strategy holds a stock that pays a dividend. Without corporate-action
evidence the run cannot credit that cash; with it, you choose how and when it
is credited. You need this article when distributions matter to your returns.

ledgr can post a fully evidenced ordinary cash distribution without enabling
the broader availability model. The research preset uses the vendor-normalized
gross amount and credits it at the ex-dividend session close.

That timing is a model assumption. It prevents a missing-asset dip on the
ex-date, but makes cash spendable before a broker might actually pay it.

## What The Print Lines Say

Every run print and summary ends with two lines about this evidence.
`Corporate actions: NOT SUPPLIED` means the snapshot holds no corporate-action
facts, so returns may omit distributions; `MODELED` means supplied evidence was
applied under the settings the summary lists. `Price basis: UNDECLARED` means
the bars do not say how they were adjusted: if they already include dividends
and ledgr also credits them, the same distribution is counted twice. A
declared basis such as `split_adjusted` removes that doubt. Both lines matter
whenever the instruments pay distributions and the claim is about total
return; for learning the mechanics on demo data they can be read and set
aside.

## A Minimal Evidenced Dividend

```{r}
#| label: dividend-data

times <- as.POSIXct("2020-01-01 16:00:00", tz = "UTC") + 86400 * 0:3
bars <- data.frame(
  instrument_id = "AAA",
  ts_utc = times,
  open = c(100, 90, 45, 46),
  high = c(100, 90, 45, 46),
  low = c(100, 90, 45, 46),
  close = c(100, 90, 45, 46),
  volume = 1000
)
terms <- data.frame(
  fact_id = "example-dividend",
  subtype = "ordinary_cash_dividend",
  parent_instrument_id = "AAA",
  entitlement_time = times[[2]],
  effective_time = times[[4]],
  knowledge_time = times[[1]],
  complete = TRUE,
  provenance_tier = "snapshot_bound",
  gross_cash_per_parent_unit = 1.25,
  gross_cash_validated = TRUE,
  recipient_identity_validated = FALSE,
  recipient_quantity_validated = FALSE,
  source = "vignette-example"
)

store <- tempfile(fileext = ".duckdb")
snapshot <- ledgr_snapshot_from_df(
  bars,
  facts = ledgr_facts(ledgr_facts_equity_corporate_actions(terms)),
  price_basis = "split_adjusted",
  db_path = store,
  snapshot_id = "cash_distribution_demo"
)
opening <- ledgr_opening(
  cash = 1000,
  positions = c(AAA = 2),
  cost_basis = c(AAA = 100)
)
hold <- function(ctx, params) ctx$hold()
```

The later halving of the price is deliberately present. The canonical cash
term is already normalized to the sealed snapshot's unit basis; ledgr does not
reapply a vendor adjustment.

```{r}
#| label: research-preset

research <- ledgr_experiment(
  snapshot,
  hold,
  opening = opening,
  corporate_action_policy = ledgr_corporate_actions_research(),
  cost_model = ledgr_cost_zero()
)
research_run <- ledgr_run(research, run_id = "dividend-research")
summary(research_run)
```

The annualized return and Sharpe ratio come from annualizing a few bars and
mean nothing here; the total return and the corporate-action record are the
result. In that record each setting appears twice, as its value and as the
versioned identity that fixes what the value means, and the counts below it
show which choices this run exercised. The modeled result reports one
`cash_amount.gross` choice, one `cash_posting.effective_close` choice and gross
cash posted of 2.5. It does
not claim a payment date, withholding, investor tax or broker-net amount.

This release does not claim corporate-action completeness, broker-exact
settlement, net cash, tax correctness, or exact recipient exposure.

## Strict Refusal

The strict preset uses the same facts and refuses to model the cash effect.

```{r}
#| label: strict-preset

strict <- ledgr_experiment(
  snapshot,
  hold,
  opening = opening,
  corporate_action_policy = ledgr_corporate_actions_strict(),
  cost_model = ledgr_cost_zero()
)
strict_result <- tryCatch(
  ledgr_run(strict, run_id = "dividend-strict"),
  ledgr_corporate_action_unsupported = function(error) {
    paste("Strict policy refused:", conditionMessage(error))
  }
)
strict_result
```

## Where This Fits

This article keeps one local `AAA` holding so the gross cash amount and the
later price change are easy to inspect. The full reusable point-in-time model
is in `vignette("point-in-time-inputs", package = "ledgr")`; use it when the
same study also needs sessions, membership, status, lifetime or quantity
events. A dense study that only needs evidenced cash distributions does not
have to adopt every fact family.

`AAA` is local to this article, not an alias for a shared `DEMO_*` instrument.
When you want to model a dividend on data that also carries sessions,
membership and trading status, load the shared `ledgr_demo_pit_inputs` bundle.
It contains a cash dividend, and this check finds its terms:

```{r}
#| label: shared-cash-dividend

data("ledgr_demo_pit_inputs", package = "ledgr")
shared_cash_case <- subset(
  ledgr_demo_pit_inputs$cases,
  type == "cash_dividend"
)
shared_cash_terms <- subset(
  ledgr_demo_pit_inputs$corporate_actions,
  parent_instrument_id == shared_cash_case$instrument_id[[1L]] &
    as.Date(entitlement_time) == shared_cash_case$date[[1L]]
)
shared_cash_terms[
  , c("subtype", "parent_instrument_id", "gross_cash_per_parent_unit")
]
```

## Where Next

- Read `vignette("survivorship-bias", package = "ledgr")` for what replacing
  historical membership with today's survivors does to a result.
- Read `vignette("corporate-action-adapter-authoring", package = "ledgr")` if
  you are translating vendor evidence into the canonical fact family.
- Read `vignette("point-in-time-inputs", package = "ledgr")` to place this fact
  inside the complete sealed data model.
- Use `ledgr_corporate_action_summary()` when you need the full policy and
  settlement report rather than the compact run print shown above.

```{r}
#| label: cleanup
#| include: false

close(research_run)
ledgr_snapshot_close(snapshot)
unlink(store)
```
