---
title: "Authoring A Corporate-Action Adapter"
format:
  html:
    toc: true
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{Authoring A Corporate-Action Adapter}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 90)
options(cli.unicode = FALSE)
pkg_root <- if (file.exists("DESCRIPTION")) "." else if (file.exists("../DESCRIPTION")) ".." else NA_character_
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
adapter_path <- file.path(pkg_root, "vignettes", "fictional-corporate-action-adapter.R")
source(adapter_path)
```

An adapter has one job: translate a source's vocabulary and units into
ledgr's canonical, vendor-neutral corporate-action facts. It must not make
portfolio policy decisions. Sealing then validates and binds those facts to
the snapshot; execution policy is selected later by the experiment.

The boundary is:

```
private source rows -> adapter -> canonical facts -> sealed snapshot
                                              |
                                              +-> experiment policy -> run
```

## The Canonical Header

Every row needs a stable `fact_id`, a closed `subtype`, and a
`parent_instrument_id`. Four clocks have distinct meanings:

- `entitlement_time` determines which holding is entitled;
- `effective_time` records when the source says the event takes effect;
- `knowledge_time` is the first instant the terms may be used without
  look-ahead; and
- `payment_time` is optional evidence, not the research preset's posting
  clock.

`complete` says whether the required terms are present. An incomplete row
must carry a declared `refusal_reason`; it is evidence of an unsupported
event, not an absent event.

## Terms, Units, And Validation Flags

Cash is expressed as `gross_cash_per_parent_unit`. A security leg uses
`recipient_instrument_id` and `recipient_quantity_per_parent_unit`. Each leg
has its own validation flag. An adapter sets a flag only after it has checked
the source meaning and converted the term to the unit basis of the sealed
bars.

ledgr does not redo a vendor adjustment. The snapshot therefore declares a
`price_basis`. Corporate-action settlement requires `split_adjusted`; bars
that already include distributions are refused because posting cash again
would double count it.

The fictional source below reports its dividend as 2.5 old parent units and
declares a 0.5 parent-unit multiplier for a later split. The canonical term is
therefore 1.25 in the same unit basis as the sealed bars. The source adapter,
not ledgr execution, owns that conversion.

## Provenance

`provenance_tier = "snapshot_bound"` is the portable tier. It binds the
canonical facts to the sealed snapshot without requiring a provider build.
`build_and_vintage_bound` is stronger: it additionally requires
`upstream_build_id` and `bar_vintage_id`. The tier records evidence strength;
it does not select a settlement capability.

Keep the original source locator or record key in `fact_id` or source-owned
lineage before translation. Never place licensed identifiers or raw payloads
in public examples or error messages.

## A Fictional Source

This example intentionally uses source names that do not resemble the
canonical names. That makes the adapter boundary visible instead of merely
copying a data frame.

```{r}
#| label: fictional-source

times <- as.POSIXct("2020-01-01 16:00:00", tz = "UTC") + 86400 * 0:2
records <- data.frame(
  record_key = c("notice-1", "notice-2"),
  effect_code = c("PAYMENT", "CHILD_GRANT"),
  subject_key = c("AAA", "AAA"),
  rights_at = c(times[[2]], times[[3]]),
  changes_at = c(times[[2]], times[[3]]),
  seen_at = c(times[[1]], times[[1]]),
  settles_at = as.POSIXct(c(NA, NA), origin = "1970-01-01", tz = "UTC"),
  terms_ready = c(TRUE, TRUE),
  why_refused = c(NA_character_, NA_character_),
  lineage_level = c("snapshot_bound", "snapshot_bound"),
  source_build = c(NA_character_, NA_character_),
  price_release = c(NA_character_, NA_character_),
  cash_units = c(2.5, NA_real_),
  parent_unit_multiplier = c(0.5, 1),
  cash_checked = c(TRUE, FALSE),
  destination_key = c(NA_character_, "BBB"),
  destination_checked = c(FALSE, TRUE),
  share_units = c(NA_real_, 0.5),
  share_units_checked = c(FALSE, TRUE),
  stringsAsFactors = FALSE
)

facts <- fictional_corporate_action_adapter(records)
facts
```

The adapter owns the mapping from `PAYMENT` and `CHILD_GRANT`, the clock
interpretation, unit normalization, and the validation flags. ledgr owns the
canonical validator. Unknown source codes fail in the adapter; invalid
canonical combinations fail in `ledgr_facts_equity_corporate_actions()`.

## Seal The Result

The physical axis must contain any recipient whose value may be needed later,
even when that recipient is not a member of the strategy universe. Physical
axis closure does not grant membership and does not make the recipient visible
to features before the event.

```{r}
#| label: seal-fictional-facts

bars <- do.call(rbind, lapply(c("AAA", "BBB"), function(id) {
  data.frame(
    instrument_id = id,
    ts_utc = times,
    open = if (id == "AAA") 10:12 else 20:22,
    high = if (id == "AAA") 10:12 else 20:22,
    low = if (id == "AAA") 10:12 else 20:22,
    close = if (id == "AAA") 10:12 else 20:22,
    volume = 1000
  )
}))
store <- tempfile(fileext = ".duckdb")
snapshot <- ledgr_snapshot_from_df(
  bars,
  instruments_df = data.frame(instrument_id = c("AAA", "BBB")),
  facts = ledgr_facts(facts),
  price_basis = "split_adjusted",
  db_path = store
)
info <- ledgr_snapshot_info(snapshot)
c(
  status = info$status[[1]],
  bar_count = info$bar_count[[1]],
  instrument_count = info$instrument_count[[1]]
)
```

The snapshot hash now covers the canonical facts and their declared
provenance. Reopening verifies that identity; a consumer does not need the
private source or adapter to reproduce the sealed input.

## Refusal Is A Result

An adapter should preserve a known event whose terms cannot be validated.
Set `complete = FALSE`, provide the applicable refusal reason, and leave
unvalidated terms unavailable. Do not infer a ratio, fabricate a mark, or
turn an unsupported event into absence.

At execution time, strict policy stops on a relevant unsupported effect.
Research policy may report a quantity effect without changing account state.
Neither path claims exact recipient exposure, broker settlement, withholding,
tax treatment, or completeness of the upstream source.

```{r}
#| label: cleanup
#| include: false

ledgr_snapshot_close(snapshot)
unlink(store)
```
