---
title: "Custom Indicators And External Features"
output:
  rmarkdown::html_vignette:
    toc: false
  github_document:
    toc: false
    html_preview: false
vignette: >
  %\VignetteIndexEntry{Custom Indicators And External Features}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 90)
demo_data_path <- if (file.exists("data/ledgr_demo_bars.rda")) {
  "data/ledgr_demo_bars.rda"
} else {
  file.path("..", "data", "ledgr_demo_bars.rda")
}
if (!exists("ledgr_demo_bars") && file.exists(demo_data_path)) {
  load(demo_data_path)
}
```

```{r, message=FALSE}
library(ledgr)
library(dplyr)
data("ledgr_demo_bars", package = "ledgr")
```

Custom indicators are ledgr's extension point for derived market data. They are
useful when the built-in indicators and TTR-backed indicators do not express
the feature you want.

They are also the highest-risk feature boundary. A custom indicator can keep a
strategy pulse-safe, or it can hide future information in an ordinary-looking
feature value. This article explains the authoring contract.

## The Indicator Object

`ledgr_indicator()` creates a feature definition. The important fields are:

| Field | Meaning |
| --- | --- |
| `id` | Stable feature ID used in `ctx$feature(instrument_id, feature_id)`. |
| `fn` | Scalar function for one bounded historical window. |
| `series_fn` | Optional vectorized function for one instrument's full bar series. |
| `requires_bars` | Minimum lookback requirement for the indicator definition. |
| `stable_after` | First row where the output is considered usable. |
| `params` | Named deterministic parameter list included in the fingerprint. |
| `source` | Source label: `"ledgr"`, `"TTR"`, or `"custom"`. |

Use `params` for intentional configuration. Do not close over mutable session
objects when the value should be part of the feature definition.

## Scalar Indicators

The scalar path is the simplest contract:

```{r scalar-indicator}
range_3 <- ledgr_indicator(
  id = "range_3",
  fn = function(window, params) {
    mean(window$high - window$low)
  },
  requires_bars = 3,
  stable_after = 3,
  params = list()
)
```

The engine calls `fn(window, params)` on a bounded historical window ending at
the current bar. Before `stable_after`, ledgr returns `NA_real_` for that
feature. After warmup, the scalar result must be one finite numeric value.

This path is easy to reason about because the function receives only historical
rows up to the current decision point. It is the right first implementation for
most custom features.

## Vectorized Indicators

`series_fn` is the fast path for indicators that are naturally computed over a
whole series. When both `fn` and `series_fn` are supplied, the engine uses
`series_fn` for full-series feature computation and keeps `fn` as the scalar
definition for the same feature. They should be equivalent after warmup.

```{r vectorized-indicator, eval=FALSE}
sma_3_custom <- ledgr_indicator(
  id = "sma_3_custom",
  fn = function(window, params) {
    mean(utils::tail(window$close, params$n))
  },
  series_fn = function(bars, params) {
    stats::filter(
      bars$close,
      rep(1 / params$n, params$n),
      sides = 1
    ) |>
      as.numeric()
  },
  requires_bars = 3,
  stable_after = 3,
  params = list(n = 3)
)
```

The `series_fn(bars, params)` contract is strict:

- `bars` contains one instrument's bars in ascending timestamp order;
- the return value must be an atomic numeric vector;
- the return length must equal `nrow(bars)`;
- output position `i` belongs to input row `i`;
- warmup rows before `stable_after` are normalized to `NA_real_`;
- post-warmup `NA`, `NaN`, and infinite values are errors.

The example uses `sides = 1`, so row `i` is computed from row `i` and earlier
rows only. That is the causal alignment a vectorized feature must preserve.

Output validation proves shape and value validity. It does not prove causal
correctness. Because `series_fn` receives the full bar series, a badly written
vectorized function can still use future rows internally while returning a
correctly shaped vector.

That is why scalar `fn` is often the safer first version. Add `series_fn` when
the feature logic is stable and the alignment is obvious.

## Warmup And Stability

`requires_bars` and `stable_after` are related but not identical.

`requires_bars` says how much history the indicator definition needs.
`stable_after` says when the output is usable in the feature series. It must be
greater than or equal to `requires_bars`.

For a three-bar moving average, both are usually `3`. For indicators with a
longer settling period, `stable_after` can be larger. ledgr treats rows before
`stable_after` as warmup and exposes them as `NA_real_`.

Warmup `NA` is expected. Post-warmup `NA`, `NaN`, or infinite values mean the
feature did not satisfy its contract.

## Fingerprints

Indicator definitions are fingerprinted so runs can later verify that the
registered feature definition still matches the one recorded with the run.

The fingerprint includes the feature ID, scalar function, vectorized function
when present, `requires_bars`, `stable_after`, and deterministic `params`.

Fingerprints are an identity check, not a semantic proof. They help ledgr answer
"is this the same feature definition?" They do not prove that a custom
`series_fn` avoided lookahead or that an external data source was historically
available at the simulated decision time.

## Deterministic Parameters And Unsafe Calls

`params` must be a named list of deterministic values. Use strings for dates and
timestamps when they are part of the feature definition. Do not pass open
connections, environments, external pointers, or live session objects.

Indicator functions are checked for common unsafe patterns. Examples include:

- global assignment with `<<-`;
- wall-clock calls such as `Sys.time()` and `Sys.Date()`;
- randomness such as `runif()`, `rnorm()`, and `sample()`;
- dynamic lookup and execution helpers such as `get()`, `eval()`, and
  `assign()`;
- environment reads such as `Sys.getenv()`.

These checks are guardrails. They do not replace careful review of the feature
logic.

## Adapter Helpers

`ledgr_adapter_r()` wraps a function that operates on the close series of the
bounded window. It is useful for simple package or base R functions:

```{r adapter-r, eval=FALSE}
median_close <- ledgr_adapter_r(
  stats::median,
  id = "median_close_5",
  requires_bars = 5
)
```

The adapter stores the adapted function identity and arguments in indicator
parameters. It still creates an ordinary `ledgr_indicator`.

`ledgr_adapter_csv()` adapts a CSV of precomputed values:

```{r adapter-csv, eval=FALSE}
csv_indicator <- ledgr_adapter_csv(
  "features/my_values.csv",
  value_col = "my_value",
  id = "my_value"
)
```

The CSV must identify timestamp, instrument, and value columns. This is useful
for external feature pipelines, but it moves availability discipline outside
ledgr. The CSV values must already respect the simulated decision times. ledgr
can hash and look up the values; it cannot know whether the upstream pipeline
used future information.

## Register And Read

Custom indicators are registered with the experiment just like built-in
indicators:

```{r register-read}
bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% c("DEMO_01", "DEMO_02"),
    between(
      ts_utc,
      ledgr_utc("2019-01-01"),
      ledgr_utc("2019-02-28")
    )
  )

snapshot <- ledgr_snapshot_from_df(
  bars,
  snapshot_id = paste0("custom-indicators-", Sys.getpid())
)

features <- list(range_3)

strategy <- function(ctx, params) {
  targets <- ctx$flat()

  for (id in ctx$universe) {
    value <- ctx$feature(id, "range_3")
    if (is.finite(value) && value < params$max_range) {
      targets[id] <- params$qty
    }
  }

  targets
}

exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = strategy,
  features = features,
  opening = ledgr_opening(cash = 10000)
)

ledgr_feature_id(features)
```

Inside the strategy, `ctx$feature(id, "range_3")` reads the exact feature ID
from the pulse context for one instrument. Unknown feature IDs fail loudly.
Warmup for a known feature is represented by `NA_real_`.

```{r custom-cleanup, include=FALSE}
ledgr_snapshot_close(snapshot)
```

## What To Remember

Custom indicators let external feature logic enter ledgr's deterministic pulse
engine. Keep the boundary explicit:

- prefer scalar `fn` until the logic is stable;
- add `series_fn` only when full-series alignment is clear;
- treat `series_fn` and CSV adapters as review points for leakage;
- keep all intentional configuration in deterministic `params`;
- register every feature before `ledgr_run()`;
- use `ledgr_feature_id()` to confirm the exact ID a strategy should read.
