---
title: "Importing And Sealing Market Data"
format:
  html:
    toc: false
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{Importing And Sealing Market Data}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

options(width = 90)
options(cli.unicode = FALSE)
pkg_root <- if (file.exists("DESCRIPTION")) {
  "."
} else if (file.exists("../DESCRIPTION")) {
  ".."
} else {
  NA_character_
}
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
default_output_hook <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(
  output = function(x, options) {
    x <- gsub(
      "Database:(\\s+)[^\\r\\n]+[.]duckdb",
      "Database:\\1<temporary DuckDB path>",
      x,
      perl = TRUE
    )
    x <- gsub("[ \t]+(?=\n)", "", x, perl = TRUE)
    default_output_hook(x, options)
  }
)
```

You have market data in memory or in a file. This article takes the shortest
path from that data to a sealed ledgr snapshot you can inspect and reopen.
Point-in-time facts are a separate concern; add them only when your research
question needs them.

By the end, you will have sealed a small bar panel, inspected its public
metadata, seen the equivalent CSV and Yahoo entry points, and made one explicit
decision about malformed observations.

::: {.ledgr-callout .ledgr-callout-note}
**Running this yourself**

The runnable examples use temporary DuckDB files. In real work, choose a
project path such as `artifacts/ledgr_store.duckdb`.
:::

```{r}
#| label: library
#| message: false

library(ledgr)
library(dplyr)
data("ledgr_demo_bars", package = "ledgr")
```

## The Required Bar Shape

Bars are the only required input. Each row identifies one physical instrument
at one UTC observation time and supplies open, high, low, and close. Volume is
optional.

```{r}
#| label: bar-shape

bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% c("DEMO_01", "DEMO_02"),
    between(ts_utc, ledgr_utc("2019-01-01"), ledgr_utc("2019-01-31"))
  )

bars |>
  select(instrument_id, ts_utc, open, high, low, close, volume) |>
  slice_head(n = 4)
```

The `(instrument_id, ts_utc)` pair must be unique. Timestamps are whole-second
UTC instants. If your source uses local exchange times, normalize them before
calling ledgr rather than leaving the timezone implicit.

## Seal A Snapshot

`ledgr_snapshot_from_df()` validates, stores, and seals the data in one call.

```{r}
#| label: seal-snapshot

db_path <- ledgr_temp_store(file.path(tempdir(), "ledgr_import_demo.duckdb"))

snapshot <- ledgr_snapshot_from_df(
  bars,
  db_path = db_path,
  snapshot_id = "import_demo"
)
```

The returned handle is already sealed. Inspect the stable public fields rather
than the raw metadata envelope.

```{r}
#| label: inspect-snapshot

ledgr_snapshot_info(snapshot) |>
  select(
    status,
    snapshot_hash,
    bar_count,
    instrument_count,
    start_date,
    end_date
  )
```

Sealing gives you validated bar keys, an immutable snapshot hash, and a stable
physical instrument master. It does not invent missing sessions, infer
historical universe membership, or repair prices.

If the source evidence changes, create a new snapshot. Do not append bars to a
sealed snapshot or reseal different data under the same identity.

## Choose The Smallest Honest Input Set

Most users should start with bars and add facts only when the question requires
them.

| Research shape | Minimum sealed inputs | Read next |
|---|---|---|
| Dense static panel | Bars | Continue below |
| Described physical panel | Bars plus instruments | Constructor help |
| Session-aware gaps | Bars plus sessions | Point-in-time inputs |
| Point-in-time universe | Bars, sessions, membership | Point-in-time inputs |
| Trading or lifetime restrictions | Bars, sessions, status or lifetime | Point-in-time inputs |
| Equity economic events | Bars, instruments, corporate actions | Point-in-time inputs |

Use `vignette("point-in-time-inputs", package = "ledgr")` when bars alone
cannot say what existed, what was eligible, or what was known. That article
owns the complete data dictionary and entity diagram.

## Start From A CSV Instead

`ledgr_snapshot_from_csv()` has the same snapshot contract as the data-frame
form. It requires `instrument_id`, `ts_utc`, `open`, `high`, `low`, and
`close`; `volume` is optional.

```{r}
#| label: csv-snapshot
#| eval: false

snapshot <- ledgr_snapshot_from_csv(
  "data/daily_bars.csv",
  db_path = "artifacts/ledgr_store.duckdb",
  snapshot_id = "eod_2019_h1"
)
```

Supply a separate instrument file when you have stable metadata. It needs an
`instrument_id`; `symbol`, `currency`, `asset_class`, `multiplier`, and
`tick_size` are optional.

```{r}
#| label: csv-with-instruments
#| eval: false

snapshot <- ledgr_snapshot_from_csv(
  "data/daily_bars.csv",
  instruments_csv_path = "data/instruments.csv",
  db_path = "artifacts/ledgr_store.duckdb",
  snapshot_id = "eod_2019_h1"
)
```

## Or Fetch Yahoo Data

The Yahoo adapter downloads bars through `quantmod` and seals the result.

```{r}
#| label: yahoo-snapshot
#| eval: false

snapshot <- ledgr_snapshot_from_yahoo(
  symbols = c("SPY", "QQQ"),
  from = "2019-01-01",
  to = "2019-06-30",
  db_path = "artifacts/ledgr_store.duckdb",
  snapshot_id = "yahoo_2019_h1"
)
```

::: {.ledgr-callout .ledgr-callout-warning}
**Yahoo is a convenience source**

Yahoo and the remote endpoint are outside ledgr's reproducibility boundary.
The adapter seals the OHLCV values returned by `quantmod`; it does not rebuild
adjusted OHLC bars from adjusted close. Prepare and seal your own panel when
the research requires a different adjustment policy.
:::

## Refuse Or Quarantine A Bad Observation

A malformed bar is not a missing bar. ledgr refuses it by default rather than
guess what it meant. This row has a high below its open.

```{r}
#| label: invalid-bars

session_dates <- as.Date("2019-01-01") + 0:4
invalid_bars <- tibble(
  instrument_id = "IMPORT_01",
  ts_utc = session_dates,
  open = 100:104,
  high = 101:105,
  low = 99:103,
  close = 100:104,
  volume = 1000
)
invalid_bars$high[[3]] <- invalid_bars$open[[3]] - 5
```

Quarantine is meaningful only when an independent calendar says the
observation was expected. Otherwise a dropped row is indistinguishable from a
date that never existed.

```{r}
#| label: quarantine-sessions

session_facts <- ledgr_facts_sessions(
  tibble(
    session_date = session_dates,
    status = "open",
    session_open = "14:30:00",
    session_close = "21:00:00",
    knowledge_time = ledgr_utc("2018-12-01")
  ),
  venue_id = "IMPORT_VENUE",
  timezone = "UTC"
)

strict <- tryCatch(
  ledgr_snapshot_from_df(
    invalid_bars,
    instruments_df = tibble(instrument_id = "IMPORT_01"),
    facts = ledgr_facts(session_facts),
    db_path = ledgr_temp_store(file.path(tempdir(), "ledgr_strict.duckdb"))
  ),
  error = identity
)

conditionMessage(strict)
```

Nothing was sealed. If you have decided that the source row should remain in
the audit record but must not reach runtime, ask for quarantine explicitly.

```{r}
#| label: quarantine-snapshot

quarantined <- ledgr_snapshot_from_df(
  invalid_bars,
  instruments_df = tibble(instrument_id = "IMPORT_01"),
  facts = ledgr_facts(session_facts),
  db_path = ledgr_temp_store(file.path(tempdir(), "ledgr_quarantine.duckdb")),
  invalid_observations = "quarantine"
)

ledgr_snapshot_info(quarantined)$bar_count
```

Four of five observations reached runtime. The invalid source row remains
bound into the snapshot evidence; quarantine is not deletion or repair.

## Reopen The Same Artifact

Record the path and snapshot id, close the live handle, and reopen with hash
verification in a later session.

```{r}
#| label: reopen-snapshot

snapshot_id <- snapshot$snapshot_id
ledgr_snapshot_close(snapshot)

snapshot <- ledgr_snapshot_open(
  db_path,
  snapshot_id = snapshot_id,
  verify = TRUE
)

ledgr_snapshot_info(snapshot) |>
  select(status, snapshot_hash, bar_count)
```

For backup, recovery, live handles, and stored runs, continue with
`vignette("experiment-store", package = "ledgr")`.

## Cleanup

```{r}
#| label: cleanup

ledgr_snapshot_close(quarantined)
ledgr_snapshot_close(snapshot)
```

## Where Next

- `vignette("point-in-time-inputs", package = "ledgr")` adds sessions,
  membership, trading status, lifetime, and economic events.
- `vignette("strategy-development", package = "ledgr")` uses a sealed
  snapshot in a complete backtest.
- `vignette("experiment-store", package = "ledgr")` covers stored runs,
  handles, backup, and recovery.
