---
title: "Data Model And Point-In-Time Inputs"
format:
  html:
    toc: false
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{Data Model And Point-In-Time Inputs}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

options(width = 90)
options(cli.unicode = FALSE)
# Local renders load the package from source via pkgload (working dir is either
# project root or vignettes/); installed builds fall through to library(ledgr).
pkg_root <- if (file.exists("DESCRIPTION")) "." else if (file.exists("../DESCRIPTION")) ".." else NA_character_
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
default_output_hook <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(
  output = function(x, options) {
    x <- gsub("Database:(\\s+)[^\\r\\n]+[.]duckdb", "Database:\\1<temporary DuckDB path>", x, perl = TRUE)
    x <- gsub("[ \t]+(?=\n)", "", x, perl = TRUE)
    default_output_hook(x, options)
  }
)
```

This article focuses on bringing data into ledgr and sealing it into a
snapshot. Runs, labels, reopening, and recovery evidence live in
`vignette("experiment-store", package = "ledgr")`.

::: {.ledgr-callout .ledgr-callout-note}
**Running this yourself**

This article is evaluated when rendered. It writes to temporary DuckDB stores
so package builds and local previews do not leave project artifacts behind. In
real work, use a project-local path such as `artifacts/ledgr_store.duckdb`.
:::

::: {.ledgr-callout .ledgr-callout-warning}
**Pre-CRAN compatibility**

ledgr is pre-CRAN. Store schemas, config hashes, provenance formats, and
experimental APIs may change before the first CRAN release. Treat stores
created with pre-CRAN ledgr as research artifacts for the version that produced
them, and expect to rerun experiments after upgrading.
:::

The examples use `dplyr` for data preparation and compact display. It is a
suggested package used by the vignettes, not part of the experiment-store
contract.

```{r}
#| label: library
#| message: false

library(ledgr)
library(dplyr)
data("ledgr_demo_bars", package = "ledgr")
data("ledgr_demo_pit_inputs", package = "ledgr")
```

## Snapshot Lifecycle And Data Input

Market data and derived data have different lifecycle rules in ledgr. A sealed
snapshot freezes the real market-data input and its hash. If you need more
instruments, more dates, corrected bars, or tick-derived bars, create a new
snapshot. Indicators, runs, labels, tags, comparisons, and telemetry are derived
from sealed market data and can be added later without mutating the snapshot.

Snapshot lifecycle anti-patterns:

- appending bars to a sealed snapshot in place;
- resealing different data under the same snapshot ID;
- deleting snapshots that stored runs still reference;
- mixing live ticks into a backtest snapshot;
- filling data gaps with undocumented synthetic corrections.

If the evidence changes, create a new snapshot. That is what makes later
comparison meaningful.

This vignette uses `ledgr_temp_store()` so it can run without writing into your
project directory. For real research, use `artifacts/ledgr_store.duckdb` and a
snapshot ID you will recognize later.

```{r}
#| label: snapshot

db_path <- ledgr_temp_store(file.path(tempdir(), "ledgr_store_demo.duckdb"))

bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% c("DEMO_01", "DEMO_02"),
    between(
      ts_utc,
      ledgr_utc("2019-01-01"),
      ledgr_utc("2019-06-30")
    )
  )

snapshot <- ledgr_snapshot_from_df(
  bars,
  db_path = db_path,
  snapshot_id = "store_demo_snapshot"
)
```

## Data Model And Point-In-Time Decision Map

Bars are the only required input. Every other table makes a claim that bars
alone cannot make: what instruments mean, which dates are sessions, which
instruments belonged to a research universe, whether they could trade or still
existed, and which evidenced economic events occurred.

### Choose The Smallest Honest Input Set

| Research shape | Minimum sealed inputs | Runtime declaration |
|---|---|---|
| Dense static panel | Bars | A strategy and ordinary experiment settings |
| Described physical panel | Bars plus instruments | Same as dense; metadata becomes inspectable |
| Session-aware data gaps | Bars plus sessions | A stale-mark valuation policy |
| Point-in-time universe | Bars, sessions and one membership shape | A matching universe selector and stale-mark policy |
| Tradeability or lifetime restrictions | Bars, sessions, and status or lifetime facts | A stale-mark policy; membership only when selection needs it |
| Demonstrated full equity bundle | Bars, instruments, sessions, complete membership, status, lifetime and corporate actions | Universe, valuation and settlement policy choices |

An availability family currently requires a complete declared session calendar
and a stale-mark policy. Corporate-action facts do not activate availability by
themselves. Adding more fact families is not automatically more honest: include
only evidence that your source can support.

### Rows, Keys And Required Columns

This table is the package-level data dictionary. Constructor help remains the
versioned authority for exhaustive optional columns and accepted values.

| Input | Row grain and identity | Required core columns or shape | Scope and time | Constructor help |
|---|---|---|---|---|
| Bars | One row per `(instrument_id, ts_utc)`; the pair is unique | `instrument_id`, `ts_utc`, `open`, `high`, `low`, `close`; `volume` optional | Physical instrument; observation clock | [`ledgr_snapshot_from_df()`](../reference/ledgr_snapshot_from_df.html) |
| Instruments | One row per unique `instrument_id` | `instrument_id`; symbol, currency, asset class, multiplier and tick size are optional | Physical master; identity is stable, not time-varying | [`ledgr_snapshot_from_df()`](../reference/ledgr_snapshot_from_df.html) |
| Sessions | One row per civil date in one declared venue range | `session_date`, `status`; open rows also need `session_open`, `session_close` | `venue_id`; knowledge clock plus local calendar time | [`ledgr_facts_sessions()`](../reference/ledgr_facts.html) |
| Membership intervals | One assertion per instrument interval | `instrument_id`, `effective_from`, `member`; optional `effective_to` | `universe_id`; half-open effective interval plus knowledge clock | [`ledgr_facts_membership_intervals()`](../reference/ledgr_facts.html) |
| Complete membership snapshots | One set header per effective/knowledge/source group plus member rows | `effective_from` and row-wise `instrument_id` or one `members` list-column; `complete` is declared | `universe_id`; omission means non-membership only in a complete set | [`ledgr_facts_membership_snapshots()`](../reference/ledgr_facts.html) |
| Trading status | One sourced assertion per instrument interval and precedence | `instrument_id`, `effective_from`, `status`, `source`; optional `effective_to`, `precedence` and supersession fields | Physical instrument; effective and knowledge clocks | [`ledgr_facts_trading_status()`](../reference/ledgr_facts.html) |
| Lifetime | One assertion per instrument interval | `instrument_id`, `effective_from`, `assertion`; optional `effective_to`, `terminal_event` | Physical instrument; effective and knowledge clocks | [`ledgr_facts_lifetime()`](../reference/ledgr_facts.html) |
| Equity corporate actions | One identified economic fact | `subtype`, `parent_instrument_id`, `complete`, `provenance_tier`; terms and refusal reason are conditional | Physical parent and optional recipient; four distinct clocks | [`ledgr_facts_equity_corporate_actions()`](../reference/ledgr_facts.html) |

Every interval is half-open: `effective_from` is included and `effective_to` is
excluded. Three cross-cutting rules apply:

- identifiers stay stable for the history being sealed;
- timestamps are whole-second UTC instants, even when sessions begin as local
  civil dates and wall times;
- knowledge is declared per family. It is not universally required to precede
  effectiveness: status and lifetime evidence may arrive late, while session
  facts obey their stricter constructor rule.

Corporate actions distinguish entitlement (who qualifies), effective time
(when the market state changes), knowledge time (when the evidence is usable),
and payment time (when cash is due). Do not collapse those clocks into one
vendor date during preparation.

### Entity Relationships

The physical instrument master, venue calendar and research universe are
different scopes. The diagram shows keys and cardinalities, not the order in
which helpers happen to be called.

::: {.ledgr-diagram .ledgr-input-entities}

```{mermaid}
%%| echo: false
erDiagram
  SNAPSHOT ||--o{ INSTRUMENT : contains
  INSTRUMENT ||--o{ BAR : observed_as
  SNAPSHOT ||--o{ SESSION : seals
  SNAPSHOT ||--o{ MEMBERSHIP : seals
  INSTRUMENT ||--o{ MEMBERSHIP : referenced_by
  INSTRUMENT ||--o{ TRADING_STATUS : constrained_by
  INSTRUMENT ||--o{ LIFETIME : described_by
  INSTRUMENT ||--o{ CORPORATE_ACTION : parent_or_recipient
  SNAPSHOT {
    string snapshot_id PK
    string snapshot_hash
  }
  INSTRUMENT {
    string instrument_id PK
  }
  BAR {
    string instrument_id FK
    datetime ts_utc PK
  }
  SESSION {
    string venue_id PK
    date session_date PK
  }
  MEMBERSHIP {
    string universe_id PK
    string instrument_id FK
    datetime effective_from
    datetime knowledge_time
  }
  TRADING_STATUS {
    string instrument_id FK
    datetime effective_from
    datetime knowledge_time
  }
  LIFETIME {
    string instrument_id FK
    datetime effective_from
    datetime knowledge_time
  }
  CORPORATE_ACTION {
    string fact_id PK
    string parent_instrument_id FK
    string recipient_instrument_id FK
  }
```

:::

Corporate-action parent and optional recipient identifiers resolve against the
physical instrument master, not against membership in a research universe. A
security can therefore be physically known without being an eligible member.
Trading-status and lifetime facts also resolve through that master; neither is
a substitute for universe membership.

### What Sealing Establishes

After a successful seal, ledgr can rely on stable physical identifiers,
whole-second bar timestamps, unique bar keys, fact referential integrity,
validated family structure, and an immutable snapshot hash. Reopening with
`verify = TRUE` rechecks that artifact identity.

Sealing does not imply a dense instrument-by-session rectangle, current
universe membership, continuously fresh prices, inferred corporate-action
terms, or knowledge before effectiveness. Those remain explicit data or policy
questions at execution time.

### One Composable Bundle

`ledgr_demo_pit_inputs` is plain data, not a prebuilt snapshot. Its synthetic
`DEMO_VENUE` calendar uses the `America/New_York` timezone. Matching
session-close bars, the instrument master and five fact inputs travel together;
they are not a fact overlay for an unrelated observation panel. The case
manifest locates a weekday closure, an open-session observation gap, a
delisting, a late-known halt and a cash dividend.

```{r}
#| label: pit-input-shape

pit <- ledgr_demo_pit_inputs
names(pit)
pit$cases
```

The recipe carries the scope and knowledge choices needed by the public
constructors. The small `constructor_args()` helper removes only the recipe's
inspectable `enabled` marker; it does not rewrite a fact row.

```{r}
#| label: pit-facts

constructor_args <- function(name) {
  args <- pit$recipe$constructors[[name]]
  args[names(args) != "enabled"]
}

session_facts <- do.call(
  ledgr_facts_sessions,
  c(list(df = pit$sessions), constructor_args("sessions"))
)
membership_facts <- do.call(
  ledgr_facts_membership_snapshots,
  c(list(df = pit$membership), constructor_args("membership"))
)
lifetime_facts <- do.call(
  ledgr_facts_lifetime,
  c(list(df = pit$lifetime), constructor_args("lifetime"))
)
status_facts <- do.call(
  ledgr_facts_trading_status,
  c(list(df = pit$trading_status), constructor_args("trading_status"))
)
corporate_action_facts <- ledgr_facts_equity_corporate_actions(
  pit$corporate_actions
)
pit_facts <- ledgr_facts(
  session_facts,
  membership_facts,
  lifetime_facts,
  status_facts,
  corporate_action_facts
)
```

The bundle makes three important boundaries executable. Its halt is already
effective before it becomes knowable. Sub-second observations fail before
sealing, and a fact that names an instrument outside the physical master fails
the public pre-seal validation with a classed error. The seal independently
retains its lower-level `ledgr_snapshot_fact_referential_integrity` guard; the
public constructor normally refuses the bad bundle before reaching it.

```{r}
#| label: pit-boundary-failures

halt_clock <- pit$trading_status[
  pit$trading_status$status == "halted",
  c("effective_from", "knowledge_time")
]
data.frame(
  boundary = "halt knowledge arrives after effectiveness",
  observed = halt_clock$knowledge_time > halt_clock$effective_from
)

capture_condition_class <- function(value) {
  result <- tryCatch(force(value), error = identity)
  if (inherits(result, "condition")) return(class(result)[[1L]])
  if (inherits(result, "ledgr_snapshot")) ledgr_snapshot_close(result)
  "NO_ERROR"
}

subsecond_bars <- pit$bars[1:2, , drop = FALSE]
subsecond_bars$ts_utc[[1L]] <- subsecond_bars$ts_utc[[1L]] + 0.5

orphan_lifetime <- pit$lifetime
orphan_lifetime$instrument_id[[1L]] <- "NOT_IN_MASTER"
orphan_lifetime_facts <- do.call(
  ledgr_facts_lifetime,
  c(list(df = orphan_lifetime), constructor_args("lifetime"))
)
orphan_facts <- ledgr_facts(
  session_facts,
  membership_facts,
  orphan_lifetime_facts,
  status_facts,
  corporate_action_facts
)

data.frame(
  boundary = c("sub-second bar", "fact outside instrument master"),
  condition_class = c(
    capture_condition_class(ledgr_snapshot_from_df(subsecond_bars)),
    capture_condition_class(ledgr_snapshot_from_df(
      pit$bars,
      instruments_df = pit$instruments,
      facts = orphan_facts,
      price_basis = pit$recipe$price_basis
    ))
  )
)
```

Pass those objects through the ordinary public workflow without filtering rows
or rewriting timestamps.

```{r}
#| label: pit-snapshot-run

pit_path <- ledgr_temp_store(file.path(tempdir(), "ledgr_pit_demo.duckdb"))
pit_snapshot <- ledgr_snapshot_from_df(
  pit$bars,
  instruments_df = pit$instruments,
  facts = pit_facts,
  db_path = pit_path,
  snapshot_id = "pit_demo_snapshot",
  price_basis = pit$recipe$price_basis
)
pit_snapshot_id <- pit_snapshot$snapshot_id
ledgr_snapshot_close(pit_snapshot)

pit_snapshot <- ledgr_snapshot_open(
  pit_path,
  pit_snapshot_id,
  verify = TRUE
)
flat_strategy <- function(ctx, params) ctx$flat()
pit_experiment <- ledgr_experiment(
  pit_snapshot,
  flat_strategy,
  universe = ledgr_universe_members(pit$recipe$scopes$universe_id),
  valuation_policy = ledgr_valuation_stale(2),
  cost_model = ledgr_cost_zero()
)
pit_run <- ledgr_run(pit_experiment, run_id = "pit-demo-run")
pit_completion <- ledgr_run_completion(pit_run)
unlist(pit_completion[c("completion_status", "complete_performance")])
```

The verified reopen proves the sealed artifact can stand on its own. The
`DONE` result proves the same bundle is executable; it does not imply that its
flat strategy exercised every teaching event economically.

After snapshot creation, store operations take `snapshot`, not `db_path`. In a
new R session, recover the handle with `ledgr_snapshot_open(db_path,
snapshot_id)`.

If your market data starts in CSV, seal the CSV into the same kind of durable
store. `ledgr_snapshot_from_csv()` is the file-input form of
`ledgr_snapshot_from_df()`: one contract, two input types. Both require
`instrument_id`, `ts_utc`, `open`, `high`, `low`, and `close`, with `volume`
optional. ledgr imports only those canonical bar columns. Other columns are
ignored and do not become part of the sealed snapshot or its hash. `ts_utc`
accepts a date-only value, an ISO 8601 UTC datetime with a trailing `Z`, and
the same datetime without the `Z`.

```{r}
#| label: csv-snapshot
#| eval: false

snapshot <- ledgr_snapshot_from_csv(
  "data/daily_bars.csv",
  db_path = "artifacts/ledgr_store.duckdb",
  snapshot_id = "eod_2019_h1"
)
```

Instrument metadata and point-in-time facts take the same arguments they take
on the data-frame adapter. Supply instruments as a second file with
`instrument_id` and the optional `symbol`, `currency`, `asset_class`,
`multiplier`, and `tick_size` columns; without it, ledgr generates instruments
from the ids the bars carry.

```{r}
#| label: csv-snapshot-instruments
#| eval: false

snapshot <- ledgr_snapshot_from_csv(
  "data/daily_bars.csv",
  instruments_csv_path = "data/instruments.csv",
  db_path = "artifacts/ledgr_store.duckdb",
  snapshot_id = "eod_2019_h1"
)
```

In any later session, recover the handle without re-sealing the data:

```{r}
#| label: csv-snapshot-load
#| eval: false

snapshot <- ledgr_snapshot_open(
  "artifacts/ledgr_store.duckdb",
  snapshot_id = "eod_2019_h1"
)
```

CSV and local data validation happens while the snapshot is created and sealed,
before a strategy can run. Missing columns, unparseable timestamps, duplicate
`instrument_id`/`ts_utc` rows, and OHLC violations are snapshot import problems.
They are not strategy execution errors.

Yahoo imports follow the same lifecycle, but the adapter downloads bars before
sealing the snapshot:

```{r}
#| label: yahoo-snapshot
#| eval: false

snapshot <- ledgr_snapshot_from_yahoo(
  symbols = c("SPY", "QQQ"),
  from = "2019-01-01",
  to = "2019-06-30",
  db_path = "artifacts/ledgr_store.duckdb",
  snapshot_id = "yahoo_2019_h1"
)
```

The returned handle is already sealed. Calling `ledgr_snapshot_seal(snapshot)`
again is an idempotent verification step: on a snapshot handle it returns an
invisible structured list with `$hash` and `$snapshot`; on a low-level DBI
connection plus `snapshot_id` it returns the hash string. Use
`ledgr_snapshot_info(snapshot)` to inspect `status`, `snapshot_hash`,
`bar_count`, `instrument_count`,
`start_date`, `end_date`, and raw `meta_json`. The dates are ISO UTC values.
`meta_json` is envelope metadata; snapshot identity comes from normalized bars
and instruments, not from human descriptions.

::: {.ledgr-callout .ledgr-callout-warning}
**Yahoo data boundary**

Yahoo support is a convenience adapter, not a data-vendor guarantee. It uses
`quantmod::getSymbols()` and therefore requires the suggested `quantmod`
package and network access. Package startup or S3 method-overwrite messages
printed while quantmod loads are not ledgr snapshot warnings. The adapter seals
the Yahoo `.Open`, `.High`, `.Low`, `.Close`, and `.Volume` columns as returned
by quantmod; it does not rewrite OHLC values from Yahoo's adjusted-close
column. If your research requires split/dividend-adjusted OHLC bars, prepare
those bars explicitly and seal them with `ledgr_snapshot_from_df()` or
`ledgr_snapshot_from_csv()`.
:::

```{r}
#| label: yahoo-seal-check
#| eval: false

yahoo_info <- ledgr_snapshot_info(snapshot)
yahoo_seal <- ledgr_snapshot_seal(snapshot)
yahoo_hash <- yahoo_seal$hash
stopifnot(identical(yahoo_info$snapshot_hash[[1]], yahoo_hash))
```

Snapshot metadata uses these public field names:

| Field | Meaning |
|---|---|
| `status` | snapshot lifecycle state, usually `SEALED` after helper creation |
| `snapshot_hash` | hash of normalized bars and instruments |
| `bar_count` | current count of rows in `snapshot_bars` |
| `instrument_count` | current count of rows in `snapshot_instruments` |
| `start_date`, `end_date` | seal-time date range parsed from metadata |
| `meta_json` | raw JSON envelope containing user metadata plus seal metadata |

Seal metadata inside `meta_json` may use internal names such as `n_bars` and
`n_instruments`. The structured columns from `ledgr_snapshot_info()` are
`bar_count` and `instrument_count`; use those names in programmatic code.

## Invalid Observations: Refuse First, Quarantine Deliberately

A malformed bar is not a data gap. ledgr refuses to seal one by default rather
than guess what it meant. The alternative is available, but you have to ask for
it by name.

This example corrupts one bar so its high sits below its open.

```{r}
#| label: invalid-bars

session_dates <- as.Date("2019-01-01") + 0:4

invalid_bars <- tibble(
  instrument_id = "DEMO_01",
  ts_utc = session_dates,
  open = 100:104,
  high = 101:105,
  low = 99:103,
  close = 100:104,
  volume = 1000
)
invalid_bars$high[[3]] <- invalid_bars$open[[3]] - 5

invalid_bars
```

Quarantine is a decision about an *expected* observation, so it requires a
declared session calendar. Without one there is no assertion that the session
happened, and a dropped row would be indistinguishable from a date that never
existed.

```{r}
#| label: quarantine-sessions

session_facts <- ledgr_facts_sessions(
  tibble(
    session_date = session_dates,
    status = "open",
    session_open = "14:30:00",
    session_close = "21:00:00",
    knowledge_time = ledgr_utc("2018-12-01")
  ),
  venue_id = "DEMO",
  timezone = "UTC"
)
```

The default refuses the entire seal:

```{r}
#| label: strict-refusal

strict <- tryCatch(
  ledgr_snapshot_from_df(
    invalid_bars,
    instruments_df = tibble(instrument_id = "DEMO_01"),
    facts = ledgr_facts(session_facts),
    db_path = ledgr_temp_store(file.path(tempdir(), "ledgr_strict_demo.duckdb")),
    snapshot_id = "strict_demo"
  ),
  error = function(e) e
)

class(strict)[1:2]
conditionMessage(strict)
```

Nothing was sealed. One bad row stopped the whole snapshot, which is the point:
you find out before the evidence is frozen, not afterwards. To proceed you must
say so explicitly.

```{r}
#| label: quarantine-seal

quarantined <- ledgr_snapshot_from_df(
  invalid_bars,
  instruments_df = tibble(instrument_id = "DEMO_01"),
  facts = ledgr_facts(session_facts),
  db_path = ledgr_temp_store(
    file.path(tempdir(), "ledgr_quarantine_demo.duckdb")
  ),
  snapshot_id = "quarantine_demo",
  invalid_observations = "quarantine"
)

ledgr_snapshot_info(quarantined, "quarantine_demo")$bar_count
```

Four bars of five reached runtime. The fifth was not deleted: it is hashed as
part of the sealed evidence, so the snapshot still records what the source
actually supplied. The expected session stays on the calendar, which is what
keeps the gap visible instead of silently shortening the history.

Quarantine is a statement about one observation, not a repair. It interpolates
no price and asserts nothing about whether the instrument was tradeable.
Correcting the data means a new snapshot, as it does for every other change to
sealed evidence.

```{r}
#| label: quarantine-cleanup

ledgr_snapshot_close(quarantined)
```

## Backup Conventions

The store is an ordinary DuckDB file. Back it up when no ledgr process has it
open.

::: {.ledgr-callout .ledgr-callout-warning}
**Back up closed stores**

Close run and snapshot handles, then copy or sync the closed store file. A
simple project pattern is:

```r
dir.create("backups", showWarnings = FALSE)
file.copy(
  "artifacts/ledgr_store.duckdb",
  file.path("backups", paste0("ledgr_store_", Sys.Date(), ".duckdb")),
  overwrite = TRUE
)
```

For larger projects, use the same closed-file rule with your normal backup or
sync tool. Do not rely on the phrase "ordinary backup discipline" without a
specific copy/sync pattern for the store file.
:::


## Cleanup

```{r cleanup}
close(pit_run)
ledgr_snapshot_close(pit_snapshot)
ledgr_snapshot_close(snapshot)
```

## Where Next

- `vignette("survivorship-bias", package = "ledgr")` shows how membership,
  sessions, and stable instrument identity become one point-in-time workflow.
- `vignette("experiment-store", package = "ledgr")` shows how sealed
  snapshots are used by committed runs and recovery workflows.
- `vignette("research-workflow", package = "ledgr")` puts snapshots in the
  larger research project workflow.
- `vignette("strategy-development", package = "ledgr")` uses a sealed
  snapshot in a complete backtest.
