---
title: "Execution Semantics"
format:
  html:
    toc: false
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{Execution Semantics}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

options(width = 90)
options(cli.unicode = FALSE)
pkg_root <- if (file.exists("DESCRIPTION")) "." else if (file.exists("../DESCRIPTION")) ".." else NA_character_
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
```

```{=html}
<style>
.ledgr-diagram {
  margin: 1.25rem auto 1.5rem auto;
  text-align: center;
}

.ledgr-diagram .mermaid {
  display: inline-block;
  max-width: 100%;
}

.ledgr-diagram .mermaid svg {
  display: block;
  height: auto !important;
  margin-left: auto;
  margin-right: auto;
}

.ledgr-execution-diagram .mermaid svg {
  max-width: 920px !important;
}
</style>
```

This article explains what happens after a strategy returns target holdings.
It is about fill timing and ledger state, not strategy design or validation.

## Prerequisites

```{r}
#| label: prerequisites

library(ledgr)
library(dplyr)
data("ledgr_demo_bars", package = "ledgr")
```

## The Short Version

ledgr strategies emit desired holdings at each pulse. They do not emit orders.
The fold core compares the desired target to current state, creates the required
position delta, and fills that delta at the next bar's open.

::: {.ledgr-diagram .ledgr-execution-diagram}

```{mermaid}
%%| echo: false
%%{init: {"theme": "base", "flowchart": {"nodeSpacing": 22, "rankSpacing": 24, "curve": "linear"}, "themeVariables": {"fontFamily": "system-ui, -apple-system, Segoe UI, sans-serif", "fontSize": "22px", "primaryColor": "#f8fafc", "primaryTextColor": "#1f2937", "primaryBorderColor": "#64748b", "lineColor": "#64748b", "tertiaryColor": "#eef2ff", "tertiaryTextColor": "#1f2937", "tertiaryBorderColor": "#64748b"}}}%%

flowchart LR
  pulse["Pulse t data"]
  target["Strategy target"]
  delta["Target delta"]
  fill["Next-open fill<br/>at t + 1"]
  ledger["Ledger state"]

  pulse --> target --> delta --> fill --> ledger
```

:::

That one-bar delay is the no-lookahead boundary. The strategy can react to the
current pulse, but it cannot fill at a price from the same pulse.

## Targets Are Holdings

Returning `10` means "I want to hold 10 units", not "buy 10 units every bar."
If the current position already equals the target, there is no new fill.

```{r}
#| label: target-holdings

bars <- ledgr_demo_bars |>
  filter(
    instrument_id == "DEMO_01",
    between(ts_utc, ledgr_utc("2018-01-01"), ledgr_utc("2018-01-10"))
  )

demo_qty <- 10

hold_then_flat <- function(ctx, params) {
  targets <- ctx$flat()
  ts <- ledgr_utc(ctx$ts_utc)
  if (ts >= ledgr_utc("2018-01-03") && ts < ledgr_utc("2018-01-08")) {
    targets["DEMO_01"] <- demo_qty
  }
  targets
}

bt <- ledgr_backtest(
  data = bars,
  strategy = hold_then_flat,
  initial_cash = 10000,
  run_id = "execution_target_holdings"
)

ledgr_results(bt, what = "fills") |>
  select(ts_utc, instrument_id, side, qty, price, fee)
```

The strategy asked for a position across several pulses, but only position
changes created fills: one opening fill and one closing fill.

## Next-Open Fill Timing

The signal pulse and execution pulse are different. The target changes on one
bar; the fill uses the next bar's open.

```{r}
#| label: next-open-evidence

signal_rows <- bars |>
  filter(ts_utc %in% c(ledgr_utc("2018-01-03"), ledgr_utc("2018-01-04"))) |>
  select(ts_utc, open, close)

fills <- ledgr_results(bt, what = "fills") |>
  select(ts_utc, side, qty, price)

signal_rows
fills
```

The opening fill occurs at the next available bar after the strategy changed
target. That is why a strategy should treat `ctx$close(id)` and feature values
as pulse-known information, not as executable prices.

## Costs Are Part Of The Fill

::: {.callout-important}
## Public cost API

The stable public transaction-cost model API is planned for v0.1.9.x / v0.2.0.
The example below documents fill behavior for readers inspecting execution
results. Do not treat this list interface as the stable public cost API.
:::

The fill model shown here is the next-open model with spread and fixed
commission fields.

```{r}
#| label: fill-costs

cost_bt <- ledgr_backtest(
  data = bars,
  strategy = hold_then_flat,
  initial_cash = 10000,
  run_id = "execution_cost_example",
  fill_model = list(type = "next_open", spread_bps = 5, commission_fixed = 1)
)

ledgr_results(cost_bt, what = "fills") |>
  select(ts_utc, side, qty, price, fee)
```

## Final-Bar Targets Cannot Fill

A target change on the final pulse is valid strategy output, but there is no
later bar where ledgr can simulate the next-open fill. ledgr warns and leaves
the ledger unchanged for that final target change.

```{r}
#| label: final-bar-no-fill

final_bar_strategy <- function(ctx, params) {
  targets <- ctx$flat()
  if (ledgr_utc(ctx$ts_utc) == ledgr_utc("2018-01-10")) {
    targets["DEMO_01"] <- demo_qty
  }
  targets
}

last_bar_warning <- FALSE
final_bar_bt <- withCallingHandlers(
  ledgr_backtest(
    data = bars,
    strategy = final_bar_strategy,
    initial_cash = 10000,
    run_id = "execution_final_bar_warning"
  ),
  warning = function(w) {
    if (grepl("LEDGR_LAST_BAR_NO_FILL", conditionMessage(w), fixed = TRUE)) {
      last_bar_warning <<- TRUE
      invokeRestart("muffleWarning")
    }
  }
)

last_bar_warning
ledgr_results(final_bar_bt, what = "fills")
```

If that final target matters, extend the snapshot by one executable bar and run
again. Do not suppress the warning and treat the missing fill as a completed
trade.

## Warmup Gates Belong In The Strategy

::: {.callout-note}
## Definition

Warmup is the early part of a feature series where an indicator has not yet seen
enough history to produce a usable value. For example, a 20-bar moving average
is not usable on the first bar of a snapshot.
:::

Feature warmup is not an execution rule. The strategy decides whether a feature
vector is usable. Different strategies may want different warmup behavior, so
ledgr exposes warmup state instead of silently imposing a trading rule. For
active aliases, `passed_warmup()` is the standard guard:

```{r}
#| label: warmup-guard
#| eval: false

values <- ctx$features("DEMO_01")
targets <- ctx$flat()

if (passed_warmup(values) && values[["fast"]] > values[["slow"]]) {
  targets["DEMO_01"] <- params$qty
}

targets
```

`passed_warmup()` keeps the rule inactive while mapped features are still
ordinary early `NA`. A typo in an alias or a feature that can never become usable
is a different problem; ledgr should fail or report diagnostics rather than
silently treating missing features as no signal.

## Zero Fills And Zero Trades Mean Different Things

Zero fills means no execution occurred. Non-empty fills with zero trades means
the run opened or adjusted a position but did not close a round trip inside the
sample.

```{r}
#| label: fills-vs-trades

ledgr_results(bt, what = "fills") |>
  select(ts_utc, side, qty, price)

ledgr_results(bt, what = "trades") |>
  select(any_of(c("entry_ts_utc", "exit_ts_utc", "entry_ts", "exit_ts", "qty", "pnl", "realized_pnl")))
```

Start with fills when debugging execution. Trades are derived from filled round
trips, not from target changes.

## Where Next

- For strategy authoring and `passed_warmup()` patterns, read
  `vignette("strategy-development", package = "ledgr")`.
- For feature maps and warmup diagnostics, read
  `vignette("indicators", package = "ledgr")`.
- For ledger, fills, trades, equity, and metrics, read
  `vignette("metrics-and-accounting", package = "ledgr")`.
- For candidate sweeps and promotion context, read
  `vignette("sweeps", package = "ledgr")`.

```{r}
#| label: cleanup
#| include: false

close(bt)
close(cost_bt)
close(final_bar_bt)
```
