---
title: "Experiment Store"
output:
  rmarkdown::html_vignette:
    toc: false
  github_document:
    toc: false
    html_preview: false
vignette: >
  %\VignetteIndexEntry{Experiment Store}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 90)
options(cli.unicode = FALSE)
default_output_hook <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(
  output = function(x, options) {
    default_output_hook(gsub("[ \t]+(?=\n)", "", x, perl = TRUE), options)
  }
)
demo_data_path <- if (file.exists("data/ledgr_demo_bars.rda")) {
  "data/ledgr_demo_bars.rda"
} else {
  file.path("..", "data", "ledgr_demo_bars.rda")
}
if (!exists("ledgr_demo_bars") && file.exists(demo_data_path)) {
  load(demo_data_path)
}
```

The experiment store is the DuckDB file behind a sealed snapshot. It keeps
market data, run artifacts, provenance, labels, tags, archive state, and compact
telemetry together.

```text
snapshot handle -> run experiments -> list / inspect / compare / reopen
```

`run_id` is immutable. Labels, tags, and archive state are mutable metadata.

The examples use `dplyr` and `tibble` for data preparation and compact display.
They are suggested packages used by the vignettes, not part of the experiment
store contract.

```{r library, message=FALSE}
library(ledgr)
library(dplyr)
library(tibble)
data("ledgr_demo_bars", package = "ledgr")
```

## Create A Durable Snapshot

Market data and derived data have different lifecycle rules in ledgr. A sealed
snapshot freezes the real market-data input and its hash. If you need more
instruments, more dates, corrected bars, or tick-derived bars, create a new
snapshot. Indicators, runs, labels, tags, comparisons, and telemetry are derived
from sealed market data and can be added later without mutating the snapshot.

This vignette uses `tempfile()` so it can run without writing into your project
directory. For real research, use a stable path such as `"research.duckdb"` and
a snapshot ID you will recognize later.

```{r snapshot}
db_path <- tempfile("ledgr_store_", fileext = ".duckdb")

bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% c("DEMO_01", "DEMO_02"),
    between(
      ts_utc,
      ledgr::ledgr_utc("2019-01-01"),
      ledgr::ledgr_utc("2019-06-30")
    )
  )

snapshot <- ledgr_snapshot_from_df(
  bars,
  db_path = db_path,
  snapshot_id = "store_demo_snapshot"
)
```

After snapshot creation, store operations take `snapshot`, not `db_path`. In a
new R session, recover the handle with `ledgr_snapshot_load(db_path,
snapshot_id)`.

If your market data starts in CSV, seal the CSV into the same kind of durable
store. The CSV must contain at least `instrument_id`, `ts_utc`, `open`, `high`,
`low`, and `close`.

```{r csv-snapshot, eval=FALSE}
snapshot <- ledgr_snapshot_from_csv(
  "data/daily_bars.csv",
  db_path = "research.duckdb",
  snapshot_id = "eod_2019_h1"
)
```

In any later session, recover the handle without re-sealing the data:

```{r csv-snapshot-load, eval=FALSE}
snapshot <- ledgr_snapshot_load("research.duckdb", snapshot_id = "eod_2019_h1")
```

## Bridge A Low-Level CSV Import

The high-level CSV helper above is the normal path. The lower-level path is
useful when you want to create the snapshot row, import one or more CSV files,
inspect the sealed metadata, and then load the sealed artifact in a separate
step.

The order is important:

1. create the snapshot envelope;
2. import bars into that CREATED snapshot;
3. seal it to validate bars and write the snapshot hash;
4. load it with `verify = TRUE`;
5. pass the loaded snapshot to `ledgr_experiment()` and `ledgr_run()`.

```{r low-level-csv-bridge}
csv_db_path <- tempfile("ledgr_csv_bridge_", fileext = ".duckdb")
csv_bars_path <- tempfile("ledgr_csv_bars_", fileext = ".csv")

csv_bars <- bars |>
  filter(ts_utc <= ledgr::ledgr_utc("2019-01-15")) |>
  mutate(ts_utc = format(ts_utc, "%Y-%m-%dT%H:%M:%SZ", tz = "UTC"))

utils::write.csv(csv_bars, csv_bars_path, row.names = FALSE)

csv_con <- ledgr_db_init(csv_db_path)
csv_snapshot_id <- ledgr_snapshot_create(
  csv_con,
  snapshot_id = "csv_bridge_snapshot",
  meta = list(description = "low-level CSV bridge demo")
)

ledgr_snapshot_import_bars_csv(
  csv_con,
  csv_snapshot_id,
  bars_csv_path = csv_bars_path,
  instruments_csv_path = NULL,
  auto_generate_instruments = TRUE
)

csv_hash <- ledgr_snapshot_seal(csv_con, csv_snapshot_id)
csv_info <- ledgr_snapshot_info(csv_con, csv_snapshot_id)
DBI::dbDisconnect(csv_con, shutdown = TRUE)

csv_hash
csv_info |>
  select(
    snapshot_id,
    status,
    snapshot_hash,
    bar_count,
    instrument_count,
    start_date,
    end_date,
    meta_json
  )
```

`bar_count` and `instrument_count` are live counts from the sealed snapshot
tables. The raw `meta_json` is envelope metadata on the snapshot row; seal-time
metadata inside it uses `n_bars` and `n_instruments`. Snapshot identity does not
come from that metadata. `snapshot_hash` identifies the normalized bars and
instruments only, so adding a human description to `meta_json` does not change
the artifact hash.

Load the sealed snapshot before constructing the experiment. This is the same
handle you would use in a later R session.

```{r low-level-csv-run}
csv_snapshot <- ledgr_snapshot_load(
  csv_db_path,
  snapshot_id = csv_snapshot_id,
  verify = TRUE
)

csv_strategy <- function(ctx, params) {
  targets <- ctx$flat()
  targets["DEMO_01"] <- params$qty
  targets
}

csv_exp <- ledgr_experiment(
  snapshot = csv_snapshot,
  strategy = csv_strategy,
  opening = ledgr_opening(cash = 10000)
)

csv_bt <- ledgr_run(csv_exp, params = list(qty = 1), run_id = "csv_bridge_run")
tail(ledgr_results(csv_bt, what = "equity"), 3)
```

## Run Variants

```{r experiment}
features <- list(ledgr_ind_sma(20))

trend_strategy <- function(ctx, params) {
  targets <- ctx$flat()
  for (id in ctx$universe) {
    sma <- ctx$feature(id, "sma_20")
    if (is.finite(sma) && ctx$close(id) > sma) {
      targets[id] <- params$qty
    }
  }
  targets
}

exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = trend_strategy,
  features = features,
  opening = ledgr_opening(cash = 10000)
)

bt_small <- exp |>
  ledgr_run(params = list(qty = 5), run_id = "trend_qty_5")

bt_large <- exp |>
  ledgr_run(params = list(qty = 15), run_id = "trend_qty_15")
```

## Discover Runs

`ledgr_run_list()` is the store discovery view.

```{r list}
ledgr_run_list(snapshot)
```

Use labels and tags for mutable human-facing organization.

```{r label-tags}
snapshot <- snapshot |>
  ledgr_run_label("trend_qty_5", "Baseline quantity") |>
  ledgr_run_tag("trend_qty_5", c("baseline", "trend")) |>
  ledgr_run_tag("trend_qty_15", c("trend", "larger-size"))

ledgr_run_list(snapshot)
```

Tags and labels do not alter snapshot hashes, strategy hashes, parameter hashes,
config hashes, or result artifacts.

The returned objects are still tibbles. When you need a custom view, convert to
a tibble and select the columns you want.

```{r list-deeper}
ledgr_run_list(snapshot) |>
  as_tibble() |>
  select(run_id, label, tags, status, final_equity, execution_mode)
```

## Inspect And Compare

```{r info}
info <- ledgr_run_info(snapshot, "trend_qty_5")
info
```

`ledgr_run_info()` is the detailed metadata view. It includes execution mode,
compact telemetry, status, identity hashes, and reproducibility tier.

```{r compare}
ledgr_compare_runs(snapshot, run_ids = c("trend_qty_5", "trend_qty_15"))
```

Comparison is read-only and does not rerun strategies. `n_trades` counts
closed, realised trade observations, not every fill. A run can have fills but no
closed trades yet, in which case win rate is not defined.

## Inspect Stored Strategy Source

Completed runs keep strategy provenance in the experiment store. This is one of
the most useful audit artifacts: you can inspect the source text that produced a
run without reopening the backtest handle and without rerunning the strategy.

Use `trust = FALSE` for safe inspection. It returns stored source text,
parameters, hashes, dependency metadata, and warnings without parsing,
evaluating, or executing the source.

```{r extract-source}
stored_strategy <- ledgr_extract_strategy(snapshot, "trend_qty_5", trust = FALSE)
stored_strategy
```

The source text is just data in this mode.

```{r extract-source-text}
writeLines(stored_strategy$strategy_source_text)
```

Hash verification proves stored-text identity, not code safety. Use
`trust = TRUE` only when you already trust the experiment store and
intentionally want ledgr to parse and evaluate the stored text into a function
object. Legacy/pre-provenance runs and strategy types without capturable source
may report `strategy_source_text = NA`; those runs remain inspectable through
`ledgr_run_info()` and stored result tables, but their strategy function cannot
be recovered from provenance alone.

## Reopen A Completed Run In A Later Session

`ledgr_run_open()` reconstructs a completed run handle from stored artifacts. It
does not recompute the strategy. This is useful when you want full result
tables or plots after restarting R.

```{r open}
reopened <- ledgr_run_open(snapshot, "trend_qty_5")
summary(reopened)
tail(ledgr_results(reopened, what = "equity"), 3)
close(reopened)
```

Only completed runs can be reopened. Failed or incomplete runs remain
inspectable through `ledgr_run_info()`.

## Archive Without Deleting

```{r archive}
snapshot <- snapshot |>
  ledgr_run_archive("trend_qty_15", reason = "larger position kept for reference")

ledgr_run_list(snapshot)
```

Archiving hides a run from default listings without deleting artifacts.

`ledgr_run()` and `ledgr_run_open()` return live handles for durable run
artifacts. The artifacts are already durable when a run completes, and ordinary
result inspection opens and closes read connections per operation. Use
`close(bt)` as explicit resource cleanup in long sessions, tests, explicit-open
workflows, and lazy result cursors. Close the snapshot handle when the workflow
is finished.

```{r cleanup}
close(bt_small)
close(bt_large)
ledgr_snapshot_close(csv_snapshot)
ledgr_snapshot_close(snapshot)
unlink(csv_bars_path)
```

## What's Next?

For fills, trades, equity rows, and metric definitions, read
`vignette("metrics-and-accounting", package = "ledgr")`. For strategy authoring,
read `vignette("strategy-development", package = "ledgr")`.
