---
title: "Experiment Store"
output:
  github_document:
    toc: false
    html_preview: false
vignette: >
  %\VignetteIndexEntry{Experiment Store}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 90)
options(cli.unicode = FALSE)
default_output_hook <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(
  output = function(x, options) {
    default_output_hook(gsub("[ \t]+(?=\n)", "", x, perl = TRUE), options)
  }
)
demo_data_path <- if (file.exists("data/ledgr_demo_bars.rda")) {
  "data/ledgr_demo_bars.rda"
} else {
  file.path("..", "data", "ledgr_demo_bars.rda")
}
if (!exists("ledgr_demo_bars") && file.exists(demo_data_path)) {
  load(demo_data_path)
}
```

The experiment store is the DuckDB file behind a sealed snapshot. It keeps
market data, run artifacts, provenance, labels, tags, archive state, and compact
telemetry together.

```text
snapshot handle -> run experiments -> list / inspect / compare / reopen
```

`run_id` is immutable. Labels, tags, and archive state are mutable metadata.

```{r library}
library(ledgr)
data("ledgr_demo_bars", package = "ledgr")
```

## Create A Durable Snapshot

```{r snapshot}
db_path <- tempfile("ledgr_store_", fileext = ".duckdb")

bars <- subset(
  ledgr_demo_bars,
  instrument_id %in% c("DEMO_01", "DEMO_02") &
    ts_utc >= as.POSIXct("2019-01-01", tz = "UTC") &
    ts_utc <= as.POSIXct("2019-06-30", tz = "UTC")
)

snapshot <- ledgr_snapshot_from_df(
  bars,
  db_path = db_path,
  snapshot_id = "store_demo_snapshot"
)
```

After snapshot creation, store operations take `snapshot`, not `db_path`. In a
new R session, recover the handle with `ledgr_snapshot_load(db_path,
snapshot_id)`.

## Run Variants

```{r experiment}
features <- list(ledgr_ind_sma(20))

trend_strategy <- function(ctx, params) {
  targets <- ctx$flat()
  for (id in ctx$universe) {
    sma <- ctx$feature(id, "sma_20")
    if (is.finite(sma) && ctx$close(id) > sma) {
      targets[id] <- params$qty
    }
  }
  targets
}

exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = trend_strategy,
  features = features,
  opening = ledgr_opening(cash = 10000)
)

bt_small <- exp |>
  ledgr_run(params = list(qty = 5), run_id = "trend_qty_5")

bt_large <- exp |>
  ledgr_run(params = list(qty = 15), run_id = "trend_qty_15")
```

## Discover Runs

`ledgr_run_list()` is the store discovery view.

```{r list}
ledgr_run_list(snapshot)[
  c("run_id", "label", "tags", "status", "final_equity", "execution_mode")
]
```

Use labels and tags for mutable human-facing organization.

```{r label-tags}
snapshot <- snapshot |>
  ledgr_run_label("trend_qty_5", "Baseline quantity") |>
  ledgr_run_tag("trend_qty_5", c("baseline", "trend")) |>
  ledgr_run_tag("trend_qty_15", c("trend", "larger-size"))

ledgr_run_list(snapshot)[c("run_id", "label", "tags")]
```

Tags and labels do not alter snapshot hashes, strategy hashes, parameter hashes,
config hashes, or result artifacts.

## Inspect And Compare

```{r info}
info <- ledgr_run_info(snapshot, "trend_qty_5")
info
```

`ledgr_run_info()` is the detailed metadata view. It includes execution mode,
compact telemetry, status, identity hashes, and reproducibility tier.

```{r compare}
ledgr_compare_runs(snapshot, run_ids = c("trend_qty_5", "trend_qty_15"))[
  c("run_id", "final_equity", "total_return", "max_drawdown", "n_trades", "win_rate")
]
```

Comparison is read-only and does not rerun strategies.

## Reopen A Completed Run

```{r open}
reopened <- ledgr_run_open(snapshot, "trend_qty_5")
summary(reopened)
tail(ledgr_results(reopened, what = "equity"), 3)
close(reopened)
```

Only completed runs can be reopened. Failed or incomplete runs remain
inspectable through `ledgr_run_info()`.

## Archive Without Deleting

```{r archive}
snapshot <- snapshot |>
  ledgr_run_archive("trend_qty_15", reason = "larger position kept for reference")

ledgr_run_list(snapshot)[c("run_id", "status", "archived", "archive_reason")]
```

Archiving hides a run from default listings without deleting artifacts.

```{r cleanup}
close(bt_small)
close(bt_large)
ledgr_snapshot_close(snapshot)
```
