---
title: "Experiment Store"
format:
  html:
    toc: false
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{Experiment Store}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

options(width = 90)
options(cli.unicode = FALSE)
# Local renders load the package from source via pkgload (working dir is either
# project root or vignettes/); installed builds fall through to library(ledgr).
pkg_root <- if (file.exists("DESCRIPTION")) "." else if (file.exists("../DESCRIPTION")) ".." else NA_character_
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
default_output_hook <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(
  output = function(x, options) {
    x <- gsub("Database:(\\s+)[^\\r\\n]+[.]duckdb", "Database:\\1<temporary DuckDB path>", x, perl = TRUE)
    x <- gsub("[ \t]+(?=\n)", "", x, perl = TRUE)
    default_output_hook(x, options)
  }
)
```

The experiment store keeps committed run evidence: run records, labels,
comparison surfaces, recovery metadata, and reopened results. For snapshot
creation and data-input boundaries, read
`vignette("data-input-and-snapshots", package = "ledgr")`.

::: {.ledgr-callout .ledgr-callout-note}
**Running this yourself**

This article is evaluated when rendered. It writes to temporary DuckDB stores
so package builds and local previews do not leave project artifacts behind. In
real work, use a project-local path such as `artifacts/ledgr_store.duckdb`.
:::

::: {.ledgr-callout .ledgr-callout-warning}
**Pre-CRAN compatibility**

ledgr is pre-CRAN. Store schemas, config hashes, provenance formats, and
experimental APIs may change before the first CRAN release. Treat stores
created with pre-CRAN ledgr as research artifacts for the version that produced
them, and expect to rerun experiments after upgrading.
:::

The examples use `dplyr` for data preparation and compact display. It is a
suggested package used by the vignettes, not part of the experiment-store
contract.

```{r}
#| label: library
#| message: false

library(ledgr)
library(dplyr)
data("ledgr_demo_bars", package = "ledgr")
```

## Temporary Snapshot Setup

This article uses a small temporary snapshot so the store examples are
self-contained. The full snapshot lifecycle is covered in
`vignette("data-input-and-snapshots", package = "ledgr")`.

```{r}
#| label: snapshot

db_path <- file.path(tempdir(), "ledgr_store_demo.duckdb")
if (file.exists(db_path)) {
  unlink(db_path)
}

bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% c("DEMO_01", "DEMO_02"),
    between(
      ts_utc,
      ledgr_utc("2019-01-01"),
      ledgr_utc("2019-06-30")
    )
  )

snapshot <- ledgr_snapshot_from_df(
  bars,
  db_path = db_path,
  snapshot_id = "store_demo_snapshot"
)
```

After snapshot creation, store operations take `snapshot`, not `db_path`. In a
new R session, recover the handle with
`ledgr_snapshot_open(db_path, snapshot_id)`.

## Record Two Variants For Comparison

```{r}
#| label: experiment

features <- list(ledgr_ind_sma(20))

trend_strategy <- function(ctx, params) {
  targets <- ctx$flat()
  for (id in ctx$universe) {
    sma <- ctx$feature(id, "sma_20")
    if (is.finite(sma) && ctx$close(id) > sma) {
      targets[id] <- params$qty
    }
  }
  targets
}

exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = trend_strategy,
  features = features,
  opening = ledgr_opening(cash = 10000),
  cost_model = ledgr_cost_zero()
)

bt_small <- exp |>
  ledgr_run(params = list(qty = 5), run_id = "trend_qty_5")

bt_large <- exp |>
  ledgr_run(params = list(qty = 15), run_id = "trend_qty_15")
```

## Discover Runs

`ledgr_run_list()` is the store discovery view.

```{r}
#| label: list

ledgr_run_list(snapshot)
```

Use labels and tags for mutable human-facing organization.

```{r}
#| label: label-tags

snapshot <- snapshot |>
  ledgr_run_label("trend_qty_5", "Baseline quantity") |>
  ledgr_run_tag("trend_qty_5", c("baseline", "trend")) |>
  ledgr_run_tag("trend_qty_15", c("trend", "larger-size"))

ledgr_run_list(snapshot)
```

Tags and labels do not alter snapshot hashes, strategy hashes, parameter hashes,
config hashes, or result artifacts.

The returned objects are still tibbles. When you need a custom view, convert to
a tibble and select the columns you want.

```{r}
#| label: list-deeper

ledgr_run_list(snapshot) |>
  as_tibble() |>
  select(run_id, label, tags, status, final_equity, execution_mode)
```

## Inspect And Compare

```{r}
#| label: info

info <- ledgr_run_info(snapshot, "trend_qty_5")
info
```

`ledgr_run_info()` is the detailed metadata view. It includes execution mode,
compact telemetry, status, identity hashes, and reproducibility tier.

Useful fields include:

| Field | Meaning |
|---|---|
| `run_id`, `status`, `label`, `tags`, `archived` | mutable and immutable run organization fields |
| `snapshot_id`, `snapshot_hash` | sealed data identity |
| `strategy_source_hash`, `strategy_params_hash`, `config_hash` | strategy, parameter, and run-configuration identity |
| `reproducibility_level` | strategy preflight tier recorded with the run |
| `execution_mode`, `elapsed_sec`, `pulse_count` | execution telemetry |
| `persist_features`, `feature_cache_hits`, `feature_cache_misses` | compact feature-engine telemetry |
| `error_msg` | failure diagnostic for non-completed runs |

```{r}
#| label: compare

comparison <- ledgr_run_compare(snapshot, run_ids = c("trend_qty_5", "trend_qty_15"))
comparison
```

Comparison is read-only and does not rerun strategies. `n_trades` counts
closed, realised trade observations, not every fill. A run can have fills but no
closed trades yet, in which case win rate is not defined.

`ledgr_run_compare()` starts from the durable snapshot handle because it reads
stored run artifacts. When you want the comparison to use an experiment's metric
assumptions, pass that context explicitly:

```{r}
#| label: compare-exp-context
#| eval: false

comparison <- ledgr_run_compare(
  snapshot,
  run_ids = c("trend_qty_5", "trend_qty_15"),
  metric_context = ledgr_metric_context(exp)
)
```

The printed comparison formats some columns for reading. Programmatic code gets
raw numeric columns from the tibble:

```{r}
#| label: compare-raw

comparison |>
  select(run_id, final_equity, total_return, sharpe_ratio, max_drawdown, n_trades)
```

For report writing, coerce the comparison to a data frame or tibble before
formatting percentages yourself:

```{r}
#| label: compare-report-ready

comparison_report <- comparison |>
  as_tibble() |>
  select(run_id, final_equity, total_return, sharpe_ratio, max_drawdown)

comparison_report
```

After selecting a run, reopen it and inspect the underlying result tables rather
than parsing the printed comparison:

```{r}
#| label: compare-followup

best_run_id <- comparison |>
  arrange(desc(total_return)) |>
  pull(run_id) |>
  first()

best_bt <- ledgr_run_open(snapshot, best_run_id)
tail(ledgr_results(best_bt, what = "equity"), 3)
close(best_bt)
```

## Inspect Stored Strategy Source

Completed runs keep strategy provenance in the experiment store. This is one of
the most useful audit artifacts: you can inspect the source text that produced a
run without reopening the backtest handle and without rerunning the strategy.
The full trust and tier model lives in `vignette("reproducibility", package =
"ledgr")`; this section shows the store workflow.

Use `trust = FALSE` for safe inspection. It returns stored source text,
parameters, hashes, dependency metadata, and warnings without parsing,
evaluating, or executing the source.

```{r}
#| label: extract-source

stored_strategy <- ledgr_run_strategy(snapshot, "trend_qty_5", trust = FALSE)
stored_strategy
```

The source text is just data in this mode.

```{r}
#| label: extract-source-text

writeLines(stored_strategy$strategy_source_text)
```

Hash verification proves stored-text identity, not code safety. Use
`trust = TRUE` only when you already trust the experiment store and
intentionally want ledgr to parse and evaluate the stored text into a function
object. Legacy/pre-provenance runs remain inspectable through `ledgr_run_info()`
and stored result tables, but their strategy function cannot be recovered from
provenance alone.

When a run ID is missing, store lookup helpers fail with class
`ledgr_run_not_found`:

```{r}
#| label: missing-run
#| eval: false

ledgr_run_info(snapshot, "missing_run")
```

Trusted recovery can be used to rerun a stored strategy only after you have
decided that evaluating the stored source is acceptable:

```{r}
#| label: trusted-rerun
#| eval: false

recovered <- ledgr_run_strategy(snapshot, "trend_qty_5", trust = TRUE)

rerun_exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = recovered$strategy_function,
  features = features,
  opening = ledgr_opening(cash = 10000),
  cost_model = ledgr_cost_zero()
)

ledgr_run(
  rerun_exp,
  params = recovered$strategy_params,
  run_id = "trend_qty_5_rerun"
)
```

## Reopen A Completed Run In A Later Session

`ledgr_run_open()` reconstructs a completed run handle from stored artifacts. It
does not recompute the strategy. This is useful when you want full result
tables or plots after restarting R.

```{r}
#| label: open

reopened <- ledgr_run_open(snapshot, "trend_qty_5")
summary(reopened)
tail(ledgr_results(reopened, what = "equity"), 3)
close(reopened)
```

Only completed runs can be reopened. Failed or incomplete runs remain
inspectable through `ledgr_run_info()`.

Store-level helpers such as `ledgr_run_info()`, `ledgr_run_list()`, and
`ledgr_run_compare()` use the snapshot handle and remain available after a
completed run handle is closed. Result-table helpers such as `ledgr_results()`
need a live or reopened backtest handle.

## Recovery

Most ledgr recovery workflows should start with the high-level store helpers
above:

- use `ledgr_snapshot_open()` to reopen the sealed snapshot store;
- use `ledgr_run_open()` to reopen a completed run handle;
- use `ledgr_run_info()` and `ledgr_run_strategy()` to inspect stored metadata
  and strategy provenance.

The lower-level recovery pair remains public for restart inspection and
maintainer workflows that need to work directly against the store connection.

`ledgr_db_init(db_path)` opens a DBI connection to a ledgr DuckDB store and
ensures the ledgr schema exists. In normal workflows, ordinary users usually do
not need it because snapshot and run helpers open, verify, and close the
required connections for their own operations.

`ledgr_state_reconstruct(run_id, con)` reconstructs ledgr's expected simulated
state for one stored run from ledger-backed evidence. It returns reconstructed
state artifacts such as positions, cash, equity, fills, and trades from the
stored run records. It is useful when you are inspecting a restart boundary,
debugging stored evidence, or building a low-level tool that already owns a DBI
connection.

```{r}
#| label: low-level-recovery
#| eval: false

con <- ledgr_db_init(db_path)
state <- ledgr_state_reconstruct("trend_qty_5", con)
DBI::dbDisconnect(con, shutdown = TRUE)
```

This pair is intentionally not a broker or migration layer. It does not perform
broker reconciliation, prove live restart safety, migrate old schemas,
repair a sealed snapshot, or recover strategy dependencies that were never captured in
the run provenance. Treat it as low-level inspection over ledgr's own stored
evidence, not as an escape hatch around the sealed-data and provenance
contracts.

## Archive Without Deleting

```{r}
#| label: archive

snapshot <- snapshot |>
  ledgr_run_archive("trend_qty_15", reason = "larger position kept for reference")

ledgr_run_list(snapshot)
```

Archiving hides a run from default listings without deleting artifacts.

## Current Feature Persistence Boundary

Run metadata records whether feature persistence was enabled, and pulse
inspection lets you view registered feature values at one decision time. Public
feature inspection is intentionally scoped to feature contracts, warmup
feasibility, and pulse-time feature views:

- `ledgr_feature_contracts(features)` shows declared feature requirements;
- `ledgr_feature_contract_check(snapshot, features)` checks whether those
  requirements are achievable in a sealed snapshot;
- `ledgr_pulse_snapshot()` plus `ledgr_pulse_features()` or
  `ledgr_pulse_wide()` inspects one pulse.

A full persisted feature-series retrieval API remains outside the current
experiment-store surface; use precompute and sweep provenance when you need
feature-set identity at sweep scale.

External point-in-time regressors are a separate future data surface. The
public roadmap keeps that work out of v0.1.8.5 and tracks it for a later cycle
so vintage semantics, lineage, ASOF lookup, and leakage prevention can be
designed explicitly rather than smuggled into CSV bars or active aliases.

## Resource Cleanup

`ledgr_run()` and `ledgr_run_open()` return live handles for durable run
artifacts. The artifacts are already durable when a run completes, and ordinary
result inspection opens and closes read connections per operation. Use
`close(bt)` as explicit resource cleanup in long sessions, tests, explicit-open
workflows, and lazy result cursors. Close snapshot handles when the workflow is
finished.

## Task Intent Map

Use this map when you know the task but not the function name:

| Intent | Start here |
|---|---|
| Seal in-memory bars | `ledgr_snapshot_from_df()` |
| Seal a local CSV | `ledgr_snapshot_from_csv()` |
| Fetch and seal Yahoo bars | `ledgr_snapshot_from_yahoo()` |
| Control low-level CSV create/import/seal lifecycle | `?ledgr_snapshot_import_bars_csv` |
| Reopen an existing store | `ledgr_snapshot_open()` |
| List stored runs | `ledgr_run_list()` |
| Compare durable runs | `ledgr_run_compare()` |

Yahoo data is a convenience source. The sealed snapshot is the ledgr artifact;
the remote Yahoo endpoint remains outside ledgr's reproducibility boundary.

```{r}
#| label: cleanup
#| include: false

close(bt_small)
close(bt_large)
ledgr_snapshot_close(snapshot)
```

## What's Next?

For fills, trades, equity rows, and metric definitions, read
`vignette("metrics-and-accounting", package = "ledgr")`. For strategy authoring,
read `vignette("strategy-development", package = "ledgr")`.
