---
title: "Experiment Store"
output:
  github_document:
    toc: false
    html_preview: false
vignette: >
  %\VignetteIndexEntry{Experiment Store}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 90)
options(cli.unicode = FALSE)
default_output_hook <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(
  output = function(x, options) {
    default_output_hook(gsub("[ \t]+(?=\n)", "", x, perl = TRUE), options)
  }
)
```

The experiment store is a durable DuckDB file that keeps sealed market data,
run metadata, result artifacts, provenance, and compact telemetry together.

The basic workflow is:

```text
sealed snapshot -> many runs -> list / inspect / compare / reopen
```

`run_id` is the immutable experiment key. Labels, tags, and archive state are
mutable metadata layered on top of that key.

## Create One Store

Use one durable `db_path` for the snapshot and all runs you want to compare.

```{r data}
library(ledgr)

db_path <- tempfile(fileext = ".duckdb")

bars <- data.frame(
  ts_utc = rep(as.POSIXct("2020-01-01", tz = "UTC") + 86400 * 0:6, 2),
  instrument_id = rep(c("AAA", "BBB"), each = 7),
  open = c(100, 101, 102, 103, 104, 105, 106,
           80,  80,  81,  81,  82,  83,  83),
  high = c(101, 102, 103, 104, 105, 106, 107,
           81,  81,  82,  82,  83,  84,  84),
  low = c(99, 100, 101, 102, 103, 104, 105,
          79, 79, 80, 80, 81, 82, 82),
  close = c(100, 101, 102, 103, 104, 105, 106,
            80, 80, 81, 81, 82, 83, 83),
  volume = 1000
)

snapshot <- ledgr_snapshot_from_df(
  bars,
  db_path = db_path,
  snapshot_id = "demo_snapshot"
)
```

The snapshot is sealed. Reusing it means each run is evaluated against the same
data artifact.

## Run Variants

Here two parameter sets use the same strategy function and the same snapshot.

```{r runs}
trend_strategy <- function(ctx, params) {
  targets <- ctx$flat()
  for (id in ctx$universe) {
    if (ctx$close(id) > params$threshold[[id]]) {
      targets[id] <- params$qty
    }
  }
  targets
}

bt_small <- ledgr_backtest(
  snapshot = snapshot,
  strategy = trend_strategy,
  strategy_params = list(threshold = c(AAA = 101, BBB = 80), qty = 1),
  db_path = db_path,
  run_id = "trend_small"
)

bt_large <- ledgr_backtest(
  snapshot = snapshot,
  strategy = trend_strategy,
  strategy_params = list(threshold = c(AAA = 102, BBB = 81), qty = 3),
  db_path = db_path,
  run_id = "trend_large"
)
```

## List Runs

`ledgr_run_list()` is the discovery view. It is read-only.

```{r list}
ledgr_run_list(db_path)[, c(
  "run_id", "label", "tags", "status", "final_equity", "total_return",
  "execution_mode", "reproducibility_level"
)]
```

`run_id` should be stable and script-friendly. Use labels for human names.

```{r label-tags}
ledgr_run_label(db_path, "trend_small", "Lower threshold, one share")
ledgr_run_tag(db_path, "trend_small", c("baseline", "trend"))
ledgr_run_tag(db_path, "trend_large", c("trend", "higher-size"))

ledgr_run_list(db_path)[, c("run_id", "label", "tags")]
```

Tags are mutable grouping metadata. They do not change identity hashes, stored
artifacts, strategy provenance, or comparison semantics.

## Inspect One Run

`ledgr_run_info()` gives the detailed metadata for one run.

```{r info}
info <- ledgr_run_info(db_path, "trend_small")
info
```

Important fields:

- `execution_mode`: how the run wrote artifacts, for example `audit_log` or
  `db_live`;
- `elapsed_sec`, `pulse_count`, and cache counts: compact telemetry retained
  after the R session ends;
- `reproducibility_level`: whether ledgr captured enough strategy metadata for
  source inspection or recovery;
- `strategy_source_hash` and `strategy_params_hash`: identity metadata for the
  strategy and parameters.

Older stores may contain legacy/pre-provenance runs. ledgr reads them, but the
provenance fields can be missing.

## Compare Runs

`ledgr_compare_runs()` builds on stored run metadata and result artifacts. It
does not rerun strategies.

```{r compare}
ledgr_compare_runs(db_path, run_ids = c("trend_small", "trend_large"))[, c(
  "run_id", "final_equity", "total_return", "max_drawdown",
  "n_trades", "win_rate", "strategy_params_hash"
)]
```

This is the lightweight comparison surface in v0.1.6. Parameter sweeps are
future scope.

## Reopen A Completed Run

`ledgr_run_open()` returns a normal `ledgr_backtest` handle over existing
artifacts. It does not recompute the run.

```{r open}
reopened <- ledgr_run_open(db_path, "trend_small")
summary(reopened)
ledgr_results(reopened, what = "equity")
close(reopened)
```

Only completed runs can be opened. Failed or incomplete runs should be inspected
with `ledgr_run_info()`.

## Archive Instead Of Delete

Archiving hides a run from the default list but keeps the artifacts available.

```{r archive}
ledgr_run_archive(db_path, "trend_large", reason = "superseded by smaller baseline")

ledgr_run_list(db_path)[, c("run_id", "archived")]
ledgr_run_list(db_path, include_archived = TRUE)[, c("run_id", "archived", "archive_reason")]
```

Archive is non-destructive and idempotent. Hard delete is not a v0.1.6 API.

```{r cleanup}
close(bt_small)
close(bt_large)
close(snapshot)
```
