---
title: "Experiment Store"
format:
  html:
    toc: false
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{Experiment Store}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

options(width = 90)
options(cli.unicode = FALSE)
# Local renders load the package from source via pkgload (working dir is either
# project root or vignettes/); installed builds fall through to library(ledgr).
pkg_root <- if (file.exists("DESCRIPTION")) "." else if (file.exists("../DESCRIPTION")) ".." else NA_character_
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
default_output_hook <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(
  output = function(x, options) {
    x <- gsub("Database:(\\s+)[^\\r\\n]+[.]duckdb", "Database:\\1<temporary DuckDB path>", x, perl = TRUE)
    x <- gsub("[ \t]+(?=\n)", "", x, perl = TRUE)
    default_output_hook(x, options)
  }
)
```

Every committed run is stored with its evidence. By the end of this article
you can find your runs again, label and compare them, see what produced each
one, and reopen them in a later session. For snapshot creation and data-input
boundaries, read `vignette("data-input-and-snapshots", package = "ledgr")`.

::: {.ledgr-callout .ledgr-callout-note}
**Running this yourself**

This article is evaluated when rendered. It writes to temporary DuckDB stores
so package builds and local previews do not leave project artifacts behind. In
real work, use a project-local path such as `artifacts/ledgr_store.duckdb`.
:::

::: {.ledgr-callout .ledgr-callout-warning}
**Pre-CRAN compatibility**

ledgr is pre-CRAN. Store schemas, config hashes, provenance formats, and
experimental APIs may change before the first CRAN release. Treat stores
created with pre-CRAN ledgr as research artifacts for the version that produced
them, and expect to rerun experiments after upgrading.
:::

The examples use `dplyr` for data preparation and compact display. It is a
suggested package used by the vignettes, not part of the experiment-store
contract.

```{r}
#| label: library
#| message: false

library(ledgr)
library(dplyr)
data("ledgr_demo_bars", package = "ledgr")
```

## Temporary Snapshot Setup

This article uses a small snapshot in a temporary store from
`ledgr_temp_store()`, so the store examples are self-contained. A real project
passes a persistent path, so its runs outlive the R session. The full snapshot
lifecycle is covered in
`vignette("data-input-and-snapshots", package = "ledgr")`.

```{r}
#| label: snapshot

db_path <- ledgr_temp_store(file.path(tempdir(), "ledgr_store_demo.duckdb"))

bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% c("DEMO_01", "DEMO_02"),
    between(
      ts_utc,
      ledgr_utc("2019-01-01"),
      ledgr_utc("2019-06-30")
    )
  )

snapshot <- ledgr_snapshot_from_df(
  bars,
  db_path = db_path,
  snapshot_id = "store_demo_snapshot"
)
```

After snapshot creation, store operations take `snapshot`, not `db_path`. In a
new R session, recover the handle with
`ledgr_snapshot_open(db_path, snapshot_id)`.

## Record Two Variants For Comparison

```{r}
#| label: experiment

features <- list(ledgr_ind_sma(20))

trend_strategy <- function(ctx, params) {
  above <- ctx$vec$close > ledgr_signal_feature(ctx, "sma_20")
  ctx |>
    ledgr_selection(where = above, missing = "exclude") |>
    ledgr_target_quantity(ctx, params$qty)
}

exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = trend_strategy,
  features = features,
  opening = ledgr_opening(cash = 10000),
  cost_model = ledgr_cost_zero()
)

bt_small <- exp |>
  ledgr_run(params = list(qty = 5), run_id = "trend_qty_5")

bt_large <- exp |>
  ledgr_run(params = list(qty = 15), run_id = "trend_qty_15")
```

## Discover Runs

`ledgr_run_list()` is the store discovery view.

```{r}
#| label: list

ledgr_run_list(snapshot)
```

Both runs finished with status `DONE`. `complete_performance` and
`achieved_end_utc` are `NA` because these runs had no point-in-time rules that
could stop them early, and the footer's note about `INCOMPLETE` metrics applies
only to runs whose status says so; neither run here does.

Use labels and tags for mutable human-facing organization.

```{r}
#| label: label-tags

snapshot <- snapshot |>
  ledgr_run_label("trend_qty_5", "Baseline quantity") |>
  ledgr_run_tag("trend_qty_5", c("baseline", "trend")) |>
  ledgr_run_tag("trend_qty_15", c("trend", "larger-size"))

ledgr_run_list(snapshot)
```

Tags and labels do not alter snapshot hashes, strategy hashes, parameter hashes,
config hashes, or result artifacts.

The returned objects are still tibbles. When you need a custom view, convert to
a tibble and select the columns you want.

```{r}
#| label: list-deeper

ledgr_run_list(snapshot) |>
  as_tibble() |>
  select(run_id, label, tags, status, final_equity, execution_mode)
```

## Inspect And Compare

```{r}
#| label: info

info <- ledgr_run_info(snapshot, "trend_qty_5")
info
```

`ledgr_run_info()` is the detailed metadata view. It includes execution
mode, compact telemetry, status, identity hashes, and reproducibility tier. The
hashes identify every input and are the same whenever the same inputs run;
`Elapsed Sec` and the cache counts describe how this execution went on this
machine and can differ between otherwise identical runs.

Useful fields include:

| Field | Meaning |
|---|---|
| `run_id`, `status`, `label`, `tags`, `archived` | mutable and immutable run organization fields |
| `snapshot_id`, `snapshot_hash` | sealed data identity |
| `strategy_source_hash`, `strategy_params_hash`, `config_hash` | strategy, parameter, and run-configuration identity |
| `reproducibility_level` | strategy preflight tier recorded with the run |
| `execution_mode`, `elapsed_sec`, `pulse_count` | execution telemetry |
| `persist_features`, `feature_cache_hits`, `feature_cache_misses` | compact feature-engine telemetry |
| `error_msg` | failure diagnostic for non-completed runs |

```{r}
#| label: compare

comparison <- ledgr_run_compare(snapshot, run_ids = c("trend_qty_5", "trend_qty_15"))
comparison
```

The two runs differ only in quantity: tripling it roughly triples the profit
and the drawdown while the Sharpe ratio barely moves. Both Sharpe ratios are
annualized from half a year of two demo instruments, so they describe this
sample, not the rule. Comparison is read-only and does not rerun strategies.
`n_trades` counts closed, realised trade observations, not every fill. A run can have fills but no
closed trades yet, in which case win rate is not defined.

`ledgr_run_compare()` starts from the durable snapshot handle because it reads
stored run artifacts. When you want the comparison to use an experiment's metric
assumptions, pass that context explicitly:

```{r}
#| label: compare-exp-context

exp_comparison <- ledgr_run_compare(
  snapshot,
  run_ids = c("trend_qty_5", "trend_qty_15"),
  metric_context = ledgr_metric_context(exp)
)
```

The printed comparison formats some columns for reading, and a `select()` on
the comparison keeps that print. The columns underneath are raw numbers.
Convert with `as_tibble()` before code or a report reads them, and format
percentages yourself:

```{r}
#| label: compare-raw

comparison_report <- comparison |>
  as_tibble() |>
  select(run_id, final_equity, total_return, sharpe_ratio, max_drawdown, n_trades)

comparison_report
```

After selecting a run, reopen it and inspect the underlying result tables rather
than parsing the printed comparison:

```{r}
#| label: compare-followup

best_run_id <- comparison |>
  arrange(desc(total_return)) |>
  pull(run_id) |>
  first()

best_bt <- ledgr_run_open(snapshot, best_run_id)
tail(ledgr_results(best_bt, what = "equity"), 3)
close(best_bt)
```

## Inspect Stored Strategy Source

Completed runs keep strategy provenance in the experiment store: the source
text, parameters, hashes and dependency metadata that produced each run.
`ledgr_run_strategy(snapshot, run_id, trust = FALSE)` reads them without
executing anything; `vignette("reproducibility", package = "ledgr")` teaches
that reader and when trusted recovery is appropriate.

When a run ID is missing, store lookup helpers fail with class
`ledgr_run_not_found`, so code can tell a typo from other store errors:

```{r}
#| label: missing-run

tryCatch(
  ledgr_run_info(snapshot, "missing_run"),
  ledgr_run_not_found = function(error) class(error)[[1]]
)
```

## Reopen A Completed Run In A Later Session

`ledgr_run_open()` reconstructs a completed run handle from stored artifacts. It
does not recompute the strategy. This is useful when you want full result
tables or plots after restarting R.

```{r}
#| label: open

reopened <- ledgr_run_open(snapshot, "trend_qty_5")
summary(reopened)
tail(ledgr_results(reopened, what = "equity"), 3)
close(reopened)
```

Only completed runs can be reopened. Failed or incomplete runs remain
inspectable through `ledgr_run_info()`.

Store-level helpers such as `ledgr_run_info()`, `ledgr_run_list()`, and
`ledgr_run_compare()` use the snapshot handle and remain available after a
completed run handle is closed. Result-table helpers such as `ledgr_results()`
need a live or reopened backtest handle.

## Recovery

Most ledgr recovery workflows should start with the high-level store helpers
above:

- use `ledgr_snapshot_open()` to reopen the sealed snapshot store;
- use `ledgr_run_open()` to reopen a completed run handle;
- use `ledgr_run_info()` and `ledgr_run_strategy()` to inspect stored metadata
  and strategy provenance.

The lower-level recovery pair is public for restart inspection and for tools
that need to work directly against the store connection.

`ledgr_db_init(db_path)` opens a DBI connection to a ledgr DuckDB store and
ensures the ledgr schema exists. In normal workflows, ordinary users usually do
not need it because snapshot and run helpers open, verify, and close the
required connections for their own operations.

`ledgr_state_reconstruct(run_id, con)` reconstructs ledgr's expected simulated
state for one stored run from ledger-backed evidence. It returns a list of
positions, cash, profit and loss (`pnl`) and an equity curve (`equity_curve`)
rebuilt from the stored run records. It is useful when you are inspecting a restart boundary,
debugging stored evidence, or building a low-level tool that already owns a DBI
connection.

```{r}
#| label: low-level-recovery

con <- ledgr_db_init(db_path)
state <- ledgr_state_reconstruct("trend_qty_5", con)
DBI::dbDisconnect(con, shutdown = TRUE)

names(state)
```

This pair is intentionally not a broker or migration layer. It does not perform
broker reconciliation, prove live restart safety, migrate old schemas,
repair a sealed snapshot, or recover strategy dependencies that were never captured in
the run provenance. Treat it as low-level inspection over ledgr's own stored
evidence, not as an escape hatch around the sealed-data and provenance
contracts.

## Archive Without Deleting

```{r}
#| label: archive

snapshot <- snapshot |>
  ledgr_run_archive("trend_qty_15", reason = "larger position kept for reference")

ledgr_run_list(snapshot)
```

Archiving hides a run from default listings without deleting artifacts.

## Back Up A Closed Store

The experiment store is an ordinary DuckDB file. Back it up only after every
run and snapshot handle using it has been closed.

::: {.ledgr-callout .ledgr-callout-warning}
**Back up closed stores**

Close live handles, then copy or sync the store with the rest of the project's
research artifacts.

```r
dir.create("backups", showWarnings = FALSE)
file.copy(
  "artifacts/ledgr_store.duckdb",
  file.path("backups", paste0("ledgr_store_", Sys.Date(), ".duckdb")),
  overwrite = TRUE
)
```

Do not copy a file while another ledgr process has it open. For a larger
project, keep the same closed-file rule and use your normal versioned backup or
sync system.
:::

## Resource Cleanup

`ledgr_run()` and `ledgr_run_open()` return live handles for durable run
artifacts. The artifacts are already durable when a run completes, and ordinary
result inspection opens and closes read connections per operation. Use
`close(bt)` as explicit resource cleanup in long sessions, tests, and
explicit-open workflows. Close snapshot handles when the workflow is finished.

```{r}
#| label: cleanup
#| include: false

close(bt_small)
close(bt_large)
ledgr_snapshot_close(snapshot)
```

## Where Next

- `vignette("reproducibility", package = "ledgr")` explains what a stored run
  can and cannot reproduce, and the strategy tiers.
- `vignette("metrics-and-accounting", package = "ledgr")` covers fills, trades,
  equity rows, and metric definitions.
- `vignette("data-input-and-snapshots", package = "ledgr")` covers snapshot
  creation and sealed-data boundaries.
