---
title: "Experiment Store"
output:
  rmarkdown::html_vignette:
    toc: false
  github_document:
    toc: false
    html_preview: false
vignette: >
  %\VignetteIndexEntry{Experiment Store}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 90)
options(cli.unicode = FALSE)
default_output_hook <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(
  output = function(x, options) {
    default_output_hook(gsub("[ \t]+(?=\n)", "", x, perl = TRUE), options)
  }
)
demo_data_path <- if (file.exists("data/ledgr_demo_bars.rda")) {
  "data/ledgr_demo_bars.rda"
} else {
  file.path("..", "data", "ledgr_demo_bars.rda")
}
if (!exists("ledgr_demo_bars") && file.exists(demo_data_path)) {
  load(demo_data_path)
}
```

The experiment store is the DuckDB file behind a sealed snapshot. It keeps
market data, run artifacts, provenance, labels, tags, archive state, and compact
telemetry together.

```text
snapshot handle -> run experiments -> list / inspect / compare / reopen
```

`run_id` is immutable. Labels, tags, and archive state are mutable metadata.

The examples use `dplyr` and `tibble` for data preparation and compact display.
They are suggested packages used by the vignettes, not part of the experiment
store contract.

```{r library, message=FALSE}
library(ledgr)
library(dplyr)
library(tibble)
data("ledgr_demo_bars", package = "ledgr")
```

## Create A Durable Snapshot

Market data and derived data have different lifecycle rules in ledgr. A sealed
snapshot freezes the real market-data input and its hash. If you need more
instruments, more dates, corrected bars, or tick-derived bars, create a new
snapshot. Indicators, runs, labels, tags, comparisons, and telemetry are derived
from sealed market data and can be added later without mutating the snapshot.

This vignette uses `tempfile()` so it can run without writing into your project
directory. For real research, use a stable path such as `"research.duckdb"` and
a snapshot ID you will recognize later.

```{r snapshot}
db_path <- tempfile("ledgr_store_", fileext = ".duckdb")

bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% c("DEMO_01", "DEMO_02"),
    between(
      ts_utc,
      ledgr::ledgr_utc("2019-01-01"),
      ledgr::ledgr_utc("2019-06-30")
    )
  )

snapshot <- ledgr_snapshot_from_df(
  bars,
  db_path = db_path,
  snapshot_id = "store_demo_snapshot"
)
```

After snapshot creation, store operations take `snapshot`, not `db_path`. In a
new R session, recover the handle with `ledgr_snapshot_load(db_path,
snapshot_id)`.

If your market data starts in CSV, seal the CSV into the same kind of durable
store. The CSV must contain at least `instrument_id`, `ts_utc`, `open`, `high`,
`low`, and `close`.

```{r csv-snapshot, eval=FALSE}
snapshot <- ledgr_snapshot_from_csv(
  "data/daily_bars.csv",
  db_path = "research.duckdb",
  snapshot_id = "eod_2019_h1"
)
```

In any later session, recover the handle without re-sealing the data:

```{r csv-snapshot-load, eval=FALSE}
snapshot <- ledgr_snapshot_load("research.duckdb", snapshot_id = "eod_2019_h1")
```

Yahoo imports follow the same lifecycle, but the adapter downloads bars before
sealing the snapshot:

```{r yahoo-snapshot, eval=FALSE}
snapshot <- ledgr_snapshot_from_yahoo(
  symbols = c("SPY", "QQQ"),
  from = "2019-01-01",
  to = "2019-06-30",
  db_path = "research.duckdb",
  snapshot_id = "yahoo_2019_h1"
)
```

The returned handle is already sealed. Use `ledgr_snapshot_info(snapshot)` to
inspect `status`, `snapshot_hash`, `bar_count`, `instrument_count`,
`start_date`, `end_date`, and raw `meta_json`. The dates are ISO UTC values.
`meta_json` is envelope metadata; snapshot identity comes from normalized bars
and instruments, not from human descriptions.

## Run Variants

```{r experiment}
features <- list(ledgr_ind_sma(20))

trend_strategy <- function(ctx, params) {
  targets <- ctx$flat()
  for (id in ctx$universe) {
    sma <- ctx$feature(id, "sma_20")
    if (is.finite(sma) && ctx$close(id) > sma) {
      targets[id] <- params$qty
    }
  }
  targets
}

exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = trend_strategy,
  features = features,
  opening = ledgr_opening(cash = 10000)
)

bt_small <- exp |>
  ledgr_run(params = list(qty = 5), run_id = "trend_qty_5")

bt_large <- exp |>
  ledgr_run(params = list(qty = 15), run_id = "trend_qty_15")
```

## Discover Runs

`ledgr_run_list()` is the store discovery view.

```{r list}
ledgr_run_list(snapshot)
```

Use labels and tags for mutable human-facing organization.

```{r label-tags}
snapshot <- snapshot |>
  ledgr_run_label("trend_qty_5", "Baseline quantity") |>
  ledgr_run_tag("trend_qty_5", c("baseline", "trend")) |>
  ledgr_run_tag("trend_qty_15", c("trend", "larger-size"))

ledgr_run_list(snapshot)
```

Tags and labels do not alter snapshot hashes, strategy hashes, parameter hashes,
config hashes, or result artifacts.

The returned objects are still tibbles. When you need a custom view, convert to
a tibble and select the columns you want.

```{r list-deeper}
ledgr_run_list(snapshot) |>
  as_tibble() |>
  select(run_id, label, tags, status, final_equity, execution_mode)
```

## Inspect And Compare

```{r info}
info <- ledgr_run_info(snapshot, "trend_qty_5")
info
```

`ledgr_run_info()` is the detailed metadata view. It includes execution mode,
compact telemetry, status, identity hashes, and reproducibility tier.

Useful fields include:

| Field | Meaning |
|---|---|
| `run_id`, `status`, `label`, `tags`, `archived` | mutable and immutable run organization fields |
| `snapshot_id`, `snapshot_hash`, `data_hash` | sealed data identity |
| `strategy_source_hash`, `strategy_params_hash`, `config_hash` | strategy, parameter, and run-configuration identity |
| `reproducibility_level` | strategy preflight tier recorded with the run |
| `execution_mode`, `elapsed_sec`, `pulse_count` | execution telemetry |
| `persist_features`, `feature_cache_hits`, `feature_cache_misses` | compact feature-engine telemetry |
| `error_msg` | failure diagnostic for non-completed runs |

```{r compare}
comparison <- ledgr_compare_runs(snapshot, run_ids = c("trend_qty_5", "trend_qty_15"))
comparison
```

Comparison is read-only and does not rerun strategies. `n_trades` counts
closed, realised trade observations, not every fill. A run can have fills but no
closed trades yet, in which case win rate is not defined.

The printed comparison formats some columns for reading. Programmatic code gets
raw numeric columns from the tibble:

```{r compare-raw}
comparison |>
  select(run_id, final_equity, total_return, sharpe_ratio, max_drawdown, n_trades)
```

After selecting a run, reopen it and inspect the underlying result tables rather
than parsing the printed comparison:

```{r compare-followup}
best_run_id <- comparison |>
  arrange(desc(total_return)) |>
  pull(run_id) |>
  first()

best_bt <- ledgr_run_open(snapshot, best_run_id)
tail(ledgr_results(best_bt, what = "equity"), 3)
close(best_bt)
```

## Inspect Stored Strategy Source

Completed runs keep strategy provenance in the experiment store. This is one of
the most useful audit artifacts: you can inspect the source text that produced a
run without reopening the backtest handle and without rerunning the strategy.
The full trust and tier model lives in `vignette("reproducibility", package =
"ledgr")`; this section shows the store workflow.

Use `trust = FALSE` for safe inspection. It returns stored source text,
parameters, hashes, dependency metadata, and warnings without parsing,
evaluating, or executing the source.

```{r extract-source}
stored_strategy <- ledgr_extract_strategy(snapshot, "trend_qty_5", trust = FALSE)
stored_strategy
```

The source text is just data in this mode.

```{r extract-source-text}
writeLines(stored_strategy$strategy_source_text)
```

Hash verification proves stored-text identity, not code safety. Use
`trust = TRUE` only when you already trust the experiment store and
intentionally want ledgr to parse and evaluate the stored text into a function
object. Legacy/pre-provenance runs remain inspectable through `ledgr_run_info()`
and stored result tables, but their strategy function cannot be recovered from
provenance alone.

When a run ID is missing, store lookup helpers fail with class
`ledgr_run_not_found`:

```{r missing-run, eval=FALSE}
ledgr_run_info(snapshot, "missing_run")
```

Trusted recovery can be used to rerun a stored strategy only after you have
decided that evaluating the stored source is acceptable:

```{r trusted-rerun, eval=FALSE}
recovered <- ledgr_extract_strategy(snapshot, "trend_qty_5", trust = TRUE)

rerun_exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = recovered$strategy_function,
  features = features,
  opening = ledgr_opening(cash = 10000)
)

ledgr_run(
  rerun_exp,
  params = recovered$strategy_params,
  run_id = "trend_qty_5_rerun"
)
```

## Reopen A Completed Run In A Later Session

`ledgr_run_open()` reconstructs a completed run handle from stored artifacts. It
does not recompute the strategy. This is useful when you want full result
tables or plots after restarting R.

```{r open}
reopened <- ledgr_run_open(snapshot, "trend_qty_5")
summary(reopened)
tail(ledgr_results(reopened, what = "equity"), 3)
close(reopened)
```

Only completed runs can be reopened. Failed or incomplete runs remain
inspectable through `ledgr_run_info()`.

Store-level helpers such as `ledgr_run_info()`, `ledgr_run_list()`, and
`ledgr_compare_runs()` use the snapshot handle and remain available after a
completed run handle is closed. Result-table helpers such as `ledgr_results()`
need a live or reopened backtest handle.

## Archive Without Deleting

```{r archive}
snapshot <- snapshot |>
  ledgr_run_archive("trend_qty_15", reason = "larger position kept for reference")

ledgr_run_list(snapshot)
```

Archiving hides a run from default listings without deleting artifacts.

## Bridge A Low-Level CSV Import

This is advanced import material. The high-level CSV helper above is the normal
path. The lower-level path is useful when you want to create the snapshot row,
import one or more CSV files, inspect the sealed metadata, and then load the
sealed artifact in a separate step. A future article, "Data Input And Snapshot
Creation", may move this bridge out of the store workflow.

The order is important:

1. create the snapshot envelope;
2. import bars into that CREATED snapshot;
3. seal it to validate bars and write the snapshot hash;
4. load it with `verify = TRUE`;
5. pass the loaded snapshot to `ledgr_experiment()` and `ledgr_run()`.

```{r low-level-csv-bridge}
csv_db_path <- tempfile("ledgr_csv_bridge_", fileext = ".duckdb")
csv_bars_path <- tempfile("ledgr_csv_bars_", fileext = ".csv")

csv_bars <- bars |>
  filter(ts_utc <= ledgr::ledgr_utc("2019-01-15")) |>
  mutate(ts_utc = format(ts_utc, "%Y-%m-%dT%H:%M:%SZ", tz = "UTC"))

utils::write.csv(csv_bars, csv_bars_path, row.names = FALSE)

csv_con <- ledgr_db_init(csv_db_path)
csv_snapshot_id <- ledgr_snapshot_create(
  csv_con,
  snapshot_id = "csv_bridge_snapshot",
  meta = list(description = "low-level CSV bridge demo")
)

ledgr_snapshot_import_bars_csv(
  csv_con,
  csv_snapshot_id,
  bars_csv_path = csv_bars_path,
  instruments_csv_path = NULL,
  auto_generate_instruments = TRUE
)

csv_hash <- ledgr_snapshot_seal(csv_con, csv_snapshot_id)
csv_info <- ledgr_snapshot_info(csv_con, csv_snapshot_id)
DBI::dbDisconnect(csv_con, shutdown = TRUE)

csv_hash
csv_info |>
  select(
    snapshot_id,
    status,
    snapshot_hash,
    bar_count,
    instrument_count,
    start_date,
    end_date,
    meta_json
  )
```

`bar_count` and `instrument_count` are live counts from the sealed snapshot
tables. The raw `meta_json` is envelope metadata on the snapshot row; seal-time
metadata inside it uses `n_bars` and `n_instruments`. Snapshot identity does not
come from that metadata. `snapshot_hash` identifies the normalized bars and
instruments only, so adding a human description to `meta_json` does not change
the artifact hash.

Load the sealed snapshot before constructing the experiment. This is the same
handle you would use in a later R session.

```{r low-level-csv-run}
csv_snapshot <- ledgr_snapshot_load(
  csv_db_path,
  snapshot_id = csv_snapshot_id,
  verify = TRUE
)

csv_strategy <- function(ctx, params) {
  targets <- ctx$flat()
  targets["DEMO_01"] <- params$qty
  targets
}

csv_exp <- ledgr_experiment(
  snapshot = csv_snapshot,
  strategy = csv_strategy,
  opening = ledgr_opening(cash = 10000)
)

csv_bt <- ledgr_run(csv_exp, params = list(qty = 1), run_id = "csv_bridge_run")
tail(ledgr_results(csv_bt, what = "equity"), 3)
```

## Current Feature Persistence Boundary

Run metadata records whether feature persistence was enabled, and pulse
inspection lets you view registered feature values at one decision time. Public
feature inspection in v0.1.7.9 is intentionally scoped to feature contracts,
warmup feasibility, and pulse-time feature views:

- `ledgr_feature_contracts(features)` shows declared feature requirements;
- `ledgr_feature_contract_check(snapshot, features)` checks whether those
  requirements are achievable in a sealed snapshot;
- `ledgr_pulse_snapshot()` plus `ledgr_pulse_features()` or
  `ledgr_pulse_wide()` inspects one pulse.

A full persisted feature-series retrieval API is deferred to the v0.1.8
precompute and sweep-result design.

`ledgr_run()` and `ledgr_run_open()` return live handles for durable run
artifacts. The artifacts are already durable when a run completes, and ordinary
result inspection opens and closes read connections per operation. Use
`close(bt)` as explicit resource cleanup in long sessions, tests, explicit-open
workflows, and lazy result cursors. Close snapshot handles when the workflow is
finished.

```{r cleanup}
close(bt_small)
close(bt_large)
close(csv_bt)
ledgr_snapshot_close(csv_snapshot)
ledgr_snapshot_close(snapshot)
unlink(csv_bars_path)
```

## What's Next?

For fills, trades, equity rows, and metric definitions, read
`vignette("metrics-and-accounting", package = "ledgr")`. For strategy authoring,
read `vignette("strategy-development", package = "ledgr")`.
