---
title: "Indicators And Features"
format:
  html:
    toc: false
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{Indicators And Features}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(width = 90)
# Local renders load ledgr from source; installed builds use the library call
# in the prerequisites chunk.
pkg_root <- if (file.exists("DESCRIPTION")) "." else if (file.exists("../DESCRIPTION")) ".." else NA_character_
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
demo_data_path <- if (file.exists("data/ledgr_demo_bars.rda")) {
  "data/ledgr_demo_bars.rda"
} else {
  file.path("..", "data", "ledgr_demo_bars.rda")
}
if (!exists("ledgr_demo_bars") && file.exists(demo_data_path)) {
  load(demo_data_path)
}
```

```{=html}
<style>
.ledgr-diagram {
  margin: 1.25rem auto 1.5rem auto;
  text-align: center;
}
.ledgr-diagram .mermaid {
  display: inline-block;
  max-width: 760px;
  width: 100%;
}
.ledgr-diagram .node text,
.ledgr-diagram .edgeLabel {
  font-size: 18px !important;
}
</style>
```
Your strategy needs a moving average, a return, or an RSI at each decision.
By the end of this article you can declare those features once, predict their
IDs and warmup, and read their pulse-known values inside a strategy without
ever reaching into future data.

An **indicator** is a declared feature computation. ledgr computes indicators
into pulse-known values; scalar accessors, mapped accessors, long tables, and
wide tables are different views of that same pulse-known data. That model is
the same for built-in ledgr indicators, TTR-backed indicators, and custom
indicators.

## Prerequisites

```{r}
#| label: packages
#| message: false

library(ledgr)
library(dplyr)
data("ledgr_demo_bars", package = "ledgr")
```

## Start With Built-In Features

Use two demo instruments and two built-in features. The economic idea will be
small on purpose:

> Own an instrument only when its recent return is positive enough and today's
> close is above its moving average.

That rule needs one momentum feature and one trend feature. A feature map gives
your strategy code readable aliases while preserving ledgr's exact engine
feature IDs.

```{r}
#| label: built-in-feature-map

features <- ledgr_feature_map(
  ret_5 = ledgr_ind_returns(5),
  sma_10 = ledgr_ind_sma(10)
)
```

A **feature ID** is ledgr's engine-facing name for a computed value. An
**alias** is the strategy-facing name you choose in a feature map. Use aliases
such as `ret_5` or `sma_10` when you write strategy logic with
`ctx$features()`. Use feature IDs such as `return_5` or `sma_10` when you need
the explicit engine contract, for example with `ctx$feature()` or when
inspecting stored feature metadata.

Feature objects appear in three registration and inspection places:

| Surface | Accepted feature shape | How names are used |
|---|---|---|
| `ledgr_experiment(features = ...)` | indicator, list, named list, or feature map | registers feature definitions for the run |
| `ledgr_feature_contracts()` / `ledgr_feature_contract_check()` | static indicator, list, named list, or feature map | reports aliases and engine IDs |
| `ledgr_pulse_snapshot(features = ...)` | static list or feature map | computes pulse-known values for inspection |

The strategy context then exposes the computed values through accessors:

| Accessor | Name type | Use |
|---|---|---|
| `ctx$vec$feature(feature_id)` | engine feature ID string | reads one feature across the whole decision axis |
| `ctx$feature(id, feature_id)` | engine feature ID string | reads one scalar value by exact ID |
| `ctx$features(id, feature_map)` | feature map | returns a named vector keyed by alias |

The canonical strategy workflow is: register features on
`ledgr_experiment()`, resolve the engine ID once, then read the whole decision
axis with `ctx$vec$feature()`. Keep `ctx$feature()` and `ctx$features()` for
one-instrument inspection and for active aliases, whose concrete engine IDs
vary by candidate.

## Feature Lifecycle: From Declaration To Lookup

The feature path has five steps:

::: {.ledgr-diagram .ledgr-feature-lifecycle}

```{mermaid}
%%| echo: false

flowchart LR
  declare["declare<br/>indicator or map"]
  register["register<br/>experiment"]
  compute["compute<br/>pulse-known values"]
  access["access<br/>ctx feature methods"]
  decide["decide<br/>target holdings"]

  declare --> register --> compute --> access --> decide
```

:::

1. You declare features in any of several shapes: individual indicators,
   built-in helpers, TTR adapters, CSV/R adapters, feature maps, or
   active-alias parameterizations. All resolve to the same lifecycle below.
2. `ledgr_experiment()` stores the declaration. Static lists and feature maps
   are ready immediately. Active-alias features are materialized for concrete
   feature-grid values before candidate execution.
3. Optional `ledgr_precompute_features()` resolves a parameter grid, computes
   each candidate's concrete feature set, deduplicates shared indicator
   fingerprints, and records candidate feature-set hashes.
4. During `ledgr_run()` or `ledgr_sweep()`, the engine computes the
   registered feature values at each pulse without looking past the current
   bar.
5. Strategy code normally reads one feature across the decision axis with
   `ctx$vec$feature()`. Scalar accessors remain useful for inspection and for
   active aliases that have no whole-universe alias accessor yet.

A **fingerprint** identifies the feature definition, not just the name. If the
calculation, parameters, warmup rule, adapter, or selected output changes, the
fingerprint changes even when the feature ID stays readable. Feature IDs
identify values inside the pulse context. Fingerprints identify the feature
definition used to compute those values. For multi-output sources such as TTR
`BBands` or `MACD`, each selected output is an ordinary indicator with its own
feature ID and output-specific fingerprint.

Feature IDs in one map must be unique. An alias does not make two different
definitions with the same engine ID safe, and an alias may not equal another
entry's feature ID. It may equal its own feature ID. These fail-closed rules
keep scalar engine-ID reads and mapped alias reads from naming different
series. An alias never changes the underlying feature ID or fingerprint.

A **bundle** is an authoring convenience for declaring several indicator
outputs at once. The engine receives ordinary single-output feature definitions
after the bundle is flattened. The multi-output bundle helper follows the same
lifecycle. It is not a second feature system.

The same idea works for crossover rules. An SMA crossover registers two
separate indicators: one short moving average and one long moving average. The
economic meaning is "fast trend above slow trend" rather than "close above one
trend line." Each moving average has its own feature ID, warmup, and stored
values.

```{r}
#| label: sma-crossover-contracts

crossover_features <- ledgr_feature_map(
  sma_fast = ledgr_ind_sma(10),
  sma_slow = ledgr_ind_sma(30)
)
crossover_ids <- ledgr_feature_id(crossover_features)

ledgr_feature_contracts(crossover_features)
```

In a strategy, the crossover condition is just a comparison of the two mapped
aliases after warmup:

```{r}
#| label: sma-crossover-snippet

crossover_strategy <- function(ctx, params) {
  fast <- ctx$vec$feature(crossover_ids[["sma_fast"]])
  slow <- ctx$vec$feature(crossover_ids[["sma_slow"]])
  ready <- is.finite(fast) & is.finite(slow)

  targets <- ctx$flat()
  targets[ready & fast > slow] <- params$qty
  targets
}
```

## Inspect One Pulse

Create a small sealed snapshot and inspect one decision pulse before running a
full backtest. This keeps the runtime data visible before the article introduces
more metadata.

```{r}
#| label: pulse-setup

bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% c("DEMO_01", "DEMO_02"),
    between(
      ts_utc,
      ledgr_utc("2019-01-01"),
      ledgr_utc("2019-06-30")
    )
  )

snapshot <- ledgr_snapshot_from_df(
  bars,
  snapshot_id = "indicators_demo"
)

pulse <- ledgr_pulse_snapshot(
  snapshot,
  universe = c("DEMO_01", "DEMO_02"),
  ts_utc = ledgr_utc("2019-03-01"),
  features = features
)
```

At this timestamp, ledgr has computed the two features for each instrument in
the universe through the same feature path a dense run uses over snapshot
history through this timestamp. That includes each declaration's parameters,
`stable_after`, and `series_fn` when present. The long pulse view shows the
result directly: one row per instrument and feature. Without a feature map,
`alias` is `NA`. With the map, rows are filtered to the mapped features and
aliases are filled.

```{r}
#| label: pulse-long

ledgr_pulse_features(pulse, features)
```

The wide pulse view is useful for debugging and ML-style row-per-observation
workflows. It contains one OHLCV block and one feature block for each
instrument. OHLCV columns use `{instrument_id}__ohlcv_{field}`. Feature columns
use `{instrument_id}__feature_{feature_id}`. A feature map filters and orders
feature columns, using aliases as the wide feature keys.

```{r}
#| label: pulse-wide

ledgr_pulse_wide(pulse, features)
```

`ledgr_pulse_features()` and `ledgr_pulse_wide()` work on interactive pulse
snapshots and on the `ctx` object inside an ordinary strategy function. They are
inspection views over the same pulse-known data used by `ctx$feature()` and
`ctx$features()`.

## Access Features In A Strategy

The long and wide pulse views are useful when you want to inspect the computed
data, compare instruments, or think in model-like rows. They are not always the
clearest shape for strategy code. A strategy often wants to ask a smaller
question: "what are the current values for this instrument?"

That is why ledgr also exposes the same pulse data through scalar and mapped
accessors. The table views and the accessors are not competing APIs; they are
different views over the same pulse-known data.

The explicit scalar accessor is useful when you want to show or debug one
value. It uses the engine ID, not the alias:

```{r}
#| label: scalar-access

ids <- ledgr_feature_id(features)
pulse$feature("DEMO_01", ids[["ret_5"]])
```

Mapped access returns a named numeric vector keyed by alias for one instrument
at one pulse:

```{r}
#| label: mapped-access

x <- pulse$features("DEMO_01", features)
x
ledgr_passed_warmup(x)
```

**Warmup** is the period before a known feature has enough prior bars to
produce a usable value. A warmup `NA` is not an unknown feature; it is a known
feature that is not ready yet.

::: {.ledgr-callout .ledgr-callout-tip}
**Try it**

Change the scalar accessor to `pulse$feature("DEMO_01", "returns_5")`. Why does
that fail while `ids[["ret_5"]]` works?
:::

A strategy decides for every instrument at once. `ctx$vec$feature()` reads one
feature for the whole universe in a single call, one value per instrument in
the order of `ctx$universe`. It takes the engine ID, which is what `ids`
already holds:

```{r}
#| label: strategy-definition

strategy <- function(ctx, params) {
  ret_5 <- ctx$vec$feature(ids[["ret_5"]])
  sma_10 <- ctx$vec$feature(ids[["sma_10"]])
  ready <- !is.na(ret_5) & !is.na(sma_10)
  qualifies <- ready & ret_5 > params$min_return & ctx$vec$close > sma_10

  targets <- ctx$flat()
  targets[qualifies] <- params$qty
  targets
}
```

That pattern keeps the signal logic readable:

- `features` is where feature identity and aliases live; `ids` turns an alias
  into the engine ID.
- `ready` is the warmup gate: an instrument qualifies only once both of its
  features are usable. For one instrument, `ledgr_passed_warmup()` asks the
  same question of the vector `ctx$features()` returns.
- The rest of the condition is the economic rule.
- The strategy still returns ordinary target quantities.

The rule reads the feature vectors by hand because reading them is this
section's lesson. `ledgr_signal_feature(ctx, ids[["ret_5"]])` reads the same
vector for the helper pipeline that
`vignette("strategy-development", package = "ledgr")` shows.

`ctx$features()` in a loop over `ctx$universe` would give the same targets, but
it repeats per-instrument work at every pulse. Keep the scalar and mapped
accessors for inspecting one instrument.

The strategy reads `ids` from the code around it, so it closes over a value
built from the feature map. The run records the registered feature definitions,
but the stored strategy source refers to `ids` only by name. Keep the code that
builds `features` and `ids` with your research record;
`vignette("reproducibility", package = "ledgr")` explains why such a strategy is
Tier 2.

## Run The Example

The experiment registers the indicator objects. ledgr computes those features
for every instrument at each pulse, then gives the strategy only the pulse-time
values.

```{r}
#| label: run-example

exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = strategy,
  features = features,
  opening = ledgr_opening(cash = 10000),
  cost_model = ledgr_cost_zero()
)

bt <- exp |>
  ledgr_run(
    params = list(min_return = 0, qty = 10),
    run_id = "indicators_demo_run"
  )

ledgr_results(bt, what = "fills")

close(pulse)
close(bt)
```

The fill table is the first execution check: it shows which instruments
actually crossed from a target into a position, at which next-open price. An
empty table would be a prompt to inspect warmup and signal values before trying
more parameters.

## Read The Feature Contracts

After you have seen the feature values at a pulse, the contract table is easier
to read. The feature contracts are what ledgr will compute for every instrument
in the run. `alias` is for your strategy code. `feature_id` is the stable engine ID. Warmup metadata tells you when a known feature may still be `NA`.

```{r}
#| label: feature-contracts

ledgr_feature_contracts(features)
```

Plain lists remain valid too. For a named list, names become aliases in the
contract table. For an unnamed list, `alias` is `NA`.

```{r}
#| label: plain-list-contracts

plain_features <- list(ledgr_ind_returns(5), ledgr_ind_sma(10))
ledgr_feature_contracts(plain_features)
```

## Parameter Grids Register Every Needed Feature

If a parameter grid changes a lookback, register every lookback variant before
the run. ledgr does not create indicators dynamically from `params`; the run
only computes the feature contracts registered on the experiment.

```{r}
#| label: parameter-grid-features

swept_features <- ledgr_feature_map(
  ret_5 = ledgr_ind_returns(5),
  ret_10 = ledgr_ind_returns(10),
  ret_20 = ledgr_ind_returns(20)
)

ledgr_feature_id(swept_features)
```

The feature set must cover the whole grid: `lookback = 20` means `return_20`
must already be registered. A missing feature ID is an unknown-feature error,
not warmup. In short, all feature parameter values must be registered before
`ledgr_run()`;
do not create `ledgr_ind_returns(params$lookback)` lazily inside the strategy.

For exploratory sweeps over ledgr-owned indicator parameters, prefer active
aliases. Declare the varying constructor argument with `ledgr_param()`; each
sweep candidate then resolves its own concrete indicator behind a stable alias:

```{r}
#| label: active-alias-declaration

active_features <- ledgr_feature_map(
  fast = ledgr_ind_sma(ledgr_param("fast_n")),
  slow = ledgr_ind_sma(ledgr_param("slow_n"))
)

ledgr_snapshot_close(snapshot)
```

`vignette("sweeps", package = "ledgr")` composes the feature and strategy grids
for this map, runs the sweep and reads the candidate rows.

For single-output indicators, the feature-map alias is the strategy-facing name
returned by `ctx$features(id)`. Active aliases currently have no alias-aware
whole-universe accessor, so an active-alias strategy must use that scalar mapped
read per instrument. Fixed feature IDs should use `ctx$vec$feature()` instead.
Bundle entries are intentionally flat; the TTR companion explains why bundles
use `prefix` or `naming` rather than one outer alias.

For TTR-backed declarations, multi-output bundles, and adapter warmup rules,
read `vignette("ttr-indicators", package = "ledgr")`.

## Troubleshoot Warmup And Zero Trades

Warmup problems are easiest to diagnose by connecting four facts:

1. `ledgr_feature_contracts(features)` tells you how many bars each feature
   needs before it can produce a usable value.
2. `ledgr_feature_contract_check(snapshot, features)` joins those contracts to
   the actual per-instrument bar counts in the snapshot.
3. `ledgr_pulse_features(pulse, features)` shows the current pulse-known
   values for the instruments and aliases you registered.
4. `summary(bt)` prints `Warmup Diagnostics` when a completed run has
   registered features that can never become usable for an instrument because
   available bars are below the feature contract.

```{r}
#| label: feature-contract-check

warmup_check_snapshot <- ledgr_snapshot_from_df(
  bars |>
    group_by(instrument_id) |>
    filter(instrument_id != "DEMO_02" | row_number() <= 4L) |>
    ungroup(),
  snapshot_id = "indicators_warmup_check"
)

ledgr_feature_contract_check(warmup_check_snapshot, features) |>
  select(alias, instrument_id, available_bars, stable_after, warmup_achievable)

ledgr_snapshot_close(warmup_check_snapshot)
```

The `warmup_achievable` column is `FALSE` when an instrument does not have
enough available bars to satisfy a feature's `stable_after` contract.

Normal early warmup is temporary: a feature is `NA` near the beginning of an
instrument's sample and later becomes finite. Impossible warmup is different:
the instrument never has enough available bars for that feature. In that case,
zero trades can be a valid completed run plus a useful diagnostic, not a failed
run.

### Diagnose A Successful Run With Zero Trades

A completed run with zero trades is not automatically wrong. It means ledgr
accepted the strategy outputs and the ledger reached the end of the sample, but
no closed round trips were recorded.

Use this checklist before changing the strategy:

1. Start with `summary(bt)`. If `Closed Trades` is zero, `win_rate` and
   `avg_trade` should be `NA`, not zero. If a registered feature can never
   become usable because the sample is too short, `summary(bt)` also prints a
   `Warmup Diagnostics` note with the feature ID, instrument ID, required bars,
   and available bars.
2. Inspect `ledgr_results(bt, what = "fills")`. Empty fills mean nothing ever
   executed. Non-empty fills with empty trades mean positions opened but did
   not close.
3. Confirm the feature IDs with `ledgr_feature_id(features)`. A helper such as
   `ledgr_signal_return(ctx, lookback = 60)` reads `return_60`; that indicator must
   be registered before the run.
4. Compare feature contracts with sample length before assuming the strategy is
   broken. `ledgr_feature_contracts(features)` shows `requires_bars` and
   `stable_after`. If an instrument has fewer available bars than a feature's
   `requires_bars`, that feature cannot become usable for that instrument.
5. Inspect a late pulse with `ledgr_pulse_snapshot()`. If the feature is still
   `NA` for every instrument near the end of the sample, the issue is no longer
   ordinary early warmup. Check the lookback length, sample length, universe,
   and feature registration.
6. If the helper pipeline returns a `ledgr_empty_selection` on a late diagnostic
   pulse, inspect the signal values directly. Ordinary early warmup should have
   passed by then; a late all-missing signal usually points to sample length,
   universe, or feature-registration issues.

Timestamp checks should compare normalized UTC values, not local string
representations. The chunk is a fragment for use inside a strategy or on a
pulse you have built, so it is not executed here:

```{r}
#| label: timestamp-troubleshooting
#| eval: false

target_ts <- ledgr_utc("2020-01-05T00:00:00Z")
ledgr_utc(ctx$ts_utc) == target_ts

intraday_time <- format(ledgr_utc(ctx$ts_utc), "%H:%M:%S", tz = "UTC")
intraday_time >= "14:30:00" && intraday_time <= "21:00:00"
```

If fills are empty, distinguish zero signals from zero sizing. In your own run,
use the same snapshot, universe, timestamp, features, strategy, and params you
are debugging. The `...` placeholders stand for those values, so this fragment
is not executed here:

```{r}
#| label: zero-sizing-troubleshooting
#| eval: false

pulse <- ledgr_pulse_snapshot(
  snapshot,
  universe = ...,
  ts_utc = ...,
  features = features
)
target <- strategy(pulse, params)
target
all(target == 0)
```

Custom fill-model contract errors are different from zero-trade outcomes; see
the fill-model reference documentation for the required fill fields.

#### Four Warmup-Adjacent Cases

Warmup is per instrument. One instrument can have a usable value while another
is still `NA` because it has fewer bars or a different data history.

Ordinary feature warmup is local to the beginning of each instrument's usable
sample. A known feature is `NA` for early pulses, then becomes usable once the
feature contract has enough bars.

Impossible warmup is different: every value for an instrument/feature remains
`NA` because the instrument never reaches the feature contract. That is the
case reported by the `Warmup Diagnostics` note in `summary(bt)`.

Current-bar absence is a third failure mode. If ledgr cannot construct the pulse
sequence because a current bar is absent for an instrument in the requested
universe, the run fails before strategy evaluation for that incomplete pulse;
this is a pulse construction error, not a feature warmup value.

## Expected Sessions And Certified Gap Behavior

The dense lifecycle above is the entrance. Availability-aware research adds a
stricter question: which sessions should have contained an observation? A
missing observation on a declared open session invalidates every bounded
window that contains it. ledgr does not compress the window, carry a stale
value into it, or call the gap ordinary startup warmup.

```{r}
#| label: strict-gap-sma-sources
#| eval: !expr requireNamespace("TTR", quietly = TRUE)

strict_smas <- ledgr_feature_map(
  built_in = ledgr_ind_sma(2),
  public_ttr = ledgr_ind_ttr("SMA", input = "close", n = 2),
  custom = ledgr_indicator(
    "custom_sma_2",
    function(window) mean(window$close),
    requires_bars = 2,
    gap_contract = "strict_window"
  )
)

ledgr_feature_contracts(strict_smas)
```

In the strict-gap table in
`vignette("missing-data-and-sessions", package = "ledgr")`, all three follow
the same `NA, 11, NA, NA, 15, 17` availability path. That is session
equivalence, not a claim that different formulas are numerically identical.

<!-- strict-gap-support:start -->
| Form | Declaration | Availability-aware status |
| --- | --- | --- |
| Built-in SMA | `ledgr_ind_sma(n)` | Supported |
| Built-in returns | `ledgr_ind_returns(n)` | Supported |
| Custom bounded window | `ledgr_indicator(..., gap_contract = "strict_window")` | Supported when the declaration is truthful |
| Public TTR SMA | `ledgr_ind_ttr("SMA", input = "close", n = n)` | Supported for this exact single-output shape |
| Built-in EMA or RSI | `ledgr_ind_ema(n)` / `ledgr_ind_rsi(n)` | Unsupported |
| TTR EMA or RSI | `ledgr_ind_ttr("EMA", ...)` / `ledgr_ind_ttr("RSI", ...)` | Unsupported |
| TTR output bundle | `ledgr_ind_ttr_outputs(...)` | Unsupported |
| Other TTR signatures | any shape not matching the public TTR SMA row | Unsupported |
<!-- strict-gap-support:end -->

Unsupported declarations fail with `ledgr_indicator_gap_unsupported` before
the strategy executes. They are never silently evaluated under dense
semantics.

For result-table interpretation after a zero-trade run, read
`vignette("metrics-and-accounting", package = "ledgr")`.


## Where Next

- `vignette("leakage", package = "ledgr")` shows how a feature can leak future
  information and which boundaries ledgr enforces.
- `vignette("ttr-indicators", package = "ledgr")` covers TTR-backed indicators
  and multi-output bundles.
- `vignette("custom-indicators", package = "ledgr")` covers writing your own
  indicators.
- `vignette("strategy-development", package = "ledgr")` uses feature values
  inside complete strategies.
