---
title: "Indicators And Feature IDs"
output:
  rmarkdown::html_vignette:
    toc: false
  github_document:
    toc: false
    html_preview: false
    md_extensions: -smart
vignette: >
  %\VignetteIndexEntry{Indicators And Feature IDs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(width = 90)
article_utc <- function(x) as.POSIXct(x, tz = "UTC")
demo_data_path <- if (file.exists("data/ledgr_demo_bars.rda")) {
  "data/ledgr_demo_bars.rda"
} else {
  file.path("..", "data", "ledgr_demo_bars.rda")
}
if (!exists("ledgr_demo_bars") && file.exists(demo_data_path)) {
  load(demo_data_path)
}
```

```{r packages, message=FALSE}
library(ledgr)
library(dplyr)
data("ledgr_demo_bars", package = "ledgr")
```

Indicators are how ledgr turns sealed market data into pulse-time features.
This article teaches built-in ledgr indicators and TTR-backed indicators under
one mental model.

The important contract is the same for built-in indicators and TTR-backed
indicators:

1. Define indicator objects before the run.
2. Register those objects with the experiment.
3. Ask ledgr for their exact feature IDs.
4. Read those IDs from `ctx$feature()` inside the strategy.
5. Treat warmup `NA` as "known feature, not usable yet."

The strategy never receives the whole feature table. At each pulse it receives
only the current feature values that could have been known at that time.

## Built-In Indicators

Built-in indicators cover common ledgr-native features. They are ordinary
indicator objects with deterministic IDs and warmup requirements.

```{r built-in-indicators}
builtins <- list(
  sma_20 = ledgr_ind_sma(20),
  ema_20 = ledgr_ind_ema(20),
  rsi_14 = ledgr_ind_rsi(14),
  ret_5 = ledgr_ind_returns(5)
)

ledgr_feature_id(builtins)
```

The names in `builtins` are for your R code. The strings returned by
`ledgr_feature_id()` are ledgr's feature IDs. Strategies use the returned
feature IDs, not guessed strings.

For example:

```r
ctx$feature("DEMO_01", "sma_20")
ctx$feature("DEMO_01", "return_5")
```

## TTR-Backed Indicators

`ledgr_ind_ttr()` is the adapter for supported indicators from the suggested
`TTR` package. TTR stays outside the core engine:

```text
TTR -> ledgr_ind_ttr() -> ledgr_indicator -> deterministic pulse engine
```

The engine sees a normal `ledgr_indicator`. That means TTR-backed indicators
follow the same feature-ID and warmup rules as built-in indicators.

```{r ttr-backed-indicators, eval=requireNamespace("TTR", quietly = TRUE)}
ttr_features <- list(
  ttr_rsi = ledgr_ind_ttr("RSI", input = "close", n = 14),
  bb_up = ledgr_ind_ttr("BBands", input = "close", output = "up", n = 20),
  macd = ledgr_ind_ttr(
    "MACD",
    input = "close",
    output = "macd",
    nFast = 12,
    nSlow = 26,
    nSig = 9,
    percent = FALSE
  ),
  macd_signal = ledgr_ind_ttr(
    "MACD",
    input = "close",
    output = "signal",
    nFast = 12,
    nSlow = 26,
    nSig = 9,
    percent = FALSE
  )
)

ledgr_feature_id(ttr_features)
```

The examples produce IDs such as `ttr_bbands_20_up`,
`ttr_macd_12_26_9_false_macd`, and
`ttr_macd_12_26_9_false_signal`.

Some TTR functions return several columns. For those functions, choose one
column with `output` before asking ledgr for the feature ID. `BBands` exposes
`dn`, `mavg`, `up`, and `pctB`. `MACD` exposes `macd` and `signal`; ledgr also
supports a derived `histogram`.

The two MACD examples above use matching explicit arguments. Explicit arguments
become part of the feature ID, so combine MACD outputs in one strategy only when
their argument sets match the computation you intend. If one MACD output uses
`percent = FALSE`, the paired `signal` output should usually set
`percent = FALSE` too.

## Warmup Is General

Every ledgr indicator declares `requires_bars` and `stable_after`.
`requires_bars` is the first row where the indicator may produce a non-`NA`
value. `stable_after` is the first row ledgr treats as stable.

Built-in and supported TTR-backed indicators use the same rule: before warmup
has passed, the feature value is `NA`.

```{r warmup-builtins}
data.frame(
  alias = names(builtins),
  feature_id = ledgr_feature_id(builtins),
  requires_bars = vapply(builtins, function(x) x$requires_bars, integer(1)),
  stable_after = vapply(builtins, function(x) x$stable_after, integer(1)),
  row.names = NULL
)
```

Warmup `NA` is expected. Unknown feature IDs are different:
`ctx$feature()` fails loudly when the ID was not registered with the experiment.

TTR warmup inference is inspectable:

```{r ttr-warmup-rules, message=FALSE, eval=requireNamespace("TTR", quietly = TRUE)}
ledgr_ttr_warmup_rules() |>
  select(ttr_fn, input, formula)
```

For MACD, ledgr verifies the supported warmup rules against direct TTR output.
With current TTR behavior, the `macd` output is first valid at `nSlow`, while
`signal` and the derived ledgr `histogram` output are first valid at
`nSlow + nSig - 1`. The same rule is verified for both `percent = TRUE` and
`percent = FALSE`.

## Use Features In A Strategy

Keep feature definitions and feature IDs together. A named list gives readable
names to your R code, while `ledgr_feature_id()` gives the exact IDs used by the
pulse context.

```{r strategy-feature-setup, eval=requireNamespace("TTR", quietly = TRUE)}
features <- list(
  ret_5 = ledgr_ind_returns(5),
  rsi_14 = ledgr_ind_ttr("RSI", input = "close", n = 14),
  bb_up = ledgr_ind_ttr("BBands", input = "close", output = "up", n = 20)
)

feature_ids <- setNames(ledgr_feature_id(features), names(features))
feature_ids
```

Inside the strategy, read the current values for each instrument and guard the
decision until all requested features have passed warmup.

```{r strategy-definition, eval=requireNamespace("TTR", quietly = TRUE)}
read_features <- function(ctx, id, feature_ids) {
  vapply(
    feature_ids,
    function(feature_id) ctx$feature(id, feature_id),
    numeric(1)
  )
}

strategy <- function(ctx, params) {
  targets <- ctx$flat()

  for (id in ctx$universe) {
    x <- read_features(ctx, id, feature_ids)

    if (
      all(!is.na(x)) &&
        x[["ret_5"]] > 0 &&
        x[["rsi_14"]] > 50 &&
        ctx$close(id) > x[["bb_up"]]
    ) {
      targets[id] <- params$qty
    }
  }

  targets
}
```

That pattern keeps the signal logic readable:

- `feature_ids` is where feature identity lives.
- `read_features()` is only a local convenience for the current API.
- `all(!is.na(x))` is the warmup gate.
- The condition after the warmup gate is the trading idea.

## Run The Example

Use two demo instruments so the strategy body has to work over
`ctx$universe`, not one hardcoded instrument.

```{r run-example, eval=requireNamespace("TTR", quietly = TRUE)}
bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% c("DEMO_01", "DEMO_02"),
    between(
      ts_utc,
      article_utc("2019-01-01"),
      article_utc("2019-06-30")
    )
  )

snapshot <- ledgr_snapshot_from_df(
  bars,
  snapshot_id = paste0("indicators-vignette-", Sys.getpid())
)

exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = strategy,
  features = features,
  opening = ledgr_opening(cash = 10000)
)

run_id <- paste0("indicators-demo-", Sys.getpid())

bt <- exp |>
  ledgr_run(params = list(qty = 10), run_id = run_id)

as.data.frame(ledgr_results(bt, what = "fills"))

close(bt)
ledgr_snapshot_close(snapshot)
```

The experiment registers the indicator objects. The strategy reads feature
values by ID at each pulse. ledgr handles feature computation before execution
and then gives the strategy only the pulse-time values.

## Unsupported Or Custom Indicators

When a TTR function is not in the warmup rules table, provide `requires_bars`
explicitly:

```{r unsupported-ttr, eval=requireNamespace("TTR", quietly = TRUE)}
ledgr_ind_ttr(
  "DEMA",
  input = "close",
  n = 10,
  requires_bars = 20
)$id
```

For non-TTR sources or more specialized logic, use `ledgr_indicator()` directly
with a `series_fn`. That is the adapter escape hatch: external logic remains at
the boundary, while the engine keeps the same deterministic indicator contract.

## What's Next?

For strategy authoring, read
`vignette("strategy-development", package = "ledgr")`. For accounting and
summary metrics, read `vignette("metrics-and-accounting", package = "ledgr")`.
For TTR-specific output names and supported warmup inference, see
`?ledgr_ind_ttr` and `?ledgr_ttr_warmup_rules`.
