---
title: "Indicators And Feature IDs"
output:
  rmarkdown::html_vignette:
    toc: false
  github_document:
    toc: false
    html_preview: false
    md_extensions: -smart
vignette: >
  %\VignetteIndexEntry{Indicators And Feature IDs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(width = 90)
demo_data_path <- if (file.exists("data/ledgr_demo_bars.rda")) {
  "data/ledgr_demo_bars.rda"
} else {
  file.path("..", "data", "ledgr_demo_bars.rda")
}
if (!exists("ledgr_demo_bars") && file.exists(demo_data_path)) {
  load(demo_data_path)
}
```

```{r packages, message=FALSE}
library(ledgr)
library(dplyr)
data("ledgr_demo_bars", package = "ledgr")
```

Indicators are how ledgr turns sealed market data into pulse-time features.
This article teaches the runtime shape first, then the accessor APIs.

The central model is:

> ledgr computes feature contracts into pulse-known data; the accessors are
> different views into that same pulse.

That model is the same for built-in ledgr indicators, TTR-backed indicators,
and custom indicators.

## Start With Built-In Features

Use two demo instruments and two built-in features. The economic idea will be
small on purpose:

> Own an instrument only when its recent return is positive enough and today's
> close is above its moving average.

That rule needs one momentum feature and one trend feature. A feature map gives readable aliases to your R code while preserving ledgr's exact engine feature IDs.

```{r built-in-feature-map}
features <- ledgr_feature_map(
  ret_5 = ledgr_ind_returns(5),
  sma_10 = ledgr_ind_sma(10)
)
```

There are two names in play:

- the **alias** is the readable name you choose, such as `ret_5` or `sma_10`;
- the **feature ID** is ledgr's stable engine name, such as `return_5` or
  `sma_10`.

Use aliases when you write strategy logic with `ctx$features()`. Use feature IDs
when you need the explicit engine contract, for example with `ctx$feature()` or
when inspecting stored feature metadata.

The same idea works for crossover rules. An SMA crossover registers two
separate indicators: one short moving average and one long moving average. The
economic meaning is "fast trend above slow trend" rather than "close above one
trend line." Each moving average has its own feature ID, warmup, and stored
values.

```{r sma-crossover-contracts}
crossover_features <- ledgr_feature_map(
  sma_fast = ledgr_ind_sma(10),
  sma_slow = ledgr_ind_sma(30)
)

ledgr_feature_contracts(crossover_features)
```

In a strategy, the crossover condition is just a comparison of the two mapped
aliases after warmup:

```{r sma-crossover-snippet, eval=FALSE}
x <- ctx$features(id, crossover_features)
if (passed_warmup(x) && x[["sma_fast"]] > x[["sma_slow"]]) {
  targets[id] <- params$qty
}
```

## Inspect One Pulse

Create a small sealed snapshot and inspect one decision pulse before running a
full backtest. This keeps the runtime data visible before the article introduces
more metadata.

```{r pulse-setup}
bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% c("DEMO_01", "DEMO_02"),
    between(
      ts_utc,
      ledgr::ledgr_utc("2019-01-01"),
      ledgr::ledgr_utc("2019-06-30")
    )
  )

snapshot <- ledgr_snapshot_from_df(
  bars,
  snapshot_id = paste0("indicators-vignette-", Sys.getpid())
)

pulse <- ledgr_pulse_snapshot(
  snapshot,
  universe = c("DEMO_01", "DEMO_02"),
  ts_utc = ledgr::ledgr_utc("2019-03-01"),
  features = features
)
```

At this timestamp, ledgr has computed the same two features for each instrument
in the universe. The long pulse view shows that directly: one row per
instrument and feature. Without a feature map, `alias` is `NA`. With the map,
rows are filtered to the mapped features and aliases are filled.

```{r pulse-long}
ledgr_pulse_features(pulse, features)
```

The wide pulse view is useful for debugging and future model-style workflows. It
contains one OHLCV block and one feature block for each instrument. OHLCV
columns use `{instrument_id}__ohlcv_{field}`. Feature columns use
`{instrument_id}__feature_{feature_id}`. A feature map can filter and order
feature columns, but it does not rename wide columns to aliases.

```{r pulse-wide}
ledgr_pulse_wide(pulse, features)
```

`ledgr_pulse_features()` and `ledgr_pulse_wide()` work on interactive pulse
snapshots and on the `ctx` object inside an ordinary strategy function. They are
inspection views over the same pulse-known data used by `ctx$feature()` and
`ctx$features()`.

## Access Features In A Strategy

The long and wide pulse views are useful when you want to inspect the computed
data, compare instruments, or think in model-like rows. They are not always the
clearest shape for strategy code. A strategy often wants to ask a smaller
question: "what are the current values for this instrument?"

That is why ledgr also exposes the same pulse data through scalar and mapped
accessors. The table views and the accessors are not competing APIs; they are
different views over the same pulse-known data.

The explicit scalar accessor is useful when you want to show or debug one
value. It uses the engine ID, not the alias:

```{r scalar-access}
ids <- ledgr_feature_id(features)
pulse$feature("DEMO_01", ids[["ret_5"]])
```

Mapped access returns a named numeric vector keyed by alias for one instrument
at one pulse:

```{r mapped-access}
x <- pulse$features("DEMO_01", features)
x
passed_warmup(x)
```

Inside a strategy, loop over `ctx$universe` so the rule works for every
instrument in the run.

Read the body economically:

- start from `ctx$flat()`, so the default desired state is no positions;
- inspect each instrument's current mapped features;
- skip the instrument while either feature is still in warmup;
- buy `params$qty` only when recent return is above the threshold and price is
  above the moving average;
- because the strategy starts flat on every pulse, an instrument is sold when
  the condition stops being true.

```{r strategy-definition}
strategy <- function(ctx, params) {
  targets <- ctx$flat()

  for (id in ctx$universe) {
    x <- ctx$features(id, features)

    if (
      passed_warmup(x) &&
        x[["ret_5"]] > params$min_return &&
        ctx$close(id) > x[["sma_10"]]
    ) {
      targets[id] <- params$qty
    }
  }

  targets
}
```

That pattern keeps the signal logic readable:

- `features` is where feature identity and aliases live.
- `ctx$features()` reads the current mapped values for one instrument.
- `passed_warmup()` is the warmup gate for the mapped feature vector.
- The condition after the warmup gate is the economic rule.
- The strategy still returns ordinary target quantities.

## Run The Example

The experiment registers the indicator objects. ledgr computes those features
for every instrument at each pulse, then gives the strategy only the pulse-time
values.

```{r run-example}
exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = strategy,
  features = features,
  opening = ledgr_opening(cash = 10000)
)

run_id <- paste0("indicators-demo-", Sys.getpid())

bt <- exp |>
  ledgr_run(params = list(min_return = 0, qty = 10), run_id = run_id)

ledgr_results(bt, what = "fills")

close(pulse)
close(bt)
ledgr_snapshot_close(snapshot)
```

## Read The Feature Contracts

After you have seen the feature values at a pulse, the contract table is easier
to read. The feature contracts are what ledgr will compute for every instrument
in the run. `alias` is for your strategy code. `feature_id` is the stable engine ID. Warmup metadata tells you when a known feature may still be `NA`.

```{r feature-contracts}
ledgr_feature_contracts(features)
```

Plain lists remain valid too. For a named list, names become aliases in the
contract table. For an unnamed list, `alias` is `NA`.

```{r plain-list-contracts}
plain_features <- list(ledgr_ind_returns(5), ledgr_ind_sma(10))
ledgr_feature_contracts(plain_features)
```

## Parameter Grids Register Every Needed Feature

If a parameter grid changes a lookback, register every lookback variant before
the run. ledgr does not create indicators dynamically from `params`; the run
only computes the feature contracts registered on the experiment.

```{r parameter-grid-features, eval=FALSE}
swept_features <- ledgr_feature_map(
  ret_5 = ledgr_ind_returns(5),
  ret_10 = ledgr_ind_returns(10),
  ret_20 = ledgr_ind_returns(20)
)

feature_ids <- ledgr_feature_id(swept_features)

parameterized_strategy <- function(ctx, params) {
  targets <- ctx$flat()
  feature_id <- feature_ids[[paste0("ret_", params$lookback)]]

  for (id in ctx$universe) {
    ret <- ctx$feature(id, feature_id)
    if (is.finite(ret) && ret > params$min_return) {
      targets[id] <- params$qty
    }
  }

  targets
}

grid <- ledgr_param_grid(
  lookback = c(5, 10, 20),
  min_return = 0,
  qty = 10
)
```

The important rule is that the feature set covers the whole grid:
`lookback = 20` means `return_20` must already be registered. A missing feature
ID is an unknown-feature error, not warmup.
The alias names in `swept_features` must also match the lookup key pattern used
by the strategy, here `paste0("ret_", params$lookback)`.
In short, all feature parameter values must be registered before `ledgr_run()`;
do not create `ledgr_ind_returns(params$lookback)` lazily inside the strategy.

## TTR-Backed Indicators

`ledgr_ind_ttr()` is the adapter for supported indicators from the suggested
`TTR` package. TTR stays outside the core engine:

```text
TTR -> ledgr_ind_ttr() -> ledgr_indicator -> deterministic pulse engine
```

The engine sees a normal `ledgr_indicator`. That means TTR-backed indicators
follow the same feature-ID, warmup, and pulse-view rules as built-in
indicators. The examples below are skipped when `TTR` is not installed. In your
own project, install TTR before creating TTR-backed indicators:

```{r install-ttr, eval=FALSE}
install.packages("TTR")
```

```{r ttr-backed-indicators, eval=requireNamespace("TTR", quietly = TRUE)}
ttr_features <- ledgr_feature_map(
  ret_5 = ledgr_ind_returns(5),
  ttr_rsi = ledgr_ind_ttr("RSI", input = "close", n = 14),
  bb_up = ledgr_ind_ttr("BBands", input = "close", output = "up", n = 20),
  macd = ledgr_ind_ttr(
    "MACD",
    input = "close",
    output = "macd",
    nFast = 12,
    nSlow = 26,
    nSig = 9,
    percent = FALSE
  ),
  macd_signal = ledgr_ind_ttr(
    "MACD",
    input = "close",
    output = "signal",
    nFast = 12,
    nSlow = 26,
    nSig = 9,
    percent = FALSE
  )
)

ledgr_feature_contracts(ttr_features)
ledgr_feature_id(ttr_features)
```

This mixed feature map combines a built-in return feature with TTR-backed RSI,
BBands, and MACD features. The examples produce IDs such as `return_5`,
`ttr_rsi_14`, `ttr_bbands_20_up`,
`ttr_macd_12_26_9_false_macd`, and
`ttr_macd_12_26_9_false_signal`.

RSI is a common mean-reversion input. One compact rule is: buy when RSI is below
30, then return to flat when the condition is no longer true. The experiment
registers the RSI indicator before the run; the strategy only reads the
pulse-time value.

```{r rsi-mean-reversion, eval=requireNamespace("TTR", quietly = TRUE)}
rsi_features <- ledgr_feature_map(
  rsi_14 = ledgr_ind_ttr("RSI", input = "close", n = 14)
)

rsi_strategy <- function(ctx, params) {
  targets <- ctx$flat()
  for (id in ctx$universe) {
    x <- ctx$features(id, rsi_features)
    if (passed_warmup(x) && x[["rsi_14"]] < params$oversold) {
      targets[id] <- params$qty
    }
  }
  targets
}

rsi_snapshot <- ledgr_snapshot_from_df(
  bars,
  snapshot_id = paste0("rsi-vignette-", Sys.getpid())
)

rsi_exp <- ledgr_experiment(
  snapshot = rsi_snapshot,
  strategy = rsi_strategy,
  features = rsi_features,
  opening = ledgr_opening(cash = 10000)
)

rsi_bt <- ledgr_run(
  rsi_exp,
  params = list(oversold = 30, qty = 10),
  run_id = paste0("rsi-demo-", Sys.getpid())
)

ledgr_results(rsi_bt, what = "fills")
close(rsi_bt)
ledgr_snapshot_close(rsi_snapshot)
```

Some TTR functions return several columns. For those functions, choose one
column with `output` before asking ledgr for the feature ID. `BBands` exposes
`dn`, `mavg`, `up`, and `pctB`. `MACD` exposes `macd` and `signal`; ledgr also
supports a derived `histogram`.

```{r ttr-output-examples, eval=requireNamespace("TTR", quietly = TRUE)}
ledgr_feature_contracts(ledgr_feature_map(
  bb_dn = ledgr_ind_ttr("BBands", input = "close", output = "dn", n = 20),
  bb_mavg = ledgr_ind_ttr("BBands", input = "close", output = "mavg", n = 20),
  bb_up = ledgr_ind_ttr("BBands", input = "close", output = "up", n = 20),
  bb_pctB = ledgr_ind_ttr("BBands", input = "close", output = "pctB", n = 20)
))
```

The two MACD examples above use matching explicit arguments. Explicit arguments
become part of the feature ID, so combine MACD outputs in one strategy only when
their argument sets match the computation you intend. If one MACD output uses
`percent = FALSE`, the paired `signal` output should usually set
`percent = FALSE` too.

TTR warmup inference is inspectable:

```{r ttr-warmup-rules, message=FALSE, eval=requireNamespace("TTR", quietly = TRUE)}
ledgr_ttr_warmup_rules() |>
  select(ttr_fn, input, formula)
```

For MACD, ledgr verifies the supported warmup rules against direct TTR output.
TTR computes the signal EMA internally even when you select only the `macd`
column. In a pulse-by-pulse backtest, all supported MACD outputs are therefore
first callable at `nSlow + nSig - 1`. The same rule is verified for `macd`,
`signal`, the derived ledgr `histogram`, and both `percent = TRUE` and
`percent = FALSE`.

To debug a TTR-backed feature at one decision time, use an active snapshot
handle, choose a timestamp late enough for the indicator warmup, and pass the
same TTR feature map to `ledgr_pulse_snapshot()`. A completed backtest proves
the run succeeded, but it does not replace the snapshot handle needed for
interactive pulse inspection.

```{r ttr-pulse-snapshot, eval=FALSE}
ttr_pulse <- ledgr_pulse_snapshot(
  snapshot,
  universe = c("DEMO_01", "DEMO_02"),
  ts_utc = ledgr::ledgr_utc("2019-06-03"),
  features = ttr_features
)

ledgr_pulse_features(ttr_pulse, ttr_features)
close(ttr_pulse)
```

## Troubleshoot Warmup And Zero Trades

Warmup problems are easiest to diagnose by connecting three facts:

1. `ledgr_feature_contracts(features)` tells you how many bars each feature
   needs before it can produce a usable value.
2. `ledgr_feature_contract_check(snapshot, features)` joins those contracts to
   the actual per-instrument bar counts in the snapshot.
3. `ledgr_pulse_features(pulse, features)` shows the current pulse-known
   values for the instruments and aliases you registered.
4. `summary(bt)` prints `Warmup Diagnostics` when a completed run has
   registered features that can never become usable for an instrument because
   available bars are below the feature contract.

```{r feature-contract-check}
warmup_check_snapshot <- ledgr_snapshot_from_df(
  bars |>
    filter(!(instrument_id == "DEMO_02" & ts_utc > ledgr_utc("2019-01-25"))),
  snapshot_id = paste0("warmup-check-", Sys.getpid())
)

ledgr_feature_contract_check(warmup_check_snapshot, features)

ledgr_snapshot_close(warmup_check_snapshot)
```

The `warmup_achievable` column is `FALSE` when an instrument does not have
enough available bars to satisfy a feature's `stable_after` contract.

Normal early warmup is temporary: a feature is `NA` near the beginning of an
instrument's sample and later becomes finite. Impossible warmup is different:
the instrument never has enough available bars for that feature. In that case,
zero trades can be a valid completed run plus a useful diagnostic, not a failed
run.

For result-table interpretation after a zero-trade run, read
`vignette("metrics-and-accounting", package = "ledgr")`.

## Unsupported Or Custom Indicators

When a TTR function is not in the warmup rules table, provide `requires_bars`
explicitly:

```{r unsupported-ttr, eval=requireNamespace("TTR", quietly = TRUE)}
ledgr_ind_ttr(
  "DEMA",
  input = "close",
  n = 10,
  requires_bars = 20
)$id
```

For non-TTR sources or more specialized logic, use `ledgr_indicator()` directly
with a `series_fn`. That is the adapter escape hatch: external logic remains at
the boundary, while the engine keeps the same deterministic indicator contract.

## What's Next?

For strategy authoring, read
`vignette("strategy-development", package = "ledgr")`. For accounting and
summary metrics, read `vignette("metrics-and-accounting", package = "ledgr")`.
For formal help on the inspection views, see `?ledgr_feature_contracts`,
`?ledgr_feature_contract_check`, `?ledgr_pulse_features`, and
`?ledgr_pulse_wide`. For TTR-specific output names and supported warmup
inference, see `?ledgr_ind_ttr` and
`?ledgr_ttr_warmup_rules`.
