---
title: "Indicators And Feature IDs"
format:
  html:
    toc: false
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{Indicators And Feature IDs}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(width = 90)
# Local renders load ledgr from source; installed builds use the library call
# in the prerequisites chunk.
pkg_root <- if (file.exists("DESCRIPTION")) "." else if (file.exists("../DESCRIPTION")) ".." else NA_character_
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
demo_data_path <- if (file.exists("data/ledgr_demo_bars.rda")) {
  "data/ledgr_demo_bars.rda"
} else {
  file.path("..", "data", "ledgr_demo_bars.rda")
}
if (!exists("ledgr_demo_bars") && file.exists(demo_data_path)) {
  load(demo_data_path)
}
```

```{=html}
<style>
.ledgr-diagram {
  margin: 1.25rem auto 1.5rem auto;
  text-align: center;
}
.ledgr-diagram .mermaid {
  display: inline-block;
  max-width: 760px;
  width: 100%;
}
.ledgr-diagram .node text,
.ledgr-diagram .edgeLabel {
  font-size: 18px !important;
}
</style>
```

Indicators are how ledgr turns sealed market data into pulse-time features.
This article teaches the runtime shape first, then the accessor APIs.

::: {.ledgr-callout .ledgr-callout-note}
**Definition**

An indicator is a declared feature computation. ledgr computes indicators into
pulse-known values; scalar accessors, mapped accessors, long tables, and wide
tables are different views of that same pulse-known data.
:::

That model is the same for built-in ledgr indicators, TTR-backed indicators,
and custom indicators.

## Prerequisites

```{r}
#| label: packages
#| message: false

library(ledgr)
library(dplyr)
data("ledgr_demo_bars", package = "ledgr")
```

## Start With Built-In Features

Use two demo instruments and two built-in features. The economic idea will be
small on purpose:

> Own an instrument only when its recent return is positive enough and today's
> close is above its moving average.

That rule needs one momentum feature and one trend feature. A feature map gives
your strategy code readable aliases while preserving ledgr's exact engine
feature IDs.

```{r}
#| label: built-in-feature-map

features <- ledgr_feature_map(
  ret_5 = ledgr_ind_returns(5),
  sma_10 = ledgr_ind_sma(10)
)
```

::: {.ledgr-callout .ledgr-callout-note}
**Definition**

A feature ID is ledgr's engine-facing name for a computed value. An alias is
the strategy-facing name you choose in a feature map. The alias makes strategy
code readable; the feature ID remains the durable engine identity.
:::

Use aliases such as `ret_5` or `sma_10` when you write strategy logic with
`ctx$features()`. Use feature IDs such as `return_5` or `sma_10` when you need
the explicit engine contract, for example with `ctx$feature()` or when
inspecting stored feature metadata.

Feature objects appear in three registration and inspection places:

| Surface | Accepted feature shape | How names are used |
|---|---|---|
| `ledgr_experiment(features = ...)` | indicator, list, named list, or feature map | registers feature definitions for the run |
| `ledgr_feature_contracts()` / `ledgr_feature_contract_check()` | static indicator, list, named list, or feature map | reports aliases and engine IDs |
| `ledgr_pulse_snapshot(features = ...)` | static list or feature map | computes pulse-known values for inspection |

The strategy context then exposes the computed values through accessors:

| Accessor | Name type | Use |
|---|---|---|
| `ctx$feature(id, feature_id)` | engine feature ID string | reads one scalar value by exact ID |
| `ctx$features(id, feature_map)` | feature map | returns a named vector keyed by alias |

The canonical workflow is: register features on `ledgr_experiment()`, then read
pulse-known values through `ctx$feature()` or `ctx$features()` inside the
strategy.

## Feature Lifecycle: From Declaration To Lookup

The feature path has five steps:

::: {.ledgr-diagram .ledgr-feature-lifecycle}

```{mermaid}
%%| echo: false

flowchart LR
  declare["declare<br/>indicator or map"]
  register["register<br/>experiment"]
  compute["compute<br/>pulse-known values"]
  access["access<br/>ctx feature methods"]
  decide["decide<br/>target holdings"]

  declare --> register --> compute --> access --> decide
```

:::

1. You declare features in any of several shapes: individual indicators,
   built-in helpers, TTR adapters, CSV/R adapters, feature maps, or
   active-alias parameterizations. All resolve to the same lifecycle below.
2. `ledgr_experiment()` stores the declaration. Static lists and feature maps
   are ready immediately. Active-alias features are materialized for concrete
   feature-grid values before candidate execution.
3. Optional `ledgr_precompute_features()` resolves a parameter grid, computes
   each candidate's concrete feature set, deduplicates shared indicator
   fingerprints, and records candidate feature-set hashes.
4. During `ledgr_run()` or `ledgr_sweep()`, the fold core computes the
   registered feature values at each pulse without looking past the current
   bar.
5. Strategy code reads the current pulse-known values with `ctx$feature()` by
   engine feature ID, or with `ctx$features()` by feature-map alias.

::: {.ledgr-callout .ledgr-callout-note}
**Definition**

A fingerprint identifies the feature definition, not just the name. If the
calculation, parameters, warmup rule, adapter, or selected output changes, the
fingerprint changes even when the feature ID stays readable.
:::

Feature IDs identify values inside the pulse context. Fingerprints identify the
feature definition used to compute those values. For multi-output sources such
as TTR `BBands` or `MACD`, each selected output is an ordinary indicator with
its own feature ID and output-specific fingerprint.

If two feature declarations produce the same engine feature ID, ledgr treats
that as one feature name in the pulse context. Use distinct IDs or aliases when
you need to compare two different definitions. A feature-map alias never changes
the underlying engine feature ID or fingerprint; it only gives your strategy a
readable name for mapped access.

::: {.ledgr-callout .ledgr-callout-note}
**Definition**

A bundle is an authoring convenience for declaring several indicator outputs at
once. The engine receives ordinary single-output feature definitions after the
bundle is flattened.
:::

The multi-output bundle helper follows the same lifecycle. It is not a second
feature system.

The same idea works for crossover rules. An SMA crossover registers two
separate indicators: one short moving average and one long moving average. The
economic meaning is "fast trend above slow trend" rather than "close above one
trend line." Each moving average has its own feature ID, warmup, and stored
values.

```{r}
#| label: sma-crossover-contracts

crossover_features <- ledgr_feature_map(
  sma_fast = ledgr_ind_sma(10),
  sma_slow = ledgr_ind_sma(30)
)

ledgr_feature_contracts(crossover_features)
```

In a strategy, the crossover condition is just a comparison of the two mapped
aliases after warmup:

```{r}
#| label: sma-crossover-snippet
#| eval: false

x <- ctx$features(id, crossover_features)
if (ledgr_passed_warmup(x) && x[["sma_fast"]] > x[["sma_slow"]]) {
  targets[id] <- params$qty
}
```

## Inspect One Pulse

Create a small sealed snapshot and inspect one decision pulse before running a
full backtest. This keeps the runtime data visible before the article introduces
more metadata.

```{r}
#| label: pulse-setup

bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% c("DEMO_01", "DEMO_02"),
    between(
      ts_utc,
      ledgr_utc("2019-01-01"),
      ledgr_utc("2019-06-30")
    )
  )

snapshot <- ledgr_snapshot_from_df(
  bars,
  snapshot_id = paste0("indicators-vignette-", Sys.getpid())
)

pulse <- ledgr_pulse_snapshot(
  snapshot,
  universe = c("DEMO_01", "DEMO_02"),
  ts_utc = ledgr_utc("2019-03-01"),
  features = features
)
```

At this timestamp, ledgr has computed the same two features for each instrument
in the universe. The long pulse view shows that directly: one row per
instrument and feature. Without a feature map, `alias` is `NA`. With the map,
rows are filtered to the mapped features and aliases are filled.

```{r}
#| label: pulse-long

ledgr_pulse_features(pulse, features)
```

The wide pulse view is useful for debugging and ML-style row-per-observation
workflows. It contains one OHLCV block and one feature block for each
instrument. OHLCV columns use `{instrument_id}__ohlcv_{field}`. Feature columns
use `{instrument_id}__feature_{feature_id}`. A feature map filters and orders
feature columns, using aliases as the wide feature keys.

```{r}
#| label: pulse-wide

ledgr_pulse_wide(pulse, features)
```

`ledgr_pulse_features()` and `ledgr_pulse_wide()` work on interactive pulse
snapshots and on the `ctx` object inside an ordinary strategy function. They are
inspection views over the same pulse-known data used by `ctx$feature()` and
`ctx$features()`.

## Access Features In A Strategy

The long and wide pulse views are useful when you want to inspect the computed
data, compare instruments, or think in model-like rows. They are not always the
clearest shape for strategy code. A strategy often wants to ask a smaller
question: "what are the current values for this instrument?"

That is why ledgr also exposes the same pulse data through scalar and mapped
accessors. The table views and the accessors are not competing APIs; they are
different views over the same pulse-known data.

The explicit scalar accessor is useful when you want to show or debug one
value. It uses the engine ID, not the alias:

```{r}
#| label: scalar-access

ids <- ledgr_feature_id(features)
pulse$feature("DEMO_01", ids[["ret_5"]])
```

Mapped access returns a named numeric vector keyed by alias for one instrument
at one pulse:

```{r}
#| label: mapped-access

x <- pulse$features("DEMO_01", features)
x
ledgr_passed_warmup(x)
```

::: {.ledgr-callout .ledgr-callout-note}
**Definition**

Warmup is the period before a known feature has enough prior bars to produce a
usable value. A warmup `NA` is not an unknown feature; it is a known feature
that is not ready yet.
:::

::: {.ledgr-callout .ledgr-callout-tip}
**Try it**

Change the scalar accessor to `pulse$feature("DEMO_01", "returns_5")`. Why does
that fail while `ids[["ret_5"]]` works?
:::

Inside a strategy, loop over `ctx$universe` so the rule works for every
instrument in the run.

```{r}
#| label: strategy-definition

strategy <- function(ctx, params) {
  targets <- ctx$flat()

  for (id in ctx$universe) {
    x <- ctx$features(id, features)

    if (
      ledgr_passed_warmup(x) &&
        x[["ret_5"]] > params$min_return &&
        ctx$close(id) > x[["sma_10"]]
    ) {
      targets[id] <- params$qty
    }
  }

  targets
}
```

That pattern keeps the signal logic readable:

- `features` is where feature identity and aliases live.
- `ctx$features()` reads the current mapped values for one instrument.
- `ledgr_passed_warmup()` is the warmup gate for the mapped feature vector.
- The condition after the warmup gate is the economic rule.
- The strategy still returns ordinary target quantities.

## Run The Example

The experiment registers the indicator objects. ledgr computes those features
for every instrument at each pulse, then gives the strategy only the pulse-time
values.

```{r}
#| label: run-example

exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = strategy,
  features = features,
  opening = ledgr_opening(cash = 10000),
  cost_model = ledgr_cost_zero()
)

run_id <- paste0("indicators-demo-", Sys.getpid())

bt <- exp |>
  ledgr_run(params = list(min_return = 0, qty = 10), run_id = run_id)

ledgr_results(bt, what = "fills")

close(pulse)
close(bt)
ledgr_snapshot_close(snapshot)
```

## Read The Feature Contracts

After you have seen the feature values at a pulse, the contract table is easier
to read. The feature contracts are what ledgr will compute for every instrument
in the run. `alias` is for your strategy code. `feature_id` is the stable engine ID. Warmup metadata tells you when a known feature may still be `NA`.

```{r}
#| label: feature-contracts

ledgr_feature_contracts(features)
```

Plain lists remain valid too. For a named list, names become aliases in the
contract table. For an unnamed list, `alias` is `NA`.

```{r}
#| label: plain-list-contracts

plain_features <- list(ledgr_ind_returns(5), ledgr_ind_sma(10))
ledgr_feature_contracts(plain_features)
```

## Parameter Grids Register Every Needed Feature

If a parameter grid changes a lookback, register every lookback variant before
the run. ledgr does not create indicators dynamically from `params`; the run
only computes the feature contracts registered on the experiment.

```{r}
#| label: parameter-grid-features
#| eval: false

swept_features <- ledgr_feature_map(
  ret_5 = ledgr_ind_returns(5),
  ret_10 = ledgr_ind_returns(10),
  ret_20 = ledgr_ind_returns(20)
)

feature_ids <- ledgr_feature_id(swept_features)

parameterized_strategy <- function(ctx, params) {
  targets <- ctx$flat()
  feature_id <- feature_ids[[paste0("ret_", params$lookback)]]

  for (id in ctx$universe) {
    ret <- ctx$feature(id, feature_id)
    if (is.finite(ret) && ret > params$min_return) {
      targets[id] <- params$qty
    }
  }

  targets
}

grid <- ledgr_param_grid(
  ret_5 = list(lookback = 5, min_return = 0, qty = 10),
  ret_10 = list(lookback = 10, min_return = 0, qty = 10),
  ret_20 = list(lookback = 20, min_return = 0, qty = 10)
)
```

The feature set must cover the whole grid: `lookback = 20` means `return_20`
must already be registered. A missing feature ID is an unknown-feature error,
not warmup.
The alias names in `swept_features` must also match the lookup key pattern used
by the strategy, here `paste0("ret_", params$lookback)`.
In short, all feature parameter values must be registered before `ledgr_run()`;
do not create `ledgr_ind_returns(params$lookback)` lazily inside the strategy.

For exploratory sweeps over ledgr-owned indicator parameters, prefer active
aliases. Declare the varying constructor arguments with `ledgr_param()` and
compose feature and strategy grids explicitly:

```{r}
#| label: active-alias-sweep-example
#| eval: false

features <- ledgr_feature_map(
  fast = ledgr_ind_sma(ledgr_param("fast_n")),
  slow = ledgr_ind_sma(ledgr_param("slow_n"))
)

strategy <- ledgr_demo_sma_crossover_strategy()
exp <- ledgr_experiment(snapshot, strategy, features = features, cost_model = ledgr_cost_zero())

grid <- ledgr_grid_cross(
  features = ledgr_feature_grid(
    fast_n = c(10L, 20L),
    slow_n = c(40L, 80L),
    .filter = fast_n < slow_n
  ),
  strategy = ledgr_strategy_grid(threshold = c(0, 0.01), qty = 10)
)

precomputed <- ledgr_precompute_features(exp, grid)
results <- ledgr_sweep(exp, grid, precomputed_features = precomputed)
```

For single-output indicators, the feature-map alias is the strategy-facing name
returned by `ctx$features(id)`. Bundle entries are intentionally flat; see the
TTR bundle section below for how bundle aliases differ from single-output
aliases in mapped access.

## TTR-Backed Indicators

`ledgr_ind_ttr()` is the adapter for supported indicators from the suggested
`TTR` package. TTR stays outside the core engine:

::: {.ledgr-diagram .ledgr-ttr-adapter}

```{mermaid}
%%| echo: false

flowchart LR
  ttr["TTR"]
  adapter["ledgr_ind_ttr()"]
  indicator["ledgr_indicator"]
  engine["deterministic<br/>pulse engine"]

  ttr --> adapter --> indicator --> engine
```

:::

The engine sees a normal `ledgr_indicator`. That means TTR-backed indicators
follow the same feature-ID, warmup, and pulse-view rules as built-in
indicators. The TTR-backed examples below are skipped when `TTR` is not
installed. In your own project, install TTR with `install.packages("TTR")`
before creating TTR-backed indicators.

### Single-Output TTR Indicators

```{r}
#| label: ttr-backed-indicators
#| eval: !expr requireNamespace("TTR", quietly = TRUE)

ttr_features <- ledgr_feature_map(
  ret_5 = ledgr_ind_returns(5),
  ttr_rsi = ledgr_ind_ttr("RSI", input = "close", n = 14),
  bb_up = ledgr_ind_ttr("BBands", input = "close", output = "up", n = 20),
  macd = ledgr_ind_ttr(
    "MACD",
    input = "close",
    output = "macd",
    nFast = 12,
    nSlow = 26,
    nSig = 9,
    percent = FALSE
  ),
  macd_signal = ledgr_ind_ttr(
    "MACD",
    input = "close",
    output = "signal",
    nFast = 12,
    nSlow = 26,
    nSig = 9,
    percent = FALSE
  )
)

ledgr_feature_contracts(ttr_features)
ledgr_feature_id(ttr_features)
```

This mixed feature map combines a built-in return feature with TTR-backed RSI,
BBands, and MACD features. The MACD ID embeds the explicit arguments because
they are part of the calculation identity.

### Native RSI vs TTR RSI

ledgr also includes a native RSI helper. It does not require TTR and follows the
same ID and warmup contract as other built-in indicators:

```{r}
#| label: native-rsi-contract

native_rsi_features <- ledgr_feature_map(
  rsi_14 = ledgr_ind_rsi(14)
)

ledgr_feature_contracts(native_rsi_features)
ledgr_feature_id(native_rsi_features)
```

The native RSI feature ID is `rsi_14`. The TTR-backed RSI feature ID above is
`ttr_rsi_14`. Those are different feature definitions and should not be treated
as interchangeable without checking that their calculation and warmup behavior
match your research intent.

RSI is a common mean-reversion input. One compact rule is: buy when RSI is below
30, then return to flat when the condition is no longer true. The experiment
registers the RSI indicator before the run; the strategy only reads the
pulse-time value. The example below uses native RSI so it works without TTR.
Use TTR-backed RSI when you deliberately want adapter behavior.

```{r}
#| label: rsi-mean-reversion

rsi_features <- native_rsi_features

rsi_strategy <- function(ctx, params) {
  targets <- ctx$flat()
  for (id in ctx$universe) {
    x <- ctx$features(id, rsi_features)
    if (ledgr_passed_warmup(x) && x[["rsi_14"]] < params$oversold) {
      targets[id] <- params$qty
    }
  }
  targets
}

rsi_snapshot <- ledgr_snapshot_from_df(
  bars,
  snapshot_id = paste0("rsi-vignette-", Sys.getpid())
)

rsi_exp <- ledgr_experiment(
  snapshot = rsi_snapshot,
  strategy = rsi_strategy,
  features = rsi_features,
  opening = ledgr_opening(cash = 10000),
  cost_model = ledgr_cost_zero()
)

rsi_bt <- ledgr_run(
  rsi_exp,
  params = list(oversold = 30, qty = 10),
  run_id = paste0("rsi-demo-", Sys.getpid())
)

ledgr_results(rsi_bt, what = "fills")
close(rsi_bt)
ledgr_snapshot_close(rsi_snapshot)
```

Some TTR functions return several columns. For those functions, choose one
column with `output` when you need exactly one output, or use
`ledgr_ind_ttr_outputs()` to declare several outputs from one shared TTR
configuration. `BBands` exposes `dn`, `mavg`, `up`, and `pctB`. `MACD` exposes
`macd` and `signal`; ledgr also supports a derived `histogram`.

### Multi-Output TTR Indicators

```{r}
#| label: ttr-output-examples
#| eval: !expr requireNamespace("TTR", quietly = TRUE)

ledgr_feature_contracts(ledgr_feature_map(
  bb_dn = ledgr_ind_ttr("BBands", input = "close", output = "dn", n = 20),
  bb_mavg = ledgr_ind_ttr("BBands", input = "close", output = "mavg", n = 20),
  bb_up = ledgr_ind_ttr("BBands", input = "close", output = "up", n = 20),
  bb_pctB = ledgr_ind_ttr("BBands", input = "close", output = "pctB", n = 20)
))
```

### Bundle Naming Rules

For multi-output authoring, prefer the bundle helper. By default, bundle feature
IDs use a normalized prefix derived from the TTR function name. For `BBands`,
that produces `bbands_dn`, `bbands_mavg`, `bbands_up`, and `bbands_pctb`.
The helper returns a `ledgr_indicator_bundle`, but the experiment sees ordinary
single-output indicators after feature declaration is materialized.

Those bundle IDs are shorter than the hand-written single-output TTR IDs such
as `ttr_bbands_20_up`. That asymmetry is intentional: bundle defaults optimize
for readable output names. Use `naming = c(up = "ttr_bbands_20_up")` or
hand-written `ledgr_ind_ttr(output = ...)` calls when you need exact legacy IDs.

```{r}
#| label: ttr-output-bundle
#| eval: !expr requireNamespace("TTR", quietly = TRUE)

bbands_bundle <- ledgr_ind_ttr_outputs("BBands", input = "close", n = 20)
ledgr_feature_id(bbands_bundle)
ledgr_feature_contracts(bbands_bundle)
```

When a bundle is placed inside `ledgr_feature_map()`, its entries expand using
their feature IDs as aliases. A single alias on the bundle argument is ignored
because one alias cannot name several outputs. Control the generated feature IDs
with the bundle's `prefix` argument instead.

Use `outputs` as a filter. The derived or explicit prefix still applies to
selected outputs, so a subset remains collision-resistant:

```{r}
#| label: ttr-output-bundle-filter
#| eval: !expr requireNamespace("TTR", quietly = TRUE)

bbands_subset <- ledgr_ind_ttr_outputs(
  "BBands",
  input = "close",
  outputs = c("dn", "up"),
  prefix = "bb",
  n = 20
)
ledgr_feature_id(bbands_subset)
```

`naming` renames selected outputs; it is not itself an output filter. When
renaming only part of a bundle, make the filter explicit:

```{r}
#| label: ttr-output-bundle-naming
#| eval: !expr requireNamespace("TTR", quietly = TRUE)

bbands_named_subset <- ledgr_ind_ttr_outputs(
  "BBands",
  input = "close",
  outputs = c("dn", "up"),
  naming = c(dn = "lower_band", up = "upper_band"),
  n = 20
)
ledgr_feature_id(bbands_named_subset)
```

Set `prefix = NULL` only when you explicitly want raw normalized output names
such as `dn`, `up`, or `pctb`. Raw names are short and can collide when one
experiment combines several bundles or parameterizations.

### MACD Argument Consistency

The two MACD entries in `ttr_features` both set `percent = FALSE`. Explicit
arguments become part of the feature ID, so combine MACD outputs in one strategy
only when their argument sets match the computation you intend. If one MACD
output uses `percent = FALSE`, the paired `signal` output should usually set
`percent = FALSE` too.

### TTR Warmup Rules

TTR warmup inference is inspectable:

```{r}
#| label: ttr-warmup-rules
#| message: false
#| eval: !expr requireNamespace("TTR", quietly = TRUE)

ledgr_ind_ttr_warmup_rules() |>
  select(ttr_fn, input, formula)
```

For MACD, ledgr verifies the supported warmup rules against direct TTR output.
TTR computes the signal EMA internally even when you select only the `macd`
column. In a pulse-by-pulse backtest, all supported MACD outputs are therefore
first callable at `nSlow + nSig - 1`. The same rule is verified for `macd`,
`signal`, the derived ledgr `histogram`, and both `percent = TRUE` and
`percent = FALSE`.

To debug a TTR-backed feature at one decision time, use an active snapshot
handle, choose a timestamp late enough for the indicator warmup, and pass the
same TTR feature map to `ledgr_pulse_snapshot()`. A completed backtest proves
the run succeeded, but it does not replace the snapshot handle needed for
interactive pulse inspection.

```{r}
#| label: ttr-pulse-snapshot
#| eval: false

ttr_pulse <- ledgr_pulse_snapshot(
  snapshot,
  universe = c("DEMO_01", "DEMO_02"),
  ts_utc = ledgr_utc("2019-06-03"),
  features = ttr_features
)

ledgr_pulse_features(ttr_pulse, ttr_features)
close(ttr_pulse)
```

## Troubleshoot Warmup And Zero Trades

Warmup problems are easiest to diagnose by connecting four facts:

1. `ledgr_feature_contracts(features)` tells you how many bars each feature
   needs before it can produce a usable value.
2. `ledgr_feature_contract_check(snapshot, features)` joins those contracts to
   the actual per-instrument bar counts in the snapshot.
3. `ledgr_pulse_features(pulse, features)` shows the current pulse-known
   values for the instruments and aliases you registered.
4. `summary(bt)` prints `Warmup Diagnostics` when a completed run has
   registered features that can never become usable for an instrument because
   available bars are below the feature contract.

```{r}
#| label: feature-contract-check

warmup_check_snapshot <- ledgr_snapshot_from_df(
  bars |>
    filter(!(instrument_id == "DEMO_02" & ts_utc > ledgr_utc("2019-01-25"))),
  snapshot_id = paste0("warmup-check-", Sys.getpid())
)

ledgr_feature_contract_check(warmup_check_snapshot, features)

ledgr_snapshot_close(warmup_check_snapshot)
```

The `warmup_achievable` column is `FALSE` when an instrument does not have
enough available bars to satisfy a feature's `stable_after` contract.

Normal early warmup is temporary: a feature is `NA` near the beginning of an
instrument's sample and later becomes finite. Impossible warmup is different:
the instrument never has enough available bars for that feature. In that case,
zero trades can be a valid completed run plus a useful diagnostic, not a failed
run.

For result-table interpretation after a zero-trade run, read
`vignette("metrics-and-accounting", package = "ledgr")`.

## Unsupported Or Custom Indicators

When a TTR function is not in the warmup rules table, provide `requires_bars`
explicitly:

```{r}
#| label: unsupported-ttr
#| eval: !expr requireNamespace("TTR", quietly = TRUE)

ledgr_ind_ttr(
  "DEMA",
  input = "close",
  n = 10,
  requires_bars = 20
)$id
```

For non-TTR sources or more specialized logic, use `ledgr_indicator()` directly
with a `series_fn`. That is the adapter escape hatch: external logic remains at
the boundary, while the engine keeps the same deterministic indicator contract.

## Where Next

- For strategy authoring, read
  `vignette("strategy-development", package = "ledgr")`.
- For exploratory feature grids and active aliases, read
  `vignette("sweeps", package = "ledgr")`.
- For accounting and summary metrics, read
  `vignette("metrics-and-accounting", package = "ledgr")`.
- For function-level details, see `?ledgr_feature_contracts`,
  `?ledgr_feature_contract_check`, `?ledgr_pulse_features`,
  `?ledgr_pulse_wide`, `?ledgr_ind_ttr`, and `?ledgr_ind_ttr_warmup_rules`.

