---
title: "On Leakage: ledgr Design Choices"
format:
  html:
    toc: false
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{On Leakage: ledgr Design Choices}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

A backtest is valid only if each decision uses information that was knowable at
that point in time -- and nothing else. By the end of this article you will be
able to spot an obvious lead, distinguish expanding from full-sample
preprocessing, and state which leakage risks ledgr cannot certify for you.

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 90)
options(cli.unicode = FALSE)
# Local renders load the package from source via pkgload (working dir is
# either project root or vignettes/); installed builds fall through and
# rely on the library(ledgr) call in the prerequisites chunk.
pkg_root <- if (file.exists("DESCRIPTION")) "." else if (file.exists("../DESCRIPTION")) ".." else NA_character_
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
```

```{r}
#| label: library
#| message: false

library(ledgr)
library(dplyr)
data("ledgr_demo_bars", package = "ledgr")

bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% c("DEMO_01", "DEMO_02"),
    between(
      ts_utc,
      ledgr_utc("2019-01-01"),
      ledgr_utc("2019-06-30")
    )
  )
```

Leakage is what happens when that boundary is violated. It can be obvious, or it
can be hidden inside preprocessing that looks harmless.

ledgr's workflow is designed around this problem. The strategy interface, sealed
snapshots, and registered feature definitions each make a common leakage
pattern harder to write by accident -- not just documented as bad practice. They
are not a sandbox: R code can still reach data ledgr did not hand it. This
article explains where that protection applies and where your responsibility
begins.

::: {.ledgr-callout .ledgr-callout-warning}
**Leakage breaks the question**

A leaky backtest can have correct arithmetic and still answer the wrong
research question. ledgr narrows common leakage paths, but it cannot certify
your data availability or research process.
:::

::: {.ledgr-callout .ledgr-callout-note}
**Definition**

Leakage is any path by which a decision uses information that was not knowable
at the simulated decision time.
:::

## The Obvious Leak

The blunt example is `lead(close)`: tomorrow's close placed onto today's row.

```{r}
#| label: lead-close

leaky_signals <- bars |>
  group_by(instrument_id) |>
  arrange(ts_utc, .by_group = TRUE) |>
  mutate(
    tomorrow_close = lead(close),
    buy_signal = tomorrow_close > close
  )

leaky_signals |>
  select(instrument_id, ts_utc, close, tomorrow_close, buy_signal) |>
  slice_head(n = 4)
```

The table is still grouped by instrument, so `slice_head(n = 4)` shows the
first four rows of each. Every `TRUE` marks a day whose next close is higher,
which is known only the day after. The resulting `buy_signal` looks like an
ordinary column, but it answers a question the strategy could not have
answered at today's decision time.

A strategy that trades on this signal is contaminated even if the resulting
trades are not all profitable. The signal was created from tomorrow's outcome,
so the backtest is no longer measuring a decision rule that could have existed
in real time. The arithmetic may be correct; the question being measured is not.

## The Subtle Leak

Real leakage is often less cartoonish. A full-sample threshold can leak without
calling `lead()`.

```{r}
#| label: full-sample-quantile

leaky_features <- bars |>
  group_by(instrument_id) |>
  arrange(ts_utc, .by_group = TRUE) |>
  mutate(
    ret_5 = close / lag(close, 5) - 1,
    strong_return = ret_5 > quantile(ret_5, 0.75, na.rm = TRUE)
  )

leaky_features |>
  select(instrument_id, ts_utc, ret_5, strong_return) |>
  filter(!is.na(ret_5)) |>
  slice_head(n = 4)
```

There is no future row reference in the final rule. The leak happened earlier:
the January rows are compared against a return distribution that includes later
months. Future information has been baked into a normal-looking feature.

A threshold estimated on a prior training sample and then frozen for later
evaluation can be valid. The leak appears when the threshold is estimated from
the same future-inclusive sample on which early decisions are judged.

The gap is concrete. Suppose the first quarter of the sample has strong returns
and the remaining three quarters are weak:

```{r}
#| label: quantile-gap

set.seed(42)
ret_5 <- c(
  rnorm(63,  mean =  0.004, sd = 0.010),  # first quarter: strong
  rnorm(189, mean = -0.001, sd = 0.012)   # remaining three quarters: weak
)

full_sample <- unname(quantile(ret_5, 0.75, na.rm = TRUE))
expanding <- vapply(seq_along(ret_5), function(i) {
  if (i <= 20L) {
    return(NA_real_)
  }
  unname(quantile(ret_5[seq_len(i - 1L)], 0.75, na.rm = TRUE))
}, numeric(1))

first_quarter <- seq_len(63L)
tibble(
  rule = c("full sample", "expanding, prior rows only"),
  threshold_at_quarter_end = c(full_sample, expanding[[63L]]),
  first_quarter_signals = c(
    sum(ret_5[first_quarter] > full_sample),
    sum(ret_5[first_quarter] > expanding[first_quarter], na.rm = TRUE)
  )
)
```

The full-sample threshold is lower because the later weak-return period drags
the distribution down. The expanding rule uses only rows that precede each
decision, so it is the causal comparison. Here the future-inclusive rule emits
many more first-quarter signals. The exact count is less important than the
direction: later observations changed earlier decisions.

## The Strategy Boundary

::: {.ledgr-diagram .ledgr-leakage-boundary}

```{mermaid}
%%| echo: false

flowchart LR
  data["sealed data"]
  features["registered features"]
  pulse["pulse context"]
  target["target holdings"]
  ledger["ledger events"]

  data --> features --> pulse --> target --> ledger
```

:::

At execution time, ledgr strategies receive one pulse context. The strategy can
read current bars, current registered features, current positions, cash, equity,
and helper functions such as `ctx$flat()` and `ctx$hold()`.

The strategy does not receive the full future market-data table. It has no
market-data object from which it can casually index tomorrow's bar.

```{r}
#| label: pulse-strategy

strategy <- function(ctx, params) {
  up <- ctx$vec$close > ctx$vec$open
  ctx |>
    ledgr_selection(where = up) |>
    ledgr_target_quantity(ctx, params$qty)
}
```

Everything this rule reads comes from `ctx`: today's close and open for every
instrument. Nothing in it can reach tomorrow's bar.

This boundary prevents one common class of leakage. It does not cleanse
contaminated data that were already handed to ledgr, and it does not stop a
strategy from reading data it was not handed. A function can close over an
object from your session, such as the full bars table or a column of
precomputed signals, and index it by `ctx$ts_utc`. Preflight classifies such a
strategy as Tier 2 and records a note naming the captured object; it does not
refuse it, and ledgr cannot tell whether that object holds future rows.

## The Feature Boundary

Registered indicators are ledgr's feature boundary. A feature is not just a
casual column added to a data frame. It has an ID, parameters, required history,
warmup/stability rules, and a deterministic place in the pulse context.

Scalar indicator functions are evaluated on bounded historical windows ending
at the current bar. Vectorized `series_fn` functions, which compute a custom indicator over an
instrument's whole series
(`vignette("custom-indicators", package = "ledgr")`), must return a numeric
series aligned to the input bars, with valid warmup and post-warmup values.
Unknown feature IDs fail loudly at pulse time rather than silently becoming
missing values.

Those rules make accidental leakage harder than ad hoc full-sample signal
columns.

## The Sharp Edge: `series_fn`

Vectorized custom indicators are useful because many indicators are naturally
computed over a full series. They are also a residual risk. Output shape and
value checks can prove that a vector has the right length and valid values. They
cannot prove that the function avoided future rows internally.

The [custom-indicator article](custom-indicators.html) explains this boundary in
more detail. The core warning is the same: full-series custom feature code must
be written with causal discipline.

## What ledgr Enforces At These Boundaries

| Boundary | ledgr behavior |
| --- | --- |
| Strategy reads future rows | Strategy receives a pulse context, not a future table; an object captured from your session is noted by preflight, not blocked. |
| Unknown feature ID | `ctx$feature()` fails loudly instead of returning warmup. |
| Known feature in warmup | Values are explicitly `NA_real_` until usable. |
| Invalid target shape | Missing or extra target instruments are rejected. |
| Scalar indicator history | Scalar indicators are evaluated on bounded windows ending at the current bar. |
| Vectorized feature output | `series_fn` output must be aligned, numeric, and valid after warmup. |

## What Remains Your Responsibility

ledgr does not certify that the dataset, event timestamps, universe
construction, parameter search, or custom vectorized feature code are causally
clean.

Use `vignette("point-in-time-inputs", package = "ledgr")` for knowledge-time
facts and `vignette("survivorship-bias", package = "ledgr")` for membership,
lifetime, and terminal-event evidence. Those tools make declared evidence
executable; they cannot repair vendor data that never recorded when a value was
actually knowable.

| Risk | Why ledgr cannot fully solve it |
| --- | --- |
| Survivorship-biased universe | A snapshot may already exclude dead or unavailable instruments. |
| Bad availability timestamps | Event, fundamental, or macro data may be timestamped by label date rather than when it was tradable information. |
| Restated data | Vendors may revise values after the simulated decision time. |
| Research-loop leakage | Trying many ideas and reporting the best one turns the sample into training data. |
| Semantically leaky `series_fn` | A shape-valid vector can still be future-aware internally. |
| Captured session objects | A strategy can close over a data frame from your session and read rows the pulse context would not give it. |
| Split-adjusted price levels | Split-adjusted bars restate history at today's share basis, so a rule that uses an absolute price level, such as a minimum price or a lot size, sees levels that did not exist before a later split. |

## Checklist Before Trusting A Run

- Was the snapshot built from data available at the simulated decision times?
- Was the universe chosen without future survival information?
- Are event or fundamental timestamps availability timestamps?
- Are features registered as indicators rather than full-sample signal columns?
- Does the strategy fill after the information it used?
- Did parameter choices survive out-of-sample or regime checks?
- Can the run be reopened and explained from stored provenance?

::: {.ledgr-callout .ledgr-callout-tip}
**Try it**

Change the weak later regime to a strong one and rerun the expanding comparison.
Which early decisions change under the full-sample rule, and why can the
expanding rule not see that regime yet?
:::

## What To Remember

ledgr protects the strategy boundary and makes the feature boundary explicit.
That is necessary, not sufficient.

A causally honest backtest still depends on clean data availability, honest
universe construction, disciplined feature authoring, and a research process
that does not turn the test sample into training data.

## Where Next

- `vignette("metrics-and-accounting", package = "ledgr")` reads a run's ledger,
  fills, trades, equity, and metrics.
- `vignette("indicators", package = "ledgr")` explains feature IDs, aliases,
  fingerprints, and warmup.
- `vignette("walk-forward", package = "ledgr")` shows held-out evaluation
  built on top of sweeps and runs.
