---
title: "Metrics And Accounting"
output:
  rmarkdown::html_vignette:
    toc: false
  github_document:
    toc: false
    html_preview: false
vignette: >
  %\VignetteIndexEntry{Metrics And Accounting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 90)
options(cli.unicode = FALSE)
default_output_hook <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(
  output = function(x, options) {
    default_output_hook(gsub("[ \t]+(?=\n)", "", x, perl = TRUE), options)
  }
)
```

ledgr records a backtest as accounting evidence first and summary metrics
second.

The useful reading order is:

1.  ledger events say what actually filled;
2.  fills are execution rows derived from those events;
3.  trades are only the fill rows that close quantity;
4.  equity rows value the portfolio through time;
5.  summary metrics are formulas over those public result tables.

That order matters. A strategy can open a position without closing it. That run
has fills and equity exposure, but zero closed trades. In that case
`n_trades = 0` and `win_rate = NA` are correct, not missing data.

```{r library, message=FALSE}
library(ledgr)
library(dplyr)
library(tibble)
```

## A Tiny Run

Use a five-bar in-memory fixture so the accounting can be checked by hand.

```{r run}
bars <- data.frame(
  ts_utc = as.POSIXct("2020-01-01", tz = "UTC") + 86400 * 0:4,
  instrument_id = "AAA",
  open = c(100, 101, 105, 106, 106),
  high = c(100, 101, 105, 106, 106),
  low = c(100, 101, 105, 106, 106),
  close = c(100, 101, 105, 106, 106),
  volume = 1
)

one_day_strategy <- function(ctx, params) {
  targets <- ctx$flat()
  if (ledgr_utc(ctx$ts_utc) == ledgr_utc("2020-01-01")) {
    targets["AAA"] <- 1
  }
  targets
}

bt <- ledgr_backtest(
  data = bars,
  strategy = one_day_strategy,
  initial_cash = 1000,
  run_id = "accounting_example"
)
```

The strategy asks to hold one share on the first pulse and then returns to flat.
In these examples, decisions fill at the next open. The buy therefore fills on
the second bar, and the exit fills on the third bar.

## Fills And Trades

`what = "fills"` returns execution fill rows. Opening and closing fills both
appear here.

```{r fills}
fills <- ledgr_results(bt, what = "fills")
fills
```

The important columns are:

| Column | Meaning |
|----|----|
| `side` | execution direction, such as `BUY` or `SELL` |
| `qty` | absolute fill quantity |
| `price` | fill price |
| `fee` | execution fee charged on the fill |
| `action` | whether the fill opened or closed quantity |
| `realized_pnl` | profit or loss booked by closing quantity |

`what = "trades"` keeps only closed trade rows. That is the table used by
`n_trades`, `win_rate`, and `avg_trade`.

```{r trades}
trades <- ledgr_results(bt, what = "trades")
trades
```

This run has two fill rows but one closed trade row. Counting fills as trades
would double-count the round trip.

## Equity Rows

The equity curve records portfolio state through time. It combines cash,
current position value, and equity.

```{r equity}
equity <- ledgr_results(bt, what = "equity")
equity
```

Open positions affect `positions_value` and therefore equity even before any
trade closes. Realized P&L belongs to closed quantity; open-position gains and
losses stay in the equity curve until a closing fill realizes them.

## Recompute The Metrics

The summary metrics can be recomputed from public result tables.

```{r recompute}
equity_values <- equity$equity
period_returns <- equity_values[-1] / equity_values[-length(equity_values)] - 1
bars_per_year <- 252

metric_check <- tibble(
  total_return =
    equity_values[length(equity_values)] / equity_values[1] - 1,
  annualized_return =
    (1 + total_return)^(
      1 / ((length(equity_values) - 1) / bars_per_year)
    ) - 1,
  volatility =
    sd(period_returns) * sqrt(bars_per_year),
  max_drawdown =
    min(equity_values / cummax(equity_values) - 1),
  n_trades =
    nrow(trades),
  win_rate =
    if (nrow(trades) > 0) mean(trades$realized_pnl > 0) else NA_real_,
  avg_trade =
    if (nrow(trades) > 0) mean(trades$realized_pnl) else NA_real_,
  time_in_market =
    mean(abs(equity$positions_value) > 1e-6)
)

metric_check
```

Those are the same definitions used by `summary(bt)` and
`ledgr_compute_metrics(bt)`. The first public equity row is the initial equity
for return calculations. Max drawdown is the maximum peak-to-trough decline in
the public equity rows. Time in market is the share of equity rows with
absolute `positions_value > 1e-6`.

This small example uses `bars_per_year <- 252` because the bars are daily. ledgr
detects bar frequency for `ledgr_compute_metrics()` and snaps common cadences,
such as daily and weekly, to standard annualization constants. Use the detected
value if you need an external calculation to match ledgr exactly on non-daily
data.

```{r summary}
summary(bt)
```

## Zero Trades Can Be Correct

A flat strategy produces no fills and no closed trades. The result tables still
keep their schemas, so downstream code can rely on the same column names.

```{r flat}
flat_strategy <- function(ctx, params) ctx$flat()

flat_bt <- ledgr_backtest(
  data = bars,
  strategy = flat_strategy,
  initial_cash = 1000,
  run_id = "flat_accounting_example"
)

ledgr_results(flat_bt, what = "fills")
ledgr_results(flat_bt, what = "trades")
ledgr_compute_metrics(flat_bt)[c("n_trades", "win_rate", "avg_trade")]
```

`win_rate` and `avg_trade` are `NA` because there are no closed trade rows to
evaluate. That is different from a zero percent win rate, which would mean
there were trades and none of them were profitable.

## Open Positions And Final-Bar Targets

An open-only run is also valid. It has an opening fill and equity exposure, but
no closed trade rows. The unrealized result belongs in equity, not in
`realized_pnl`.

```{r open-only}
open_only_strategy <- function(ctx, params) {
  targets <- ctx$flat()
  targets["AAA"] <- 1
  targets
}

open_bt <- ledgr_backtest(
  data = bars,
  strategy = open_only_strategy,
  initial_cash = 1000,
  run_id = "open_accounting_example"
)

ledgr_results(open_bt, what = "fills")
ledgr_results(open_bt, what = "trades")
ledgr_compute_metrics(open_bt)[c("n_trades", "win_rate", "avg_trade")]
```

A final-bar target under a next-open fill model can also be valid research
input while producing no fill. There is no later bar available for execution, so
ledgr warns and leaves the ledger unchanged for that last target change.

```{r final-bar-no-fill}
final_bar_strategy <- function(ctx, params) {
  targets <- ctx$flat()
  if (ledgr_utc(ctx$ts_utc) == ledgr_utc("2020-01-05")) {
    targets["AAA"] <- 1
  }
  targets
}

warned <- FALSE
final_bar_bt <- withCallingHandlers(
  ledgr_backtest(
    data = bars,
    strategy = final_bar_strategy,
    initial_cash = 1000,
    run_id = "final_bar_accounting_example"
  ),
  warning = function(w) {
    if (grepl("LEDGR_LAST_BAR_NO_FILL", conditionMessage(w), fixed = TRUE)) {
      warned <<- TRUE
      invokeRestart("muffleWarning")
    }
  }
)

warned
ledgr_results(final_bar_bt, what = "fills")
```

## Cleanup

Closing handles releases DuckDB resources in long sessions. It is not data
safety ceremony; completed run artifacts are already durable when the run
returns.

```{r cleanup}
close(bt)
close(flat_bt)
close(open_bt)
close(final_bar_bt)
```
