---
title: "The Accounting Model"
format:
  html:
    toc: false
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{The Accounting Model}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 90)
options(cli.unicode = FALSE)
# Local renders load ledgr from source; installed builds use the library call
# in the prerequisites chunk.
pkg_root <- if (file.exists("DESCRIPTION")) "." else if (file.exists("../DESCRIPTION")) ".." else NA_character_
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
default_output_hook <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(
  output = function(x, options) {
    default_output_hook(gsub("[ \t]+(?=\n)", "", x, perl = TRUE), options)
  }
)
```

```{=html}
<style>
.ledgr-diagram {
  margin: 1.25rem auto 1.5rem auto;
  text-align: center;
}
.ledgr-diagram .mermaid {
  display: inline-block;
  max-width: 760px;
  width: 100%;
}
.ledgr-diagram .node text,
.ledgr-diagram .edgeLabel {
  font-size: 18px !important;
}
</style>
```
ledgr records a backtest as accounting evidence first and summary metrics
second.

The useful reading order is:

::: {.ledgr-diagram .ledgr-accounting-hierarchy}

```{mermaid}
%%| echo: false

flowchart TB
  ledger["ledger events<br/>source of truth"]
  fills["fills<br/>execution rows"]
  trades["trades<br/>closed round trips"]
  equity["equity rows<br/>portfolio value"]
  metrics["summary metrics<br/>formulas over results"]

  ledger --> fills
  fills --> trades
  ledger --> equity
  trades --> metrics
  equity --> metrics
```

:::

1.  ledger events say what actually filled;
2.  fills are execution rows derived from those events;
3.  trades are only the fill rows that close quantity;
4.  equity rows value the portfolio through time;
5.  summary metrics are formulas over those public result tables.

That order matters. A strategy can open a position without closing it. That run
has fills and equity exposure, but zero closed trades. In that case
`n_trades = 0` and `win_rate = NA` are correct, not missing data.

## Prerequisites

```{r}
#| label: library
#| message: false

library(ledgr)
library(dplyr)
library(tibble)
```

## Inspection Surfaces

Use the narrowest inspection surface that answers the question:

| Question | Public surface | Shape |
|---|---|---|
| What happened at a glance? | `print(bt)` | printed run header with final equity |
| What are the standard metrics? | `summary(bt)` | printed interpretation; returns `bt` invisibly |
| What metric values can code consume? | `ledgr_compute_metrics(bt)` | list-like `ledgr_metrics` object with raw numeric values |
| What rows value the portfolio? | `ledgr_results(bt, what = "equity")` | classed tibble |
| What executed? | `ledgr_results(bt, what = "fills")` | classed tibble |
| What closed quantity? | `ledgr_results(bt, what = "trades")` | classed tibble |
| What did the event ledger record? | `ledgr_results(bt, what = "ledger")` | classed tibble |
| How do stored runs compare? | `ledgr_run_compare(snapshot, run_ids = ...)` | classed comparison tibble |
| What did a sweep candidate summarize? | `ledgr_sweep()` result rows | classed sweep tibble |
| What context was stored by promotion? | `ledgr_promotion_context(bt)` or `ledgr_run_promotion_context()` | nested list |

The result-table helpers return structured objects. Their print methods may
format timestamps for readability, but `as_tibble()` gives raw columns for
programmatic use. The stable programming contract is the column meaning, not the
number of rows: a valid run can have zero fills, zero closed trades, or a final
open position.

Two things are not in the standard result tables.

`final_equity` is not a field in the `ledgr_compute_metrics()` list. Read it
from the last equity row, from `print(bt)`, from comparison rows, or from sweep
rows.

There is no committed `ledgr_results(bt, what = "features")` table. Feature
values are inspected at pulse time with `ledgr_pulse_snapshot()` or through
sweep/precompute provenance, not through a persisted feature-result table
accessor.

Metric assumptions are inspectable through `ledgr_metric_context()`. Use it on a
backtest, `ledgr_metrics` object, comparison table, sweep result table, or
promotion context to see the risk-free-rate and annualization context that
produced that surface.

## A Tiny Run

Use a five-bar in-memory fixture so the accounting can be checked by hand. OHLC
values are equal per bar so the example focuses on accounting arithmetic; real
bars use the same code with ordinary intra-bar movement.
This article uses `ledgr_backtest()` as a compact fixture helper for accounting
examples. The canonical research workflow remains:
snapshot -> `ledgr_experiment()` -> `ledgr_run()`.

```{r}
#| label: run

bars <- data.frame(
  ts_utc = as.POSIXct("2020-01-01", tz = "UTC") + 86400 * 0:4,
  instrument_id = "AAA",
  open = c(100, 101, 105, 106, 106),
  high = c(100, 101, 105, 106, 106),
  low = c(100, 101, 105, 106, 106),
  close = c(100, 101, 105, 106, 106),
  volume = 1
)

one_day_strategy <- function(ctx, params) {
  targets <- ctx$flat()
  if (ledgr_utc(ctx$ts_utc) == ledgr_utc("2020-01-01")) {
    targets["AAA"] <- 1
  }
  targets
}

bt <- ledgr_backtest(
  data = bars,
  strategy = one_day_strategy,
  initial_cash = 1000,
  run_id = "accounting_example",
  cost_model = ledgr_cost_zero()
)
```

The strategy asks to hold one share on the first pulse and then returns to flat.
In these examples, decisions fill at the next open. The buy therefore fills on
the second bar, and the exit fills on the third bar.

## Timing, Spread, And Fees

Timing and cost are separate execution steps. `ledgr_timing_next_open()` decides
where a target change can fill: the next available open after the strategy
decision. A cost model then resolves the proposed fill price and explicit fee.
Strategies do not receive cost state, and cost models do not change side,
quantity, instrument, or execution timestamp.

`ledgr_cost_spread_bps()` uses a quoted bid/ask-spread convention. A buy crosses
half the quoted spread above the reference open:
`open * (1 + spread_bps / 20000)`. A sell crosses half below it:
`open * (1 - spread_bps / 20000)`. For the same reference price, a buy/sell
round trip therefore crosses approximately `spread_bps` basis points before
explicit fees.

```{r}
#| label: quoted-spread-arithmetic

spread_bps <- 25
reference_price <- 100
buy_cross <- reference_price * (1 + spread_bps / 20000)
sell_cross <- reference_price * (1 - spread_bps / 20000)
round_trip_bps <- (buy_cross - sell_cross) / reference_price * 10000
round(round_trip_bps, 6)
```

Price transforms and explicit fees are different. `ledgr_cost_spread_bps()`
changes the fill price. `ledgr_cost_fixed_fee()` and
`ledgr_cost_notional_bps_fee()` add values to the fill `fee` column after any
price transforms have resolved.

Target risk, timing, and cost are separate layers. A `risk_chain` can transform
validated strategy target quantities before fill proposals exist. The timing
model decides which bar is used for execution. The cost model then adjusts price
and fee on the accepted fill proposal; it does not decide whether a target is
affordable or liquid.

```{r}
#| label: cost-model-shape
#| results: hide

example_cost_model <- ledgr_cost_chain(
  ledgr_cost_spread_bps(25),
  ledgr_cost_fixed_fee(0.50),
  ledgr_cost_notional_bps_fee(1)
)

ledgr_cost_steps(example_cost_model)
ledgr_cost_describe(example_cost_model)
```

::: {.ledgr-callout .ledgr-callout-important}
**What costs do not model**

Cost models are deterministic research assumptions over accepted fill proposals.
They do not implement liquidity or capacity limits, financing, transaction-cost
analysis, taxes, OMS lifecycle behavior, or broker reconciliation. Those are
separate future layers, not hidden behavior in the cost API.
:::

## Ledger Events

::: {.ledgr-callout .ledgr-callout-note}
**Definition**

A ledger event is the append-only accounting record written when execution
changes cash, positions, or run state. Fills, trades, equity, and metrics are
derived views over ledger-backed evidence.
:::

The ledger is the most literal view. The friendlier result tables below are
derived from these events.

```{r}
#| label: ledger

ledger <- ledgr_results(bt, what = "ledger")
ledger
```

## Fills And Trades

::: {.ledgr-callout .ledgr-callout-note}
**Definition**

A fill is an execution row: direction, quantity, price, fee, and action. A
trade is the subset of fill evidence that closes quantity and realizes P&L.
:::

`what = "fills"` returns execution fill rows. Opening and closing fills both
appear here.

```{r}
#| label: fills

fills <- ledgr_results(bt, what = "fills")
fills
```

The important columns are:

| Column | Meaning |
|----|----|
| `side` | execution direction, such as `BUY` or `SELL` |
| `qty` | absolute fill quantity |
| `price` | fill price |
| `fee` | execution fee charged on the fill |
| `action` | whether the fill opened or closed quantity |
| `realized_pnl` | profit or loss booked by closing quantity |

`what = "trades"` keeps only closed trade rows. That is the table used by
`n_trades`, `win_rate`, and `avg_trade`.

```{r}
#| label: trades

trades <- ledgr_results(bt, what = "trades")
trades
```

This run has two fill rows but one closed trade row. Counting fills as trades
would double-count the round trip.

## Equity Rows

::: {.ledgr-callout .ledgr-callout-note}
**Definition**

An equity row values the portfolio at one timestamp. It combines cash, current
position value, and total equity, including open-position exposure.
:::

The equity curve records portfolio state through time.

```{r}
#| label: equity

equity <- ledgr_results(bt, what = "equity")
equity
```

Open positions affect `positions_value` and therefore equity even before any
trade closes. Realized P&L belongs to closed quantity; open-position gains and
losses stay in the equity curve until a closing fill realizes them.

## Recompute The Metrics

The summary metrics can be recomputed from public result tables.

```{r}
#| label: recompute

equity_values <- equity$equity
period_returns <- equity_values[-1] / equity_values[-length(equity_values)] - 1
bars_per_year <- 252
rf_annual <- 0
rf_period_return <- (1 + rf_annual)^(1 / bars_per_year) - 1
excess_returns <- period_returns - rf_period_return

metric_check <- tibble(
  total_return =
    equity_values[length(equity_values)] / equity_values[1] - 1,
  annualized_return =
    (1 + total_return)^(
      1 / ((length(equity_values) - 1) / bars_per_year)
    ) - 1,
  volatility =
    sd(period_returns) * sqrt(bars_per_year),
  sharpe_ratio =
    mean(excess_returns) / sd(excess_returns) * sqrt(bars_per_year),
  max_drawdown =
    min(equity_values / cummax(equity_values) - 1),
  n_trades =
    nrow(trades),
  win_rate =
    if (nrow(trades) > 0) mean(trades$realized_pnl > 0) else NA_real_,
  avg_trade =
    if (nrow(trades) > 0) mean(trades$realized_pnl) else NA_real_,
  time_in_market =
    mean(abs(equity$positions_value) > 1e-6)
)

metric_check
```

Those are the same definitions used by `summary(bt)` and
`ledgr_compute_metrics(bt)`. The first public equity row is the initial equity
for return calculations. Max drawdown is the maximum peak-to-trough decline in
the public equity rows. Time in market is the share of equity rows with
absolute `positions_value > 1e-6`.

`ledgr_results()` returns persisted result tables: `equity`, `fills`, `trades`,
or `ledger`. There is no `what = "metrics"` result table. Use `summary(bt)` for
printed interpretation, or `ledgr_compute_metrics(bt)` when you need the named
metric values in code.

This small example uses `bars_per_year <- 252` because the bars are daily. ledgr
detects bar frequency for `ledgr_compute_metrics()` and snaps common cadences,
such as daily and weekly, to standard annualization constants. Use the detected
value if you need an external calculation to match ledgr exactly on non-daily
data.

::: {.ledgr-callout .ledgr-callout-tip}
**Try it**

Change `bars_per_year` in the recompute chunk from `252` to `365`. Which
metrics change? Why does `total_return` stay the same?
:::

## Zero Trades Can Be Correct

A flat strategy produces no fills and no closed trades. The result tables still
keep their schemas, so downstream code can rely on the same column names.

```{r}
#| label: flat

flat_strategy <- function(ctx, params) ctx$flat()

flat_bt <- ledgr_backtest(
  data = bars,
  strategy = flat_strategy,
  initial_cash = 1000,
  run_id = "flat_accounting_example",
  cost_model = ledgr_cost_zero()
)

ledgr_results(flat_bt, what = "fills")
ledgr_results(flat_bt, what = "trades")
ledgr_compute_metrics(flat_bt)[c("n_trades", "win_rate", "avg_trade")]
```

`win_rate` and `avg_trade` are `NA` because there are no closed trade rows to
evaluate. That is different from a zero percent win rate, which would mean
there were trades and none of them were profitable.

## Open Positions And Final-Bar Targets

An open-only run is also valid. It has an opening fill and equity exposure, but
no closed trade rows. The unrealized result belongs in equity, not in
`realized_pnl`.

```{r}
#| label: open-only

open_only_strategy <- function(ctx, params) {
  targets <- ctx$flat()
  targets["AAA"] <- 1
  targets
}

open_bt <- ledgr_backtest(
  data = bars,
  strategy = open_only_strategy,
  initial_cash = 1000,
  run_id = "open_accounting_example",
  cost_model = ledgr_cost_zero()
)

ledgr_results(open_bt, what = "fills")
ledgr_results(open_bt, what = "trades")
ledgr_compute_metrics(open_bt)[c("n_trades", "win_rate", "avg_trade")]
```

A final-bar target under a next-open fill model can also be valid research
input while producing no fill. There is no later bar available for execution, so
ledgr warns and leaves the ledger unchanged for that last target change.

```{r}
#| label: final-bar-no-fill

final_bar_strategy <- function(ctx, params) {
  targets <- ctx$flat()
  if (ledgr_utc(ctx$ts_utc) == ledgr_utc("2020-01-05")) {
    targets["AAA"] <- 1
  }
  targets
}

final_bar_bt <- ledgr_backtest(
  data = bars,
  strategy = final_bar_strategy,
  initial_cash = 1000,
  run_id = "final_bar_accounting_example",
  cost_model = ledgr_cost_zero()
)

ledgr_results(final_bar_bt, what = "fills")
```


## Cleanup

```{r cleanup}
close(bt)
close(flat_bt)
close(open_bt)
close(final_bar_bt)
```

## Where Next

- `vignette("metric-contexts-and-conventions", package = "ledgr")` covers
  metric contexts, annualization, and diagnostics.
- `vignette("execution-semantics", package = "ledgr")` explains why fills
  happen on the next bar and why the final decision bar cannot fill.
- `?ledgr_cost_spread_bps` and `?ledgr_cost_fixed_fee` describe public
  transaction-cost model declarations.
