---
title: "Metric Contexts And Conventions"
format:
  html:
    toc: false
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{Metric Contexts And Conventions}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 90)
options(cli.unicode = FALSE)
# Local renders load ledgr from source; installed builds use the library call
# in the prerequisites chunk.
pkg_root <- if (file.exists("DESCRIPTION")) "." else if (file.exists("../DESCRIPTION")) ".." else NA_character_
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
default_output_hook <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(
  output = function(x, options) {
    default_output_hook(gsub("[ \t]+(?=\n)", "", x, perl = TRUE), options)
  }
)
```

```{=html}
<style>
.ledgr-diagram {
  margin: 1.25rem auto 1.5rem auto;
  text-align: center;
}
.ledgr-diagram .mermaid {
  display: inline-block;
  max-width: 760px;
  width: 100%;
}
.ledgr-diagram .node text,
.ledgr-diagram .edgeLabel {
  font-size: 18px !important;
}
</style>
```

Metrics are only useful when their assumptions are visible. This companion to
The Accounting Model shows where ledgr stores those assumptions, how
annualization and risk-free-rate choices affect metrics, and how to inspect the
metric context that travels with runs, sweeps, and promotions.

## Prerequisites

```{r}
#| label: library
#| message: false

library(ledgr)
library(dplyr)
library(tibble)
```

## Fixture For Context Examples

This companion article needs a small completed run so metric-context examples
can inspect stored assumptions. The accounting walkthrough, result-table
hierarchy, and hand recomputation live in
`vignette("metrics-and-accounting", package = "ledgr")`.

```{r}
#| label: metric-context-fixture

bars <- data.frame(
  ts_utc = as.POSIXct("2020-01-01", tz = "UTC") + 86400 * 0:4,
  instrument_id = "AAA",
  open = c(100, 101, 105, 106, 106),
  high = c(100, 101, 105, 106, 106),
  low = c(100, 101, 105, 106, 106),
  close = c(100, 101, 105, 106, 106),
  volume = 1
)

one_day_strategy <- function(ctx, params) {
  targets <- ctx$flat()
  if (ledgr_utc(ctx$ts_utc) == ledgr_utc("2020-01-01")) {
    targets["AAA"] <- 1
  }
  targets
}

bt <- ledgr_backtest(
  data = bars,
  strategy = one_day_strategy,
  initial_cash = 1000,
  run_id = "metric_context_example",
  cost_model = ledgr_cost_zero()
)
```

## Metric Context

Metric assumptions now live in a `metric_context`. The default context is US
equity daily: zero annual risk-free rate and `252 * 1` periods per year. Use
market templates for common assumptions:

::: {.ledgr-callout .ledgr-callout-note}
**Definition**

A metric context is the assumption object behind metrics: risk-free rate,
calendar, annualization, and reserved provider slots. It makes metric
assumptions inspectable instead of hidden in summary output.
:::

### Common Contexts

```{r}
#| label: metric-context-templates

ledgr_metric_us_equity()
ledgr_metric_us_equity(risk_free_rate = 0.04)
ledgr_metric_us_equity(
  bars_per_day = 390L,
  risk_free_rate = ledgr_risk_free_rate(0.04, label = "policy rate")
)
ledgr_metric_crypto()
```

A scalar shorthand is accepted when only the annual risk-free rate changes:

```{r}
#| label: metric-context-shorthand

ledgr_metric_context(0.04)
```

Use an explicit context when the calendar matters:

```{r}
#| label: metric-context-explicit

intraday_context <- ledgr_metric_context(
  calendar = ledgr_calendar_us_equity(bars_per_day = 390L),
  risk_free_rate = ledgr_risk_free_rate(0.04, label = "manual assumption")
)
```

The full constructor fields are `risk_free_rate`, `calendar`, `benchmark`,
`market_factor`, and `mar`. The provider fields `benchmark`, `market_factor`,
and `mar` are reserved and must be `NULL`; they exist so future benchmark and
provider designs have an explicit home instead of changing metric semantics
later.

Intraday work should set `calendar` explicitly. For example, US equity minute
bars use `ledgr_calendar_us_equity(bars_per_day = 390L)`. ledgr does not infer
that policy from ticker symbols, file names, or provider names.

### Stored Context vs Sensitivity Overrides

`summary(bt)` and `ledgr_compute_metrics(bt)` use the metric context stored with
the committed run. Call-time overrides are sensitivity checks; they do not
mutate the run:

```{r}
#| label: metric-context-sensitivity

stored_run_context <- ledgr_metric_context(bt)
stored_metrics <- ledgr_compute_metrics(bt)
zero_rf_metrics <- ledgr_compute_metrics(bt, risk_free_rate = 0)

ledgr_metric_context(stored_metrics)
ledgr_metric_context(zero_rf_metrics)
identical( # confirm the override did not mutate the stored context on `bt`
  ledgr_metric_context_hash(stored_run_context),
  ledgr_metric_context_hash(ledgr_metric_context(bt))
)
```

Metric-context hashes include the metric-context version, annual risk-free rate,
risk-free source, risk-free `as_of`, and calendar annualization and source
fields. Human display labels are stored for inspection but do not change the
hash. If the print output is too compact for a report, inspect the nested object
directly:

```{r}
#| label: metric-context-provenance

context <- ledgr_metric_context(bt)
context$risk_free_rate$label
context$risk_free_rate$source
context$risk_free_rate$as_of
ledgr_metric_context_hash(context)
```

### Comparison, Sweep, And Promotion Contexts

The following snippets use objects from the experiment-store and sweeps
workflows. They show where metric context is carried, not a complete runnable
example.

`ledgr_run_compare()` has exactly one comparison context per table. The
snapshot-first form uses the default context unless you pass one explicitly:

```{r}
#| label: comparison-context
#| eval: false

comparison <- ledgr_run_compare(
  snapshot,
  run_ids = c("trend_qty_5", "trend_qty_15"),
  metric_context = ledgr_metric_context(exp)
)

ledgr_metric_context(comparison)
```

Sweep result tables also have exactly one metric context. Promotion context
keeps that source sweep context separate from the committed run's own context:

```{r}
#| label: promotion-contexts
#| eval: false

results <- ledgr_sweep(train_exp, grid)
candidate <- ledgr_candidate(results, 1)
test_run <- ledgr_promote(test_exp, candidate, require_same_snapshot = FALSE)

ledgr_metric_context(results)
ledgr_metric_context(ledgr_promotion_context(test_run))
ledgr_metric_context(test_run)
```

This distinction matters for train/test work: the source sweep context explains
how a candidate was ranked, while the committed run context explains the default
analysis assumptions stored with the promoted run.

## Risk Metric Contract

The standard metric contract includes `sharpe_ratio` as ledgr's first
risk-adjusted metric. It is computed from the same public equity rows as
volatility, not from hidden runner state and not from an external metrics
package.

The return series is still the adjacent public equity-row return:

```text
equity_return[t] = equity[t] / equity[t - 1] - 1
```

Sharpe-style metrics use period excess returns:

```text
excess_return[t] = equity_return[t] - rf_period_return[t]
sharpe_ratio = mean(excess_return) / sd(excess_return) * sqrt(bars_per_year)
```

The first risk-free-rate provider is a scalar annual rate expressed as a
decimal, so `0.02` means two percent per year. The default is `0`. ledgr
converts that scalar annual rate to a per-period return with the same
`bars_per_year` used for annualized return and volatility:

```text
rf_period_return = (1 + rf_annual)^(1 / bars_per_year) - 1
```

Time-varying risk-free-rate series and real data providers such as FRED,
Treasury, ECB, or central-bank adapters are deferred to later metric-context
provider work. Future providers must feed the same pulse-aligned
`rf_period_return` vector into the formula above; they must not create a
separate Sharpe formula branch.

Sharpe returns `NA_real_` for short samples, flat equity, or near-zero
volatility; see `?ledgr_compute_metrics` for the exact edge-case rules.

Other risk-adjusted or benchmark-relative metrics are deferred to v0.2.x
benchmark context work: Sortino, Calmar, Omega, information ratio, alpha/beta,
benchmark-relative metrics, VaR, and tail-risk metrics.

```{r}
#| label: summary

summary(bt)
```

`summary(bt)` is a print-oriented view. It returns the backtest handle
invisibly, not a metrics object. Use `ledgr_compute_metrics()` for scripted
workflows:

```{r}
#| label: metric-object

metrics <- ledgr_compute_metrics(bt)
metrics[c("total_return", "sharpe_ratio", "n_trades", "win_rate")]
ledgr_metric_context(metrics)
```

The raw metrics object keeps metric-kernel attributes for provenance. Those
attributes are part of the programmatic object, not the printed metric table.
Use named fields such as `metrics$sharpe_ratio` or the subset above when you
need a compact report.

`ledgr_run_compare()` is also programmatic: it returns a tibble-like
`ledgr_comparison` object with raw numeric metric columns for filtering and
ranking. Its print method only curates the displayed columns. Comparison metrics
are recomputed from stored equity and fill tables and use the same closed-trade
semantics as `ledgr_compute_metrics()`.

For reports, convert the comparison object and keep the raw numeric columns:

```{r}
#| label: comparison-report
#| eval: false

comparison |>
  as.data.frame() |>
  select(run_id, final_equity, total_return, sharpe_ratio, max_drawdown)
```

## Diagnose A Successful Run With Zero Trades

A completed run with zero trades is not automatically wrong. It means ledgr
accepted the strategy outputs and the ledger reached the end of the sample, but
no closed round trips were recorded.

Use this checklist before changing the strategy:

1. Start with `summary(bt)`. If `Total Trades` is zero, `win_rate` and
   `avg_trade` should be `NA`, not zero. If a registered feature can never
   become usable because the sample is too short, `summary(bt)` also prints a
   `Warmup Diagnostics` note with the feature ID, instrument ID, required bars,
   and available bars.
2. Inspect `ledgr_results(bt, what = "fills")`. Empty fills mean nothing ever
   executed. Non-empty fills with empty trades mean positions opened but did
   not close.
3. Confirm the feature IDs with `ledgr_feature_id(features)`. A helper such as
   `ledgr_signal_return(ctx, lookback = 60)` reads `return_60`; that indicator must
   be registered before the run.
4. Compare feature contracts with sample length before assuming the strategy is
   broken. `ledgr_feature_contracts(features)` shows `requires_bars` and
   `stable_after`. If an instrument has fewer available bars than a feature's
   `requires_bars`, that feature cannot become usable for that instrument.
5. Inspect a late pulse with `ledgr_pulse_snapshot()`. If the feature is still
   `NA` for every instrument near the end of the sample, the issue is no longer
   ordinary early warmup. Check the lookback length, sample length, universe,
   and feature registration.
6. If the helper pipeline returns a `ledgr_empty_selection` on a late diagnostic
   pulse, inspect the signal values directly. Ordinary early warmup should have
   passed by then; a late all-missing signal usually points to sample length,
   universe, or feature-registration issues.

Timestamp checks should compare normalized UTC values, not local string
representations:

```{r}
#| label: timestamp-troubleshooting
#| eval: false

target_ts <- ledgr_utc("2020-01-05T00:00:00Z")
ledgr_utc(ctx$ts_utc) == target_ts

intraday_time <- format(ledgr_utc(ctx$ts_utc), "%H:%M:%S", tz = "UTC")
intraday_time >= "14:30:00" && intraday_time <= "21:00:00"
```

If fills are empty, distinguish zero signals from zero sizing. In your own run,
use the same snapshot, universe, timestamp, features, strategy, and params you
are debugging:

```{r}
#| label: zero-sizing-troubleshooting
#| eval: false

pulse <- ledgr_pulse_snapshot(
  snapshot,
  universe = ...,
  ts_utc = ...,
  features = features
)
target <- strategy(pulse, params)
target
all(target == 0)
```

Custom fill-model contract errors are different from zero-trade outcomes; see
the fill-model reference documentation for the required fill fields.

### Four Warmup-Adjacent Cases

Warmup is per instrument. One instrument can have a usable value while another
is still `NA` because it has fewer bars or a different data history.

Ordinary feature warmup is local to the beginning of each instrument's usable
sample. A known feature is `NA` for early pulses, then becomes usable once the
feature contract has enough bars.

Impossible warmup is different: every value for an instrument/feature remains
`NA` because the instrument never reaches the feature contract. That is the
case reported by the `Warmup Diagnostics` note in `summary(bt)`.

Current-bar absence is a third failure mode. If ledgr cannot construct the pulse
sequence because a current bar is absent for an instrument in the requested
universe, the run fails before strategy evaluation for that incomplete pulse;
this is a pulse construction error, not a feature warmup value.

A final-bar target is separate from warmup. Under the next-open fill model,
there is no later bar available to fill a target emitted on the last pulse, so
ledgr reports `LEDGR_LAST_BAR_NO_FILL` and leaves the ledger unchanged for that
target change.

### Compiled Accounting Fails Closed

Default execution uses the canonical R accounting path:
`compiled_accounting_model = NULL`. The scoped `"spot_fifo"` accelerator is an
ephemeral sweep opt-in for supported spot-FIFO workloads; committed durable runs
fail closed if you request it there. Unsupported model names raise
`ledgr_unsupported_accounting_model`. Missing compiled support raises
`ledgr_compiled_spot_fifo_unavailable`.

Those classes are the stable top-level conditions to assert on in tests. The
accelerator does not broaden ledgr's accounting model into futures, margin,
broker reconciliation, or derivative accounting.


## Cleanup

```{r cleanup}
close(bt)
```

## Where Next

- `vignette("metrics-and-accounting", package = "ledgr")` is the shorter
  accounting-first tutorial.
- `vignette("sweeps", package = "ledgr")` shows how metric contexts thread
  through parameter sweeps.
- `vignette("walk-forward", package = "ledgr")` shows how selected-candidate
  test runs retain metric context evidence.
