---
title: "Metrics And Accounting"
output:
  rmarkdown::html_vignette:
    toc: false
  github_document:
    toc: false
    html_preview: false
vignette: >
  %\VignetteIndexEntry{Metrics And Accounting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 90)
options(cli.unicode = FALSE)
default_output_hook <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(
  output = function(x, options) {
    default_output_hook(gsub("[ \t]+(?=\n)", "", x, perl = TRUE), options)
  }
)
```

ledgr records a backtest as accounting evidence first and summary metrics
second.

The useful reading order is:

1.  ledger events say what actually filled;
2.  fills are execution rows derived from those events;
3.  trades are only the fill rows that close quantity;
4.  equity rows value the portfolio through time;
5.  summary metrics are formulas over those public result tables.

That order matters. A strategy can open a position without closing it. That run
has fills and equity exposure, but zero closed trades. In that case
`n_trades = 0` and `win_rate = NA` are correct, not missing data.

```{r library, message=FALSE}
library(ledgr)
library(dplyr)
library(tibble)
```

## A Tiny Run

Use a five-bar in-memory fixture so the accounting can be checked by hand.
This article uses `ledgr_backtest()` as a compact fixture helper for accounting
examples. The canonical research workflow remains:
snapshot -> `ledgr_experiment()` -> `ledgr_run()`.

```{r run}
bars <- data.frame(
  ts_utc = as.POSIXct("2020-01-01", tz = "UTC") + 86400 * 0:4,
  instrument_id = "AAA",
  open = c(100, 101, 105, 106, 106),
  high = c(100, 101, 105, 106, 106),
  low = c(100, 101, 105, 106, 106),
  close = c(100, 101, 105, 106, 106),
  volume = 1
)

one_day_strategy <- function(ctx, params) {
  targets <- ctx$flat()
  if (ledgr_utc(ctx$ts_utc) == ledgr_utc("2020-01-01")) {
    targets["AAA"] <- 1
  }
  targets
}

bt <- ledgr_backtest(
  data = bars,
  strategy = one_day_strategy,
  initial_cash = 1000,
  run_id = "accounting_example"
)
```

The strategy asks to hold one share on the first pulse and then returns to flat.
In these examples, decisions fill at the next open. The buy therefore fills on
the second bar, and the exit fills on the third bar.

## Ledger Events

The ledger is the append-only accounting record for the run. It is the most
literal view: rows record what ledgr wrote when an execution changed cash,
positions, or run state. The friendlier result tables below are derived from
these events.

```{r ledger}
ledger <- ledgr_results(bt, what = "ledger")
ledger
```

## Fills And Trades

`what = "fills"` returns execution fill rows. Opening and closing fills both
appear here.

```{r fills}
fills <- ledgr_results(bt, what = "fills")
fills
```

The important columns are:

| Column | Meaning |
|----|----|
| `side` | execution direction, such as `BUY` or `SELL` |
| `qty` | absolute fill quantity |
| `price` | fill price |
| `fee` | execution fee charged on the fill |
| `action` | whether the fill opened or closed quantity |
| `realized_pnl` | profit or loss booked by closing quantity |

`what = "trades"` keeps only closed trade rows. That is the table used by
`n_trades`, `win_rate`, and `avg_trade`.

```{r trades}
trades <- ledgr_results(bt, what = "trades")
trades
```

This run has two fill rows but one closed trade row. Counting fills as trades
would double-count the round trip.

## Equity Rows

The equity curve records portfolio state through time. It combines cash,
current position value, and equity.

```{r equity}
equity <- ledgr_results(bt, what = "equity")
equity
```

Open positions affect `positions_value` and therefore equity even before any
trade closes. Realized P&L belongs to closed quantity; open-position gains and
losses stay in the equity curve until a closing fill realizes them.

## Recompute The Metrics

The summary metrics can be recomputed from public result tables.

```{r recompute}
equity_values <- equity$equity
period_returns <- equity_values[-1] / equity_values[-length(equity_values)] - 1
bars_per_year <- 252

metric_check <- tibble(
  total_return =
    equity_values[length(equity_values)] / equity_values[1] - 1,
  annualized_return =
    (1 + total_return)^(
      1 / ((length(equity_values) - 1) / bars_per_year)
    ) - 1,
  volatility =
    sd(period_returns) * sqrt(bars_per_year),
  max_drawdown =
    min(equity_values / cummax(equity_values) - 1),
  n_trades =
    nrow(trades),
  win_rate =
    if (nrow(trades) > 0) mean(trades$realized_pnl > 0) else NA_real_,
  avg_trade =
    if (nrow(trades) > 0) mean(trades$realized_pnl) else NA_real_,
  time_in_market =
    mean(abs(equity$positions_value) > 1e-6)
)

metric_check
```

Those are the same definitions used by `summary(bt)` and
`ledgr_compute_metrics(bt)`. The first public equity row is the initial equity
for return calculations. Max drawdown is the maximum peak-to-trough decline in
the public equity rows. Time in market is the share of equity rows with
absolute `positions_value > 1e-6`.

`ledgr_results()` returns persisted result tables: `equity`, `fills`, `trades`,
or `ledger`. There is no `what = "metrics"` result table. Use `summary(bt)` for
printed interpretation, or `ledgr_compute_metrics(bt)` when you need the named
metric values in code.

This small example uses `bars_per_year <- 252` because the bars are daily. ledgr
detects bar frequency for `ledgr_compute_metrics()` and snaps common cadences,
such as daily and weekly, to standard annualization constants. Use the detected
value if you need an external calculation to match ledgr exactly on non-daily
data.

## Risk Metric Contract

The v0.1.7.7 standard metric contract adds `sharpe_ratio` as the first
risk-adjusted metric. It is a ledgr-owned metric computed from the same public
equity rows as volatility, not from hidden runner state and not from an external
metrics package.

The return series is still the adjacent public equity-row return:

```text
equity_return[t] = equity[t] / equity[t - 1] - 1
```

Sharpe-style metrics use period excess returns:

```text
excess_return[t] = equity_return[t] - rf_period_return[t]
sharpe_ratio = mean(excess_return) / sd(excess_return) * sqrt(bars_per_year)
```

The first risk-free-rate provider is a scalar annual rate expressed as a
decimal, so `0.02` means two percent per year. The default is `0`. ledgr
converts that scalar annual rate to a per-period return with the same
`bars_per_year` used for annualized return and volatility:

```text
rf_period_return = (1 + rf_annual)^(1 / bars_per_year) - 1
```

Time-varying risk-free-rate series and real data providers such as FRED,
Treasury, ECB, or central-bank adapters are deferred. Future providers must
feed the same pulse-aligned `rf_period_return` vector into the formula above;
they must not create a separate Sharpe formula branch.

The metric is intentionally conservative around edge cases. Short samples,
invalid adjacent equity returns, flat equity, constant-return series, zero or
near-zero excess-return volatility, and all-missing return inputs return
`NA_real_` rather than an infinite or misleading Sharpe value.

Other risk-adjusted or benchmark-relative metrics are deferred in this release:
Sortino, Calmar, Omega, information ratio, alpha/beta, benchmark-relative
metrics, VaR, and tail-risk metrics.

```{r summary}
summary(bt)
```

## Zero Trades Can Be Correct

A flat strategy produces no fills and no closed trades. The result tables still
keep their schemas, so downstream code can rely on the same column names.

```{r flat}
flat_strategy <- function(ctx, params) ctx$flat()

flat_bt <- ledgr_backtest(
  data = bars,
  strategy = flat_strategy,
  initial_cash = 1000,
  run_id = "flat_accounting_example"
)

ledgr_results(flat_bt, what = "fills")
ledgr_results(flat_bt, what = "trades")
ledgr_compute_metrics(flat_bt)[c("n_trades", "win_rate", "avg_trade")]
```

`win_rate` and `avg_trade` are `NA` because there are no closed trade rows to
evaluate. That is different from a zero percent win rate, which would mean
there were trades and none of them were profitable.

## Open Positions And Final-Bar Targets

An open-only run is also valid. It has an opening fill and equity exposure, but
no closed trade rows. The unrealized result belongs in equity, not in
`realized_pnl`.

```{r open-only}
open_only_strategy <- function(ctx, params) {
  targets <- ctx$flat()
  targets["AAA"] <- 1
  targets
}

open_bt <- ledgr_backtest(
  data = bars,
  strategy = open_only_strategy,
  initial_cash = 1000,
  run_id = "open_accounting_example"
)

ledgr_results(open_bt, what = "fills")
ledgr_results(open_bt, what = "trades")
ledgr_compute_metrics(open_bt)[c("n_trades", "win_rate", "avg_trade")]
```

A final-bar target under a next-open fill model can also be valid research
input while producing no fill. There is no later bar available for execution, so
ledgr warns and leaves the ledger unchanged for that last target change.

```{r final-bar-no-fill}
final_bar_strategy <- function(ctx, params) {
  targets <- ctx$flat()
  if (ledgr_utc(ctx$ts_utc) == ledgr_utc("2020-01-05")) {
    targets["AAA"] <- 1
  }
  targets
}

warned <- FALSE
final_bar_bt <- withCallingHandlers(
  ledgr_backtest(
    data = bars,
    strategy = final_bar_strategy,
    initial_cash = 1000,
    run_id = "final_bar_accounting_example"
  ),
  warning = function(w) {
    if (grepl("LEDGR_LAST_BAR_NO_FILL", conditionMessage(w), fixed = TRUE)) {
      warned <<- TRUE
      invokeRestart("muffleWarning")
    }
  }
)

warned
ledgr_results(final_bar_bt, what = "fills")
```

## Diagnose A Successful Run With Zero Trades

A completed run with zero trades is not automatically wrong. It means ledgr
accepted the strategy outputs and the ledger reached the end of the sample, but
no closed round trips were recorded.

Use this checklist before changing the strategy:

1. Start with `summary(bt)`. If `Total Trades` is zero, `win_rate` and
   `avg_trade` should be `NA`, not zero. If a registered feature can never
   become usable because the sample is too short, `summary(bt)` also prints a
   `Warmup Diagnostics` note with the feature ID, instrument ID, required bars,
   and available bars.
2. Inspect `ledgr_results(bt, what = "fills")`. Empty fills mean nothing ever
   executed. Non-empty fills with empty trades mean positions opened but did
   not close.
3. Confirm the feature IDs with `ledgr_feature_id(features)`. A helper such as
   `signal_return(ctx, lookback = 60)` reads `return_60`; that indicator must
   be registered before the run.
4. Compare feature contracts with sample length before assuming the strategy is
   broken. `ledgr_feature_contracts(features)` shows `requires_bars` and
   `stable_after`. If an instrument has fewer available bars than a feature's
   `requires_bars`, that feature cannot become usable for that instrument.
5. Inspect a late pulse with `ledgr_pulse_snapshot()`. If the feature is still
   `NA` for every instrument near the end of the sample, the issue is no longer
   ordinary early warmup. Check the lookback length, sample length, universe,
   and feature registration.
6. If the strategy suppresses `ledgr_empty_selection` warnings during a full
   run, rerun the same helper pipeline on the diagnostic pulse without
   suppression. The warning message reports the signal origin and non-missing
   count.

### Three Warmup-Adjacent Cases

Warmup is per instrument. One instrument can have a usable value while another
is still `NA` because it has fewer bars or a different data history.

Ordinary feature warmup is local to the beginning of each instrument's usable
sample. A known feature is `NA` for early pulses, then becomes usable once the
feature contract has enough bars.

Impossible warmup is different: every value for an instrument/feature remains
`NA` because the instrument never reaches the feature contract. That is the
case reported by the `Warmup Diagnostics` note in `summary(bt)`.

Current-bar absence is a third failure mode. If ledgr cannot construct the pulse
sequence because a current bar is absent for an instrument in the requested
universe, the run fails before strategy evaluation for that incomplete pulse;
this is a pulse construction error, not a feature warmup value.

A final-bar target is separate from warmup. Under the next-open fill model,
there is no later bar available to fill a target emitted on the last pulse, so
ledgr reports `LEDGR_LAST_BAR_NO_FILL` and leaves the ledger unchanged for that
target change.

## Cleanup

Closing handles releases DuckDB resources in long sessions. It is not data
safety ceremony; completed run artifacts are already durable when the run
returns.

```{r cleanup}
close(bt)
close(flat_bt)
close(open_bt)
close(final_bar_bt)
```

## What's Next?

For strategy authoring, read
`vignette("strategy-development", package = "ledgr")`. For indicators, feature
IDs, and warmup, read `vignette("indicators", package = "ledgr")`. For durable
run inspection, read `vignette("experiment-store", package = "ledgr")`.
