---
title: "Metric Contexts And Conventions"
format:
  html:
    toc: false
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{Metric Contexts And Conventions}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 90)
options(cli.unicode = FALSE)
# Local renders load ledgr from source; installed builds use the library call
# in the prerequisites chunk.
pkg_root <- if (file.exists("DESCRIPTION")) "." else if (file.exists("../DESCRIPTION")) ".." else NA_character_
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
default_output_hook <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(
  output = function(x, options) {
    default_output_hook(gsub("[ \t]+(?=\n)", "", x, perl = TRUE), options)
  }
)
```

```{=html}
<style>
.ledgr-diagram {
  margin: 1.25rem auto 1.5rem auto;
  text-align: center;
}
.ledgr-diagram .mermaid {
  display: inline-block;
  max-width: 760px;
  width: 100%;
}
.ledgr-diagram .node text,
.ledgr-diagram .edgeLabel {
  font-size: 18px !important;
}
</style>
```

Your Sharpe ratio depends on assumptions: how many bars make a year, and which
risk-free rate applies. By the end of this article you can see where ledgr
stores those assumptions, how changing them moves the metrics, and how the same
context travels with runs, sweeps, and promotions. You need it when your data
is not daily US-equity bars, or when you compare metrics across runs.

## Prerequisites

```{r}
#| label: library
#| message: false

library(ledgr)
library(dplyr)
library(tibble)
```

## Fixture For Context Examples

This companion article needs a small completed run so metric-context examples
can inspect stored assumptions. The accounting walkthrough, result-table
hierarchy, and hand recomputation live in
`vignette("metrics-and-accounting", package = "ledgr")`.

The fixture uses `ledgr_backtest()` only as the compact in-memory compatibility
surface. New project workflows should enter through a sealed snapshot,
`ledgr_experiment()`, and `ledgr_run()` as shown in the Quickstart.

```{r}
#| label: metric-context-fixture

bars <- data.frame(
  ts_utc = as.POSIXct("2020-01-01", tz = "UTC") + 86400 * 0:4,
  instrument_id = "AAA",
  open = c(100, 101, 105, 106, 106),
  high = c(100, 101, 105, 106, 106),
  low = c(100, 101, 105, 106, 106),
  close = c(100, 101, 105, 106, 106),
  volume = 1
)

one_day_strategy <- function(ctx, params) {
  targets <- ctx$flat()
  if (ledgr_utc(ctx$ts_utc) == ledgr_utc("2020-01-01")) {
    targets["AAA"] <- 1
  }
  targets
}

bt <- ledgr_backtest(
  data = bars,
  strategy = one_day_strategy,
  initial_cash = 1000,
  run_id = "metric_context_example",
  cost_model = ledgr_cost_zero()
)
```

## Metric Context

Metric assumptions now live in a `metric_context`. The default context is US
equity daily: zero annual risk-free rate and `252 * 1` periods per year. Use
market templates for common assumptions:

::: {.ledgr-callout .ledgr-callout-note}
**Definition**

A metric context is the assumption object behind metrics: risk-free rate,
calendar, annualization, and reserved provider slots. It makes metric
assumptions inspectable instead of hidden in summary output.
:::

### Common Contexts

```{r}
#| label: metric-context-templates

ledgr_metric_us_equity()
ledgr_metric_us_equity(risk_free_rate = 0.04)
ledgr_metric_us_equity(
  bars_per_day = 390L,
  risk_free_rate = ledgr_risk_free_rate(0.04, label = "policy rate")
)
ledgr_metric_crypto()
```

A scalar shorthand is accepted when only the annual risk-free rate changes:

```{r}
#| label: metric-context-shorthand

ledgr_metric_context(0.04)
```

Use an explicit context when the calendar matters:

```{r}
#| label: metric-context-explicit

intraday_context <- ledgr_metric_context(
  calendar = ledgr_calendar_us_equity(bars_per_day = 390L),
  risk_free_rate = ledgr_risk_free_rate(0.04, label = "manual assumption")
)
```

The full constructor fields are `risk_free_rate`, `calendar`, `benchmark`,
`market_factor`, and `mar`. The provider fields `benchmark`, `market_factor`,
and `mar` are reserved and must be `NULL`; they exist so future benchmark and
provider designs have an explicit home instead of changing metric semantics
later.

Intraday work should set `calendar` explicitly. For example, US equity minute
bars use `ledgr_calendar_us_equity(bars_per_day = 390L)`. ledgr does not infer
that policy from ticker symbols, file names, or provider names.

When run, sweep, comparison, or walk-forward metrics see clearly subdaily
timestamps under a daily annualization context, ledgr emits the classed warning
`ledgr_metric_context_cadence_mismatch`. The warning does not change metric
values, metric recipes, stored evidence, or identity. It is a prompt to rerun
the analysis with an explicit cadence, not first-class intraday execution
support.

### Stored Context vs Sensitivity Overrides

`summary(bt)` and `ledgr_compute_metrics(bt)` use the metric context stored with
the committed run. Call-time overrides are sensitivity checks; they do not
mutate the run:

```{r}
#| label: metric-context-sensitivity

stored_run_context <- ledgr_metric_context(bt)
stored_metrics <- ledgr_compute_metrics(bt)
zero_rf_metrics <- ledgr_compute_metrics(bt, risk_free_rate = 0)

ledgr_metric_context(stored_metrics)
ledgr_metric_context(zero_rf_metrics)
identical( # confirm the override did not mutate the stored context on `bt`
  ledgr_metric_context_hash(stored_run_context),
  ledgr_metric_context_hash(ledgr_metric_context(bt))
)
```

Metric-context hashes include the metric-context version, annual risk-free rate,
risk-free source, risk-free `as_of`, and calendar annualization and source
fields. Human display labels are stored for inspection but do not change the
hash. If the print output is too compact for a report, inspect the nested object
directly:

```{r}
#| label: metric-context-provenance

context <- ledgr_metric_context(bt)
context$risk_free_rate$label
context$risk_free_rate$source
context$risk_free_rate$as_of
ledgr_metric_context_hash(context)
```

### Comparison, Sweep, And Promotion Contexts

The following snippets use objects from the experiment-store and sweeps
workflows. They show where metric context is carried, not a complete runnable
example.

`ledgr_run_compare()` has exactly one comparison context per table. The
snapshot-first form uses the default context unless you pass one explicitly:

```{r}
#| label: comparison-context
#| eval: false

comparison <- ledgr_run_compare(
  snapshot,
  run_ids = c("trend_qty_5", "trend_qty_15"),
  metric_context = ledgr_metric_context(exp)
)

ledgr_metric_context(comparison)
```

Sweep result tables also have exactly one metric context. Promotion context
keeps that source sweep context separate from the committed run's own context:

```{r}
#| label: promotion-contexts
#| eval: false

results <- ledgr_sweep(train_exp, grid)
candidate <- ledgr_candidate(results, 1)
test_run <- ledgr_promote(test_exp, candidate, require_same_snapshot = FALSE)

ledgr_metric_context(results)
ledgr_metric_context(ledgr_promotion_context(test_run))
ledgr_metric_context(test_run)
```

This distinction matters for train/test work: the source sweep context explains
how a candidate was ranked, while the committed run context explains the default
analysis assumptions stored with the promoted run.

## Risk Metric Contract

The standard metric contract includes `sharpe_ratio` as ledgr's first
risk-adjusted metric. It is computed from the same public equity rows as
volatility, not from hidden runner state and not from an external metrics
package.

The return series is still the adjacent public equity-row return:

```text
equity_return[t] = equity[t] / equity[t - 1] - 1
```

Sharpe-style metrics use period excess returns:

```text
excess_return[t] = equity_return[t] - rf_period_return[t]
sharpe_ratio = mean(excess_return) / sd(excess_return) * sqrt(bars_per_year)
```

The first risk-free-rate provider is a scalar annual rate expressed as a
decimal, so `0.02` means two percent per year. The default is `0`. ledgr
converts that scalar annual rate to a per-period return with the same
`bars_per_year` used for annualized return and volatility:

```text
rf_period_return = (1 + rf_annual)^(1 / bars_per_year) - 1
```

Time-varying risk-free-rate series and real data providers such as FRED,
Treasury, ECB, or central-bank adapters are deferred to later metric-context
provider work. Future providers must feed the same pulse-aligned
`rf_period_return` vector into the formula above; they must not create a
separate Sharpe formula branch.

Sharpe returns `NA_real_` for short samples, flat equity, or near-zero
volatility; see `?ledgr_compute_metrics` for the exact edge-case rules.

Other risk-adjusted or benchmark-relative metrics are deferred to v0.2.x
benchmark context work: Sortino, Calmar, Omega, information ratio, alpha/beta,
benchmark-relative metrics, VaR, and tail-risk metrics.

```{r}
#| label: summary

summary(bt)
```

`summary(bt)` is a print-oriented view. It returns the backtest handle
invisibly, not a metrics object. Use `ledgr_compute_metrics()` for scripted
workflows:

```{r}
#| label: metric-object

metrics <- ledgr_compute_metrics(bt)
metrics[c("total_return", "sharpe_ratio", "n_trades", "win_rate")]
ledgr_metric_context(metrics)
```

The raw metrics object keeps metric-kernel attributes for provenance. Those
attributes are part of the programmatic object, not the printed metric table.
Use named fields such as `metrics$sharpe_ratio` or the subset above when you
need a compact report.

`ledgr_run_compare()` is also programmatic: it returns a tibble-like
`ledgr_comparison` object with raw numeric metric columns for filtering and
ranking. Its print method only curates the displayed columns. Comparison metrics
are recomputed from stored equity and fill tables and use the same closed-trade
semantics as `ledgr_compute_metrics()`.

For reports, convert the comparison object and keep the raw numeric columns:

```{r}
#| label: comparison-report
#| eval: false

comparison |>
  as.data.frame() |>
  select(run_id, final_equity, total_return, sharpe_ratio, max_drawdown)
```

## Zero Trades In The Metrics

A completed run with zero closed trades reports `win_rate` and `avg_trade` as
`NA`, not zero, because there is no closed trade to average. To find out why a
run made no trades, use the checklist in
`vignette("indicators", package = "ledgr")`.

## Cleanup

```{r}
#| label: cleanup
#| include: false

close(bt)
```

## Where Next

- `vignette("execution-semantics", package = "ledgr")` explains when and at
  what price a target fills.
- `vignette("metrics-and-accounting", package = "ledgr")` is the shorter
  accounting-first tutorial.
- `vignette("sweeps", package = "ledgr")` shows how metric contexts thread
  through parameter sweeps.
