---
title: "Missing Data and Session Calendars"
format:
  html:
    toc: true
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{Missing Data and Session Calendars}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 90)
options(cli.unicode = FALSE)
pkg_root <- if (file.exists("DESCRIPTION")) "." else if (file.exists("../DESCRIPTION")) ".." else NA_character_
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
```

Your price file has holes in it. An instrument was halted for a day, a vendor
dropped a session, a listing began halfway through your window, a venue closed
for a holiday your file knows nothing about. What should a backtester do with
a day on which one of your instruments has no price?

The tempting answer is to carry the previous close forward so that every
instrument has a row on every day. That answer quietly invents trades: once a
carried price is in the panel, nothing distinguishes it from a real one, and
your backtest will fill orders at prices no venue ever printed.

ledgr takes the other path. It never manufactures a price. Instead it asks you
to say which sessions were supposed to exist, and then tells you, session by
session and instrument by instrument, what it knew and what it did not.

By the end of this article you will have run a backtest over a panel with a
real three-session hole in it, seen why nothing filled during the hole, and
chosen the point at which an old price stops being good enough to value a
position you still hold.

## The Big Picture: Four Questions, Not One

"Missing" feels like one condition. In a backtest it is at least four, and
they have different economic consequences. ledgr keeps them apart, and each
one has its own home in the data you supply.

| The question | Answered by | Consequence when the answer is no |
| --- | --- | --- |
| Was the venue open at all? | the session calendar | there is no pulse; nothing is expected of anyone |
| Did this instrument trade? | whether a bar row exists | the instrument is unpriced; its last mark starts ageing |
| Was it allowed to trade? | trading-status facts | you may hold or exit it, but not resize it |
| Was it listed at all? | lifetime facts | it is not on the decision axis |

The first two questions cannot be told apart from a price file alone. A closed
market and a vendor omission both look like "no row". That is why the session
calendar comes first: it is the clock that says what *should* have happened,
independent of what your data happens to contain.

A fifth case, a row that is present but wrong, is handled at the end.

::: {.ledgr-callout .ledgr-callout-warning}
**Do not pad your panel before sealing it**

If you fill the holes with the previous close so that every instrument has a
row on every session, ledgr accepts it. It cannot tell your invented rows from
real ones, and it will execute against them. Supply the panel with its holes
intact and declare a session calendar instead. The rest of this article is
about what that buys you.
:::

## Prerequisites

```{r}
#| label: prerequisites
#| message: false

library(ledgr)
library(dplyr)
library(tibble)
library(qlcal)
```

## Relationship To The Composable Bundle

The [Data Model And Point-In-Time Inputs](data-input-and-snapshots.qmd) article owns
the complete input decision map and the reusable point-in-time bundle. This
article keeps its smaller `AAA`/`BBB` fixture because a three-session hole is
the clearest detector for mark ageing and the no-fill rule. Those identifiers
are local to this article; they do not describe extra history for the shared
`DEMO_*` instruments.

The shared bundle carries the same kind of missing-observation input. This
executable touchpoint locates it and confirms that its bar row is genuinely
absent rather than padded:

```{r}
#| label: shared-missing-observation

data("ledgr_demo_pit_inputs", package = "ledgr")
shared_gap <- subset(
  ledgr_demo_pit_inputs$cases,
  type == "missing_observation"
)
shared_gap_has_bar <- with(
  ledgr_demo_pit_inputs$bars,
  any(
    instrument_id == shared_gap$instrument_id[[1L]] &
      as.Date(ts_utc) == shared_gap$date[[1L]]
  )
)
data.frame(shared_gap, observation_row_present = shared_gap_has_bar)
```

## Step 1: Declare What Sessions Existed

The expected-session clock has to come from somewhere other than your prices.
If you derive sessions from the data you happen to have, a day on which your
whole universe is missing becomes invisible, and that is exactly the day you
most need to see.

`qlcal`, a suggested dependency, ships exchange calendars, and
`ledgr_facts_sessions_qlcal()` turns one into a ledgr fact family: a set of
assertions about a venue that become part of the snapshot's identity. The
window below is two weeks of NYSE trading in January 2020.

```{r}
#| label: session-calendar

window_start <- as.Date("2020-01-13")
window_end <- as.Date("2020-01-24")

sessions <- ledgr_facts_sessions_qlcal(
  calendar = getCalendar("UnitedStates/NYSE"),
  from = window_start,
  to = window_end,
  venue_id = "XNYS",
  timezone = "America/New_York",
  session_open = "09:30:00",
  session_close = "16:00:00"
)
```

Read any fact family back with `ledgr_facts_history()`. Here the `status`
column is the lesson, not the row count.

```{r}
#| label: session-rows

session_rows <- ledgr_facts_history(
  ledgr_facts(sessions),
  family = "sessions",
  scope_id = "XNYS",
  from = window_start,
  to = window_end
)$rows

session_rows |>
  as_tibble() |>
  select(session_date, status) |>
  print(n = 12)
```

Twelve calendar days, nine of them open. The weekend is there, and so is
Martin Luther King Jr. Day on 2020-01-20, each marked `closed` rather than
left out. Stating the closures is what lets ledgr distinguish "the market was
shut" from "this instrument has no row today".

::: {.ledgr-callout .ledgr-callout-note}
**No calendar package required**

`ledgr_facts_sessions()` takes a plain data frame with `session_date`,
`status`, `session_open`, and `session_close`. Reach for `qlcal` when it
covers your venue, and build the frame yourself when it does not, or when your
venue's history needs corrections the package does not carry.
:::

The nine open sessions also give you the timestamps your bars should carry.
Take them from the calendar rather than typing them out, so that the panel and
the clock cannot drift apart.

```{r}
#| label: trading-sessions

trading_sessions <- session_rows |>
  as_tibble() |>
  filter(status == "open") |>
  pull(session_close)

length(trading_sessions)
```

## Step 2: Build a Panel With a Real Hole

Two instruments over those nine sessions. `AAA` trades on all of them. `BBB`
is absent for three sessions in the middle: no row at all, not a zero-volume
row and not a carried price.

```{r}
#| label: panel

prices <- tibble(instrument_id = c("AAA", "BBB"), base_price = c(100, 50)) |>
  cross_join(tibble(ts_utc = trading_sessions)) |>
  group_by(instrument_id) |>
  mutate(close = base_price + row_number()) |>
  ungroup() |>
  mutate(open = close, high = close, low = close, volume = 1000)

hole_dates <- as.Date(c("2020-01-16", "2020-01-17", "2020-01-21"))
missing_sessions <- trading_sessions[as.Date(trading_sessions) %in% hole_dates]

bars <- prices |>
  filter(!(instrument_id == "BBB" & ts_utc %in% missing_sessions)) |>
  select(instrument_id, ts_utc, open, high, low, close, volume)

bars |> count(instrument_id)
```

`BBB` is missing on 16, 17 and 21 January. Nine rows for `AAA` and six for
`BBB`: that asymmetry is the point, because you are handing ledgr an honest
panel rather than a rectangular one.

## What The Calendar Means For Indicators

Dense mode has an implicit clock: the dates present in the rectangular bar
panel. If every instrument is absent on one date, that date is
indistinguishable from a market closure and there is no pulse. Availability
mode instead advances on the independently declared expected-session
calendar. A declared open session remains a pulse even when one or every
instrument lacks an observation.

The small matrix below is the executable oracle used by ledgr's strict-gap
tests. It has one closed date, one missing open-session observation for
`AAA`, and a `BBB` series that starts late. Both moving averages use a
two-session window.

<!-- strict-gap-table:start -->
| Date | Session | AAA bar | AAA SMA(2) | BBB bar | BBB SMA(2) | Interpretation |
| --- | --- | ---: | ---: | ---: | ---: | --- |
| 2020-01-01 | open | 10 | NA | absent | NA | AAA warmup; BBB not yet observed |
| 2020-01-02 | open | 12 | 11 | absent | NA | AAA ready; BBB not yet observed |
| 2020-01-03 | closed | no pulse | no pulse | no pulse | no pulse | venue closure |
| 2020-01-04 | open | missing | NA | absent | NA | expected observation missing; BBB still unseen |
| 2020-01-05 | open | 14 | NA | 20 | NA | AAA window still contains the gap; BBB warmup |
| 2020-01-06 | open | 16 | 15 | 22 | 21 | both windows recover |
| 2020-01-07 | open | 18 | 17 | 24 | 23 | both windows remain available |
<!-- strict-gap-table:end -->

These unavailable windows overlap on 5 January, but for different reasons.
`AAA` has enough historical observations in total and is unavailable because
its current window includes the missing open session. `BBB` is unavailable
because its late-start series has not completed warmup. A scalar `NA` alone
does not distinguish those histories; the calendar and source observations
do.

ledgr never substitutes a stale valuation mark into this feature window. If a
strategy already owns `AAA`, `ledgr_valuation_stale(1)` may use the 2 January
close to value that holding at the 4 January pulse. The `AAA` SMA remains
`NA`, because valuation continuity is not feature evidence. On 6 January the
window contains two real observations again and the SMA recovers to 15.

A strategy therefore guards the value it intends to use as well as the
instrument's current trading permission:

```r
signal <- ctx$features(id)[["signal"]]
if (is.finite(signal) && id %in% ctx$tradable()) {
  # It is now safe to use `signal` in a target decision.
}
```

This is stricter than a generic warmup check in one useful way: it says what
the strategy actually needs at this pulse. See
`vignette("indicators", package = "ledgr")` for the built-in, public TTR, and
custom indicator forms that are certified for this strict-window behavior.

## Step 3: Say Which Instruments Were Tradable

A missing row says an instrument did not trade. It does not say why. Trading
status is where you record the reason when you know it, and it changes what a
strategy is allowed to ask for.

Each row below asserts a status from `effective_from` until `effective_to`,
with the higher `precedence` winning where two rows overlap. Both instruments
are active for the whole window, and `AAA` is additionally halted for the
whole of 23 January. An open-ended assertion leaves `effective_to` empty.

```{r}
#| label: trading-status

window_open <- as.POSIXct(window_start, tz = "UTC")
halt_from <- as.POSIXct("2020-01-23", tz = "UTC")
halt_to <- as.POSIXct("2020-01-24", tz = "UTC")

status <- tribble(
  ~instrument_id, ~effective_from, ~effective_to, ~status,  ~precedence,
  "AAA",          window_open,     NA,            "active", 0L,
  "BBB",          window_open,     NA,            "active", 0L,
  "AAA",          halt_from,       halt_to,       "halted", 1L
) |>
  mutate(source = "vignette-example") |>
  ledgr_facts_trading_status(knowledge = "assume_effective")
```

`knowledge = "assume_effective"` says each assertion is known from the moment
it takes effect, which is the right reading for a reference file you treat as
settled history. When your source records when you *learned* something, use
`knowledge = "evidenced"` with a `knowledge_time` column instead, so that a
late correction cannot leak backwards into earlier decisions.

::: {.ledgr-callout .ledgr-callout-warning}
**Declaring status is all or nothing**

Once any trading-status fact exists, an instrument with no applicable
assertion resolves to `unknown` and is restricted, not assumed active. Silence
about a halt is not evidence that there was none. Declare the ordinary
`active` intervals as well as the exceptions, as the three rows above do.
:::

## Step 4: Seal the Snapshot

Sealing binds bars and facts together under one hash. Declaring sessions also
moves the run out of the dense mode, which requires every instrument on every
session, and into the availability-aware mode the rest of this article uses.

```{r}
#| label: seal

snapshot <- ledgr_snapshot_from_df(
  bars,
  instruments_df = tibble(instrument_id = c("AAA", "BBB")),
  facts = ledgr_facts(sessions, status),
  db_path = tempfile(fileext = ".duckdb")
)
```

For sealing in general, see
`vignette("data-input-and-snapshots", package = "ledgr")`.

## Step 5: Write a Strategy That Reads Availability

An availability-aware strategy receives universe-aligned planes on `ctx$vec`.
Four of them carry everything this article is about:

- `priced` — does this instrument have a mark your valuation policy accepts?
- `mark_age` — how many sessions old is that mark?
- `mark_source` — `current_close` when it traded today, `stale_close` when
  you are carrying an older one, `expired_close` once it is too old to use.
- `target_restricted`, with `target_restriction_reason` — may you resize it?

The strategy below buys ten units of everything it is allowed to buy.
`ctx$hold()` starts from the positions already held, while `ctx$tradable()`
returns the current members that are admissible and priced. Anything halted,
unlisted, or unpriced therefore stays at its current quantity without a
strategy-side eligibility reconstruction.

```{r}
#| label: strategy

buy_what_is_tradable <- function(ctx, params) {
  targets <- ctx$hold()
  targets[ctx$tradable()] <- 10
  targets
}
```

Asking for a new size on a restricted instrument fails the pulse rather than
trading silently. That is the engine refusing to do what the venue would not
have allowed.

## Step 6: Choose How Stale Is Too Stale

When an instrument has no price today and you still hold it, something has to
decide whether yesterday's close is good enough to value the position.
`ledgr_valuation_stale(max_sessions)` is the only place that decision lives,
and ledgr ships no default, because there is no defensible universal answer.

Start strict: a mark may be at most one session old.

```{r}
#| label: run-tight

tight <- ledgr_experiment(
  snapshot,
  buy_what_is_tradable,
  opening = ledgr_opening(cash = 100000),
  cost_model = ledgr_cost_zero(),
  valuation_policy = ledgr_valuation_stale(max_sessions = 1)
)

tight_run <- ledgr_run(tight, run_id = "tolerance-one")
tight_completion <- ledgr_run_completion(tight_run)
close(tight_run)
```

The call returns without an error, so look at what the run actually claims.

```{r}
#| label: run-tight-status

tight_completion
```

The run is `INCOMPLETE`, and it says why: `valuation_horizon_exhausted`. On
the second session of the hole, `BBB`'s last close became older than the one
session you allowed, the portfolio could no longer be valued, and the run
stopped. It did not truncate silently and it did not carry the price forward
to keep going. An incomplete result with a named stop reason is usable
evidence about a shorter window; a complete-looking result built on invented
prices is not.

Now allow a mark to be up to five sessions old, and run the same panel again.

```{r}
#| label: run-loose

loose <- ledgr_experiment(
  snapshot,
  buy_what_is_tradable,
  opening = ledgr_opening(cash = 100000),
  cost_model = ledgr_cost_zero(),
  valuation_policy = ledgr_valuation_stale(max_sessions = 5)
)

loose_run <- ledgr_run(loose, run_id = "tolerance-five")

ledgr_run_completion(loose_run)
```

Same data, same strategy, a different answer about whether the result covers
the window you asked for. The valuation policy is part of run identity, so a
reader can always see which tolerance produced which result.

## Reading Back What Happened to One Instrument

After a run, `ledgr_run_explain()` answers "what did you know about this
instrument across the run, and what did you do about it?". Omit the timestamp
to see the full pulse history for `BBB`:

```{r}
#| label: explain-hole

ledgr_run_explain(loose_run, "BBB") |>
  select(mark_source, mark_age, execution_outcome, execution_reason)
```

`BBB` is carried at a `stale_close` two sessions old, and nothing executes.
The mark is doing valuation work only. It is never an execution price, and no
fill is ever produced on a session where an instrument has no observation.

Supplying one timestamp keeps the focused form. Ask about `AAA` on the session
it was halted:

```{r}
#| label: explain-halt

halted_day <- as.Date(halt_from)
halt_session <- trading_sessions[as.Date(trading_sessions) == halted_day]

ledgr_run_explain(loose_run, "AAA", ts_utc = halt_session) |>
  select(mark_source, target_restricted, target_restriction_reason)
```

`AAA` has a perfectly good price that session, so it is valued normally, but
`trading_halted` blocks a resize. Being priced and being tradable are separate
properties, and this is the case that shows why.

```{r}
#| label: close-loose
#| include: false

close(loose_run)
```

::: {.ledgr-callout .ledgr-callout-tip}
**Try it**

Set `max_sessions = 2` and run again. Does the run complete? Then widen the
hole to four sessions and re-seal. Which of the two knobs, your data or your
tolerance, decides whether you get a full-window result?
:::

## Rows That Are Present but Wrong

The last case is a row that exists and is invalid, such as a non-positive
close. By default this fails the seal, which is what you want while you are
still cleaning a source.

```{r}
#| label: quarantine-default

broken <- bars |>
  add_row(
    instrument_id = "BBB",
    ts_utc = missing_sessions[1],
    open = 54, high = 54, low = 54, close = -1, volume = 1000
  ) |>
  arrange(instrument_id, ts_utc)

tryCatch(
  ledgr_snapshot_from_df(
    broken,
    instruments_df = tibble(instrument_id = c("AAA", "BBB")),
    facts = ledgr_facts(sessions, status),
    db_path = tempfile(fileext = ".duckdb")
  ),
  ledgr_availability_validation_failed = function(error) {
    conditionMessage(error)
  }
)
```

With sessions declared you can instead admit the panel and have the offending
row set aside with a recorded reason. That keeps the exclusion inspectable
inside the snapshot rather than buried in a cleaning script you ran once.

```{r}
#| label: quarantine-admitted

quarantined <- ledgr_snapshot_from_df(
  broken,
  instruments_df = tibble(instrument_id = c("AAA", "BBB")),
  facts = ledgr_facts(sessions, status),
  invalid_observations = "quarantine",
  db_path = tempfile(fileext = ".duckdb")
)

ledgr_snapshot_quarantine(quarantined) |>
  select(supplied_instrument_id, supplied_ts_utc, reason)
```

```{r}
#| label: cleanup
#| include: false

ledgr_snapshot_close(quarantined)
ledgr_snapshot_close(snapshot)
```

The quarantined row is not in the panel, so that session behaves exactly like
the vendor hole earlier: unpriced, ageing, no fill. The difference is that the
snapshot records why it is missing.

::: {.ledgr-callout .ledgr-callout-important}
**Two accessors are still missing**

Quarantined rows are read from the store directly above because there is no
accessor for them yet, and `ledgr_run_explain()` answers one timestamp per
call, so a full per-session history still needs a loop. The evidence is
persisted and hash-bound in both cases; the convenience readers are roadmap
items.
:::

## What Stays Your Decision

ledgr represents absence. It does not interpret it. Three judgments remain
yours, and each is expressed as data you supply rather than as a price you
patch.

- **Why an observation is missing.** A vendor omission and a genuine no-trade
  day look identical in a price file. When the difference matters to your
  result, say which it was with trading-status or lifetime facts.
- **How stale is acceptable.** `ledgr_valuation_stale(max_sessions)` is a
  claim about your instruments and your holding period. Short-horizon work on
  liquid names tolerates far less than a quarterly rebalance on small caps.
- **Whether an incomplete run is good enough.** An `INCOMPLETE` result with a
  named stop reason is a real answer about a shorter window. Reporting it as a
  full-window result is the error.

What is not yours to decide is whether a missing price may become a traded
price. It may not, and no setting changes that.

## Where Next

- For sealing, adapters, and snapshot identity, see
  `vignette("data-input-and-snapshots", package = "ledgr")`.
- For what a universe that changes over time does to your results, see
  `vignette("survivorship-bias", package = "ledgr")`.
- For when orders fill and at which bar, see
  `vignette("execution-semantics", package = "ledgr")`.
- For the valuation and availability policy arguments in detail, see
  `?ledgr_valuation_stale`.
