---
title: "Preparing Point-In-Time Inputs"
format:
  html:
    toc: false
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{Preparing Point-In-Time Inputs}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

options(width = 90)
options(cli.unicode = FALSE)
pkg_root <- if (file.exists("DESCRIPTION")) {
  "."
} else if (file.exists("../DESCRIPTION")) {
  ".."
} else {
  NA_character_
}
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
default_output_hook <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(
  output = function(x, options) {
    x <- gsub(
      "Database:(\\s+)[^\\r\\n]+[.]duckdb",
      "Database:\\1<temporary DuckDB path>",
      x,
      perl = TRUE
    )
    x <- gsub("[ \t]+(?=\n)", "", x, perl = TRUE)
    default_output_hook(x, options)
  }
)
```

Bars tell you what was observed. They do not tell you whether a venue was
open, which instruments belonged to a research universe, whether an instrument
could trade, or when evidence became knowable.

This article adds those claims one at a time. By the end, you will have sealed
and run one point-in-time bundle, seen a late-known halt change the public
decision view, and have a reference map for every accepted input family.

```{r}
#| label: library
#| message: false

library(ledgr)
library(dplyr)
data("ledgr_demo_pit_inputs", package = "ledgr")

pit <- ledgr_demo_pit_inputs
pit$cases
```

The bundle is synthetic and self-contained. Its bars, physical instrument
master, calendar, and facts describe the same five `DEMO_*` instruments over
the same window. Do not attach its facts to an unrelated observation panel.

## Start With The Missing Observation

The manifest says `DEMO_04` has no bar on an otherwise open date.

```{r}
#| label: missing-observation

gap <- pit$cases |>
  filter(type == "missing_observation")

tibble(
  instrument_id = gap$instrument_id,
  session_date = gap$date,
  observation_present = with(
    pit$bars,
    any(
      instrument_id == gap$instrument_id &
        as.Date(ts_utc) == gap$date
    )
  )
)
```

A price file alone cannot distinguish this gap from a venue closure. The
independent session calendar makes that distinction.

```{r}
#| label: session-facts

session_facts <- ledgr_facts_sessions(
  pit$sessions,
  venue_id = "DEMO_VENUE",
  knowledge = "evidenced",
  timezone = "America/New_York"
)

pit$sessions |>
  filter(session_date %in% c(as.Date("2020-01-06"), gap$date)) |>
  select(session_date, status, session_close)
```

The closure is declared `closed`; the gap date remains `open`. ledgr can now
advance the expected-session clock without manufacturing a price. Declaring
any availability family requires a complete session calendar. Under
`knowledge = "evidenced"`, every session row needs a `knowledge_time`: an open
session must be known by its open, and a closure before its civil day starts.
Use `"assume_effective"` only when that assumption is part of the research
design.

## Add A Point-In-Time Universe

Universe membership answers a different question: which physical instruments
were eligible for this research population? This bundle declares one complete
membership set.

```{r}
#| label: membership-facts

membership_facts <- ledgr_facts_membership_snapshots(
  pit$membership,
  universe_id = "demo_members",
  complete = TRUE,
  knowledge = "evidenced"
)

pit$membership |>
  select(effective_from, knowledge_time, instrument_id)
```

Because the set is complete, omission means non-membership for that effective
set. Use interval assertions instead when your source supplies individual
membership changes rather than complete lists.

For membership under `knowledge = "evidenced"`, a missing knowledge time is
kept for audit but cannot affect resolution. Membership evidence may become
knowable after it took effect; `"assume_effective"` deliberately removes that
distinction.

This bundle keeps all five instruments in the set. To see membership changes
alter a result, continue with
`vignette("survivorship-bias", package = "ledgr")`.

## Effective Time Is Not Knowledge Time

`DEMO_02` was halted from 3 January, but the bundle says that evidence became
knowable only on 7 January. The fact must not change decisions made earlier.

```{r}
#| label: halt-clock

halt_case <- pit$cases |>
  filter(type == "halt")
halt_rows <- pit$trading_status |>
  filter(instrument_id == halt_case$instrument_id[[1L]])
halt <- halt_rows |>
  filter(status == "halted")

halt_rows |>
  select(
    instrument_id,
    effective_from,
    effective_to,
    knowledge_time,
    status,
    precedence
  )
```

The open-ended active assertion has precedence zero. The overlapping halt has
precedence one and wins only after it is known.

```{r}
#| label: status-facts

status_facts <- ledgr_facts_trading_status(
  pit$trading_status,
  knowledge = "evidenced"
)
```

This separation is the core point-in-time rule: effective time says when the
world changed; knowledge time says when the backtest may use the evidence.
Status evidence may arrive late. A status row with missing evidenced knowledge
stays audit-only rather than becoming an active restriction.

## Add Lifetime And Economic Events When Needed

Lifetime facts distinguish an instrument that ceased to exist from one that
merely lacks a bar. Corporate-action facts record evidenced economic events.

```{r}
#| label: remaining-facts

lifetime_facts <- ledgr_facts_lifetime(
  pit$lifetime,
  knowledge = "evidenced"
)

corporate_action_facts <- ledgr_facts_equity_corporate_actions(
  pit$corporate_actions
)

inactive <- pit$lifetime |>
  filter(assertion == "known_inactive")
inactive_id <- inactive$instrument_id[[1L]]

tibble(
  instrument_id = inactive_id,
  last_observation = max(
    pit$bars$ts_utc[pit$bars$instrument_id == inactive_id]
  ),
  known_inactive_from = inactive$effective_from[[1L]],
  knowledge_time = inactive$knowledge_time[[1L]],
  terminal_event = inactive$terminal_event[[1L]]
)

pit$corporate_actions |>
  select(
    subtype,
    parent_instrument_id,
    entitlement_time,
    effective_time,
    knowledge_time,
    payment_time
  )
```

Lifetime follows the same late-knowledge rule as status: a missing evidenced
knowledge time leaves the row audit-only rather than changing the run.

Corporate actions have four clocks because entitlement, market effect,
knowledge, and payment need not coincide. Do not collapse them into one vendor
date during preparation. The flat run below holds no instrument, so it verifies
the terms and clocks rather than a cash effect. The held-position consequence
is demonstrated in `vignette("corporate-action-cash", package = "ledgr")`.
Corporate-action facts alone do not activate the availability model.

## Seal And Run The Combined Evidence

Combine the fact families, then pass the plain bundle through the ordinary
snapshot API. The scope and knowledge choices are explicit in the code above;
they are not hidden inside the bundle's recipe.

```{r}
#| label: seal-pit-snapshot

pit_facts <- ledgr_facts(
  session_facts,
  membership_facts,
  status_facts,
  lifetime_facts,
  corporate_action_facts
)

pit_path <- ledgr_temp_store(file.path(tempdir(), "ledgr_pit_demo.duckdb"))
pit_snapshot <- ledgr_snapshot_from_df(
  pit$bars,
  instruments_df = pit$instruments,
  facts = pit_facts,
  db_path = pit_path,
  snapshot_id = "pit_demo_snapshot",
  price_basis = "split_adjusted"
)

pit_snapshot_id <- pit_snapshot$snapshot_id
ledgr_snapshot_close(pit_snapshot)
pit_snapshot <- ledgr_snapshot_open(
  pit_path,
  snapshot_id = pit_snapshot_id,
  verify = TRUE
)
```

`price_basis = "split_adjusted"` declares how this synthetic panel was
prepared; it does not adjust prices. Do not copy that declaration onto raw or
fully adjusted vendor bars.

Availability-aware runs require an explicit stale-mark policy. The flat
strategy below makes no trades; its purpose is to show what the decision view
knows at the three halt boundaries. `ledgr_valuation_stale(2)` permits a held
instrument's last real close to value it for at most two declared sessions; it
does not create an execution price.

```{r}
#| label: run-pit-snapshot

flat_strategy <- function(ctx, params) ctx$flat()
pit_experiment <- ledgr_experiment(
  pit_snapshot,
  flat_strategy,
  universe = ledgr_universe_members("demo_members"),
  valuation_policy = ledgr_valuation_stale(2),
  cost_model = ledgr_cost_zero()
)
pit_run <- ledgr_run(pit_experiment, run_id = "pit-demo-run")

halt_view <- bind_rows(
  ledgr_run_explain(
    pit_run,
    halt$instrument_id[[1]],
    halt$effective_from[[1]]
  ) |>
    mutate(moment = "effective, not yet known"),
  ledgr_run_explain(
    pit_run,
    halt$instrument_id[[1]],
    halt$knowledge_time[[1]]
  ) |>
    mutate(moment = "known"),
  ledgr_run_explain(
    pit_run,
    halt$instrument_id[[1]],
    halt$effective_to[[1]]
  ) |>
    mutate(moment = "interval ended")
)

halt_view |>
  select(moment, target_restricted, target_restriction_reason)

unlist(ledgr_run_completion(pit_run)[
  c("completion_status", "complete_performance")
])
```

The first decision remains unrestricted because the halt was not yet known.
The second is restricted for `trading_halted`. The excluded interval end is
active again.

## Data Dictionary

Use this section as a reference after you understand why each family exists.
Constructor help remains authoritative for exhaustive optional columns and
accepted values.

| Input | Row grain and identity | Required core columns or shape | Scope and time | What it unlocks | Constructor help |
|---|---|---|---|---|---|
| Bars | One row per unique `(instrument_id, ts_utc)` | `instrument_id`, `ts_utc`, `open`, `high`, `low`, `close`; `volume` optional | Physical instrument; observation clock | Prices, valuation and execution inputs | `ledgr_snapshot_from_df()` |
| Instruments | One row per unique `instrument_id` | `instrument_id`; descriptive fields optional | Physical master; stable identity | Metadata and referential integrity | `ledgr_snapshot_from_df()` |
| Sessions | One row per civil date in a venue range | `session_date`, `status`, `knowledge_time` (unless `assume_effective`); open rows also require `session_open`, `session_close` | `venue_id`; local calendar plus knowledge time | Expected-session gaps and mark ageing | `ledgr_facts_sessions()` |
| Membership intervals | One assertion per instrument interval | `instrument_id`, `effective_from`, `member`; optional `effective_to` | `universe_id`; half-open effective interval plus knowledge time | Point-in-time eligibility changes | `ledgr_facts_membership_intervals()` |
| Complete membership snapshots | One set per effective/knowledge/source group | `effective_from` plus row-wise `instrument_id` or one `members` list-column; `complete` is declared | `universe_id`; omission means non-membership only when complete | Complete point-in-time populations | `ledgr_facts_membership_snapshots()` |
| Trading status | One sourced assertion per interval and precedence | `instrument_id`, `effective_from`, `status`, `source`; optional `effective_to`, `precedence` and supersession fields | Physical instrument; effective and knowledge times | Hold/exit restrictions such as halts | `ledgr_facts_trading_status()` |
| Lifetime | One assertion per instrument interval | `instrument_id`, `effective_from`, `assertion`; optional `effective_to`, `terminal_event` | Physical instrument; effective and knowledge times | Listing and terminal-state restrictions | `ledgr_facts_lifetime()` |
| Equity corporate actions | One identified economic fact | `subtype`, `parent_instrument_id`, `complete`, `provenance_tier`; terms and `refusal_reason` are conditional | Physical parent/recipient; entitlement, effective, knowledge and payment times | Supported settlement or explicit refusal | `ledgr_facts_equity_corporate_actions()` |

Every interval is half-open: `effective_from` is included and `effective_to`
is excluded. Across families:

- identifiers stay stable for the history being sealed;
- observation and fact timestamps are whole-second UTC instants; and
- knowledge is declared per family rather than assumed to precede effect;
  for membership, status and lifetime, missing evidenced knowledge remains
  audit-only unless `assume_effective` is declared. Sessions instead require a
  valid knowledge time under the evidenced policy.

## Entity Relationships

The physical instrument master, venue calendar, and research universe are
different scopes.

::: {.ledgr-diagram .ledgr-input-entities}

```{mermaid}
%%| echo: false
erDiagram
  SNAPSHOT ||--o{ INSTRUMENT : contains
  INSTRUMENT ||--o{ BAR : observed_as
  SNAPSHOT ||--o{ SESSION : seals
  SNAPSHOT ||--o{ MEMBERSHIP : seals
  INSTRUMENT ||--o{ MEMBERSHIP : referenced_by
  INSTRUMENT ||--o{ TRADING_STATUS : constrained_by
  INSTRUMENT ||--o{ LIFETIME : described_by
  INSTRUMENT ||--o{ CORPORATE_ACTION : parent_or_recipient
  SNAPSHOT {
    string snapshot_id PK
    string snapshot_hash
  }
  INSTRUMENT {
    string instrument_id PK
  }
  BAR {
    string instrument_id PK,FK
    datetime ts_utc PK
  }
  SESSION {
    string venue_id PK
    date session_date PK
  }
  MEMBERSHIP {
    string universe_id
    string instrument_id FK
    datetime effective_from
    datetime knowledge_time
  }
  TRADING_STATUS {
    string instrument_id FK
    datetime effective_from
    datetime knowledge_time
  }
  LIFETIME {
    string instrument_id FK
    datetime effective_from
    datetime knowledge_time
  }
  CORPORATE_ACTION {
    string fact_id PK
    string parent_instrument_id FK
    string recipient_instrument_id FK
  }
```

:::

Corporate-action parent and optional recipient ids resolve against the physical
master, not universe membership. Trading-status and lifetime facts also refer
to physical instruments; neither makes an instrument a universe member.

## What A Seal Establishes

After a successful seal, ledgr can rely on stable physical ids, whole-second
bar timestamps, unique bar keys, fact referential integrity, validated family
structure, and an immutable snapshot hash.

A seal does not imply a dense instrument-by-session rectangle, current
membership, continuously fresh prices, inferred corporate-action terms, or
knowledge before effectiveness. Those remain explicit evidence or policy
questions.

## Cleanup

```{r}
#| label: cleanup

close(pit_run)
ledgr_snapshot_close(pit_snapshot)
```

## Where Next

- `vignette("missing-data-and-sessions", package = "ledgr")` follows a real
  price gap through valuation and execution.
- `vignette("survivorship-bias", package = "ledgr")` shows the consequence of
  replacing historical membership with today's survivors.
- `vignette("corporate-action-cash", package = "ledgr")` shows a cash
  distribution changing portfolio accounting.
- `vignette("data-input-and-snapshots", package = "ledgr")` covers CSV and
  Yahoo imports, quarantine, and reopening.
- `vignette("strategy-development", package = "ledgr")` continues from sealed
  inputs to a strategy and its first run.
