---
title: "Quickstart"
format:
  html:
    toc: false
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{Quickstart}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 90)
options(cli.unicode = FALSE)
pkg_root <- if (file.exists("DESCRIPTION")) "." else if (file.exists("../DESCRIPTION")) ".." else NA_character_
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
```

You want the smallest useful ledgr loop before reading the tool chapters.

This quickstart creates a sealed snapshot, runs one strategy, sweeps a small
grid, and extracts one candidate for review. It is not a validation protocol
and it is not a production deployment path. It is the shortest useful path from
demo data to inspectable evidence.

## Load A Small Dataset

```{r}
#| label: prerequisites
#| message: false

library(ledgr)
library(dplyr)

data("ledgr_demo_bars", package = "ledgr")
```

Use two demo instruments over the first half of 2019. The code writes to a
temporary DuckDB store so the rendered article leaves no project artifacts
behind.

```{r}
#| label: data

store_path <- tempfile(fileext = ".duckdb")

bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% c("DEMO_01", "DEMO_02"),
    between(ts_utc, ledgr_utc("2019-01-01"), ledgr_utc("2019-06-30"))
  )

snapshot <- ledgr_snapshot_from_df(
  bars,
  db_path = store_path,
  snapshot_id = "quickstart_demo"
)
```

## Declare The Experiment

The strategy reads stable aliases named `fast` and `slow`. The concrete SMA
windows can vary later in the sweep.

```{r}
#| label: experiment

features <- ledgr_feature_map(
  fast = ledgr_ind_sma(ledgr_param("fast_n")),
  slow = ledgr_ind_sma(ledgr_param("slow_n"))
)

strategy <- ledgr_demo_sma_crossover_strategy()

exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = strategy,
  features = features,
  opening = ledgr_opening(cash = 10000),
  cost_model = ledgr_cost_zero()
)
```

`cost_model = ledgr_cost_zero()` is explicit. If you want modeled costs, use a
cost model such as `ledgr_cost_spread_bps()` or
`ledgr_cost_notional_bps_fee()` rather than hiding costs inside the strategy.

## Run One Case

Run one ordinary parameter set before sweeping. This catches setup mistakes
while the result is still easy to inspect.

```{r}
#| label: one-run

single_run <- ledgr_run(
  exp,
  params = list(qty = 5, threshold = 0),
  feature_params = list(fast_n = 5L, slow_n = 20L),
  run_id = "quickstart_run",
  seed = 2026L
)

summary(single_run)
head(ledgr_results(single_run, what = "fills"), 3)
```

If this run has no fills, impossible prices, or surprising exposure, stop here.
Sweeps amplify a setup; they do not repair it.

## Sweep A Tiny Grid

Build feature parameters and strategy parameters separately, then cross them
into candidate rows.

```{r}
#| label: sweep

feature_grid <- ledgr_feature_grid(
  fast_n = c(5L, 10L),
  slow_n = c(20L, 40L),
  .filter = fast_n < slow_n
)

strategy_grid <- ledgr_strategy_grid(
  qty = 5,
  threshold = c(0, 0.01)
)

grid <- ledgr_grid_cross(features = feature_grid, strategy = strategy_grid)

sweep <- ledgr_sweep(exp, grid, seed = 2026L)

sweep |>
  select(candidate_id, status, total_return, sharpe_ratio) |>
  arrange(desc(sharpe_ratio))
```

The sweep table is evidence, not an automatic recommendation. If you rank rows,
make the ranking rule visible at the call site.

```{r}
#| label: candidate

candidate <- ledgr_candidate(
  sweep |> arrange(desc(sharpe_ratio)),
  1L
)

candidate$candidate_id
```

Promotion is the step that turns one selected candidate into a durable run. In
a real project, use a deliberate `run_id` and a note that explains why this row
was selected.

```{r}
#| label: promotion-sketch
#| eval: false

promoted <- ledgr_promote(
  exp,
  candidate,
  run_id = "quickstart_promoted",
  note = "Highest Sharpe ratio in the tiny demo sweep."
)
```

## Where Next

Read `vignette("research-workflow", package = "ledgr")` for the full project
loop. Read `vignette("sweeps", package = "ledgr")` when the grid itself is the
thing you need to understand. Read `vignette("walk-forward", package =
"ledgr")` when you are ready to separate train-window selection from
test-window evaluation.

```{r}
#| label: cleanup
#| include: false

ledgr_snapshot_close(snapshot)
```
