---
title: "Quickstart"
format:
  html:
    toc: false
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{Quickstart}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 90)
options(cli.unicode = FALSE)
pkg_root <- if (file.exists("DESCRIPTION")) "." else if (file.exists("../DESCRIPTION")) ".." else NA_character_
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
```

You want the smallest useful ledgr loop before reading the tool chapters.

This quickstart creates a sealed snapshot, runs one strategy, sweeps a small
grid, and extracts one candidate for review. It is not a validation protocol
and it is not a production deployment path. It is the shortest useful path from
demo data to inspectable evidence.

## Load A Small Dataset

```{r}
#| label: prerequisites
#| message: false

library(ledgr)
library(dplyr)

data("ledgr_demo_bars", package = "ledgr")
```

Use two demo instruments over the first half of 2019. The code writes to a
temporary DuckDB store so the rendered article leaves no project artifacts
behind.

```{r}
#| label: data

store_path <- tempfile(fileext = ".duckdb")

bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% c("DEMO_01", "DEMO_02"),
    between(ts_utc, ledgr_utc("2019-01-01"), ledgr_utc("2019-06-30"))
  )

snapshot <- ledgr_snapshot_from_df(
  bars,
  db_path = store_path,
  snapshot_id = "quickstart_demo"
)
```

## Declare The Experiment

Start with a visible rule: hold five units when the current close is above its
20-session moving average, otherwise hold zero. The strategy reads the whole
decision axis at once.

```{r}
#| label: experiment

features <- list(ledgr_ind_sma(20))

strategy <- function(ctx, params) {
  sma <- ctx$vec$feature("sma_20")
  targets <- ctx$flat()
  selected <- which(
    is.finite(sma) &
      (ctx$vec$close / sma - 1) > params$threshold
  )
  targets[selected] <- params$qty
  targets
}

exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = strategy,
  features = features,
  opening = ledgr_opening(cash = 10000),
  cost_model = ledgr_cost_zero()
)
```

`cost_model = ledgr_cost_zero()` is explicit. If you want modeled costs, use a
cost model such as `ledgr_cost_spread_bps()` or
`ledgr_cost_notional_bps_fee()` rather than hiding costs inside the strategy.

## Run One Case

Run one ordinary parameter set before sweeping. This catches setup mistakes
while the result is still easy to inspect.

```{r}
#| label: one-run

single_run <- ledgr_run(
  exp,
  params = list(qty = 5, threshold = 0),
  run_id = "quickstart_run",
  seed = 2026L
)

single_run
```

If this run has no fills, impossible prices, or surprising exposure, stop here.
Sweeps amplify a setup; they do not repair it.

The printed warnings are deliberate. `Corporate actions: NOT SUPPLIED` means
this demo supplied no distribution evidence; `Price basis: UNDECLARED` means
ledgr cannot rule out dividend double counting in the bars. That is acceptable
for learning the mechanics, not for making an equity-research claim. See
`vignette("corporate-action-cash", package = "ledgr")` for the modeled path.

## Sweep A Tiny Grid

Vary only the strategy threshold. The feature remains the same concrete SMA
used by the single run, so each candidate is easy to explain.

```{r}
#| label: sweep

grid <- ledgr_strategy_grid(
  qty = 5,
  threshold = c(0, 0.01)
)

sweep <- ledgr_sweep(exp, grid, seed = 2026L)

sweep |>
  select(candidate_id, status, total_return, sharpe_ratio, params) |>
  arrange(desc(sharpe_ratio))
```

The sweep table is evidence, not an automatic recommendation. If you rank rows,
make the ranking rule visible at the call site.

```{r}
#| label: candidate

candidate <- ledgr_candidate(
  sweep |> arrange(desc(sharpe_ratio)),
  1L
)

candidate$candidate_id
```

Promotion is the step that turns one selected candidate into a durable run. In
a real project, use a deliberate `run_id` and a note that explains why this row
was selected.

```{r}
#| label: promotion-sketch

promoted <- ledgr_promote(
  exp,
  candidate,
  run_id = "quickstart_promoted",
  note = "Highest Sharpe ratio in the tiny demo sweep."
)

promoted
```

This promotion is safe to execute here because the article uses a temporary
store. It records the selection; it does not turn the tiny in-sample sweep into
validation.

## Where Next

Read `vignette("sweeps", package = "ledgr")` for the full project loop: grids,
candidate review, promotion, and reopening the promoted run. Read `vignette("walk-forward", package =
"ledgr")` when you are ready to separate train-window selection from
test-window evaluation.

```{r}
#| label: cleanup
#| include: false

close(single_run)
close(promoted)
ledgr_snapshot_close(snapshot)
```
