---
title: "Quickstart"
format:
  html:
    toc: false
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{Quickstart}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 90)
options(cli.unicode = FALSE)
pkg_root <- if (file.exists("DESCRIPTION")) "." else if (file.exists("../DESCRIPTION")) ".." else NA_character_
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
```

You want the smallest useful ledgr loop before reading the tool chapters.

In a few minutes you will have run one strategy on sealed demo data, tried a
small grid of variants, and kept one candidate as a run you can reopen. That is
the shortest useful path from demo data to inspectable evidence; it is not a
validation protocol and not a production deployment path.

## Load A Small Dataset

```{r}
#| label: prerequisites
#| message: false

library(ledgr)
library(dplyr)

data("ledgr_demo_bars", package = "ledgr")
```

Use two demo instruments over the first half of 2019. The code writes to a
temporary DuckDB store so the rendered article leaves no project artifacts
behind.

```{r}
#| label: data

store_path <- tempfile(fileext = ".duckdb")

bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% c("DEMO_01", "DEMO_02"),
    between(ts_utc, ledgr_utc("2019-01-01"), ledgr_utc("2019-06-30"))
  )

snapshot <- ledgr_snapshot_from_df(
  bars,
  db_path = store_path,
  snapshot_id = "quickstart_demo"
)
```

## Declare The Experiment

Start with a visible rule: hold five units when the current close is above its
20-session moving average, otherwise hold zero. The strategy reads the whole
decision axis at once. Until the average has 20 closes to work with it does not
exist, and the strategy holds nothing.

```{r}
#| label: experiment

features <- list(ledgr_ind_sma(20))

strategy <- function(ctx, params) {
  sma <- ctx$vec$feature("sma_20")
  if (!ledgr_passed_warmup(sma)) return(ctx$flat())
  above <- ctx$vec$close / sma - 1 > params$threshold
  targets <- ctx$flat()
  targets[above] <- params$qty
  targets
}

exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = strategy,
  features = features,
  opening = ledgr_opening(cash = 10000),
  cost_model = ledgr_cost_zero()
)
```

`cost_model = ledgr_cost_zero()` is explicit. If you want modeled costs, use a
cost model such as `ledgr_cost_spread_bps()` or
`ledgr_cost_notional_bps_fee()` rather than hiding costs inside the strategy.

## Run One Case

Run one ordinary parameter set before sweeping. This catches setup mistakes
while the result is still easy to inspect.

```{r}
#| label: one-run

single_run <- ledgr_run(
  exp,
  params = list(qty = 5, threshold = 0),
  run_id = "quickstart_run",
  seed = 2026L
)

single_run
```

If this run has no fills, impossible prices, or surprising exposure, stop here.
Sweeps amplify a setup; they do not repair it.

The last two lines say this demo supplied no dividend evidence and did not
declare how its prices were adjusted, which is fine for learning the mechanics
but not for an equity-research claim;
`vignette("corporate-action-cash", package = "ledgr")` explains both.

## Sweep A Tiny Grid

A sweep runs the same experiment once per parameter combination and keeps
every result; `vignette("sweeps", package = "ledgr")` covers it in full. Vary
only the strategy threshold. The feature remains the same concrete SMA used by
the single run, so each candidate is easy to explain.

```{r}
#| label: sweep

grid <- ledgr_strategy_grid(
  qty = 5,
  threshold = c(0, 0.01)
)

sweep <- ledgr_sweep(exp, grid, seed = 2026L)

sweep |>
  mutate(threshold = vapply(params, function(p) p$threshold, numeric(1))) |>
  select(candidate_id, threshold, status, total_return, sharpe_ratio) |>
  arrange(desc(sharpe_ratio))
```

Each row is one candidate: the `threshold` it ran with, its status and its
result. The Sharpe ratio is annualized from half a year of two demo
instruments, so at this size it describes the sample, not the rule. The sweep
table is evidence, not an automatic recommendation. If you rank rows, make the
ranking rule visible at the call site.

```{r}
#| label: candidate

candidate <- ledgr_candidate(
  sweep |> arrange(desc(sharpe_ratio)),
  1L
)

candidate
```

The candidate carries its parameters and the hashes of the inputs it ran on, so
the choice can be traced back to exactly what produced it.

Promotion is the step that turns one selected candidate into a durable run. In
a real project, use a deliberate `run_id` and a note that explains why this row
was selected.

```{r}
#| label: promotion-sketch

promoted <- ledgr_promote(
  exp,
  candidate,
  run_id = "quickstart_promoted",
  note = "Highest Sharpe ratio in the tiny demo sweep."
)

promoted
```

This promotion is safe to execute here because the article uses a temporary
store. It records the selection; it does not turn the tiny in-sample sweep into
validation.

## Where Next

- `vignette("data-input-and-snapshots", package = "ledgr")` seals your own
  data instead of the demo bars.
- `vignette("strategy-development", package = "ledgr")` teaches the strategy
  contract behind the rule you just ran.
- `vignette("research-workflow", package = "ledgr")` follows a research idea
  through code iterations, sweeps, and promotion.

```{r}
#| label: cleanup
#| include: false

close(single_run)
close(promoted)
ledgr_snapshot_close(snapshot)
```
