---
title: "On Reproducibility: Provenance and Strategy Tiers"
output:
  rmarkdown::html_vignette:
    toc: false
  github_document:
    toc: false
    html_preview: false
vignette: >
  %\VignetteIndexEntry{On Reproducibility: Provenance and Strategy Tiers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 90)
```

```{r library, message=FALSE}
library(ledgr)
library(dplyr)
data("ledgr_demo_bars", package = "ledgr")

bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% c("DEMO_01", "DEMO_02"),
    between(
      ts_utc,
      ledgr_utc("2019-01-01"),
      ledgr_utc("2019-06-30")
    )
  )
```

ledgr treats a backtest result as an experiment artifact. The question is not
only "what was the return?" The question is:

```text
which sealed data, which strategy, which parameters, which features,
which opening state, and which execution assumptions produced this run?
```

This article explains the reproducibility model behind that question.

## The Experiment Model

A ledgr run is produced from explicit inputs. The experiment fixes:

- a sealed snapshot;
- a strategy function;
- registered feature definitions;
- an opening state;
- a universe and execution options.

`ledgr_run()` then supplies strategy parameters and an immutable `run_id`. The
run derives fills, ledger events, equity, trades, metrics, and comparison tables
from those inputs. The sealed snapshot is the evidence base. The strategy
declares desired holdings at each pulse. The feature list declares which derived
values are available to the strategy. The opening state declares cash and any
starting positions.

That shape matters because each part has a different reproducibility role.
Market data are sealed. Parameters are stored and hashed. Strategy source is
captured when possible. Results are derived from ledger events rather than
remembered from an in-memory session.

```{r experiment-shape, eval=FALSE}
snapshot <- ledgr_snapshot_from_df(bars, snapshot_id = "research_snapshot")

features <- list(ledgr_ind_returns(5))

strategy <- function(ctx, params) {
  targets <- ctx$flat()

  for (id in ctx$universe) {
    ret_5 <- ctx$feature(id, "return_5")
    if (is.finite(ret_5) && ret_5 > params$min_return) {
      targets[id] <- params$qty
    }
  }

  targets
}

exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = strategy,
  features = features,
  opening = ledgr_opening(cash = 10000)
)

bt <- ledgr_run(
  exp,
  params = list(min_return = 0, qty = 10),
  run_id = "qty_10"
)
```

## The Provenance Model

For completed runs, ledgr stores run provenance alongside the result tables. The
provenance record includes the captured strategy source where available, source
hash, parameter JSON and parameter hash, dependency-version metadata, ledgr
version, R version, and reproducibility tier.

Those fields do not make every run perfectly replayable. They make the claim
inspectable. A run should be explainable later: what source text ledgr captured,
which parameters were supplied, whether the source hash still verifies, and what
reproducibility tier ledgr assigned before execution.

```{r run-info, eval=FALSE}
ledgr_run_info(snapshot, "qty_10")
```

## Extract Stored Strategy Source

`ledgr_extract_strategy()` inspects stored strategy provenance for a run. The
default is intentionally read-only:

```{r extract-safe, eval=FALSE}
stored <- ledgr_extract_strategy(snapshot, "qty_10", trust = FALSE)
stored
writeLines(stored$strategy_source_text)
```

`trust = FALSE` returns source text and metadata without parsing, evaluating, or
executing the stored source. In this mode, the source text is just data.

Use `trust = TRUE` only when you explicitly trust the experiment store and
intentionally want ledgr to parse and evaluate the stored text into a function
object.

```{r extract-trust, eval=FALSE}
trusted <- ledgr_extract_strategy(snapshot, "qty_10", trust = TRUE)
trusted$strategy_function
```

Hash verification proves stored-text identity, not code safety. A verified hash
means the stored text matches the stored hash. It does not mean the source is
safe to evaluate, economically sensible, or independent from external state.

Legacy/pre-provenance runs and strategy types without capturable source may
report `strategy_source_text = NA`. Those runs can still be inspected through
`ledgr_run_info()` and result tables, but the strategy function cannot be
recovered from provenance alone.

## Stored Source Is Not Full Reproducibility

Stored source is a strong audit artifact, but it is only one part of
reproducibility. A strategy may call external packages. It may close over data
objects. It may rely on package versions, system libraries, or runtime state
outside ledgr's database.

That is why ledgr classifies strategies before execution.

## Reproducibility Tiers

Tier 1 is self-contained under ledgr's static preflight rules. The strategy can
be understood from stored source and explicit parameters, using base/recommended
R references and ledgr's exported public namespace.

```{r tier-1, eval=FALSE}
tier_1_strategy <- function(ctx, params) {
  targets <- ctx$flat()

  for (id in ctx$universe) {
    targets[id] <- params$qty
  }

  targets
}

ledgr_strategy_preflight(tier_1_strategy)
```

Tier 2 is inspectable but needs environment management outside ledgr. Examples
include package-qualified calls outside the active R distribution and resolved
immutable non-function objects captured from the strategy environment.

```{r tier-2-package, eval=FALSE}
tier_2_strategy <- function(ctx, params) {
  jsonlite::toJSON(params, auto_unbox = TRUE)
  ctx$flat()
}

ledgr_strategy_preflight(tier_2_strategy)
```

The `jsonlite::toJSON()` call is written this way on purpose. Namespace
qualification tells ledgr which package supplies the function. That makes the
dependency visible in the preflight result and keeps the strategy inspectable.
The run can proceed, but ledgr cannot preserve the installed `jsonlite` version
or its system requirements by itself.

Resolved external scalar values are also Tier 2, not Tier 3. They are visible
to the preflight because they exist in the strategy closure, but ledgr does not
turn them into replayable run parameters. Prefer putting values that define the
research question into `params`, especially for sweeps.

Captured mutable environments are also resolved, so they remain Tier 2 under
the current policy. That does not make mutation reproducible. If an external
environment can change between runs or across workers, treat the value as an
explicit parameter or make the environment immutable before running the
experiment.

Tier 2 is allowed for ordinary runs and future sweep mode. It is not fully
reproducible by ledgr alone. Users own package installation, package version
parity, system libraries, and any other runtime environment needed by their
strategy.

Common environment-management approaches in R projects include `renv`, Docker,
`{rix}` (<https://github.com/ropensci/rix>), and `{uvr}`
(<https://github.com/nbafrank/uvr>). ledgr does not require those tools and this
article does not teach them. The point is simpler: if a strategy is Tier 2,
ledgr can preserve the run evidence, but the user must preserve the surrounding
environment.

Tier 3 is external state ledgr cannot recover or execute safely. Common
examples are unqualified helper functions from the interactive session,
wall-clock or process-environment calls such as `Sys.time()`, `Sys.Date()`, and
`Sys.getenv()`, and global assignment with `<<-`.

```{r tier-3, eval=FALSE}
my_helper <- function(ctx) ctx$flat()

tier_3_strategy <- function(ctx, params) {
  my_helper(ctx)
}

ledgr_strategy_preflight(tier_3_strategy)
```

Tier 3 strategies fail before execution. There is no `force = TRUE` override on
`ledgr_run()` or `ledgr_sweep()`; move external values into `params`, qualify
package calls, or use ledgr's exported helpers instead.

Preflight rejection is the first boundary. A covered Tier 3 strategy stops
before fold execution, before output-handler side effects, and before later
determinism hashing can become the first user-facing error. The condition class
chain includes `ledgr_strategy_tier3` and `ledgr_strategy_preflight_error`.

The most common hard rejections are:

| Pattern | Example | Why it fails |
|---|---|---|
| wall-clock access | `Sys.time()` or `do.call("Sys.time", list())` | runtime date/time is not stored run input |
| process environment | `Sys.getenv("TOKEN")` | external process state is not stored run input |
| dynamic evaluation | `get("x")`, `eval(expr)`, `assign("x", 1)` | preflight cannot recover the value path as stored metadata |
| global assignment | `x <<- 1` | strategy mutates state outside the run artifact |
| context mutation | `attr(ctx, "secret") <- 1` | strategy mutates ledgr's execution context |
| unresolved helper | `my_helper(ctx)` | helper source is not stored as part of the strategy |

Recommended-R functions such as `stats::median()` remain Tier 1-compatible
when called explicitly or resolved through R's base/recommended namespace. They
are not package dependencies outside the active R distribution.

Ambient strategy RNG calls such as `runif(1)` are a separate case. They are
allowed as Tier 2 because ledgr's execution seed contract can make the strategy
run repeatable, but they still deserve scrutiny because the random draw is a
decision input. This is different from custom-indicator RNG restrictions:
feature generation must be deterministic for a given snapshot and feature
definition.

## Why `params` Is The Boundary

Use `params` for strategy variation. Parameters are canonicalized, hashed, and
stored with the run. Hidden globals are not.

```{r params-boundary, eval=FALSE}
strategy <- function(ctx, params) {
  targets <- ctx$flat()

  for (id in ctx$universe) {
    if (ctx$close(id) > params$threshold) {
      targets[id] <- params$qty
    }
  }

  targets
}
```

The same rule matters for future sweep workers. Sweep mode can send explicit
parameter combinations to workers. It cannot reliably send an arbitrary
interactive session.

## Hidden Mutable State

Static analysis is not proof of semantic reproducibility. Patterns such as
`<<-`, mutable captured environments, dynamic dispatch, and dynamically
constructed calls can make a strategy order-dependent or worker-dependent even
when some symbols resolve.

```{r hidden-mutable-state, eval=FALSE}
counter <- 0

bad_strategy <- function(ctx, params) {
  counter <<- counter + 1
  ctx$flat()
}
```

Avoid this pattern. Store intentional strategy variation in `params`, and let
ledgr record decisions and state changes through the run artifacts.

## What To Remember

Reproducibility in ledgr is a chain:

```text
sealed snapshot -> experiment inputs -> preflight tier -> run provenance
-> ledger events -> derived results -> stored source inspection
```

Tier 1 is the cleanest path. Tier 2 is allowed but requires user-managed
environment parity. Tier 3 fails because ledgr cannot recover what the strategy
depended on.
