---
title: "Design Philosophy: From Research to Production"
format:
  html:
    toc: false
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{Design Philosophy: From Research to Production}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 90)
# Local renders load the package from source via pkgload (working dir is
# either project root or vignettes/); installed builds fall through and
# rely on the library(ledgr) call in the prerequisites chunk.
pkg_root <- if (file.exists("DESCRIPTION")) "." else if (file.exists("../DESCRIPTION")) ".." else NA_character_
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
```

ledgr is built around one design premise: strategies should use the same contract
across backtest, paper, and live modes. Not a translation of research code into
production code. Not a reimplementation. The same strategy function, the same
logic, the same event-sourced ledger model.

This article explains the arc ledgr is designed to cover, how the event-sourced
model enables it, and where v0.1.x sits on that path.

## The Arc

::: {.ledgr-diagram .ledgr-research-production-arc}

```{mermaid}
%%| echo: false

flowchart LR
  research["research"]
  paper["paper trading"]
  live["live trading"]

  research --> paper --> live
```

:::

Most backtesting libraries stop at the first arrow. The strategy exits the
research environment as a CSV of returns and is re-implemented in a production
system that has nothing to do with the backtest. The results differ. The bugs
differ. The audit trail is gone.

ledgr is designed so that the strategy that produced the backtested results uses
the same contract in production. The event-sourced ledger is what makes that
continuity possible.

## The Ledger Is The Bridge

In ledgr, results are never computed directly from price arrays. Every decision
-- a target position, a fill, a cash change -- is recorded as an immutable event.
Equity, trades, and metrics are derived from that ledger after the fact.

::: {.ledgr-diagram .ledgr-ledger-bridge}

```{mermaid}
%%| echo: false

flowchart LR
  data["data"]
  snapshot["sealed snapshot"]
  pulses["pulses"]
  ledger["event ledger"]
  results["results"]

  data --> snapshot --> pulses --> ledger --> results
```

:::

This is not just a correctness choice. It is an architectural choice that makes
the research-to-production arc coherent. Backtest and paper fills share the same
ledger event schema, so the reconstruction logic, result views, and audit trail
work identically across both modes. Live trading extends the event stream with
broker lifecycle events -- submissions, acknowledgments, partial fills,
rejections -- without changing the strategy contract. Safety gates, reconciliation,
and operational controls are adapter concerns; the strategy itself does not change.

## The Experiment Store

Before a strategy is deployed it needs to be validated -- not just against one
parameter set on one data slice, but across many combinations and market regimes,
with full provenance.

The ledgr experiment store makes this durable. A **sealed snapshot** pins the
market data permanently. A **`run_id`** is an immutable experiment key. Strategy
identity is captured from source text and parameters. Every run is auditable and
discoverable after the R session ends.

This is a concrete user-facing workflow:

```{r}
#| label: experiment-store-api
#| eval: false

snapshot <- ledgr_snapshot_open(db_path, "snapshot_id")

runs <- ledgr_run_list(snapshot)

info <- ledgr_run_info(snapshot, "sma_20_production_candidate")

bt <- ledgr_run_open(snapshot, "sma_20_production_candidate")
ledgr_results(bt, what = "equity")

snapshot <- snapshot |>
  ledgr_run_label("sma_20_production_candidate", "approved-baseline") |>
  ledgr_run_archive("discarded-parameter-test", reason = "bad regime fit")
```

`run_id` is the immutable experiment key. `label`, tags, and archive state are
mutable metadata only; they do not change the snapshot hash, strategy source
hash, strategy parameter hash, config hash, or ledger artifacts. Older runs
created before provenance capture remain inspectable as legacy/pre-provenance
runs, but they cannot be upgraded into fully reproducible experiments after the
fact.

The research workflow before deployment has two phases:

**Commit**. Full provenance run. Validate named candidates with
durable artifacts: sealed snapshot hash, strategy source hash, parameter hash,
config hash, ledgr and R version, dependency versions, compact telemetry, and
result artifacts. Use `ledgr_run_compare()` to compare named variants and
`ledgr_run_strategy()` to inspect stored strategy source.

**Explore**. Fast parameter sweep mode builds on the same experiment object and
parity contracts. Use sweep mode to evaluate parameter-grid candidates without
committing each candidate as a durable run, then promote a selected candidate
when it should become an auditable stored run.

## The Edge Device

DuckDB runs anywhere R runs, including ARM edge hardware such as a Raspberry Pi
or a small cloud VPS. A validated strategy can be deployed to an edge device with
an R instance, a DuckDB experiment store, and a broker adapter.

The device maintains its own ledger, appending live fills to the same schema the
backtest used. If the device restarts, `ledgr_state_reconstruct()` rebuilds
current positions and cash from the ledger events. No in-memory state is trusted
across restarts. The ledger reconstructs ledgr's expected state. In paper and
live modes, that expected state must still be reconciled against broker-reported
orders, positions, cash, and fills before trading resumes.

This makes the deployment target simpler than traditional production systems.
There is no separate database, no separate execution engine, no translation layer.
R, DuckDB, and a broker adapter are sufficient for systematic EOD and low-frequency
intraday strategies.

## The Strategy Contract

The sweep-to-production path works cleanly for strategies written as
self-contained `function(ctx, params)` functions with explicit, JSON-safe
parameters and no hidden mutable state:

```r
sma_strategy <- function(ctx, params) {
  targets <- ctx$flat()
  for (id in ctx$universe) {
    values <- c(sma = ctx$feature(id, paste0("ttr_sma_", params$window)))
    if (ledgr_passed_warmup(values) && ctx$close(id) > values[["sma"]]) {
      targets[id] <- params$quantity
    }
  }
  targets
}
```

This is Tier 1 reproducibility: the strategy is fully self-contained, its
parameters are hashable, and its source is capturable. Tier 1 strategies earn
full experiment-store identity -- source hash, parameter hash, provenance metadata
-- and are the natural fit for sweep mode and edge deployment.

ledgr supports less constrained strategies too, but the reproducibility tier is
always visible in run provenance. The trust boundary is explicit, not hidden.

## Cost And Timing Are Explicit

Production-shaped research needs execution assumptions that are visible at the
run boundary. ledgr makes timing and transaction costs explicit parts of
experiment construction:

```r
experiment <- ledgr_experiment(
  snapshot,
  strategy = sma_strategy,
  params = list(window = 20, quantity = 10),
  timing_model = ledgr_timing_next_open(),
  cost_model = ledgr_cost_chain(
    ledgr_cost_spread_bps(5),
    ledgr_cost_fixed_fee(1)
  )
)
```

Use `ledgr_cost_zero()` when a zero-cost baseline is intentional. That choice is
still recorded as a cost model, with its own `cost_model_hash` and
`cost_plan_json`, so a no-cost run is not confused with an omitted-cost run.

`ledgr_cost_spread_bps()` uses a quoted-spread convention: the configured basis
points describe the full quoted spread, and ledgr applies half of that spread to
each side of the trade. A buy pays above the execution-bar open; a sell receives
below it. Fixed and notional fees add explicit costs without changing quantity,
side, instrument, or execution timestamp.

## What v0.1.x Delivers Today

v0.1.x is the correctness-first research layer. It already covers:

- sealed snapshots, hash verification, and deterministic replay across machines
  and R sessions;
- project-local DuckDB stores with run discovery, labels, tags, archival,
  comparison, reopening, and strategy-source inspection;
- deterministic pulse execution with no-lookahead `ctx`, full target holdings,
  next-open fills, final-bar no-fill warnings, and an append-only ledger;
- accounting surfaces for ledger events, fills, trades, equity rows, summary
  metrics, and explicit metric contexts;
- built-in indicators, TTR-backed indicators, multi-output bundles, feature
  maps, warmup diagnostics, pulse inspection, and active aliases;
- feature and strategy grids, sweep execution, candidate rows, compact saved
  sweeps, retained return series, promotion context, and explicit
  selection-is-not-validation framing;
- public cost-model constructors, timing-model identity, required explicit
  costs, classed target-risk transforms with risk-chain identity,
  reproducibility tiers, strategy preflight, stored strategy source, and a
  deterministic demo dataset for documentation and examples.

The current research layer also includes the first walk-forward surface:

- walk-forward evaluation runs over the existing sweep and run surfaces,
  consuming cost identity, saved-sweep retention infrastructure, and risk-chain
  identity;
- the next planned validation-toolkit work is scoped for v0.1.9.6 and may add
  DSR, PBO/CSCV over retained return panels, and deterministic candidate
  clustering;
- crypto-readiness evidence and target-construction helper extensions remain
  separate future packets.

The target-risk layer is intentionally narrow: it transforms target quantities
before timing and cost. It does not implement affordability enforcement,
liquidity/capacity policy, margin, shorting or borrow policy, OMS lifecycle
behavior, or broker-grade controls.

Paper trading adapters are planned for v0.3.0, observability tooling for
v0.4.0, and small-scale live trading for v1.0.0. OMS state machine semantics
remain earlier prerequisite work before paper or live adapters.

The path from a validated experiment-store entry to a running edge device is
shorter than it looks. The research work done in v0.1.x is not throwaway
scaffolding -- it is the foundation the production system builds on.

## Where Next

- `vignette("research-workflow", package = "ledgr")` shows the current
  project-local research loop.
- `vignette("walk-forward", package = "ledgr")` shows held-out evaluation over
  sweep and run surfaces.
- `vignette("risk-and-cost", package = "ledgr")` explains the target-risk,
  timing, cost, liquidity, and OMS boundaries.
- `vignette("reproducibility", package = "ledgr")` covers strategy source,
  tiers, and trust boundaries.
