---
title: "Design Philosophy: From Research to Production"
format:
  html:
    toc: false
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{Design Philosophy: From Research to Production}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 90)
# Local renders load the package from source via pkgload (working dir is
# either project root or vignettes/); installed builds fall through and
# rely on the library(ledgr) call in the prerequisites chunk.
pkg_root <- if (file.exists("DESCRIPTION")) "." else if (file.exists("../DESCRIPTION")) ".." else NA_character_
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
```

ledgr currently ships a research engine. Its durable experiments, pulse
contract, and event-derived accounting are designed so later paper and live
systems do not have to discard the research model. Those later systems have not
shipped.

This article starts with the research workflow you can use now, then shows the
longer arc the architecture is intended to support. A design direction is not a
claim that paper or live execution already exists.

## The Arc

::: {.ledgr-diagram .ledgr-research-production-arc}

```{mermaid}
%%| echo: false

flowchart LR
  research["shipped research"]
  paper["planned paper trading"]
  live["planned live trading"]

  research --> paper --> live
```

:::

The first box is the product today. ledgr seals inputs, runs research strategies,
stores accounting evidence, and reopens results. The arrows are an architectural
goal: future execution modes should reuse a compatible strategy and accounting
boundary instead of treating a CSV of returns as the production handoff.

## The Ledger Is The Bridge

In ledgr, fills and other accounting effects are appended to a ledger. Equity,
trades, and metrics are derived from that evidence. A target vector is a pulse
decision, not itself a promise that every intermediate choice becomes an
immutable ledger row.

::: {.ledgr-diagram .ledgr-ledger-bridge}

```{mermaid}
%%| echo: false

flowchart LR
  data["data"]
  snapshot["sealed snapshot"]
  pulses["pulses"]
  ledger["event ledger"]
  results["results"]

  data --> snapshot --> pulses --> ledger --> results
```

:::

This is both a correctness choice and a useful future boundary. A paper or live
system would still need broker order state, partial fills, rejection handling,
reconciliation, safety gates, recovery, and observability. Those are not
adapters that can be assumed away, and ledgr does not ship them today.

## The Experiment Store

Before a strategy is trusted, its research evidence should be inspectable -- not
just one remembered number from one R session.

The ledgr experiment store makes this durable. A **sealed snapshot** pins the
market data permanently. A **`run_id`** is an immutable experiment key. Strategy
identity is captured from source text and parameters. Every run is auditable and
discoverable after the R session ends.

This is a concrete user-facing workflow:

```{r}
#| label: experiment-store-api
#| eval: false

snapshot <- ledgr_snapshot_open(db_path, "snapshot_id")

runs <- ledgr_run_list(snapshot)

info <- ledgr_run_info(snapshot, "sma_20_production_candidate")

bt <- ledgr_run_open(snapshot, "sma_20_production_candidate")
ledgr_results(bt, what = "equity")

snapshot <- snapshot |>
  ledgr_run_label("sma_20_production_candidate", "approved-baseline") |>
  ledgr_run_archive("discarded-parameter-test", reason = "bad regime fit")
```

`run_id` is the immutable experiment key. `label`, tags, and archive state are
mutable metadata only; they do not change the snapshot hash, strategy source
hash, strategy parameter hash, config hash, or ledger artifacts. Older runs
created before provenance capture remain inspectable as legacy/pre-provenance
runs, but they cannot be upgraded into fully reproducible experiments after the
fact.

The current research workflow has two phases:

**Commit**. Full provenance run. Record named candidates with
durable artifacts: sealed snapshot hash, strategy source hash, parameter hash,
config hash, ledgr and R version, dependency versions, compact telemetry, and
result artifacts. Use `ledgr_run_compare()` to compare named variants and
`ledgr_run_strategy()` to inspect stored strategy source.

**Explore**. Fast parameter sweep mode builds on the same experiment object and
parity contracts. Use sweep mode to evaluate parameter-grid candidates without
committing each candidate as a durable run, then promote a selected candidate
when it should become an auditable stored run.

## The Planned Operational Boundary

R and DuckDB can run on small servers and edge hardware, but portability of the
research store is not a trading system. A future paper or live mode would need
broker order state, partial fills, rejection handling, reconciliation, safety
gates, recovery, and observability. The ledger and reconstruction model are
useful foundations for that work; they are not substitutes for it.

## The Strategy Contract

The research-and-promotion path works cleanly for strategies written as
self-contained `function(ctx, params)` functions with explicit, JSON-safe
parameters and no hidden mutable state:

```r
sma_strategy <- function(ctx, params) {
  sma <- ctx$vec$feature("sma_20")
  targets <- ctx$flat()
  targets[which(is.finite(sma) & ctx$vec$close > sma)] <- params$quantity
  targets
}
```

This example assumes that the experiment declares `ledgr_ind_sma(20)`, whose
concrete feature ID is `sma_20`. A whole-universe read makes that fixed
dependency visible instead of rebuilding an ID from strategy parameters.

This is Tier 1 reproducibility: the strategy is fully self-contained, its
parameters are hashable, and its source is capturable. Tier 1 strategies earn
full experiment-store identity -- source hash, parameter hash, and provenance
metadata -- and are the natural fit for sweep mode and portable research
projects.

ledgr supports less constrained strategies too, but the reproducibility tier is
always visible in run provenance. The trust boundary is explicit, not hidden.

## Cost And Timing Are Explicit

Production-shaped research needs execution assumptions that are visible at the
run boundary. ledgr makes timing and transaction costs explicit parts of
experiment construction:

```r
experiment <- ledgr_experiment(
  snapshot,
  strategy = sma_strategy,
  features = list(ledgr_ind_sma(20)),
  timing_model = ledgr_timing_next_open(),
  cost_model = ledgr_cost_chain(
    ledgr_cost_spread_bps(5),
    ledgr_cost_fixed_fee(1)
  )
)

run <- ledgr_run(
  experiment,
  params = list(quantity = 10)
)
```

Use `ledgr_cost_zero()` when a zero-cost baseline is intentional. That choice is
still recorded as a cost model, with its own `cost_model_hash` and
`cost_plan_json`, so a no-cost run is not confused with an omitted-cost run.

`ledgr_cost_spread_bps()` uses a quoted-spread convention: the configured basis
points describe the full quoted spread, and ledgr applies half of that spread to
each side of the trade. A buy pays above the execution-bar open; a sell receives
below it. Fixed and notional fees add explicit costs without changing quantity,
side, instrument, or execution timestamp.

## What The Current Research Layer Delivers

The v0.2.1.0 release line is a correctness-first research layer. It covers:

- sealed snapshots, hash verification, and deterministic replay across machines
  and R sessions;
- project-local DuckDB stores with run discovery, labels, tags, archival,
  comparison, reopening, and strategy-source inspection;
- deterministic pulse execution with no-lookahead `ctx`, full target holdings,
  next-open fills, final-bar no-fill warnings, and an append-only ledger;
- accounting surfaces for ledger events, fills, trades, equity rows, summary
  metrics, and explicit metric contexts;
- built-in indicators, TTR-backed indicators, multi-output bundles, feature
  maps, warmup diagnostics, pulse inspection, and active aliases;
- feature and strategy grids, sweep execution, candidate rows, compact saved
  sweeps, retained return series, promotion context, and explicit
  selection-is-not-validation framing;
- public cost-model constructors, timing-model identity, explicit costs,
  target-risk transforms with risk-chain identity,
  reproducibility tiers, strategy preflight, stored strategy source, and a
  deterministic demo dataset for documentation and examples.

It also includes walk-forward and selection-integrity surfaces:

- walk-forward evaluation runs over the existing sweep and run surfaces,
  consuming cost identity, saved-sweep retention infrastructure, and risk-chain
  identity;
- DSR, PBO/CSCV, MinTRL, and deterministic effective-trial clustering operate
  over retained return panels;
- point-in-time inputs, availability-aware universes, missing-session rules,
  cash distributions, and bounded terminal settlement make stronger evidence
  possible when the research claim requires it.

The target-risk layer is intentionally narrow: it transforms target quantities
before timing and cost. It does not implement affordability enforcement,
liquidity/capacity policy, margin, shorting or borrow policy, OMS lifecycle
behavior, or broker-grade controls.

Paper/live execution, broker adapters, and operational observability remain
roadmap work. The research layer is intended to be reusable foundation, but no
release date or operational capability is implied by this article.

## Where Next

- `vignette("research-workflow", package = "ledgr")` shows the current
  project-local research loop.
- `vignette("walk-forward", package = "ledgr")` shows held-out evaluation over
  sweep and run surfaces.
- `vignette("risk-and-cost", package = "ledgr")` explains the target-risk,
  timing, cost, liquidity, and OMS boundaries.
- `vignette("reproducibility", package = "ledgr")` covers strategy source,
  tiers, and trust boundaries.
