---
title: "Research Workflow"
format:
  html:
    toc: false
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{Research Workflow}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

options(width = 90)
# Local renders load the package from source via pkgload (working dir is
# either project root or vignettes/); installed builds fall through and
# rely on the library(ledgr) call in the prerequisites chunk.
pkg_root <- if (file.exists("DESCRIPTION")) "." else if (file.exists("../DESCRIPTION")) ".." else NA_character_
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
```

```{=html}
<style>
.ledgr-diagram {
  margin: 1.25rem auto 1.5rem auto;
  text-align: center;
}

.ledgr-diagram .mermaid {
  display: inline-block;
  max-width: 100%;
}

.ledgr-diagram .mermaid svg {
  display: block;
  height: auto !important;
  margin-left: auto;
  margin-right: auto;
}

.ledgr-workflow-diagram .mermaid svg {
  max-width: 560px !important;
}

.ledgr-validation-diagram .mermaid svg {
  max-width: 520px !important;
}
</style>
```

You have an idea for a trading rule. Maybe a fast moving average crossing a
slow one looks useful. How do you turn that hunch into evidence you can
reopen, inspect, and explain later?

That is the job of this article. You will create a sealed data snapshot,
declare an experiment, run the strategy once, explore a small grid, inspect the
candidate table, promote one candidate with a note, and reopen the stored
artifact.

This workflow produces research evidence, not investment advice or a guarantee
that a strategy will work in the future. See the project
[disclaimer](../DISCLAIMER.md).

The loop is deliberately short:

::: {.ledgr-diagram .ledgr-workflow-diagram}

```{mermaid}
%%| echo: false
%%{init: {"theme": "base", "flowchart": {"nodeSpacing": 18, "rankSpacing": 22, "curve": "linear"}, "themeVariables": {"fontFamily": "system-ui, -apple-system, Segoe UI, sans-serif", "fontSize": "15px", "primaryColor": "#f8fafc", "primaryTextColor": "#1f2937", "primaryBorderColor": "#64748b", "lineColor": "#64748b", "tertiaryColor": "#eef2ff", "tertiaryTextColor": "#1f2937", "tertiaryBorderColor": "#64748b", "clusterBkg": "#ffffff", "clusterBorder": "#cbd5e1"}}}%%

flowchart TB
  data["Seal data"]
  experiment["Declare experiment"]
  single["Run once"]
  sweep["Sweep candidates"]
  inspect["Inspect evidence"]
  promote["Promote with note"]
  reopen["Reopen and recover"]

  data --> experiment --> single --> sweep --> inspect --> promote --> reopen
  inspect -. new research iteration .-> experiment
```

:::

Each step removes one common ambiguity:

- sealing fixes the evidence;
- experiment declaration fixes the strategy boundary;
- a single run checks that the setup behaves at all;
- sweeping compares declared candidates;
- inspection makes the selection rule visible;
- promotion records the human choice;
- reopening proves the result is a durable artifact.

The goal is not to make research slower. The goal is to make later
explanations possible.

After promotion, ledgr can show you the run metadata, the source sweep context,
the selected candidate, the strategy parameters, the strategy source metadata,
and the promotion note. Recoverability has limits, especially for Tier 2
strategies that depend on external functions or package state, but ledgr
records enough provenance to explain what caused a promoted result within the
declared reproducibility tier.

Promotion records selection; it does not prove generalization. That caveat is
part of the workflow, not a footnote at the end.

## Prerequisites

This walkthrough uses `ledgr` plus `dplyr` for compact data manipulation.
Strategies themselves do not depend on `dplyr`; inside a run they only read
from the ledgr pulse context.

```{r}
#| label: prerequisites
#| message: false

library(ledgr)
library(dplyr)

data("ledgr_demo_bars", package = "ledgr")
```

::: {.ledgr-callout .ledgr-callout-note}
**Running this yourself**

This article is evaluated when it is rendered. To keep package builds and local
previews disposable, the code writes to a temporary DuckDB store. In a real
project, replace `store_path` with a durable path such as
`artifacts/ledgr_store.duckdb`.
:::

::: {.ledgr-callout .ledgr-callout-note}
**About the demo data**

`DEMO_01` and `DEMO_02` are package-owned demo instruments. Real research
should use a sealed snapshot of the market data you intend to study. The point
here is the shape of the workflow, not the realism of the data.
:::

## Project Topology

Start with a small project shape. One script creates the evidence, one store
holds the artifacts, and one report explains the research choice.

```text
my-ledgr-project/
  artifacts/
    ledgr_store.duckdb
  reports/
    workflow_review.md
  scripts/
    research_workflow.R
```

ledgr does not require this exact layout. The point is simpler: keep the
script, durable store, and written reasoning close enough that a future review
can recover the full research decision.

In a real project, add generated stores such as `artifacts/*.duckdb` to
`.gitignore` unless you have a deliberate artifact-versioning policy. The
script and source data should explain how to recreate the store; Git should not
quietly become the database backup.

In this rendered article, use a temporary store so repeated builds do not leave
project artifacts behind. In a real project, make this a boring durable path:
future you should know where the evidence lives.

```{r}
#| label: topology

store_path <- file.path(tempdir(), "ledgr_research_workflow.duckdb")
if (file.exists(store_path)) {
  unlink(store_path)
}
```

## Fix The Evidence: Seal A Snapshot

First, choose the data window. Here you use two demo instruments over the first
half of 2019.

```{r}
#| label: snapshot-data

bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% c("DEMO_01", "DEMO_02"),
    between(
      ts_utc,
      ledgr_utc("2019-01-01"),
      ledgr_utc("2019-06-30")
    )
  )

bars |>
  slice_head(n = 4)
```

You should see a small bar table. The exact print width depends on your
console, but the important columns are the timestamp, instrument, OHLC prices,
and volume.

Now seal those bars into the project store.

```{r}
#| label: snapshot

snapshot <- ledgr_snapshot_from_df(
  bars,
  db_path = store_path,
  snapshot_id = "demo_2019_h1"
)
```

Once the data is sealed, the snapshot becomes the identity of the evidence.
Every run below refers back to that same immutable input. If you change the
evidence, create a new snapshot instead of editing this one. That is what makes
later comparison and reopening meaningful.

## Declare The Experiment Boundary

The strategy should be able to ask for `fast` and `slow`, not for long
engine-generated feature IDs. The aliases are the names the strategy will read
at each pulse. The underlying indicator declarations still remain part of the
hashed experiment configuration.

```{r}
#| label: feature-map

features <- ledgr_feature_map(
  fast = ledgr_ind_sma(ledgr_param("fast_n")),
  slow = ledgr_ind_sma(ledgr_param("slow_n"))
)
```

This is the first place where the two parameter namespaces matter:
`feature_params` will choose `fast_n` and `slow_n`; strategy `params` will
choose values such as `qty` and `threshold`.

## Choose The Strategy

Use the demo SMA crossover strategy for the workflow. It expects active aliases
named `fast` and `slow`, then returns target holdings.

```{r}
#| label: strategy-demo

strategy <- ledgr_demo_sma_crossover_strategy()
```

You do not need to write a strategy for this vignette. The following miniature
example only shows the boundary that the demo strategy follows: read
pulse-known values from `ctx`, guard warmup with `passed_warmup()`, and return
a full named numeric target vector.

```{r}
#| label: custom-strategy-bridge

custom_sma_strategy <- function(ctx, params) {
  target <- ctx$flat()

  for (instrument_id in ctx$universe) {
    values <- ctx$features(instrument_id)

    # TRUE once enough bars have passed to calculate the slow SMA.
    if (passed_warmup(values) &&
        values[["fast"]] / values[["slow"]] - 1 > params$threshold) {
      target[[instrument_id]] <- params$qty
    }
  }

  target
}
```

For full strategy-authoring patterns, use
`vignette("strategy-development", package = "ledgr")`. This article stays on
the research workflow. While the example above shows the internal mechanics,
the rest of this workflow uses the built-in strategy object created earlier.

## Sanity-Check One Run

Before you fan out into a sweep, run one parameter set. You are checking two
basic things: did anything trade, and do the derived results look plausible?
Sweeps amplify whatever your setup does, including doing nothing.

At this stage, the exact performance number matters less than the existence
and plausibility of fills, positions, and equity.

```{r}
#| label: experiment-run

exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = strategy,
  features = features,
  opening = ledgr_opening(cash = 10000)
)

single_run <- ledgr_run(
  exp,
  params = list(qty = 10, threshold = 0),
  feature_params = list(fast_n = 10L, slow_n = 40L),
  run_id = "workflow_single_run",
  seed = 2026
)

summary(single_run)
ledgr_results(single_run, what = "trades")
head(ledgr_results(single_run, what = "equity"), 3)
```

The first run should give you a completed backtest object, a trade table, and
an equity curve. You are not looking for the best result yet; you are checking
that the workflow produces evidence.

If this run has no fills, odd position changes, or implausible equity, stop
here and debug the strategy or feature declarations before sweeping.

## Compare Declared Candidates

Now expand from one run to a small grid. Keep feature parameters and strategy
parameters separate, then combine them into one executable grid.

```{r}
#| label: sweep

feature_grid <- ledgr_feature_grid(
  fast_n = c(5L, 10L),
  slow_n = c(20L, 40L),
  .filter = fast_n < slow_n
)

strategy_grid <- ledgr_strategy_grid(
  qty = c(5, 10),
  threshold = c(0, 0.01)
)

grid <- ledgr_grid_cross(features = feature_grid, strategy = strategy_grid)
precomputed <- ledgr_precompute_features(exp, grid)

sweep <- ledgr_sweep(
  exp,
  grid,
  precomputed_features = precomputed,
  seed = 2026
)
```

Precomputing features is not a separate research decision. It is an execution
optimization: the same declared feature grid is computed once and reused across
candidates.

`ledgr_sweep()` is the memory-backed exploration mode. It evaluates candidate
rows through the same fold semantics as committed runs, but it keeps compact
candidate evidence instead of writing a durable ledger and equity curve for
every row. That is why sweeps can be credible at serious research scale without
turning every exploratory candidate into a permanent artifact. Promotion is the
point where one selected candidate pays the durable-materialization cost.

`ledgr_sweep()` gives you candidate evidence. It does not choose a winner for
you. That choice belongs in the next step, where you inspect the table and make
the ranking rule visible.

::: {.ledgr-callout .ledgr-callout-tip}
**Try it**

Add a third value to `fast_n` and `slow_n`. How many candidates does the
`.filter` keep? How many would exist without the filter?
:::

## Inspect Before You Promote

Now you have a table of candidates. Before you pick one, look at both the
successful rows and the rows that did not finish. You are not validating yet;
you are learning the shape of the evidence.

For this walkthrough, use a deliberately simple ranking rule: among completed
candidates, sort by Sharpe ratio descending. In real research, make that rule
visible before treating the top row as meaningful.

```{r}
#| label: candidate-inspection

ranked <- sweep |>
  filter(status == "DONE") |>
  arrange(desc(sharpe_ratio))

candidate_columns <- c(
  "run_id", "status", "final_equity", "total_return",
  "sharpe_ratio", "params", "feature_params"
)

top_n <- ranked |>
  slice_head(n = 5) |>
  select(all_of(candidate_columns))
glimpse(top_n)

issue_columns <- c("run_id", "status", "error_class", "error_msg", "warnings")

issues <- sweep |>
  filter(status != "DONE") |>
  select(any_of(issue_columns)) |>
  as_tibble()
issues

candidate <- ledgr_candidate(ranked, 1)
```

The output should look like a candidate table rather than a final answer.

The `top_n` table shows which candidates would be selected by this ranking
choice. The `issues` table tells you whether any candidates failed, warned,
or produced diagnostics that should change how you read the sweep.

::: {.ledgr-callout .ledgr-callout-note}
**Design note**

This explicit table code keeps the selection rule visible. A future
sweep-review helper may package this review shape, but it should preserve the
same explicit selection rule instead of making ranking automatic.
:::

::: {.ledgr-callout .ledgr-callout-tip}
**Try it**

Sort by `total_return` instead of `sharpe_ratio`. Does the first candidate
change? If it does, your "best" candidate depends on the metric, not only on
the strategy.
:::

## Commit The Selection With A Note

Promotion commits one selected candidate as a named run and attaches the human
selection note.

```{r}
#| label: promotion

promoted <- ledgr_promote(
  exp,
  candidate,
  run_id = "workflow_promoted_candidate",
  note = paste(
    "Promoted from an exploratory same-snapshot sweep for workflow review.",
    "This note records the selection rationale; it is not statistical validation."
  )
)

summary(promoted)
```

The promoted run is now a committed run with its own run ID. The selection note
travels with the run.

Same-snapshot promotion is useful for audit, debugging, and durable storage of
the chosen candidate. It remains in-sample.

## Reopen The Artifact

The payoff for the project store is that you can come back later without
re-running the strategy.

```{r}
#| label: reopen

close(single_run)
close(promoted)
ledgr_snapshot_close(snapshot)

snapshot <- ledgr_snapshot_load(
  store_path,
  snapshot_id = "demo_2019_h1",
  verify = TRUE
)

ledgr_run_list(snapshot)

info <- ledgr_run_info(snapshot, "workflow_promoted_candidate")
info

reopened <- ledgr_run_open(snapshot, "workflow_promoted_candidate")
summary(reopened)
```

Reopening turns the workflow from a temporary R session into a durable research
artifact.

You can also recover the selection context and strategy provenance behind the
promoted run. Today that recovery uses two public surfaces:

- `ledgr_run_info()` and `ledgr_run_open()` for the stored run;
- `ledgr_extract_strategy()` for strategy source and parameter provenance.

::: {.ledgr-callout .ledgr-callout-warning}
**API gap**

The next few lines are intentionally lower-level. They show what ledgr records
for a promoted run. A future review helper may summarize this "what caused this
result?" record without asking users to inspect nested promotion-context fields
directly.

This API gap was carried forward from the v0.1.8.6 cycle: promotion can recover
the chosen candidate and provenance today, while higher-level review helpers
remain future workflow polish.
:::

```{r}
#| label: recover-provenance

promotion <- info$promotion_context
list(
  source = promotion$source,
  selected_candidate = promotion$selected_candidate$run_id,
  strategy_params_json = promotion$selected_candidate$params_json,
  feature_params_json = promotion$selected_candidate$feature_params_json
)

extracted <- ledgr_extract_strategy(
  snapshot,
  "workflow_promoted_candidate",
  trust = FALSE
)

extracted$strategy_params
extracted$reproducibility_level
extracted$hash_verified
```

The result is compact but load-bearing: ledgr stores which sweep candidate was
selected, which strategy parameters and feature parameters produced it, what
note justified the promotion, and which strategy source metadata was captured.

::: {.ledgr-callout .ledgr-callout-note}
**What recovery means**

Recovery is provenance, not magic. Tier 1 strategy source can usually be
inspected, hash-checked, and optionally evaluated with `trust = TRUE`. Tier 2
strategies may depend on external functions or package state, so ledgr records
the source text, hashes, parameters, dependency metadata, and warnings that
explain what was captured and what remains outside the run artifact.
:::

```{r}
#| label: cleanup
#| include: false

close(reopened)
ledgr_snapshot_close(snapshot)
```

## What Promotion Does Not Prove

You now have a promoted candidate. What does it mean?

It means you can reopen the exact evidence trail: the sealed data, feature
declarations, parameter values, run identity, candidate row, and promotion
note. It does not mean the strategy will generalize.

A single-window sweep is exploratory evidence with an audit trail. Naive
sweep-and-pick selection is a selection-bias risk because every candidate was
compared on the same evidence window. Picking the highest metric from that
window can overfit the sample even when every individual run was deterministic
and leakage-safe at the pulse boundary.

The more candidates you try, the more opportunity you create for
sample-specific luck to look like skill.

::: {.ledgr-diagram .ledgr-validation-diagram}

```{mermaid}
%%| echo: false
%%{init: {"theme": "base", "flowchart": {"nodeSpacing": 18, "rankSpacing": 22, "curve": "linear"}, "themeVariables": {"fontFamily": "system-ui, -apple-system, Segoe UI, sans-serif", "fontSize": "15px", "primaryColor": "#f8fafc", "primaryTextColor": "#1f2937", "primaryBorderColor": "#64748b", "lineColor": "#64748b", "secondaryColor": "#eef2ff", "secondaryTextColor": "#1f2937", "secondaryBorderColor": "#64748b", "tertiaryColor": "#fff7ed", "tertiaryTextColor": "#1f2937", "tertiaryBorderColor": "#fb923c"}}}%%

flowchart TB
  window["Same evidence window<br/>Candidate 1<br/>Candidate 2<br/>Candidate 3<br/>Candidate ..."]
  pick["Pick highest metric"]
  caveat["Selection recorded<br/>not validated"]

  window --> pick --> caveat
```

:::

Use the promotion note to record the human choice. Treat walk-forward and
out-of-sample evaluation as the next conceptual layer when the question becomes
selection quality rather than artifact reproducibility.

::: {.ledgr-callout .ledgr-callout-tip}
**Try it**

Write one sentence explaining why you promoted the candidate and one sentence
explaining why that promotion is not validation. If the second sentence feels
hard to write, the selection rule probably needs more work.
:::

## Write The Human Research Note

Do not leave the reasoning only in your head. Write a short report next to the
stored artifacts. A compact report should include:

- hypothesis and data window
- snapshot hash and data-source assumptions
- feature and strategy declarations
- candidate grid summary
- candidate ranking rule
- top-N candidate table
- issue and failure review
- equity and drawdown plots
- promotion note
- reason for rejecting alternatives
- selection caveat: promoted candidate is not statistically validated by promotion itself

Here is the shape of a useful entry:

```text
Hypothesis and data window:
  SMA crossover candidates may capture persistent moves in DEMO_01 and DEMO_02
  over the 2019 H1 demo window.

Promotion note:
  Promoted the top Sharpe candidate after checking that all candidate rows
  completed and no issue rows changed the interpretation.

Selection caveat:
  This is a same-window exploratory selection. Promotion records the chosen
  candidate and its provenance; it does not prove out-of-sample performance.
```

The report is where the human reasoning lives. ledgr records what happened,
which inputs were used, and which candidate was promoted; it does not certify
that the selection protocol was statistically sound.

## Next Layer: Walk-Forward Evaluation

The workflow above teaches the single-window foundation. Walk-forward and
out-of-sample evaluation are the planned next conceptual layer for separating
candidate selection from held-out evidence.

When you ask "is this candidate's evidence reproducible?", the workflow above
is the right starting point. When you ask "does this strategy generalize?",
walk-forward is the next question you want. That layer is not part of this
article; the public roadmap places walk-forward evaluation at v0.1.9.x.

This article should not imply that promotion alone validates a strategy.

## Where Next

- For full strategy-authoring patterns, use
  `vignette("strategy-development", package = "ledgr")`.
- For feature maps, indicator identity, and active aliases, use
  `vignette("indicators", package = "ledgr")`.
- For sweep mechanics, failure rows, seeds, and promotion context, use
  `vignette("sweeps", package = "ledgr")`.
- For no-lookahead pulse timing, next-open fills, and final-bar warnings, use
  `vignette("execution-semantics", package = "ledgr")`.
- For durable stores, run discovery, archive state, and reopening, use
  `vignette("experiment-store", package = "ledgr")`.
