---
title: "Design Philosophy: From Research to Production"
output:
  github_document:
    toc: false
    html_preview: false
    md_extensions: -smart
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 90)
```

ledgr is built around one design premise: strategies should use the same contract
across backtest, paper, and live modes. Not a translation of research code into
production code. Not a reimplementation. The same strategy function, the same
logic, the same event-sourced ledger model.

This article explains the arc ledgr is designed to cover, how the event-sourced
model enables it, and where v0.1.x sits on that path.

## The Arc

```text
research  ->  paper trading  ->  live trading
```

Most backtesting libraries stop at the first arrow. The strategy exits the
research environment as a CSV of returns and is re-implemented in a production
system that has nothing to do with the backtest. The results differ. The bugs
differ. The audit trail is gone.

ledgr is designed so that the strategy that produced the backtested results uses
the same contract in production. The event-sourced ledger is what makes that
continuity possible.

## The Ledger Is The Bridge

In ledgr, results are never computed directly from price arrays. Every decision
-- a target position, a fill, a cash change -- is recorded as an immutable event.
Equity, trades, and metrics are derived from that ledger after the fact.

```text
data -> sealed snapshot -> pulses -> event ledger -> results
```

This is not just a correctness choice. It is an architectural choice that makes
the research-to-production arc coherent. Backtest and paper fills share the same
ledger event schema, so the reconstruction logic, result views, and audit trail
work identically across both modes. Live trading extends the event stream with
broker lifecycle events -- submissions, acknowledgments, partial fills,
rejections -- without changing the strategy contract. Safety gates, reconciliation,
and operational controls are adapter concerns; the strategy itself does not change.

## The Experiment Store

Before a strategy is deployed it needs to be validated -- not just against one
parameter set on one data slice, but across many combinations and market regimes,
with full provenance.

The ledgr experiment store makes this durable. A **sealed snapshot** pins the
market data permanently. A **`run_id`** is an immutable experiment key. Strategy
identity is captured from source text and parameters. Every run is auditable and
discoverable after the R session ends.

In v0.1.7 this is a concrete user-facing workflow:

```{r experiment-store-api, eval=FALSE}
snapshot <- ledgr_snapshot_load(db_path, "snapshot_id")

runs <- ledgr_run_list(snapshot)

info <- ledgr_run_info(snapshot, "sma_20_production_candidate")

bt <- ledgr_run_open(snapshot, "sma_20_production_candidate")
ledgr_results(bt, what = "equity")

snapshot <- snapshot |>
  ledgr_run_label("sma_20_production_candidate", "approved-baseline") |>
  ledgr_run_archive("discarded-parameter-test", reason = "bad regime fit")
```

`run_id` is the immutable experiment key. `label`, tags, and archive state are
mutable metadata only; they do not change the snapshot hash, strategy source
hash, strategy parameter hash, config hash, or ledger artifacts. Older runs
created before provenance capture remain inspectable as legacy/pre-provenance
runs, but they cannot be upgraded into fully reproducible experiments after the
fact.

The research workflow before deployment has two phases:

**Commit** (v0.1.7). Full provenance run. Validate named candidates with
durable artifacts: sealed snapshot hash, strategy source hash, parameter hash,
config hash, ledgr and R version, dependency versions, compact telemetry, and
result artifacts. Use `ledgr_compare_runs()` to compare named variants and
`ledgr_extract_strategy()` to inspect stored strategy source.

**Explore** (future). Fast parameter sweep mode is planned for a later release.
It will build on the same experiment object and parity contracts, but it is not
part of v0.1.7.

## The Edge Device

DuckDB runs anywhere R runs, including ARM edge hardware such as a Raspberry Pi
or a small cloud VPS. A validated strategy can be deployed to an edge device with
an R instance, a DuckDB experiment store, and a broker adapter.

The device maintains its own ledger, appending live fills to the same schema the
backtest used. If the device restarts, `ledgr_state_reconstruct()` rebuilds
current positions and cash from the ledger events. No state is held in memory
across restarts. No reconciliation step is needed. The ledger is the state.

This makes the deployment target simpler than traditional production systems.
There is no separate database, no separate execution engine, no translation layer.
R, DuckDB, and a broker adapter are sufficient for systematic EOD and low-frequency
intraday strategies.

## The Strategy Contract

The sweep-to-production path works cleanly for strategies written as
self-contained `function(ctx, params)` functions with explicit, JSON-safe
parameters and no hidden mutable state:

```r
sma_strategy <- function(ctx, params) {
  targets <- ctx$flat()
  for (id in ctx$universe) {
    sma <- ctx$feature(id, paste0("ttr_sma_", params$window))
    if (!is.na(sma) && ctx$close(id) > sma) {
      targets[id] <- params$quantity
    }
  }
  targets
}
```

This is Tier 1 reproducibility: the strategy is fully self-contained, its
parameters are hashable, and its source is capturable. Tier 1 strategies earn
full experiment-store identity -- source hash, parameter hash, provenance metadata
-- and are the natural fit for sweep mode and edge deployment.

ledgr supports less constrained strategies too, but the reproducibility tier is
always visible in run provenance. The trust boundary is explicit, not hidden.

## What v0.1.x Delivers Today

v0.1.x is the research layer:

- sealed snapshots and deterministic replay across machines and R sessions;
- reproducible backtests with next-open fill semantics and an append-only ledger;
- a TTR indicator adapter with deterministic warmup, vectorized precomputation,
  and session-scoped feature caching;
- an experiment store with run discovery, provenance, labeling, and archival
  (v0.1.5+);
- experiment-first execution, run comparison, strategy-source inspection, and a
  deterministic demo dataset (v0.1.7).

Paper and live trading adapters, OMS state machine semantics, and observability
tooling follow in the v0.2.x and v0.3.x range.

The path from a validated experiment-store entry to a running edge device is
shorter than it looks. The research work done in v0.1.x is not throwaway
scaffolding -- it is the foundation the production system builds on.
