---
title: "Research Workflow"
format:
  html:
    toc: false
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{Research Workflow}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

options(width = 90)
options(cli.unicode = FALSE)
pkg_root <- if (file.exists("DESCRIPTION")) "." else if (file.exists("../DESCRIPTION")) ".." else NA_character_
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
```

```{=html}
<style>
.ledgr-diagram {
  margin: 1.25rem auto 1.5rem auto;
  text-align: center;
}

.ledgr-diagram .mermaid {
  display: inline-block;
  max-width: 100%;
}

.ledgr-diagram .mermaid svg {
  display: block;
  height: auto !important;
  margin-left: auto;
  margin-right: auto;
}

.ledgr-workflow-diagram .mermaid svg {
  max-width: 560px !important;
}
</style>
```

You have an idea for a trading rule. How do you turn that hunch into evidence
you can reopen, inspect, and explain later, while you keep changing the rule?

Most research iterations change the strategy code itself: a new condition, a
different exit, a filter you did not think of on the first try. Some
iterations ask a narrower question, such as which lookback works best. This
article follows both kinds of iteration through one project and shows where
each one hands off.

This workflow produces research evidence, not investment advice or a guarantee
that a strategy will work in the future. See the project
[disclaimer](../DISCLAIMER.md).

::: {.ledgr-diagram .ledgr-workflow-diagram}

```{mermaid}
%%| echo: false
%%{init: {"theme": "base", "flowchart": {"nodeSpacing": 18, "rankSpacing": 22, "curve": "linear"}, "themeVariables": {"fontFamily": "system-ui, -apple-system, Segoe UI, sans-serif", "fontSize": "15px", "primaryColor": "#f8fafc", "primaryTextColor": "#1f2937", "primaryBorderColor": "#64748b", "lineColor": "#64748b", "tertiaryColor": "#eef2ff", "tertiaryTextColor": "#1f2937", "tertiaryBorderColor": "#64748b", "clusterBkg": "#ffffff", "clusterBorder": "#cbd5e1"}}}%%

flowchart TB
  data["Seal data"]
  rule["Write a rule"]
  run["Run and inspect"]
  compare["Compare committed runs"]
  sweep["Sweep a parameter question"]
  select["Select and validate"]
  note["Record the decision"]

  data --> rule --> run --> compare --> select --> note
  compare -. change the code .-> rule
  compare -. parameter question .-> sweep --> select
```

:::

## Prerequisites

```{r}
#| label: prerequisites
#| message: false

library(ledgr)
library(dplyr)
data("ledgr_demo_bars", package = "ledgr")
```

::: {.ledgr-callout .ledgr-callout-note}
**Running this yourself**

This article is evaluated when it is rendered, so it writes to a temporary
store. In a real project, replace `store_path` with a durable path such as
`artifacts/ledgr_store.duckdb`.
:::

## Project Topology

Start with a small project shape. One script creates the evidence, one store
holds the artifacts, and one report explains the research choice.

```text
my-ledgr-project/
  artifacts/
    ledgr_store.duckdb
  reports/
    workflow_review.md
  scripts/
    research_workflow.R
```

ledgr does not require this exact layout. The point is simpler: keep the
script, durable store, and written reasoning close enough that a future review
can recover the full research decision.

In a real project, add generated stores such as `artifacts/*.duckdb` to
`.gitignore` unless you have a deliberate artifact-versioning policy. The
script and source data should explain how to recreate the store; Git should not
quietly become the database backup.

## Seal The Evidence

Every version of the rule below runs against the same sealed snapshot, so the
comparison is between rules, not between data sets.

```{r}
#| label: snapshot

store_path <- ledgr_temp_store()

bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% c("DEMO_01", "DEMO_02", "DEMO_03"),
    between(ts_utc, ledgr_utc("2019-01-01"), ledgr_utc("2019-12-31"))
  )

snapshot <- ledgr_snapshot_from_df(
  bars,
  db_path = store_path,
  snapshot_id = "research_2019"
)
```

## Version 1: One Rule, One Run

The first idea is plain momentum: hold `qty` shares of every instrument whose
20-bar return is positive.

```{r}
#| label: version-1

momentum_v1 <- function(ctx, params) {
  rising <- ledgr_signal_return(ctx, lookback = 20) > 0
  ctx |>
    ledgr_selection(where = rising, missing = "exclude") |>
    ledgr_target_quantity(ctx, params$qty)
}

exp_v1 <- ledgr_experiment(
  snapshot = snapshot,
  strategy = momentum_v1,
  features = list(ledgr_ind_returns(20)),
  opening = ledgr_opening(cash = 10000),
  cost_model = ledgr_cost_zero()
)

run_v1 <- ledgr_run(exp_v1, params = list(qty = 10), run_id = "momentum_v1")
run_v1
```

Before judging the idea, check that the setup behaves at all: did anything
trade, and do the numbers look plausible for ten shares of three demo
instruments? If a run has no fills or implausible equity, fix the strategy or
the feature declarations before you iterate further.

## Version 2: Change The Code

Suppose the trades show the rule buying into short rallies inside a falling
market. The next iteration is not a new parameter value; it is a new
condition. Version 2 also requires the close to sit above its 50-bar moving
average. The 50-bar average is missing for longer than the 20-bar return, and
`missing = "exclude"` keeps an instrument out until both inputs exist.

```{r}
#| label: version-2

momentum_v2 <- function(ctx, params) {
  rising <- ledgr_signal_return(ctx, lookback = 20) > 0
  above_trend <- ctx$vec$close > ledgr_signal_feature(ctx, "sma_50")
  ctx |>
    ledgr_selection(where = rising & above_trend, missing = "exclude") |>
    ledgr_target_quantity(ctx, params$qty)
}

exp_v2 <- ledgr_experiment(
  snapshot = snapshot,
  strategy = momentum_v2,
  features = list(ledgr_ind_returns(20), ledgr_ind_sma(50)),
  opening = ledgr_opening(cash = 10000),
  cost_model = ledgr_cost_zero()
)

run_v2 <- ledgr_run(exp_v2, params = list(qty = 10), run_id = "momentum_v2")
```

Each version is its own committed run with its own strategy source and hash.
You do not overwrite version 1 to try version 2; both stay in the store.

## Compare Committed Runs

Label each version with what changed, then compare them side by side. The
store and comparison surfaces are taught in
`vignette("experiment-store", package = "ledgr")`.

```{r}
#| label: compare-versions

snapshot <- snapshot |>
  ledgr_run_label("momentum_v1", "20-bar momentum") |>
  ledgr_run_label("momentum_v2", "20-bar momentum above the 50-bar trend")

ledgr_run_compare(snapshot, run_ids = c("momentum_v1", "momentum_v2"))
```

Here the trend filter did not help: version 2 finished lower with a slightly
deeper drawdown. That is still useful evidence, and it stays in the store next
to version 1. Read the comparison as evidence about two rules on one window,
not as a verdict. Picking the better of two code versions is already a selection: the
more versions you try on the same data, the more room there is for luck to
look like an improvement.

## When A Parameter Question Appears

Sometimes the next question is narrower: is 20 bars the right lookback, or 50
the right trend length? That is a parameter question, and it has its own tool.
Declare the candidates as a grid, sweep them, and promote the one you choose;
`vignette("sweeps", package = "ledgr")` teaches that step, including failure
rows and how to reopen a promoted run later.

Keep the two kinds of iteration apart. Code changes are committed runs you
compare deliberately. Parameter questions are sweeps over a declared grid.
Mixing them, for example by editing the rule between sweep candidates, makes
the evidence impossible to reconstruct.

## Select, Then Validate

Choosing a version or a promoted candidate records a decision; it does not
prove the rule generalizes. Two articles test that decision:

- `vignette("selection-integrity", package = "ledgr")` asks whether a search
  over many variants gives any reason to trust the winner;
- `vignette("walk-forward", package = "ledgr")` re-runs the selection rule on
  later, held-out windows.

## Write The Human Research Note

Do not leave the reasoning only in your head. Write a short report next to the
stored artifacts. A compact report should include:

- hypothesis and data window
- snapshot hash and data-source assumptions
- each code version and why it changed
- feature and strategy declarations
- candidate grid summary, if you swept
- the ranking or selection rule
- issue and failure review
- equity and drawdown plots
- promotion note, if you promoted
- reason for rejecting alternatives
- selection caveat: the chosen version is not statistically validated by
  selection itself

Here is the shape of a useful entry:

```text
Hypothesis and data window:
  Short-term momentum may capture persistent moves in DEMO_01 to DEMO_03
  over the 2019 demo window.

Iterations:
  v1 bought every positive 20-bar return; v2 added a 50-bar trend filter after
  v1 bought rallies inside falling markets.

Selection caveat:
  Both versions were compared on the same window. The choice is recorded, not
  validated; walk-forward evaluation is the next step.
```

The report is where the human reasoning lives. ledgr records what happened,
which inputs were used, and which runs you compared or promoted; it does not
certify that the research protocol was statistically sound.

## Where Next

- For parameter grids, candidate review and promotion, read
  `vignette("sweeps", package = "ledgr")`.
- For run discovery, labels, comparison and reopening, read
  `vignette("experiment-store", package = "ledgr")`.
- For what a committed run can and cannot reproduce, read
  `vignette("reproducibility", package = "ledgr")`.

```{r}
#| label: cleanup
#| include: false

close(run_v1)
close(run_v2)
ledgr_snapshot_close(snapshot)
```
