---
title: "Risk And Cost Execution Policy"
format:
  html:
    toc: false
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{Risk And Cost Execution Policy}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 90)
options(cli.unicode = FALSE)
options(pillar.sigfig = 5)
pkg_root <- if (file.exists("DESCRIPTION")) "." else if (file.exists("../DESCRIPTION")) ".." else NA_character_
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
```

You could bury transaction costs and position limits inside your strategy
function: compute a spread, deduct a fee, clamp a position size. ledgr does not
make that the supported path, and that refusal is the point.

Two things go wrong when execution policy hides inside signal code. First, you
stop being able to trust your own numbers: a return that quietly nets out a
guessed fee reads as gross when you compare it later, and you will not notice.
Second, your strategy stops being about the signal. Position caps, fee
arithmetic, and exposure limits turn a few lines of signal logic into a
monolith that is hard to read and harder to change. In ledgr, a strategy should
answer one question: what do I want to hold?

So ledgr splits the execution decision into four questions, each owned by a
different layer:

| Layer | Question it answers |
| --- | --- |
| strategy | what do I want to hold? (intent) |
| risk | what am I allowed to hold? (constraint) |
| timing | when does it execute? (mechanics) |
| cost | what did it cost? (friction) |

A strategy states what it wants to hold. It does not place orders, check cash
or keep the books: the layers after it decide what is allowed, when it executes
and what it costs, and ledgr records the result. ledgr has no order management
yet, so an intent that cannot execute at its fill opportunity is recorded as
not executed rather than queued, and the strategy states its intent again at
the next decision.

```text
strategy targets
  -> risk chain        (reshape quantities)
  -> next-open timing  (choose the execution bar)
  -> cost model        (adjust fill price and fee)
  -> fill and ledger event
```

That order is the execution contract. A strategy returns complete target
quantities. A risk chain may reshape those quantities before any fill exists.
The timing model chooses the execution bar. The cost model then adjusts fill
price and fee. Read as a policy, ledgr execution is four things, and the rest of
this article makes each one visible:

- **explicit**: cost and risk are declared objects, never hidden in the strategy;
- **ordered**: each layer runs in a fixed place, and within cost, price
  transforms run before fees;
- **identity-bearing**: the cost and risk you chose are part of the run's identity;
- **bounded**: risk is not portfolio optimization, and cost is not a trading venue.

In the public API, the target-risk chain is the `risk_chain` argument and the
transaction-cost model is the `cost_model` argument. They are separate on
purpose: risk changes target quantities before fills exist; cost changes fill
price and fee after timing has selected an execution bar.

## Declare The Layers

```{r}
#| label: prerequisites
#| message: false

library(ledgr)
library(dplyr)

data("ledgr_demo_bars", package = "ledgr")
```

A cost model and a risk chain are values you build and pass to the experiment.
You assemble each one from small steps.

```{r}
#| label: policy

cost <- ledgr_cost_chain(
  ledgr_cost_spread_bps(10),
  ledgr_cost_notional_bps_fee(2)
)

risk <- ledgr_risk_chain(
  ledgr_risk_long_only(),
  ledgr_risk_max_weight(0.5)
)
```

`ledgr_cost_describe()` prints the resolved chain, including the stage of each
step.

```{r}
#| label: describe

cat(ledgr_cost_describe(cost))
```

## The Composable Pieces

You build any policy from the same small set of steps. Cost steps chain with
`ledgr_cost_chain()`:

| Step | Stage | Effect |
| --- | --- | --- |
| `ledgr_cost_spread_bps(bps)` | price transform | crosses half the quoted spread |
| `ledgr_cost_fixed_fee(amount)` | fee | flat charge per fill |
| `ledgr_cost_notional_bps_fee(bps)` | fee | charge proportional to fill notional |
| `ledgr_cost_zero()` | none | explicit no-cost policy |

Cost has two stages, and the order is enforced: price transforms always run
before fee steps, so a chain that puts a fee before a price transform fails with
a classed error. That is why a notional fee is charged on the spread-adjusted
price, not the raw open.

Risk steps chain with `ledgr_risk_chain()`:

| Step | Effect |
| --- | --- |
| `ledgr_risk_long_only()` | drops short (negative) targets to zero |
| `ledgr_risk_max_weight(w)` | caps absolute per-instrument weight at `w` |
| `ledgr_risk_none()` | explicit no-op |

`ledgr_cost_zero()` and `ledgr_risk_none()` are real choices, not absence. They
say "no cost" and "no constraint" on the record, where a later reader can see
them.

::: {.ledgr-callout .ledgr-callout-important}
**More steps are planned**

Later releases add more cost and risk steps, including richer fee and exposure
models. The chain interface does not change: you compose new steps the same way
you compose these.
:::

## What Risk Does

The cleanest way to see a layer is to swap it and hold everything else fixed.
The strategy below is deliberately aggressive, written to give the risk chain
something to correct.

```{r}
#| label: demo-strategy

greedy <- function(ctx, params) {
  long_id  <- ctx$universe[[1]]
  short_id <- ctx$universe[[2]]

  target <- ctx$flat()
  # Size each position as a fraction of equity, converted to whole shares
  # at the current price.
  target[[long_id]]  <-  floor(0.90 * ctx$equity / ctx$close(long_id))
  target[[short_id]] <- -floor(0.20 * ctx$equity / ctx$close(short_id))
  target
}
```

Each target line turns a fraction of current equity into a whole-share count at
the instrument's price: the first is about 90 percent of equity held long, the
second about 20 percent held short. Together that is roughly 110 percent gross
exposure -- a concentrated, slightly levered long-short bet. Both sides are past
the limits declared above. The 90 percent long exceeds the 50 percent cap, and a
short is not long-only, so the risk chain has two separate things to fix.

```{r}
#| label: data

bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% c("DEMO_01", "DEMO_02"),
    between(ts_utc, ledgr_utc("2019-01-01"), ledgr_utc("2019-03-31"))
  )

snapshot <- ledgr_snapshot_from_df(
  bars,
  db_path = tempfile(fileext = ".duckdb"),
  snapshot_id = "risk_cost_demo"
)
```

One helper runs the same strategy under whatever policy you pass, so the only
thing that changes between runs is the layer under test.

```{r}
#| label: run-helper

run_policy <- function(cost, risk, run_id) {
  exp <- ledgr_experiment(
    snapshot = snapshot,
    strategy = greedy,
    opening = ledgr_opening(cash = 10000),
    timing_model = ledgr_timing_next_open(),
    cost_model = cost,
    risk_chain = risk
  )
  ledgr_run(exp, params = list(), run_id = run_id, seed = 2026L)
}
```

::: {.ledgr-callout .ledgr-callout-note}
**Running this yourself**

The runs below write to a temporary DuckDB store, so the article leaves no
project files behind. The greedy strategy re-targets on every bar, so the final
bar logs a `LEDGR_LAST_BAR_NO_FILL` notice. That is expected next-open behavior;
see `vignette("execution-semantics", package = "ledgr")`.
:::

Run the strategy with no risk chain, then with the chain from above, and compare
the opening-day fills.

```{r}
#| label: risk-ab

uncapped <- run_policy(ledgr_cost_zero(), ledgr_risk_none(), "uncapped")
capped   <- run_policy(ledgr_cost_zero(), risk, "capped")

opening_fills <- function(bt) {
  ledgr_results(bt, what = "fills") |>
    filter(ts_utc == min(ts_utc)) |>
    select(instrument_id, side, qty, price)
}

opening_fills(uncapped)
opening_fills(capped)
```

Without a risk chain, ledgr fills exactly what the strategy asked for: a long
in the first instrument and a short in the second. With `long_only` and
`max_weight(0.5)`, the short is gone and the long is clamped to half of equity.
The strategy code did not change. The policy did.

The clamped quantity is fractional -- 54.632 shares -- because `max_weight`
scales the target to hit exactly 50 percent of equity at the decision price,
rather than flooring to a whole share. ledgr target quantities are continuous at
this layer, and no shipped risk step rounds them. If your venue trades whole
shares or lots, round in the strategy for now, for example with `floor()` on
the quantities you return, and know that a later `max_weight` clamp can make a
whole target fractional again.

## What Cost Does

Now hold the risk chain fixed and swap the cost model. The `capped` run above
already used `ledgr_cost_zero()`, so it is the no-cost baseline. Run the same
policy with the spread-and-fee chain and compare the opening fill.

```{r}
#| label: cost-ab

with_cost <- run_policy(cost, risk, "with_cost")

first_fill <- function(bt) {
  ledgr_results(bt, what = "fills") |>
    filter(ts_utc == min(ts_utc), instrument_id == "DEMO_01") |>
    transmute(qty, price = round(price, 3), fee = round(fee, 3))
}

bind_rows(
  "no cost"      = first_fill(capped),
  "spread + fee" = first_fill(with_cost),
  .id = "policy"
)
```

Same quantity, two different fills. The spread step lifts the buy price by half
the quoted spread, and the notional fee adds a charge proportional to the fill
value. Costs do not change what you decided to hold; they change what holding it
was worth.

### Timing, Spread, And Fees

Timing and cost are separate execution steps. `ledgr_timing_next_open()` decides
where a target change can fill: the next available open after the strategy
decision. A cost model then resolves the proposed fill price and explicit fee.
Strategies do not receive cost state, and cost models do not change side,
quantity, instrument, or execution timestamp.

`ledgr_cost_spread_bps()` uses a quoted bid/ask-spread convention. A buy crosses
half the quoted spread above the reference open:
`open * (1 + spread_bps / 20000)`. A sell crosses half below it:
`open * (1 - spread_bps / 20000)`. For the same reference price, a buy/sell
round trip therefore crosses approximately `spread_bps` basis points before
explicit fees.

```{r}
#| label: quoted-spread-arithmetic

spread_bps <- 25
reference_price <- 100
buy_cross <- reference_price * (1 + spread_bps / 20000)
sell_cross <- reference_price * (1 - spread_bps / 20000)
round_trip_bps <- (buy_cross - sell_cross) / reference_price * 10000
round(round_trip_bps, 6)
```

Price transforms and explicit fees are different. `ledgr_cost_spread_bps()`
changes the fill price. `ledgr_cost_fixed_fee()` and
`ledgr_cost_notional_bps_fee()` add values to the fill `fee` column after any
price transforms have resolved.

::: {.ledgr-callout .ledgr-callout-important}
**What costs do not model**

Cost models are deterministic research assumptions over accepted fill proposals.
They do not implement liquidity or capacity limits, financing, transaction-cost
analysis, taxes, OMS lifecycle behavior, or broker reconciliation. Those are
separate future layers, not hidden behavior in the cost API.
:::

## Identity

The cost model and risk chain you chose are part of the run's identity. ledgr
hashes them, so two runs are only like-for-like comparable when their policy
matches. You can still compare different policies deliberately, as the no-cost
versus cost contrast above does; you just should not read them as sharing the
same execution assumptions.

```{r}
#| label: identity

with_cost$config$cost_model$cost_model_hash
with_cost$config$risk_chain$risk_chain_hash
```

Change the cost model or the risk chain and these hashes change with it. How
much result evidence you retain, or how you rank candidates later, does not
rewrite them. For the full identity model, see
`vignette("reproducibility", package = "ledgr")`.

## Boundaries

The four layers are deliberately narrow. Each refuses work that belongs to a
different kind of system.

::: {.ledgr-callout .ledgr-callout-warning}
**Risk is not portfolio optimization**

`ledgr_risk_max_weight()` is a small deterministic target transform, not a
solver. It clamps quantities you already chose; it does not search for an
allocation.
:::

::: {.ledgr-callout .ledgr-callout-warning}
**Cost is not liquidity or capacity**

A cost model adds spread and fees. It does not decide whether volume is
sufficient, whether an order should be refused, or how a large order moves the
market.
:::

::: {.ledgr-callout .ledgr-callout-important}
**The research engine is not a broker**

Paper trading, order lifecycle, partial fills, cancellation, broker
reconciliation, financing, taxes, and live restart safety are separate future
layers. ledgr answers what a sealed snapshot and your declared policy can
determine; it does not stand in for a live venue.
:::

These are not gaps to patch inside a strategy. They are questions a sealed
backtest cannot answer, kept outside the execution contract on purpose.

## Try It

::: {.ledgr-callout .ledgr-callout-tip}
**Try it**

Lower the cap to `ledgr_risk_max_weight(0.3)` and rerun the risk comparison
with a new run ID such as `"capped_03"`. Which opening fill changes, and by
how much? Then add a `ledgr_cost_fixed_fee()` step, use another new run ID, and
watch the fee column in the opening fill. Run IDs are durable identities, so
do not reuse the article's IDs for changed configurations.
:::

## Where Next

- For finding, comparing, and reopening stored runs, see
  `vignette("experiment-store", package = "ledgr")`.
- For fill timing, next-open semantics, and the final-bar rule, see
  `vignette("execution-semantics", package = "ledgr")`.
- For cost and risk identity in the full provenance model, see
  `vignette("reproducibility", package = "ledgr")`.

```{r}
#| label: cleanup
#| include: false

close(uncapped)
close(capped)
close(with_cost)
ledgr_snapshot_close(snapshot)
```
