---
title: "Selection Integrity"
format:
  html:
    toc: false
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{Selection Integrity}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 90)
options(cli.unicode = FALSE)
options(pillar.sigfig = 5)
pkg_root <- if (file.exists("DESCRIPTION")) "." else if (file.exists("../DESCRIPTION")) ".." else NA_character_
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
```

Sweeps let you compare many candidate parameterizations. That is useful, and it
is dangerous. The more candidates you inspect, the easier it is to mistake a
lucky in-sample winner for a robust strategy. Selection-integrity diagnostics
make that risk visible; they do not choose the winner for you.

## Question

`ledgr_sweep_pbo()` asks one question: when the retained candidate return panel
is recombined into symmetric in-sample and out-of-sample splits, how often does
the in-sample winner rank poorly out of sample?

The answer is the Probability of Backtest Overfitting (PBO). Higher values are
warning signs about the candidate family and the selection process. They are not
proof that a strategy will lose money, and low values are not proof that a
strategy will make money.

## Evidence

The diagnostic consumes retained sweep returns:

```r
sweep <- ledgr_sweep(
  experiment,
  grid,
  retain = ledgr_sweep_retention("completed")
)

pbo <- ledgr_sweep_pbo(sweep, S = 4)
```

The input is the same retained completed-candidate evidence exposed by
`ledgr_sweep_returns_panel()`: one UTC-ordered return column per completed
candidate, with the structural first `NA_real_` row verified and dropped before
the CSCV matrix is formed. It does not inspect fills, positions, promotion
records, or walk-forward folds.

## Method Shape

PBO uses Combinatorially Symmetric Cross Validation (CSCV):

1. Split the return panel into `S` contiguous subsets.
2. Choose half the subsets as in sample and the other half as out of sample.
3. Score every candidate in sample.
4. Take the in-sample winner and rank that same candidate out of sample.
5. Convert the out-of-sample rank to `lambda`.
6. Report PBO as the share of cases where `lambda <= threshold`.

The default score is mean period return. A custom metric can be supplied, but it
must return one finite numeric value per candidate column, and larger values are
treated as better.

## Interpretation

Read PBO as selection-process evidence. A high value says the candidate that
looks best in sample often fails to hold its rank out of sample under CSCV
recombination. That is a reason to distrust the apparent sweep winner, reduce
the candidate search space, gather more evidence, or add a separate
walk-forward evaluation.

The result object keeps three public tables:

- `as_tibble(pbo)` gives the one-row summary.
- `as_tibble(pbo, what = "cases")` gives each CSCV split and logit.
- `as_tibble(pbo, what = "degradation")` gives the winner in-sample versus
  out-of-sample score by split.

## Limits

PBO is not a profitability proof. It does not fix bad data, survivorship bias,
point-in-time universe mistakes, revised-data leakage, or preprocessing that
used future information before ledgr saw the inputs. It also depends on a
meaningful candidate family. If the sweep has too few candidates, too few
observations, or candidates that were already mined before the declared sweep,
interpretation weakens.

PBO is sweep-level in this release. It does not add per-fold train-sweep PBO to
walk-forward degradation tables.

## Failure Modes

The function fails closed when the evidence is not suitable:

- retained returns were not requested;
- a selected candidate failed or has no retained return rows;
- the completed candidates do not form one complete timestamp grid;
- the structural first return row is not `NA_real_`;
- `S` is odd, too large, or does not divide the post-first-row return count;
- the metric does not return one finite numeric score per candidate.

These are setup or evidence problems, not weak strategy results. Fix the panel
before interpreting the diagnostic.

## References

The method follows the CSCV/PBO shape described by Bailey, Borwein,
Lopez de Prado, and Zhu. ledgr implements the diagnostic natively over retained
return panels; the CRAN `pbo` package remains optional reference evidence, not a
runtime dependency.

## Worked Example

The synthetic retained sweep below is intentionally cautionary. Each candidate
looks good in one segment and weak elsewhere. That creates an attractive-looking
search space, but PBO flags that the in-sample winner does not usually remain
strong out of sample.

```{r}
#| label: synthetic-sweep
#| include: false

make_retained_sweep <- function(returns) {
  returns <- as.matrix(returns)
  candidate_ids <- colnames(returns)
  ts_utc <- as.POSIXct("2020-01-01", tz = "UTC") + 86400 * seq_len(nrow(returns) + 1L)
  rows <- lapply(seq_along(candidate_ids), function(j) {
    period_return <- c(NA_real_, as.numeric(returns[, j]))
    tibble::tibble(
      sweep_id = "selection-integrity-example",
      candidate_id = candidate_ids[[j]],
      ts_utc = ts_utc,
      equity = 100 * cumprod(c(1, 1 + as.numeric(returns[, j]))),
      period_return = period_return
    )
  })
  out <- tibble::tibble(
    candidate_id = candidate_ids,
    candidate_row = seq_along(candidate_ids),
    status = "DONE"
  )
  class(out) <- c("ledgr_sweep_results", class(out))
  attr(out, "sweep_id") <- "selection-integrity-example"
  attr(out, "sweep_retention") <- ledgr_sweep_retention("completed")
  attr(out, "sweep_returns") <- tibble::as_tibble(do.call(rbind, rows))
  out
}

overfit_returns <- matrix(
  -0.01,
  nrow = 12,
  ncol = 4,
  dimnames = list(NULL, paste0("candidate_", 1:4))
)
subset_id <- rep(1:4, each = 3)
for (j in 1:4) {
  overfit_returns[subset_id == j, j] <- 0.05
}
example_sweep <- make_retained_sweep(overfit_returns)
```

```{r}
#| label: pbo-summary

pbo <- ledgr_sweep_pbo(example_sweep, S = 4)
tibble::as_tibble(pbo)
```

The case table shows why the summary is high: the in-sample winner is often not
the out-of-sample winner.

```{r}
#| label: pbo-cases

tibble::as_tibble(pbo, what = "degradation")
```

`ledgr_sweep_pbo()` does not select or promote a candidate. Treat it as one
piece of evidence beside the sweep table, the walk-forward degradation table,
and the research judgment that decides what to test next.

## Minimum Track Record Length

### Question

`ledgr_sweep_min_track_record()` asks a narrower single-series question for
each retained candidate: how many return observations would this observed
Sharpe ratio need before it is statistically distinguishable from a reference
Sharpe threshold at the requested confidence level?

The answer is MinTRL. A candidate can have a positive observed Sharpe and still
need more observations before the track record clears the reference threshold.

### Evidence

The diagnostic consumes the same retained sweep return panel as PBO:

```r
min_trl <- ledgr_sweep_min_track_record(
  sweep,
  reference_sharpe = 0,
  confidence = 0.95
)
```

The retained `period_return` columns are the evidence. The structural first
`NA_real_` row is verified and dropped by the panel layer before the per-series
Sharpe, skewness, and kurtosis are computed.

### Method Shape

For each candidate, ledgr computes:

1. per-period excess returns, using `risk_free_return` when supplied;
2. observed per-period Sharpe ratio;
3. return skewness and kurtosis;
4. the Bailey/Lopez de Prado minimum track record length formula against
   `reference_sharpe` and `confidence`.

The output is measured in return observations, not calendar years. If the
observed Sharpe is not above the reference Sharpe, ledgr keeps the candidate in
the table and marks the required length as infinite instead of silently
dropping it.

### Interpretation

Read MinTRL as sample-size evidence. It answers whether the observed track
record is long enough for the selected reference Sharpe threshold. It does not
say the strategy is robust, causal, or deployable.

Use it when a candidate looks promising but short-lived. A large extra
observation count says the apparent Sharpe may mostly be a short-sample story.

### Limits

MinTRL inherits the quality of the retained return series. It does not fix
candidate mining, leakage, non-stationarity, changing market regimes, or
survivorship-biased universes. It also does not compare many candidates at once;
that broader selection-process question belongs to PBO and later DSR/effective
trial diagnostics.

### Failure Modes

The function fails closed when:

- retained returns were not requested;
- the completed candidates do not form one complete timestamp grid;
- there are fewer than four post-first-row observations;
- a return series is constant or non-finite;
- `reference_sharpe`, `confidence`, or `risk_free_return` is invalid.

These failures mean the evidence cannot support the diagnostic. They are not
strategy judgments.

### References

The MinTRL formula follows the Sharpe-ratio track-record-length expression used
by Bailey and Lopez de Prado and exposed by PerformanceAnalytics as
`MinTrackRecord()`. ledgr implements it natively; PerformanceAnalytics remains
optional reference evidence, not a runtime dependency.

### Worked Example

The first candidate below has smoother positive returns than the second. The
second candidate's Sharpe is positive, but the required track record length
shows that the evidence is still much too short at 95 percent confidence.

```{r}
#| label: min-trl-example

min_trl_returns <- cbind(
  promising = c(0.012, 0.016, 0.010, 0.018, 0.013, 0.017, 0.011, 0.015),
  noisy = c(0.040, -0.030, 0.035, -0.025, 0.030, -0.020, 0.025, -0.015)
)
min_trl_sweep <- make_retained_sweep(min_trl_returns)
min_trl <- ledgr_sweep_min_track_record(min_trl_sweep, reference_sharpe = 0)
min_trl
```

The diagnostic keeps both candidates in the table. It does not select the
candidate with the shorter required track record.

## Where Next

- For retained sweep return panels, read `vignette("sweeps", package = "ledgr")`.
- For walk-forward train/test evidence, read
  `vignette("walk-forward", package = "ledgr")`.
- For lookahead leakage examples, read `vignette("leakage", package = "ledgr")`.
