---
title: "Selection Integrity"
format:
  html:
    toc: false
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{Selection Integrity}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 90)
options(cli.unicode = FALSE)
options(pillar.sigfig = 5)
pkg_root <- if (file.exists("DESCRIPTION")) "." else if (file.exists("../DESCRIPTION")) ".." else NA_character_
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
```

Sweeps let you compare many candidate parameterizations. That is useful, and it
is dangerous. The more candidates you inspect, the easier it is to mistake a
lucky in-sample winner for a robust strategy. Selection-integrity diagnostics
make that risk visible; they do not choose the winner for you.

## Question

`ledgr_sweep_pbo()` asks one question: when the retained candidate return panel
is recombined into symmetric in-sample and out-of-sample splits, how often does
the in-sample winner rank poorly out of sample?

The answer is the Probability of Backtest Overfitting (PBO). Higher values are
warning signs about the candidate family and the selection process. They are not
proof that a strategy will lose money, and low values are not proof that a
strategy will make money.

## Evidence

The diagnostic consumes retained sweep returns:

```r
sweep <- ledgr_sweep(
  experiment,
  grid,
  retain = ledgr_sweep_retention("completed")
)

pbo <- ledgr_sweep_pbo(sweep, S = 4)
```

The input is the same retained completed-candidate evidence exposed by
`ledgr_sweep_returns_panel()`: one UTC-ordered return column per completed
candidate, with the structural first `NA_real_` row verified and dropped before
the CSCV matrix is formed. It does not inspect fills, positions, promotion
records, or walk-forward folds.

## Method Shape

PBO uses Combinatorially Symmetric Cross Validation (CSCV):

1. Split the return panel into `S` contiguous subsets.
2. Choose half the subsets as in sample and the other half as out of sample.
3. Score every candidate in sample.
4. Take the in-sample winner and rank that same candidate out of sample.
5. Convert the out-of-sample rank to `lambda`.
6. Report PBO as the share of cases where `lambda <= threshold`.

The default score is mean period return. A custom metric can be supplied, but it
must return one finite numeric value per candidate column, and larger values are
treated as better.

## Interpretation

Read PBO as selection-process evidence. A high value says the candidate that
looks best in sample often fails to hold its rank out of sample under CSCV
recombination. That is a reason to distrust the apparent sweep winner, reduce
the candidate search space, gather more evidence, or add a separate
walk-forward evaluation.

The result object keeps three public tables:

- `as_tibble(pbo)` gives the one-row summary.
- `as_tibble(pbo, what = "cases")` gives each CSCV split and logit.
- `as_tibble(pbo, what = "degradation")` gives the winner in-sample versus
  out-of-sample score by split.

## Limits

PBO is not a profitability proof. It does not fix bad data, survivorship bias,
point-in-time universe mistakes, revised-data leakage, or preprocessing that
used future information before ledgr saw the inputs. It also depends on a
meaningful candidate family. If the sweep has too few candidates, too few
observations, or candidates that were already mined before the declared sweep,
interpretation weakens.

PBO is sweep-level in this release. It does not add per-fold train-sweep PBO to
walk-forward degradation tables.

## Failure Modes

The function fails closed when the evidence is not suitable:

- retained returns were not requested;
- a selected candidate failed or has no retained return rows;
- the completed candidates do not form one complete timestamp grid;
- the structural first return row is not `NA_real_`;
- `S` is odd, too large, or does not divide the post-first-row return count;
- the metric does not return one finite numeric score per candidate.

These are setup or evidence problems, not weak strategy results. Fix the panel
before interpreting the diagnostic.

## References

The method follows the CSCV/PBO shape described by Bailey, Borwein,
Lopez de Prado, and Zhu. ledgr implements the diagnostic natively over retained
return panels; the CRAN `pbo` package remains optional reference evidence, not a
runtime dependency.

## Worked Example

The synthetic retained sweep below is intentionally cautionary. Each candidate
looks good in one segment and weak elsewhere. That creates an attractive-looking
search space, but PBO flags that the in-sample winner does not usually remain
strong out of sample.

```{r}
#| label: synthetic-sweep
#| include: false

make_retained_sweep <- function(returns) {
  returns <- as.matrix(returns)
  candidate_ids <- colnames(returns)
  ts_utc <- as.POSIXct("2020-01-01", tz = "UTC") + 86400 * seq_len(nrow(returns) + 1L)
  rows <- lapply(seq_along(candidate_ids), function(j) {
    period_return <- c(NA_real_, as.numeric(returns[, j]))
    tibble::tibble(
      sweep_id = "selection-integrity-example",
      candidate_id = candidate_ids[[j]],
      ts_utc = ts_utc,
      equity = 100 * cumprod(c(1, 1 + as.numeric(returns[, j]))),
      period_return = period_return
    )
  })
  out <- tibble::tibble(
    candidate_id = candidate_ids,
    candidate_row = seq_along(candidate_ids),
    status = "DONE"
  )
  class(out) <- c("ledgr_sweep_results", class(out))
  attr(out, "sweep_id") <- "selection-integrity-example"
  attr(out, "sweep_retention") <- ledgr_sweep_retention("completed")
  attr(out, "sweep_returns") <- tibble::as_tibble(do.call(rbind, rows))
  out
}

overfit_returns <- matrix(
  -0.01,
  nrow = 12,
  ncol = 4,
  dimnames = list(NULL, paste0("candidate_", 1:4))
)
subset_id <- rep(1:4, each = 3)
for (j in 1:4) {
  overfit_returns[subset_id == j, j] <- 0.05
}
example_sweep <- make_retained_sweep(overfit_returns)
```

```{r}
#| label: pbo-summary

pbo <- ledgr_sweep_pbo(example_sweep, S = 4)
tibble::as_tibble(pbo)
```

The case table shows why the summary is high: the in-sample winner is often not
the out-of-sample winner.

```{r}
#| label: pbo-cases

tibble::as_tibble(pbo, what = "degradation")
```

`ledgr_sweep_pbo()` does not select or promote a candidate. Treat it as one
piece of evidence beside the sweep table, the walk-forward degradation table,
and the research judgment that decides what to test next.

## Where Next

- For retained sweep return panels, read `vignette("sweeps", package = "ledgr")`.
- For walk-forward train/test evidence, read
  `vignette("walk-forward", package = "ledgr")`.
- For lookahead leakage examples, read `vignette("leakage", package = "ledgr")`.
