---
title: "Strategy Development And Comparison"
output:
  rmarkdown::html_vignette:
    toc: false
  github_document:
    toc: false
    html_preview: false
vignette: >
  %\VignetteIndexEntry{Strategy Development And Comparison}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 90)
options(cli.unicode = FALSE)
default_output_hook <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(
  output = function(x, options) {
    default_output_hook(gsub("[ \t]+(?=\n)", "", x, perl = TRUE), options)
  }
)
demo_data_path <- if (file.exists("data/ledgr_demo_bars.rda")) {
  "data/ledgr_demo_bars.rda"
} else {
  file.path("..", "data", "ledgr_demo_bars.rda")
}
if (!exists("ledgr_demo_bars") && file.exists(demo_data_path)) {
  load(demo_data_path)
}
```

The examples use `dplyr` for demo-data preparation. Strategy functions use
ledgr's pulse context rather than data-frame operations.

```{r library, message=FALSE}
library(ledgr)
library(dplyr)
data("ledgr_demo_bars", package = "ledgr")
```

A ledgr strategy is an ordinary R function, but the important idea is not the
function syntax. The important idea is the pulse.

A backtest in ledgr is a sequence of decision moments. At each pulse, ledgr
shows the strategy only what could have been known at that time. The strategy
answers with desired holdings. ledgr records the decision, applies the fill
model, and moves to the next pulse.

This matters because leakage is easy. If future information enters a historical
decision, the backtest can look profitable for the wrong reason. ledgr's
strategy interface is built to make one common mistake harder: your strategy
receives one pulse context, not the whole future. For the broader leakage model,
including feature-construction leakage and remaining user responsibilities, see
`vignette("leakage", package = "ledgr")`.

## Wrong And Right: Leakage

The tempting vectorized pattern is to compute a future-looking column first and
then trade from it. In the example below, `lead(close)` shifts tomorrow's close
onto today's row. The resulting `buy_signal` looks like an ordinary column, but
it answers a question the strategy could not have answered at today's decision
time: "will tomorrow's close be higher than today's close?" Trading from that
column lets the backtest use future market data as if it were already known.

```{r leakage-wrong, eval=FALSE}
leaky_signals <- ledgr_demo_bars |>
  group_by(instrument_id) |>
  arrange(ts_utc, .by_group = TRUE) |>
  mutate(
    tomorrow_close = lead(close),
    buy_signal = tomorrow_close > close
  )
```

The ledgr version expresses the rule at one pulse. The strategy can read the
current bar for the current instrument. Later sections add registered features
to the same pulse model. The strategy has no market-data table from which it can
casually index tomorrow's bar. That is the same information shape a live trading
strategy gets as time passes: each pulse is a new slice of the knowable
universe.

```{r leakage-right}
no_leak_bar_strategy <- function(ctx, params) {
  targets <- ctx$flat()

  for (id in ctx$universe) {
    if (ctx$close(id) > ctx$open(id)) {
      targets[id] <- 1
    }
  }

  targets
}
```

This removes one common source of leakage, but it does not certify that
snapshots, feature definitions, event timestamps, universe construction, or
parameter selection are causally clean.

## A Strategy That Does Nothing

The simplest economic policy is: hold cash and own no instruments.

```{r flat-strategy}
flat_strategy <- function(ctx, params) {
  ctx$flat()
}
```

`ctx$flat()` creates a full target vector with one entry for every instrument in
the run and every value set to zero. Economically, this means: after the next
fill opportunity, hold no positions.

This is a complete ledgr strategy. It is not useful for making money, but it is
useful for understanding the contract: at every pulse, return target holdings.

The return value is a named numeric vector. Names are instrument IDs from
`ctx$universe`, values are desired quantities. `ctx$flat()` produces the
full-universe shape with every entry at zero.

The deeper mental model is that a strategy is a **policy**, not a sequence of
orders. At each pulse, the strategy declares a desired state: "I want to hold
this many shares of each instrument." The engine compares that against current
holdings, computes the gap, and fills accordingly. The strategy never says
"buy 3 shares"; it says "I want to hold 3 shares." This distinction is what
keeps strategies free from execution-state bookkeeping, and it is what makes
a ledgr strategy composable, testable, and directly readable as financial
reasoning.

## What Is `ctx`?

`ctx` is the pulse context. It is the information packet ledgr gives your
strategy at one decision time.

It contains the current timestamp, current bars, current features, current
positions, cash, equity, and small helper functions for accessing those values.
It is deliberately not the full dataset.

| Expression | Meaning at one pulse |
|---|---|
| `ctx$ts_utc` | current decision timestamp |
| `ctx$universe` | instruments in the run |
| `ctx$open(id)`, `ctx$close(id)` | current bar values for one instrument |
| `ctx$feature(id, feature_id)` | current indicator value for one instrument by engine feature ID |
| `ctx$features(id, feature_map)` | mapped indicator values for one instrument by alias |
| `ctx$position(id)` | current simulated position |
| `ctx$cash`, `ctx$equity` | current simulated portfolio state |
| `ctx$flat()` | target zero positions unless changed |
| `ctx$hold()` | target current positions unless changed |

For the installed accessor reference, see `?ledgr_strategy_context`.

Use `ctx$flat()` when the strategy should be flat unless it sees a reason to
act. Use `ctx$hold()` when the strategy should keep existing positions unless
it sees a reason to change them.

For example, this policy starts from current holdings and only changes the book
when it sees an exit reason. Economically, it means: "keep what I already own,
unless today's bar gives me a reason to leave."

```{r hold-example}
hold_unless_down <- function(ctx, params) {
  targets <- ctx$hold()

  for (id in ctx$universe) {
    if (ctx$close(id) < ctx$open(id)) {
      targets[id] <- 0
    }
  }

  targets
}
```

This loop style is fine while the mechanics are still visible. Once you have
helpers like `signal_*()` and `select_*()`, most strategy logic is easier to
express at the whole-universe level instead of one instrument at a time.

## A First Trading Rule

Now add one small economic idea:

> If an instrument closes above its open, own one share. Otherwise own nothing.

This is still a teaching strategy, not investment advice. It shows how
observable data becomes a target.

```{r buy-if-up}
buy_if_up <- function(ctx, params) {
  targets <- ctx$flat()

  for (id in ctx$universe) {
    if (ctx$close(id) > ctx$open(id)) {
      targets[id] <- 1
    }
  }

  targets
}
```

Targets are desired quantities, not orders, signals, or portfolio weights. A
target of `1` means "after the next fill opportunity, hold one share." A target
of `0` means "hold no shares."

In these examples, decisions fill at the next open: a decision made at pulse
`t` fills at the next available bar. That keeps the strategy from deciding and
filling on the same close.

The loop is intentionally plain because this is the first example. Once the
economic idea is clear, ledgr strategies are usually easier to read when they
use helper functions that operate on the whole universe at once. The later
sections make that transition.

## Why `params` Exists

Hard-coded constants make experiments awkward. Parameters let one economic idea
run under different assumptions.

```{r buy-if-up-param}
buy_if_up_qty <- function(ctx, params) {
  targets <- ctx$flat()

  for (id in ctx$universe) {
    if (ctx$close(id) > ctx$open(id)) {
      targets[id] <- params$qty
    }
  }

  targets
}
```

Strategies use `function(ctx, params)`. `ctx` is the pulse. `params` is the
experimenter's chosen configuration for this run. Keeping them separate makes
the strategy easier to test, compare, and store.

## Prepare A Small Experiment

Use two instruments from the offline demo data so the examples run anywhere.

```{r data}
bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% c("DEMO_01", "DEMO_02"),
    between(
      ts_utc,
      ledgr::ledgr_utc("2019-01-01"),
      ledgr::ledgr_utc("2019-06-30")
    )
  )

snapshot <- ledgr_snapshot_from_df(
  bars,
  snapshot_id = "strategy_chapter_snapshot"
)
```

The snapshot seals the market data. That is the evidence base for the
experiment. Strategies and indicators can derive from it, but the underlying
bars do not change mid-research.

## Indicators And Feature IDs

Indicators are feature definitions. Before a strategy uses a feature, ask ledgr
for the exact ID.

```{r feature-ids}
features <- list(ledgr_ind_returns(5))

ledgr_feature_id(features)
```

Those strings are the names used inside `ctx$feature()`. They are exact. A typo
such as `"returns_5"` is not treated as a warmup value; it is an unknown feature
and ledgr fails loudly.

Warmup is different. A known feature can be `NA` early in the sample because
there are not enough prior bars yet. Strategy code should treat that as "no
signal yet."

## Debug One Pulse Before Running

Before running a full backtest, inspect one pulse. This is the fastest way to
understand what your strategy will see.

```{r pulse}
pulse <- ledgr_pulse_snapshot(
  snapshot,
  universe = c("DEMO_01", "DEMO_02"),
  ts_utc = ledgr::ledgr_utc("2019-03-01"),
  features = features
)

pulse$ts_utc
pulse$universe
pulse$close("DEMO_01")
pulse$feature("DEMO_01", "return_5")
pulse$hold()
```

The same pulse can also be viewed as one wide row. This is useful when you want
to see prices, portfolio state, and computed features together.

```{r pulse-wide}
ledgr_pulse_wide(pulse) |>
  glimpse()
```

The wide row and the scalar accessors are two ways of looking at the same
pulse-known data. The wide row is good for inspection and model-like thinking.
The rest of this vignette uses the non-wide accessors because they keep the
step-by-step strategy logic easier to read.

Now build the strategy logic one transformation at a time.

The economic idea:

> Rank instruments by recent return, keep the top names, split capital equally,
> and convert those weights into share quantities.

`signal_return()` is a thin helper around the same feature you inspected above:
it reads `return_N` for every instrument in the pulse and returns one
universe-wide signal object.

The helper pipeline has four stages:

| Stage | Input | Output | Question answered |
|---|---|---|---|
| signal | pulse context | numeric scores with origin metadata | What looks attractive? |
| selection | signal | logical inclusion with the same origin | What should be considered? |
| weights | selection | allocation weights with the same origin | How should capital be split? |
| target | weights and context | full-universe share quantities | What should the portfolio hold? |

Execution semantics begin only at the target stage. `signal`, `selection`, and
`weights` are research objects that help author the strategy; `target` is the
ordinary full named target vector shape the runner validates and executes.

```{r pulse-pipeline}
signal <- signal_return(pulse, lookback = 5)
signal

selection <- select_top_n(signal, n = 1)
selection

weights <- weight_equal(selection)
weights

target <- target_rebalance(weights, pulse, equity_fraction = 0.1)
target
```

`target_rebalance()` sizes with current pulse equity and current close prices,
then floors to whole shares. For the selected `DEMO_01` pulse above, 10% of
equity is allocated to the one selected instrument:

```{r floor-sizing}
raw_qty <- weights[["DEMO_01"]] * 0.1 * pulse$equity / pulse$close("DEMO_01")
c(pre_floor = raw_qty, target_qty = unclass(target)[["DEMO_01"]])
```

The general weighted sizing formula is:

```text
floor(weight * equity_fraction * ctx$equity / ctx$close(instrument_id))
```

For a raw target strategy that does not use weights, the same idea reduces to:

```text
floor(equity_fraction * ctx$equity / ctx$close(instrument_id))
```

Both formulas use decision-time close and current pulse equity. Fills still
occur at the configured later fill point, so fill value can drift from
decision-time sizing.

The helper objects are not a second execution path. They are authoring aids.
The pipeline still ends in a `ledgr_target`, which unwraps to the same target
quantity vector the runner has always consumed.

Interactive pulse snapshots and backtest handles can be closed when you are
done inspecting them. This releases DuckDB resources in long sessions; completed
run artifacts are already durable when `ledgr_run()` returns.

```{r close-pulse}
close(pulse)
```

## Turn The Idea Into A Strategy

The same transformations become an ordinary strategy function.

The full backtest replays every bar, including the earliest warmup pulses.
During those first pulses, `return_5` is `NA` for every instrument because five
prior bars do not exist yet. `select_top_n()` treats that all-missing signal as
a classed empty selection, not as a warning. That object still carries the
original universe and signal origin. `weight_equal()` turns it into empty
weights, and `target_rebalance()` turns those weights into a flat
full-universe target.

No warning suppression is needed for ordinary early warmup. A partial-selection
warning can still appear when some signal values are usable but fewer than `n`
instruments can be selected. If a run finishes with zero trades, inspect a late
pulse before assuming the empty selection was only early warmup.

```{r helper-strategy}
top_return_strategy <- function(ctx, params) {
  signal <- signal_return(ctx, lookback = params$lookback)
  selection <- select_top_n(signal, n = params$n)

  weights <- weight_equal(selection)
  target_rebalance(weights, ctx, equity_fraction = params$equity_fraction)
}
```

Read it economically:

1. `signal_return()` scores each instrument by recent return.
2. `select_top_n()` keeps the highest scores and ignores warmup `NA`.
3. `weight_equal()` splits the chosen allocation equally.
4. `target_rebalance()` converts weights into floored full target quantities.

No helper registers indicators automatically. The experiment must say which
features exist.

The empty-selection path is intentionally an object path rather than a
condition path. It lets expected warmup and "no signal today" flow through the
same helper pipeline as an ordinary selection. Diagnostics still belong at the
pulse level: when a strategy produces no fills or no closed trades, inspect a
late pulse and confirm whether the feature values are usable.

```{r experiment}
exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = top_return_strategy,
  features = features,
  opening = ledgr_opening(cash = 10000)
)
```

## Feature Maps For Readable Feature Access

The examples above keep the exact feature ID contract visible:
`ctx$feature(id, feature_id)` reads one registered feature for one instrument at
one pulse. That contract remains the foundation.

When a strategy reads several features per instrument, repeating feature ID
strings can obscure the trading idea. A feature map bundles indicator objects
with strategy-facing aliases. The same object can be registered with the
experiment and used by the strategy for pulse-time lookup.

```{r feature-map-setup}
mapped_features <- ledgr_feature_map(
  ret_5 = ledgr_ind_returns(5),
  sma_10 = ledgr_ind_sma(10)
)

ledgr_feature_id(mapped_features)
```

The strategy closes over `mapped_features`. Inside the universe loop,
`ctx$features(id, mapped_features)` returns a named numeric vector keyed by the
aliases. `passed_warmup()` is a guard for that vector: for values returned by
`ctx$features()`, it means every requested indicator is usable at this pulse.
It is not a signal pipeline transformation, and it is not a data-quality
diagnostic for arbitrary vectors. A zero-length input is a classed error because
an empty feature bundle cannot prove warmup has passed. The specific error
classes are `ledgr_empty_warmup_input` and `ledgr_invalid_warmup_input`.

```{r feature-map-strategy}
mapped_return_strategy <- function(ctx, params) {
  targets <- ctx$flat()

  for (id in ctx$universe) {
    x <- ctx$features(id, mapped_features)

    if (
      passed_warmup(x) &&
        x[["ret_5"]] > params$min_return &&
        ctx$close(id) > x[["sma_10"]]
    ) {
      targets[id] <- params$qty
    }
  }

  targets
}
```

Read that as one pulse-time decision:

1. `ctx$features()` reads the mapped feature values for one instrument.
2. `passed_warmup()` keeps the rule inactive until the mapped indicators are
   usable.
3. The condition states the trading idea.
4. The strategy still returns an ordinary target vector.

Plain `features = list(...)` remains valid. Use it when exact IDs are clearest.
Use a feature map when aliases make a feature-heavy strategy easier to read.
The modern `ledgr_experiment(features = ...)` path accepts indicators, lists,
named lists, feature maps, and feature factories. The strategy context then
uses either the exact-ID scalar accessor `ctx$feature()` or the mapped accessor
`ctx$features()`. Lower-level and legacy helpers may be narrower; when in
doubt, prefer the experiment-first workflow.

```{r feature-map-experiment}
mapped_exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = mapped_return_strategy,
  features = mapped_features,
  opening = ledgr_opening(cash = 10000)
)
```

Run it the same way as any other experiment. The strategy still returns target
quantities; the feature map only changes how the strategy reads features.

```{r run-feature-map}
bt_mapped <- mapped_exp |>
  ledgr_run(
    params = list(min_return = 0, qty = 5),
    run_id = "mapped_return"
  )

summary(bt_mapped)
```

Feature-map strategies commonly close over the feature map object. That is the
single-definition pattern: one object defines aliases, registers indicators,
and drives pulse-time lookup.

This has a provenance consequence. Tier 2 is common for strategy functions that
call package helpers or depend on external symbols; the helper-pipeline strategy
below is also tier 2. Feature maps add a more specific concern: the recovered
strategy source may reference `mapped_features`, but the alias map object itself
is not recovered as a standalone object from strategy provenance. The experiment
store records the registered feature definitions, while the human-readable alias
construction remains part of your research code. Keep the feature-map
construction code with the research record when you need to rerun the strategy
later.

## Sweeping Indicator Parameters

When a sweep changes an indicator parameter, ledgr materializes a concrete
feature set for each candidate before execution. Current strategy code has two
safe bridge patterns.

For built-in indicators, exact-ID lookup from `params` is the clearest Tier 1
pattern:

```{r parameterized-feature-id, eval=FALSE}
strategy <- function(ctx, params) {
  fast_id <- ledgr_feature_id(ledgr_ind_sma(params$fast_n))
  slow_id <- ledgr_feature_id(ledgr_ind_sma(params$slow_n))

  values <- c(
    fast = ctx$feature("AAA", fast_id),
    slow = ctx$feature("AAA", slow_id)
  )

  targets <- ctx$flat()
  if (passed_warmup(values) && values[["fast"]] > values[["slow"]]) {
    targets["AAA"] <- params$qty
  }
  targets
}
```

That keeps the strategy self-contained: it uses `ctx`, `params`, and exported
ledgr helpers. Hand-built strings such as `paste0("sma_", params$fast_n)` also
work for built-ins whose ID convention is public, but `ledgr_feature_id()` keeps
the lookup tied to the constructor.

Static feature maps remain the recommended shape when aliases do not vary by
candidate:

```{r static-feature-map-bridge, eval=FALSE}
features <- ledgr_feature_map(
  fast = ledgr_ind_sma(20),
  slow = ledgr_ind_sma(50)
)

strategy <- function(ctx, params) {
  x <- ctx$features("AAA", features)
  ctx$flat()
}
```

Do not call an external feature factory from inside a strategy:

```{r external-feature-factory-tier3, eval=FALSE}
features <- function(params) {
  ledgr_feature_map(
    fast = ledgr_ind_sma(params$fast_n),
    slow = ledgr_ind_sma(params$slow_n)
  )
}

strategy <- function(ctx, params) {
  x <- ctx$features("AAA", features(params))
  ctx$flat()
}
```

That code reads well, but `features` is a user helper function referenced from
the strategy body. Under the current preflight contract, external feature
factory functions inside strategies are Tier 3 and `ledgr_run()` rejects them
before execution. Duplicating the parameterized feature map inside the strategy
is mechanically valid, but it creates a drift risk between the experiment's
feature declaration and the strategy lookup map.

The long-term UX gap is tracked in the active parameterized feature aliases RFC.
Until that API exists, use exact feature IDs for parameterized sweeps and use
feature-map aliases for static maps.

## Run One Backtest

```{r run-one}
bt_top_1 <- exp |>
  ledgr_run(
    params = list(lookback = 5, n = 1, equity_fraction = 0.1),
    run_id = "top_return_1"
  )

summary(bt_top_1)
```

The summary is portfolio-level: total return, max drawdown, and trade count are
computed from the completed run. In ledgr, trades are closed round trips; the
fills table can contain more rows because opening fills and closing fills are
both recorded.

The annualized volatility is high because this toy strategy switches positions
often on a tiny two-instrument demo universe. Treat it as a warning about the
example, not as a property you should expect from the same idea on real data.
The drawdown is disproportionate to the final loss for the same reason: a small,
concentrated portfolio can swing hard during the run even if it ends roughly
flat.

Do not expect a teaching strategy to be good. A weak or unattractive result is
still useful evidence: ledgr records failed ideas with the same care as
successful ones, which is part of not fooling yourself.

Inspecting trades shows the actions produced by the target decisions.

```{r trades}
ledgr_results(bt_top_1, what = "trades")
```

The trade table only includes closed round trips. Small one-share rows appear
when integer sizing and price movement leave a tiny adjustment after a previous
target. Larger rows are the ordinary position exits. `realized_pnl` is the
profit or loss booked when that position closes.

If a run has zero trades, inspect fills before assuming nothing happened:

```{r fills-diagnostic}
ledgr_results(bt_top_1, what = "fills")
```

Zero fills means no execution occurred. Non-empty fills with zero trades means
positions opened but did not close. `n_trades` counts closed round trips, while
the fills table shows both opening and closing execution rows.

## Compare Parameter Variants

Now keep the economic idea fixed and change one assumption: hold the top two
instruments instead of the top one.

```{r run-two}
bt_top_2 <- exp |>
  ledgr_run(
    params = list(lookback = 5, n = 2, equity_fraction = 0.1),
    run_id = "top_return_2"
  )

ledgr_compare_runs(snapshot, run_ids = c("top_return_1", "top_return_2"))
```

Comparison is most useful when the compared runs differ for a reason you can
explain. Here the reason is simple: one run concentrates the allocation in the
single strongest recent-return instrument, and the other splits it across two.

The comparison table is not asking "which number is biggest?" in isolation. It
is asking whether the change in assumption improved the run in ways you can
defend: return, drawdown, number of closed trades, and win rate all matter.

These runs share the same sealed data, initial cash, feature set, and cost
assumptions. That keeps the teaching example narrow. A real comparison would
also ask whether the conclusion survives different samples, execution costs,
starting capital, and parameter choices.

## Compare Against A Baseline

The flat strategy is a sanity baseline, not a market benchmark. It tells you
what the result table looks like when the strategy deliberately does nothing,
and it keeps the comparison honest: if an active strategy cannot beat doing
nothing on the same sealed data, that is valuable information.

```{r baseline}
flat_exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = flat_strategy,
  opening = ledgr_opening(cash = 10000)
)

bt_flat <- flat_exp |>
  ledgr_run(params = list(), run_id = "flat_baseline")

ledgr_compare_runs(snapshot, run_ids = c("top_return_1", "top_return_2", "flat_baseline"))
```

The flat baseline has no win rate because it has no closed trades. That is not
missing data; there are no wins or losses to count.

For real Yahoo snapshots, keep benchmark strategies explicit rather than adding
a hidden benchmark helper. These simple patterns are ordinary ledgr strategies:

```{r real-data-baselines, eval=FALSE}
flat_strategy <- function(ctx, params) {
  ctx$flat()
}

buy_and_hold_strategy <- function(ctx, params) {
  targets <- ctx$hold()
  targets[params$instrument_id] <- params$qty
  targets
}

equal_weight_strategy <- function(ctx, params) {
  targets <- ctx$flat()
  capital <- ctx$equity * params$equity_fraction / length(ctx$universe)
  for (id in ctx$universe) {
    targets[id] <- floor(capital / ctx$close(id))
  }
  targets
}

single_instrument_strategy <- function(ctx, params) {
  targets <- ctx$flat()
  targets[params$instrument_id] <- params$qty
  targets
}
```

Use these as comparison baselines for a sealed real-data snapshot; they are not
benchmark APIs and they do not download benchmark returns.

## Troubleshoot Helper Pipelines

The helper pipeline is only an authoring layer:

```text
ledgr_signal -> ledgr_selection -> ledgr_weights -> ledgr_target -> target vector
```

Only the final target vector is executable. A strategy must return a full named
numeric target vector, or a `ledgr_target` that unwraps to that shape. Returning
a `ledgr_signal`, `ledgr_selection`, `ledgr_weights`, unnamed numeric vector,
data frame, list, or partial target is an invalid strategy result.

Common failures usually mean one of four things:

| Symptom | Likely cause | First check |
|---|---|---|
| unknown feature ID | the indicator was not registered with the experiment | `ledgr_feature_id(features)` |
| missing target names | the strategy did not return every `ctx$universe` instrument | compare `names(target)` to `ctx$universe` |
| non-numeric or unsupported return shape | the strategy returned a helper intermediate or object-like result | make the last line a target vector |
| zero fills or zero trades | warmup, empty selection, sizing to zero, no exit, or last-bar no-fill | inspect a late pulse and the fills table |

For helper strategies, debug one pulse before rerunning the whole experiment:

```{r helper-debug-checklist, eval=FALSE}
pulse <- ledgr_pulse_snapshot(
  snapshot,
  universe = c("DEMO_01", "DEMO_02"),
  ts_utc = ledgr_utc("2019-03-01"),
  features = features
)

signal <- signal_return(pulse, lookback = 5)
selection <- select_top_n(signal, n = 1)
weights <- weight_equal(selection)
target <- target_rebalance(weights, pulse, equity_fraction = 0.1)

signal
selection
weights
target
names(target)
pulse$universe
setdiff(pulse$universe, names(target))
setdiff(names(target), pulse$universe)
```

If `selection` inherits from `ledgr_empty_selection`, every signal value was
missing or unusable at that pulse. Early in a run this is usually ordinary
warmup. Late in a run it points to sample length, feature registration, or
universe coverage. If `target` is full-universe but every quantity is zero,
check integer flooring, `equity_fraction`, current close prices, and whether an
empty selection flowed through intentionally.

If `setdiff(pulse$universe, names(target))` is non-empty, the strategy would
fail target validation because it did not name every instrument. If
`setdiff(names(target), pulse$universe)` is non-empty, it emitted targets for
unknown instruments.

Reusable file-level helper functions are Tier 3 unless the function can be
represented as part of the strategy artifact. Inline small allocation logic or
wrap it in an approved helper surface:

```{r helper-sharing-tier3, eval=FALSE}
allocate_equal <- function(ctx, fraction) {
  targets <- ctx$flat()
  capital <- ctx$equity * fraction / length(ctx$universe)
  for (id in ctx$universe) targets[id] <- floor(capital / ctx$close(id))
  targets
}

bad_strategy <- function(ctx, params) {
  allocate_equal(ctx, params$equity_fraction)
}

ledgr_strategy_preflight(bad_strategy)

self_contained_strategy <- function(ctx, params) {
  targets <- ctx$flat()
  capital <- ctx$equity * params$equity_fraction / length(ctx$universe)
  for (id in ctx$universe) targets[id] <- floor(capital / ctx$close(id))
  targets
}
```

`ledgr_signal_strategy()` is a separate explicit compatibility wrapper for
tutorial-style signal functions. It maps an inner signal function to full target
quantities using configured long/flat/short quantities. Use it when that mapping
is really the strategy contract; otherwise prefer returning the full target
vector directly. Strategy preflight reports this wrapper as a special allowed
case because it captures an inner function and quantity settings. Tier 3
preflight failures remain hard failures: they are not downgraded to warnings or
silently accepted. Wall-clock calls such as `Sys.time()` and global assignment
with `<<-` are Tier 3 because ledgr cannot make them reproducible execution
inputs. Resolved scalar values from the strategy closure are allowed as Tier 2,
but values that define the research question should usually live in `params`.

A compact Tier 3 hard-failure example is an unresolved helper reference:

```{r tier3-hard-failure, eval=FALSE}
outside_helper <- function(ctx) ctx$flat()

tier3_strategy <- function(ctx, params) {
  outside_helper(ctx)
}

preflight <- ledgr_strategy_preflight(tier3_strategy)
preflight$tier
preflight$reason

ledgr_run(exp, params = list(), run_id = "tier3_demo")
```

`ledgr_run()` and `ledgr_sweep()` reject Tier 3 strategies before execution.
There is no force override on those public execution paths.

## When ledgr Complains

ledgr tries to fail loudly when an error would make a backtest misleading. If a
strategy returns a vector with the wrong names or length, ledgr rejects it
instead of silently treating missing instruments as zero. If a helper reads an
unregistered feature, `ctx$feature()` reports the unknown feature ID and lists
the available IDs. If `target_rebalance()` receives negative or over-allocated
weights, it fails before turning them into target quantities.

Those errors are part of the design. They are meant to catch research mistakes
while the mistake is still small enough to understand.

## Inspect Stored Source

The experiment store lets you come back later and know exactly what code and
parameters produced a result. When you return to a six-month-old run, this is
the artifact that tells you what was actually tested, not what you remember
testing.

ledgr stores strategy provenance with the run. Inspection is read-only by
default: `trust = FALSE` returns stored text and metadata without parsing or
evaluating it.

Hash verification proves stored-text identity, not safety. That is why the
read-only path is the default.

You usually do not read the hashes yourself. They are the durable fingerprints
ledgr uses to verify that recovered source and parameters are identical to the
artifacts that produced the run.

```{r extract}
ledgr_extract_strategy(snapshot, "top_return_1", trust = FALSE)
```

Use `trust = TRUE` only when you explicitly trust the experiment store and want
to recover a function object.

## Cleanup

These calls release DuckDB connections. Forgetting them in a short interactive
session is not a data-safety issue; completed run artifacts are already durable
when `ledgr_run()` returns. In long sessions and scripts, releasing resources
explicitly avoids lock and resource warnings.

```{r cleanup}
close(bt_top_1)
close(bt_top_2)
close(bt_flat)
close(bt_mapped)
ledgr_snapshot_close(snapshot)
```

## What's Next?

If you want the formal contract, read the strategy and context sections in
`inst/design/contracts.md`. If you want the indicator story, read
`vignette("indicators", package = "ledgr")`. For formal feature-map help, see
`?ledgr_strategy_context`, `?ledgr_feature_map`, and `?passed_warmup`. If you
want the deployment story, continue with `vignette("research-to-production",
package = "ledgr")`. If you want to inspect or compare durable runs, read
`vignette("experiment-store", package = "ledgr")`.
