---
title: "Strategy Development"
format:
  html:
    toc: false
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{Strategy Development}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 90)
options(cli.unicode = FALSE)
# Local renders load ledgr from source; installed builds use the library call
# in the prerequisites chunk.
pkg_root <- if (file.exists("DESCRIPTION")) "." else if (file.exists("../DESCRIPTION")) ".." else NA_character_
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
default_output_hook <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(
  output = function(x, options) {
    default_output_hook(gsub("[ \t]+(?=\n)", "", x, perl = TRUE), options)
  }
)
demo_data_path <- if (file.exists("data/ledgr_demo_bars.rda")) {
  "data/ledgr_demo_bars.rda"
} else {
  file.path("..", "data", "ledgr_demo_bars.rda")
}
if (!exists("ledgr_demo_bars") && file.exists(demo_data_path)) {
  load(demo_data_path)
}
```

```{=html}
<style>
.ledgr-diagram {
  margin: 1.25rem auto 1.5rem auto;
  text-align: center;
}
.ledgr-diagram .mermaid {
  display: inline-block;
  max-width: 760px;
  width: 100%;
}
.ledgr-diagram .node text,
.ledgr-diagram .edgeLabel {
  font-size: 18px !important;
}
</style>
```

You write a ledgr strategy as a policy. At each decision pulse, the strategy
receives `ctx`, reads only pulse-known information, and returns target holdings.
This article teaches that contract, then shows how helper objects and feature
maps make larger strategies easier to read without changing the target-vector
boundary.

## Prerequisites

The examples use `dplyr` for demo-data preparation. Strategy functions use
ledgr's pulse context rather than data-frame operations. The article assumes
basic familiarity with sealed snapshots (`vignette("experiment-store",
package = "ledgr")`) and feature IDs (`vignette("indicators", package =
"ledgr")`).

```{r}
#| label: library
#| message: false

library(ledgr)
library(dplyr)
data("ledgr_demo_bars", package = "ledgr")
```

This article moves in three steps:

1. learn the raw strategy contract: `function(ctx, params) -> target vector`;
2. inspect pulse-known data and registered features;
3. use helper objects to express larger strategies while still returning target
   holdings.

A backtest in ledgr is a sequence of decision moments. At each pulse, ledgr
shows the strategy only what could have been known at that time. The strategy
answers with desired holdings. ledgr records the decision, applies the fill
model, and moves to the next pulse.

Sweep execution can opt into the scoped spot-FIFO accelerator with
`ledgr_sweep(..., compiled_accounting_model = "spot_fifo")`. This changes the
memory-backed accounting hot frame only; the strategy still receives the same
`ctx`, returns the same full named target vector, and `NULL` remains the
canonical R default. Committed `ledgr_run()` artifacts keep the durable R path
until a separate durable compiled-integration gate lands.

This matters because leakage is easy. If future information enters a historical
decision, the backtest can look profitable for the wrong reason. ledgr's
strategy interface is built to make one common mistake harder: your strategy
receives one pulse context, not the whole future. For the broader leakage model,
including feature-construction leakage and remaining user responsibilities, see
`vignette("leakage", package = "ledgr")`.

## Wrong And Right: Leakage

The tempting vectorized pattern is to compute a future-looking column first and
then trade from it. In the example below, `lead(close)` shifts tomorrow's close
onto today's row. The resulting `buy_signal` looks like an ordinary column, but
it answers a question the strategy could not have answered at today's decision
time: "will tomorrow's close be higher than today's close?" Trading from that
column lets the backtest use future market data as if it were already known.

```{r}
#| label: leakage-wrong
#| eval: false

leaky_signals <- ledgr_demo_bars |>
  group_by(instrument_id) |>
  arrange(ts_utc, .by_group = TRUE) |>
  mutate(
    tomorrow_close = lead(close),
    buy_signal = tomorrow_close > close
  )
```

The ledgr version expresses the rule at one pulse. The strategy can read the
current bar for the current instrument. Later sections add registered features
to the same pulse model. The strategy has no market-data table from which it can
casually index tomorrow's bar. That is the same information shape a live trading
strategy gets as time passes: each pulse is a new slice of the knowable
universe.

```{r}
#| label: leakage-right

no_leak_bar_strategy <- function(ctx, params) {
  targets <- ctx$flat()

  for (id in ctx$universe) {
    if (ctx$close(id) > ctx$open(id)) {
      targets[id] <- 1
    }
  }

  targets
}
```

This removes one common source of leakage, but it does not certify that
snapshots, feature definitions, event timestamps, universe construction, or
parameter selection are causally clean.

With that boundary in mind, start with the simplest possible strategy.

## What Is `ctx`?

::: {.ledgr-callout .ledgr-callout-note}
**Definition**

`ctx` is the pulse context: the information packet ledgr gives your strategy at
one decision time. It contains pulse-known bars, features, positions, cash, and
equity. It is deliberately not the full future dataset.
:::

It contains the current timestamp, current bars, current features, current
positions, cash, equity, and small helper functions for accessing those values.
It is deliberately not the full dataset.

| Expression | Meaning at one pulse |
|---|---|
| `ctx$ts_utc` | current decision timestamp |
| `ctx$universe` | instruments in the run |
| `ctx$idx(id)` | 1-based universe position for one instrument |
| `ctx$open(id)`, `ctx$close(id)` | current bar values for one instrument |
| `ctx$vec$close` | current close values for the full universe |
| `ctx$feature(id, feature_id)` | current indicator value for one instrument by engine feature ID |
| `ctx$vec$feature(feature_id)` | current indicator values for the full universe by engine feature ID |
| `ctx$features(id, feature_map)` | mapped indicator values for one instrument by alias, using the supplied feature map |
| `ctx$position(id)` | current simulated position |
| `ctx$vec$positions` | current simulated positions aligned to `ctx$universe` |
| `ctx$cash`, `ctx$equity` | current simulated portfolio state |
| `ctx$flat()` | target zero positions unless changed |
| `ctx$hold()` | target current positions unless changed |

For the installed accessor reference, see `?ledgr_strategy_context`.

The pulse loop is the contract in motion:

::: {.ledgr-diagram .ledgr-pulse-loop}

```{mermaid}
%%| echo: false

flowchart TB
  state_t["pulse t state<br/>bars through t<br/>positions, cash, equity"]
  ctx_node["ctx<br/>pulse-known projection"]
  strategy_node["strategy(ctx, params)"]
  target_node["target vector<br/>desired holdings"]
  fill_node["next-open fill<br/>at t + 1"]
  state_next["pulse t + 1 state<br/>ledger updated"]

  state_t --> ctx_node --> strategy_node --> target_node --> fill_node --> state_next
  state_next -. next pulse .-> state_t
```

:::

`ctx` is the no-lookahead handoff. It gives the strategy the current pulse
projection, the strategy returns a target vector, and ledgr handles validation,
fill timing, ledger events, and the next pulse.

The two target starters have different economic meanings:

| Helper | Starts from | Economic meaning |
|---|---|---|
| `ctx$flat()` | zero positions | only hold what this pulse explicitly selects |
| `ctx$hold()` | current positions | keep existing positions unless changed |

For example, this policy starts from current holdings and only changes the book
when it sees an exit reason. Economically, it means: "keep what I already own,
unless today's bar gives me a reason to leave."

```{r}
#| label: hold-example

hold_unless_down <- function(ctx, params) {
  targets <- ctx$hold()

  for (id in ctx$universe) {
    if (ctx$close(id) < ctx$open(id)) {
      targets[id] <- 0
    }
  }

  targets
}
```

This loop style is fine while the mechanics are still visible. Once you have
helpers like `signal_*()` and `select_*()`, most strategy logic is easier to
express at the whole-universe level instead of one instrument at a time.

## A Strategy That Does Nothing

The simplest economic policy is: hold cash and own no instruments.

```{r}
#| label: flat-strategy

flat_strategy <- function(ctx, params) {
  ctx$flat()
}
```

`ctx$flat()` creates a full target vector with one entry for every instrument in
the run and every value set to zero. Economically, this means: after the next
fill opportunity, hold no positions.

This is a complete ledgr strategy: useful for understanding the contract, not
for making money.

The return value is a named numeric vector. Names are instrument IDs from
`ctx$universe`, values are desired quantities. `ctx$flat()` produces the
full-universe shape with every entry at zero.

::: {.ledgr-callout .ledgr-callout-note}
**Definition**

A target vector is the strategy's requested holdings for the full universe at
one pulse. It is named by instrument ID, numeric, and complete. It is not an
order list, a signal table, or a partial update.
:::

The deeper mental model is that a strategy is a **policy**, not a sequence of
orders. At each pulse, it declares a desired state: "I want to hold this many
shares of each instrument." The engine compares that against current holdings
and fills the gap.

That distinction keeps strategies free from execution-state bookkeeping.

::: {.ledgr-callout .ledgr-callout-warning}
**Affordability is not automatic**

Raw target vectors are desired holdings. ledgr does not check affordability
before filling them; if a target requires more cash than the simulated portfolio
has, the run can fill anyway and cash can go negative. Use
`ledgr_target_rebalance(equity_fraction = ...)` or size directly from `ctx$cash` and
`ctx$equity` when you need capital-aware targets. A `risk_chain` can transform
validated targets before fill timing and cost resolution -- for example
`ledgr_risk_long_only()` can clip short targets and `ledgr_risk_max_weight()`
can cap per-instrument target exposure. It is not a cash-affordability,
margin, liquidity, or broker-risk engine.
:::

## A First Trading Rule

Now add one small economic idea:

> If an instrument closes above its open, own one share. Otherwise own nothing.

This is still a teaching strategy, not investment advice. It shows how
observable data becomes a target.

```{r}
#| label: buy-if-up

buy_if_up <- function(ctx, params) {
  targets <- ctx$flat()

  for (id in ctx$universe) {
    if (ctx$close(id) > ctx$open(id)) {
      targets[id] <- 1
    }
  }

  targets
}
```

Targets are desired quantities, not orders, signals, or portfolio weights. A
target of `1` means "after the next fill opportunity, hold one share." A target
of `0` means "hold no shares."

In these examples, decisions fill at the next open: a decision made at pulse
`t` fills at the next available bar. That keeps the strategy from deciding and
filling on the same close.

The loop is intentionally plain because this is the first example. Once the
economic idea is clear, ledgr strategies are usually easier to read when they
use helper functions that operate on the whole universe at once. The later
sections make that transition.

::: {.ledgr-callout .ledgr-callout-tip}
**Try it**

Change `buy_if_up()` so it starts from `ctx$hold()` instead of `ctx$flat()`.
Which positions would persist after a down bar, and why does that change the
economic meaning of the strategy?
:::

## Why `params` Exists

::: {.ledgr-callout .ledgr-callout-note}
**Definition**

`params` is the run's strategy configuration. Put research choices you want to
compare, store, or sweep into `params`; do not hide them in globals or inside
feature declarations.
:::

Hard-coded constants make experiments awkward. Parameters let one economic idea
run under different assumptions.

```{r}
#| label: buy-if-up-param

buy_if_up_qty <- function(ctx, params) {
  targets <- ctx$flat()

  for (id in ctx$universe) {
    if (ctx$close(id) > ctx$open(id)) {
      targets[id] <- params$qty
    }
  }

  targets
}
```

Strategies use `function(ctx, params)`. `ctx` is the pulse. `params` is the
experimenter's chosen configuration for this run. Keeping them separate makes
the strategy easier to test, compare, and store.

## Prepare A Small Experiment

Use two instruments from the offline demo data so the examples run anywhere.

```{r}
#| label: data

bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% c("DEMO_01", "DEMO_02"),
    between(
      ts_utc,
      ledgr_utc("2019-01-01"),
      ledgr_utc("2019-06-30")
    )
  )

snapshot <- ledgr_snapshot_from_df(
  bars,
  snapshot_id = "strategy_chapter_snapshot"
)
```

The snapshot seals the market data. That is the evidence base for the
experiment. Strategies and indicators can derive from it, but the underlying
bars do not change mid-research.

## Indicators And Feature IDs

Indicators are feature definitions. Before a strategy uses a feature, ask ledgr
for the exact ID.

```{r}
#| label: feature-ids

features <- list(ledgr_ind_returns(5))

ledgr_feature_id(features)
```

Those strings are the names used inside `ctx$feature()`. They are exact. A typo
such as `"returns_5"` is not treated as a warmup value; it is an unknown feature
and ledgr fails loudly.

Warmup is different. A known feature can be `NA` early in the sample because
there are not enough prior bars yet. Strategy code should treat that as "no
signal yet."

## Debug One Pulse Before Running

Before running a full backtest, inspect one pulse. This is the fastest way to
understand what your strategy will see.

```{r}
#| label: pulse

pulse <- ledgr_pulse_snapshot(
  snapshot,
  universe = c("DEMO_01", "DEMO_02"),
  ts_utc = ledgr_utc("2019-03-01"),
  features = features
)

pulse$ts_utc
pulse$universe
pulse$close("DEMO_01")
pulse$feature("DEMO_01", "return_5")
pulse$hold()
```

The scalar accessors are easiest when you are writing or debugging a rule for
one instrument. Cross-sectional rules should usually switch to the vector
accessors once the contract is clear:

```{r}
#| label: pulse-vector-accessors

pulse$idx("DEMO_01")
pulse$vec$close
pulse$vec$feature("return_5")
```

`ctx$idx(id)` gives the instrument's position in `ctx$universe`. Values in
`ctx$vec$close`, `ctx$vec$positions`, and `ctx$vec$feature(feature_id)` use that
same order, so a strategy can score the whole universe without repeating scalar
lookups. By default, `ctx$idx(id)` also fails loudly for an unknown instrument;
use its `missing` argument only when a deliberate `NA` path is part of the rule.
The vector feature accessor still uses exact engine feature IDs: warmup for a
known feature is `NA`, while an unknown feature ID fails loudly.

The same pulse can also be viewed as one wide row. This is useful when you want
to see prices, portfolio state, and computed features together.

```{r}
#| label: pulse-wide

ledgr_pulse_wide(pulse) |>
  glimpse()
```

The wide row and the scalar accessors are two ways of looking at the same
pulse-known data. The wide row is good for inspection and model-like thinking.
The rest of this vignette uses the non-wide accessors because they keep the
step-by-step strategy logic easier to read.

Raw loops are the clearest way to learn the contract. Once that contract is
clear, larger strategies usually read better as a pipeline: score the universe,
select names, assign weights, then convert those weights into target
quantities.

The economic idea:

> Rank instruments by recent return, keep the top names, split capital equally,
> and convert those weights into share quantities.

`ledgr_signal_return()` is a thin helper around the same feature you inspected above:
it reads `return_N` for every instrument in the pulse and returns one
universe-wide signal object. It uses the vector accessor
`ctx$vec$feature(feature_id)` when available, then falls back to the scalar
`ctx$feature(id, feature_id)` path for compatibility.

The helper pipeline has four stages:

| Stage | Input | Output | Question answered |
|---|---|---|---|
| signal | pulse context | numeric scores with origin metadata | What looks attractive? |
| selection | signal | logical inclusion with the same origin | What should be considered? |
| weights | selection | allocation weights with the same origin | How should capital be split? |
| target | weights and context | full-universe share quantities | What should the portfolio hold? |

Execution semantics begin only at the target stage. `signal`, `selection`, and
`weights` are research objects that help author the strategy; `target` is the
ordinary full named target vector shape the runner validates and executes.

```{r}
#| label: pulse-pipeline

signal <- ledgr_signal_return(pulse, lookback = 5)
signal

selection <- ledgr_select_top_n(signal, n = 1)
selection

weights <- ledgr_weight_equal(selection)
weights

target <- ledgr_target_rebalance(weights, pulse, equity_fraction = 0.1)
target
```

`ledgr_target_rebalance()` sizes with current pulse equity and current close prices,
using `ctx$vec$close` when available, then floors to whole shares. For the
selected `DEMO_01` pulse above, 10% of equity is allocated to the one selected
instrument:

```{r}
#| label: floor-sizing

raw_qty <- weights[["DEMO_01"]] * 0.1 * pulse$equity / pulse$close("DEMO_01")
c(pre_floor = raw_qty, target_qty = unclass(target)[["DEMO_01"]])
```

The general weighted sizing formula is:

```text
floor(weight * equity_fraction * ctx$equity / ctx$close(instrument_id))
```

For a raw target strategy that does not use weights, the same idea reduces to:

```text
floor(equity_fraction * ctx$equity / ctx$close(instrument_id))
```

Both formulas use decision-time close and current pulse equity. Fills still
occur at the configured later fill point, so fill value can drift from
decision-time sizing. Residual allocation after whole-share flooring remains
cash and is reflected in the ledger-backed equity rows.

## Turn The Idea Into A Strategy

The same transformations become an ordinary strategy function.

The full backtest replays every bar, including the earliest warmup pulses.
During those first pulses, `return_5` is `NA` for every instrument because five
prior bars do not exist yet. `ledgr_select_top_n()` treats that all-missing signal as
a classed empty selection, not as a warning. That object still carries the
original universe and signal origin. `ledgr_weight_equal()` turns it into empty
weights, and `ledgr_target_rebalance()` turns those weights into a flat
full-universe target.

No warning suppression is needed for ordinary early warmup. A partial-selection
warning can still appear when some signal values are usable but fewer than `n`
instruments can be selected. If a run finishes with zero trades, inspect a late
pulse before assuming the empty selection was only early warmup.

```{r}
#| label: helper-strategy

top_return_strategy <- function(ctx, params) {
  signal <- ledgr_signal_return(ctx, lookback = params$lookback)
  selection <- ledgr_select_top_n(signal, n = params$n)

  weights <- ledgr_weight_equal(selection)
  ledgr_target_rebalance(weights, ctx, equity_fraction = params$equity_fraction)
}
```

Read it economically:

1. `ledgr_signal_return()` scores each instrument by recent return.
2. `ledgr_select_top_n()` keeps the highest scores and ignores warmup `NA`.
3. `ledgr_weight_equal()` splits the chosen allocation equally.
4. `ledgr_target_rebalance()` converts weights into floored full target quantities.

No helper registers indicators automatically. The experiment must say which
features exist.

Empty selections flow through the pipeline as objects, so expected warmup and
"no signal today" look the same to your strategy. Diagnostics still belong at
the pulse level: when a strategy produces no fills or no closed trades, inspect
a late pulse and confirm whether the feature values are usable.

```{r}
#| label: experiment

exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = top_return_strategy,
  features = features,
  opening = ledgr_opening(cash = 10000),
  cost_model = ledgr_cost_zero()
)
```

## Feature Maps For Readable Feature Access

The examples above keep the exact feature ID contract visible:
`ctx$feature(id, feature_id)` reads one registered feature for one instrument at
one pulse. That contract remains the foundation.

For cross-sectional strategies, `ctx$vec$feature(feature_id)` returns the same
feature at the same pulse for every instrument in `ctx$universe`. Warmup for a
known feature remains `NA`; an unknown feature ID fails loudly. The scalar
helper stays the clearest teaching surface, while the vector helper is the
lower-overhead surface for universe-wide scoring.

When a strategy reads several features per instrument, repeating feature ID
strings can obscure the trading idea. A feature map bundles indicator objects
with strategy-facing aliases. The same object can be registered with the
experiment and used by the strategy for pulse-time lookup.

```{r}
#| label: feature-map-setup

mapped_features <- ledgr_feature_map(
  ret_5 = ledgr_ind_returns(5),
  sma_10 = ledgr_ind_sma(10)
)

ledgr_feature_id(mapped_features)
```

The strategy closes over `mapped_features`. Inside the universe loop,
`ctx$features(id, mapped_features)` returns a named numeric vector keyed by the
aliases. `ledgr_passed_warmup()` is a guard for that vector: for values returned by
`ctx$features()`, it means every requested indicator is usable at this pulse.
It is not a signal pipeline transformation, and it is not a data-quality
diagnostic for arbitrary vectors.

```{r}
#| label: feature-map-strategy

mapped_return_strategy <- function(ctx, params) {
  targets <- ctx$flat()

  for (id in ctx$universe) {
    x <- ctx$features(id, mapped_features)

    if (
      ledgr_passed_warmup(x) &&
        x[["ret_5"]] > params$min_return &&
        ctx$close(id) > x[["sma_10"]]
    ) {
      targets[id] <- params$qty
    }
  }

  targets
}
```

Read that as one pulse-time decision:

1. `ctx$features()` reads the mapped feature values for one instrument.
2. `ledgr_passed_warmup()` keeps the rule inactive until the mapped indicators are
   usable.
3. The condition states the trading idea.
4. The strategy still returns an ordinary target vector.

Plain `features = list(...)` remains valid. Use it when exact IDs are clearest.
Use a feature map when aliases make a feature-heavy strategy easier to read.
`ledgr_experiment(features = ...)` accepts indicators, lists, named lists, and
feature maps. The strategy context then uses either the exact-ID scalar accessor
`ctx$feature()`, the exact-ID vector accessor `ctx$vec$feature()`, or the
mapped accessor `ctx$features()`. When in doubt, prefer the experiment-first
workflow.

```{r}
#| label: feature-map-experiment

mapped_exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = mapped_return_strategy,
  features = mapped_features,
  opening = ledgr_opening(cash = 10000),
  cost_model = ledgr_cost_zero()
)
```

Run it the same way as any other experiment. The strategy still returns target
quantities; the feature map only changes how the strategy reads features.

```{r}
#| label: run-feature-map

bt_mapped <- mapped_exp |>
  ledgr_run(
    params = list(min_return = 0, qty = 5),
    run_id = "mapped_return"
  )

summary(bt_mapped)
```

Feature-map strategies commonly close over the feature map object. Keep that
construction code with the research record. The experiment store records the
registered feature definitions, but recovered strategy source may still
reference the original alias-map object by name.

## Keep Feature Declaration Outside Strategy

Do not declare or rebuild features inside a strategy:

```r
strategy <- function(ctx, params) {
  features <- ledgr_feature_map(
    fast = ledgr_ind_sma(params$fast_n),
    slow = ledgr_ind_sma(params$slow_n)
  )
  x <- ctx$features("AAA", features)
  ctx$flat()
}
```

That code puts feature declaration inside the execution loop. Strategy code
should read pulse-known values from the context; experiment code owns feature
declaration. Duplicating a parameterized feature map inside the strategy creates
a drift risk between the experiment's feature declaration and the strategy
lookup map.

For exploratory sweeps over indicator parameters, use active aliases and
feature grids. The canonical walkthrough is
`vignette("sweeps", package = "ledgr")`.

## Run One Backtest

```{r}
#| label: run-one

bt_top_1 <- exp |>
  ledgr_run(
    params = list(lookback = 5, n = 1, equity_fraction = 0.1),
    run_id = "top_return_1"
  )

summary(bt_top_1)
```

The summary is portfolio-level: total return, max drawdown, and trade count are
computed from the completed run. In ledgr, trades are closed round trips; the
fills table can contain more rows because opening fills and closing fills are
both recorded.

The annualized volatility is high because this toy strategy switches positions
often on a tiny two-instrument demo universe. Treat it as a warning about the
example, not as a property you should expect from the same idea on real data.
The drawdown is disproportionate to the final loss for the same reason: a small,
concentrated portfolio can swing hard during the run even if it ends roughly
flat.

Do not expect a teaching strategy to be good. A weak or unattractive result is
still useful evidence: ledgr records failed ideas with the same care as
successful ones, which is part of not fooling yourself.

Inspecting trades shows the actions produced by the target decisions.

```{r}
#| label: trades

ledgr_results(bt_top_1, what = "trades")
```

The trade table only includes closed round trips. Small one-share rows appear
when integer sizing and price movement leave a tiny adjustment after a previous
target. Larger rows are the ordinary position exits. `realized_pnl` is the
profit or loss booked when that position closes.

If a run has zero trades, inspect fills before assuming nothing happened:

```{r}
#| label: fills-diagnostic

ledgr_results(bt_top_1, what = "fills")
```

Zero fills means no execution occurred. Non-empty fills with zero trades means
positions opened but did not close. `n_trades` counts closed round trips, while
the fills table shows both opening and closing execution rows.

If you want to compare variants, keep the strategy authoring question separate
from the research-comparison question. Use
`vignette("experiment-store", package = "ledgr")` for stored-run comparison and
`vignette("research-workflow", package = "ledgr")` for promotion and review.

## Troubleshoot Helper Pipelines

The helper pipeline is only an authoring layer:

::: {.ledgr-diagram .ledgr-helper-pipeline}

```{mermaid}
%%| echo: false

flowchart LR
  signal["ledgr_signal"]
  selection["ledgr_selection"]
  weights["ledgr_weights"]
  target_obj["ledgr_target"]
  target_vec["target vector"]

  signal --> selection --> weights --> target_obj --> target_vec
```

:::

Only the final target vector is executable. A strategy must return a full named
numeric target vector, or a `ledgr_target` that unwraps to that shape. Returning
a `ledgr_signal`, `ledgr_selection`, `ledgr_weights`, unnamed numeric vector,
data frame, list, or partial target is an invalid strategy result.

Common failures usually mean one of four things:

| Symptom | Likely cause | First check |
|---|---|---|
| unknown feature ID | the indicator was not registered with the experiment | `ledgr_feature_id(features)` |
| missing target names | the strategy did not return every `ctx$universe` instrument | compare `names(target)` to `ctx$universe` |
| non-numeric or unsupported return shape | the strategy returned a helper intermediate or object-like result | make the last line a target vector |
| zero fills or zero trades | warmup, empty selection, sizing to zero, no exit, or last-bar no-fill | inspect a late pulse and the fills table |

For helper strategies, debug one pulse before rerunning the whole experiment:

```{r}
#| label: helper-debug-checklist
#| eval: false

pulse <- ledgr_pulse_snapshot(
  snapshot,
  universe = c("DEMO_01", "DEMO_02"),
  ts_utc = ledgr_utc("2019-03-01"),
  features = features
)

signal <- ledgr_signal_return(pulse, lookback = 5)
selection <- ledgr_select_top_n(signal, n = 1)
weights <- ledgr_weight_equal(selection)
target <- ledgr_target_rebalance(weights, pulse, equity_fraction = 0.1)

signal
selection
weights
target
names(target)
pulse$universe
setdiff(pulse$universe, names(target))
setdiff(names(target), pulse$universe)
```

If `selection` inherits from `ledgr_empty_selection`, every signal value was
missing or unusable at that pulse. Early in a run this is usually ordinary
warmup. Late in a run it points to sample length, feature registration, or
universe coverage. If `target` is full-universe but every quantity is zero,
check integer flooring, `equity_fraction`, current close prices, and whether an
empty selection flowed through intentionally.

If `setdiff(pulse$universe, names(target))` is non-empty, the strategy would
fail target validation because it did not name every instrument. If
`setdiff(names(target), pulse$universe)` is non-empty, it emitted targets for
unknown instruments.

## Preflight Catches Non-Reproducible Strategy Code

Strategy functions are preflighted before execution. Keep strategy logic
self-contained, put research variation in `params`, and avoid hidden session
state such as unresolved helper functions or mutable globals.

`ledgr_signal_strategy()` is a separate compatibility wrapper for tutorial-style
signal functions. It explicitly maps an inner signal function to target
quantities. For the full tier model, read `vignette("reproducibility", package
= "ledgr")`.

::: {.ledgr-callout .ledgr-callout-note}
**Definition**

A preflight tier is ledgr's static reproducibility classification for a strategy
function. Tier 1 is self-contained, Tier 2 is inspectable with user-managed
environment parity, and Tier 3 is rejected before execution.
:::

A compact Tier 3 hard-failure example is an unresolved helper reference:

```{r}
#| label: tier3-hard-failure

tier3_strategy <- function(ctx, params) {
  outside_helper(ctx)
}

preflight <- ledgr_strategy_preflight(tier3_strategy)
preflight$tier
preflight$reason
```

`ledgr_run()` and `ledgr_sweep()` reject Tier 3 strategies before execution.
There is no force override on those public execution paths.

## When ledgr Complains

ledgr tries to fail loudly when an error would make a backtest misleading. If a
strategy returns a vector with the wrong names or length, ledgr rejects it
instead of silently treating missing instruments as zero. If a helper reads an
unregistered feature, `ctx$feature()` reports the unknown feature ID and lists
the available IDs. If `ledgr_target_rebalance()` receives negative or over-allocated
weights, it fails before turning them into target quantities.

Those errors are part of the design. They are meant to catch research mistakes
while the mistake is still small enough to understand.

## Stored Source

ledgr stores strategy provenance with committed runs. For source inspection,
hash verification, and trust boundaries, read
`vignette("reproducibility", package = "ledgr")`.

## Cleanup

These calls release DuckDB connections. Forgetting them in a short interactive
session is not a data-safety issue; completed run artifacts are already durable
when `ledgr_run()` returns. In long sessions and scripts, releasing resources
explicitly avoids lock and resource warnings.

```{r}
#| label: cleanup

close(bt_top_1)
close(bt_mapped)
close(pulse)
ledgr_snapshot_close(snapshot)
```

## Where Next

- For feature declarations, aliases, and warmup inspection, read
  `vignette("indicators", package = "ledgr")`.
- For the no-lookahead target and fill timing contract, read
  `vignette("execution-semantics", package = "ledgr")`.
- For durable run inspection and comparison, read
  `vignette("experiment-store", package = "ledgr")`.
- For function-level context details, see `?ledgr_strategy_context`,
  `?ledgr_feature_map`, and `?ledgr_passed_warmup`.
