---
title: "Strategy Authoring Tools"
format:
  html:
    toc: false
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{Strategy Authoring Tools}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 90)
options(cli.unicode = FALSE)
# Local renders load ledgr from source; installed builds use the library call
# in the prerequisites chunk.
pkg_root <- if (file.exists("DESCRIPTION")) {
  "."
} else if (file.exists("../DESCRIPTION")) {
  ".."
} else {
  NA_character_
}
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
default_output_hook <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(
  output = function(x, options) {
    default_output_hook(gsub("[ \t]+(?=\n)", "", x, perl = TRUE), options)
  }
)
demo_data_path <- if (file.exists("data/ledgr_demo_bars.rda")) {
  "data/ledgr_demo_bars.rda"
} else {
  file.path("..", "data", "ledgr_demo_bars.rda")
}
if (!exists("ledgr_demo_bars") && file.exists(demo_data_path)) {
  load(demo_data_path)
}
```

After the raw `function(ctx, params) -> target vector` contract is clear,
strategy work shifts to composing decisions you can inspect: read current
state, derive weights, turn them into complete intent, and debug one pulse
before running. This companion article teaches that workflow together with
feature aliases. For the first-pass strategy contract and leakage boundary,
read
`vignette("strategy-development", package = "ledgr")`.

## Choose An Authoring Path

At each pulse, strategy code reads current state, makes a decision from
pulse-known inputs, and returns complete portfolio intent. Two paths cover most
strategies:

| Path | Use it when | Authoring result | Executable result |
|---|---|---|---|
| derive weights, then rebalance | this pulse constructs a member allocation | relative weights over current investment members | complete quantities from `ledgr_target_rebalance()` |
| hold, then edit | existing quantities should remain unless this pulse changes them | a copy from `ctx$hold()` with named edits | the edited complete target |

The **decision axis** is `ctx$universe`: every instrument for which the
strategy must return intent. **Current investment members** are the instruments
eligible to enter selection and weighting. In an availability-aware run, the
decision axis can also include a held nonmember needed for valuation or
disposition. It stays out of ranking and weighting. Neither membership nor
`ctx$tradable()` promises execution at the next open.

Both paths finish with one named quantity for every instrument on the decision
axis. The introductory target-vector contract and the difference between
`ctx$flat()` and `ctx$hold()` live in
`vignette("strategy-development", package = "ledgr")`.

::: {.ledgr-callout .ledgr-callout-warning}
**Reserve a kept member before sizing the rest**

Do not size against the full budget and then restore a current member's
position. If two `DEMO_02` shares consume 40% of NAV, allocate only the
remaining 60%, then restore that named quantity:

```r
targets <- ledgr_target_rebalance(weights, ctx, equity_fraction = 0.6)
targets[["DEMO_02"]] <- ctx$position("DEMO_02")
```

The value `0.6` is the strategy's explicit budget decision, not a magic
default. Held *nonmembers* are reserved automatically; kept current members
are not. With NAV 100, two shares marked at 20 consume 40 and leave 60 for the
rest of the book. At a price of 10, weight 1 over that residual budget targets
six shares; weight 0.6 targets three. A strategy intending 60% of total NAV in
the new allocation therefore uses residual weight 1, not 0.6.
:::

## Prerequisites

The examples use `dplyr` for demo-data preparation. Strategy functions use
ledgr's pulse context rather than data-frame operations. The article assumes
basic familiarity with sealed snapshots (`vignette("data-input-and-snapshots",
package = "ledgr")`) and feature IDs (`vignette("indicators", package =
"ledgr")`).

```{r}
#| label: library
#| message: false

library(ledgr)
library(dplyr)
library(tibble)
data("ledgr_demo_bars", package = "ledgr")
```

## Prepare A Small Experiment

Use two instruments from the offline demo data so the examples run anywhere.

```{r}
#| label: data

bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% c("DEMO_01", "DEMO_02"),
    between(
      ts_utc,
      ledgr_utc("2019-01-01"),
      ledgr_utc("2019-06-30")
    )
  )

snapshot <- ledgr_snapshot_from_df(
  bars,
  snapshot_id = "strategy_chapter_snapshot"
)
```

The snapshot seals the market data. That is the evidence base for the
experiment. Strategies and indicators can derive from it, but the underlying
bars do not change mid-research.

## Indicators And Feature IDs

Indicators are feature definitions. Before a strategy uses a feature, ask ledgr
for the exact ID.

```{r}
#| label: feature-ids

features <- list(ledgr_ind_returns(5))

ledgr_feature_id(features)
```

Those strings are the names used inside `ctx$feature()`. They are exact. A
typo such as `"returns_5"` is not treated as a warmup value; it is an unknown
feature and ledgr fails loudly.

Warmup is different. A known feature can be `NA` early in the sample because
there are not enough prior bars yet. Strategy code should treat that as "no
signal yet."

## Debug One Pulse Before Running

Before running a full backtest, inspect one pulse. This is the fastest way to
understand what your strategy will see.

```{r}
#| label: pulse

pulse <- ledgr_pulse_snapshot(
  snapshot,
  universe = c("DEMO_01", "DEMO_02"),
  ts_utc = ledgr_utc("2019-03-01"),
  features = features
)

pulse$ts_utc
pulse$universe
pulse$close("DEMO_01")
pulse$feature("DEMO_01", "return_5")
pulse$hold()
```

The scalar accessors are easiest when you are writing or debugging a rule for
one instrument. Cross-sectional rules should usually switch to the vector
accessors once the contract is clear:

```{r}
#| label: pulse-vector-accessors

pulse$idx("DEMO_01")
pulse$vec$close
pulse$vec$feature("return_5")
```

`ctx$idx(id)` gives the instrument's position in `ctx$universe`. Values in
`ctx$vec$close`, `ctx$vec$position`, and `ctx$vec$feature(feature_id)` use that
same order, so a strategy can inspect the decision axis without repeating
scalar lookups. By default, `ctx$idx(id)` also fails loudly for an unknown
instrument; use its `missing` argument only when a deliberate `NA` path is part
of the rule.
The vector feature accessor still uses exact engine feature IDs: warmup for a
known feature is `NA`, while an unknown feature ID fails loudly.

The same pulse can also be viewed as one wide row. This is useful when you want
to see prices, portfolio state, and computed features together.

```{r}
#| label: pulse-wide

ledgr_pulse_wide(pulse) |>
  glimpse()
```

The wide row and the scalar accessors are two ways of looking at the same
pulse-known data. The rest of this article uses vector accessors because they
keep cross-sectional strategy logic both readable and inexpensive.

## Derive Weights, Then Rebalance In Context

A helper pipeline produces a relative allocation. Keep that research object
separate from the economic conversion into share quantities:

```{r}
#| label: equal-member-allocation

equal_weights <- pulse |>
  ledgr_selection() |>
  ledgr_weight_equal()

equal_target <- ledgr_target_rebalance(
  equal_weights,
  pulse,
  equity_fraction = 0.1
)

equal_weights
equal_target
```

The repeated `pulse` is meaningful. Weights say how to divide an allocation;
the current context supplies equity, prices, positions, and the complete
decision axis needed to express that allocation as quantities.

`ledgr_selection(ctx)` selects current investment members. In a dense run that
is the full decision axis. In an availability-aware run, `ctx$universe` can
also contain a held nonmember needed for valuation or disposition. That
nonmember does not enter ranking or weighting; the rebalance reserves its
marked exposure and preserves its quantity. The member set defines who can
enter the allocation; it is not a promise of execution at the next open.

## Hold, Then Edit

Use hold-and-edit when most quantities should remain unchanged. State can
decide which edit applies, but the strategy still returns complete intent:

```{r}
#| label: stateful-hold-and-edit

stateful_entry_exit <- function(ctx, params) {
  targets <- ctx$hold()
  id <- params$instrument_id
  phase <- ctx$state_prev$phase
  if (is.null(phase)) phase <- "waiting"

  if (phase == "waiting" && ctx$close(id) < params$entry_below) {
    targets[[id]] <- 1
    phase <- "holding"
  } else if (
    phase == "holding" &&
      ctx$position(id) > 0 &&
      ctx$close(id) > params$exit_above
  ) {
    targets[[id]] <- 0
    phase <- "done"
  }

  list(targets = targets, state_update = list(phase = phase))
}

stateful_exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = stateful_entry_exit,
  opening = ledgr_opening(
    cash = 10000,
    positions = c(DEMO_02 = 2),
    cost_basis = c(DEMO_02 = 180)
  ),
  cost_model = ledgr_cost_zero()
)
stateful_run <- ledgr_run(
  stateful_exp,
  params = list(
    instrument_id = "DEMO_01",
    entry_below = 90,
    exit_above = 91
  )
)

ledgr_results(stateful_run, what = "fills") |>
  select(ts_utc, instrument_id, side, qty)

close(stateful_run)
```

`DEMO_01` enters below 90 and exits above 91. The state prevents a later
re-entry, while `ctx$hold()` preserves the two opening `DEMO_02` shares without
special-case code. Returning hold does not bypass a risk chain or guarantee
that no fill occurs. It states portfolio intent at this decision; downstream
rules still apply.

## Make Missing And Zero Intent Explicit

Five cases look similar in compact code and mean different things. This
example keeps an existing `DEMO_02` position so hold is visibly different from
zero. It also injects one missing value deliberately; the sealed data itself is
unchanged.

```{r}
#| label: intent-cases

momentum_with_hold_guard <- function(ctx, params) {
  scores <- ctx$vec$feature(paste0("return_", params$lookback))
  if (anyNA(scores)) return(ctx$hold())

  weights <- ctx |>
    ledgr_signal(values = scores) |>
    ledgr_select_top_n(params$n) |>
    ledgr_weight_equal()

  ledgr_target_rebalance(
    weights,
    ctx,
    equity_fraction = params$equity_fraction
  )
}

warmup_pulse <- ledgr_pulse_snapshot(
  snapshot,
  universe = c("DEMO_01", "DEMO_02"),
  ts_utc = min(bars$ts_utc),
  features = features,
  positions = c(DEMO_01 = 0, DEMO_02 = 3)
)

prices_with_gap <- pulse$vec$close
prices_with_gap[[2L]] <- NA_real_

bare_predicate_error <- tryCatch(
  {
    ledgr_selection(pulse, where = prices_with_gap > 0)
    NA_character_
  },
  error = function(err) class(err)[[1L]]
)

guarded_selection <- ledgr_selection(
  pulse,
  where = !is.na(prices_with_gap) & prices_with_gap > 0
)

missing_score_target <- ledgr_signal(
  pulse,
  values = setNames(c(0.2, NA_real_), pulse$universe)
) |>
  ledgr_select_top_n(1) |>
  ledgr_weight_equal() |>
  ledgr_target_rebalance(pulse, equity_fraction = 0.1)

explicit_zero_target <- ledgr_weights(
  c(DEMO_01 = 1, DEMO_02 = 0),
  universe = pulse$universe
) |>
  ledgr_target_rebalance(pulse, equity_fraction = 0.1)

omitted_target <- ledgr_weights(
  c(DEMO_01 = 1),
  universe = pulse$universe
) |>
  ledgr_target_rebalance(pulse, equity_fraction = 0.1)

hold_target <- momentum_with_hold_guard(
  warmup_pulse,
  list(lookback = 5, n = 1, equity_fraction = 0.1)
)

show_target <- function(x) {
  paste(paste(names(x), c(x), sep = "="), collapse = ", ")
}

intent_cases <- tibble(
  input = c(
    "missing logical decision",
    "missing numeric score",
    "explicit zero weight",
    "omitted weight",
    "hold during warmup"
  ),
  meaning = c(
    "not a decision; guard it explicitly",
    "exclude that score from ranking",
    "validate the name and request zero",
    "leave that member unallocated",
    "preserve every current quantity"
  ),
  resulting_target = c(
    paste0(
      "error: ", bare_predicate_error, "; guarded: ",
      names(guarded_selection)[unclass(guarded_selection)]
    ),
    show_target(missing_score_target),
    show_target(explicit_zero_target),
    show_target(omitted_target),
    show_target(hold_target)
  )
)

knitr::kable(intent_cases)
close(warmup_pulse)
```

The guarded logical form turns the missing comparison into an explicit
exclusion. A missing score is already understood by ranking. Exact zero and
omission both produce zero here, but zero records a deliberate choice and is
dropped before price lookup. Hold is different: it keeps the three `DEMO_02`
shares. If missing input should pause the whole rebalance, return hold as the
warmup guard does.

## Read Planes, Not One Instrument At A Time

`ctx$vec$close` returns the aligned vector the engine already prepared for this
pulse. `ctx$close(id)` validates and looks up one instrument. Use the scalar
form after singling out an instrument; do not call it once for every member.

::: {.ledgr-callout .ledgr-callout-warning}
**Read the plane for cross-sectional work**

`ctx$close(id)`, `ctx$feature(id, feature_id)`, and `ctx$position(id)` are for
asking about one instrument you have already singled out. Reaching for them
inside `vapply()`, `sapply()`, or a `for` loop over `ctx$universe` is the most
expensive common strategy-authoring habit. When you want a value for everyone,
read `ctx$vec$close`, `ctx$vec$feature()`, or `ctx$vec$position` once. For
universes of at least 100 instruments, ledgr emits one
`ledgr_scalar_accessor_loop` warning per run when one scalar accessor reaches a
decision-axis call count in a pulse. The warning names the vector plane to use;
it is diagnostic only and changes no target, fill, result, or run identity.
:::

The optimization manual carries the measurements behind this rule. Strategy
authors only need the rule: use a plane for a cross-section and a scalar for a
chosen instrument.

## Turn The Idea Into A Strategy

The same transformations become an ordinary strategy function.

The full backtest replays every bar, including the earliest warmup pulses.
During those first pulses, `return_5` is `NA` for every instrument because
five prior bars do not exist yet. `ledgr_select_top_n()` treats that
all-missing signal as a classed empty selection, not as a warning. That object
still carries the
member set and signal origin. `ledgr_weight_equal()` turns it into empty
weights, and `ledgr_target_rebalance()` turns those weights into a flat target
over the decision axis.

No warning suppression is needed for ordinary early warmup. A partial-selection
warning can still appear when some signal values are usable but fewer than `n`
instruments can be selected. If a run finishes with zero trades, inspect a late
pulse before assuming the empty selection was only early warmup.

```{r}
#| label: helper-strategy

top_return_strategy <- function(ctx, params) {
  weights <- ctx |>
    ledgr_signal_return(lookback = params$lookback) |>
    ledgr_select_top_n(n = params$n) |>
    ledgr_weight_equal()

  ledgr_target_rebalance(
    weights,
    ctx,
    equity_fraction = params$equity_fraction
  )
}
```

Read it economically:

1. The pipeline scores current investment members by recent return.
2. It keeps the highest scores, ignores warmup `NA`, and derives equal weights.
3. The final call applies those weights in the current context and returns
   floored, complete target quantities.

No helper registers indicators automatically. The experiment must say which
features exist.

Empty selections flow through the pipeline as objects, so expected warmup and
"no signal today" look the same to your strategy. Diagnostics still belong at
the pulse level: when a strategy produces no fills or no closed trades, inspect
a late pulse and confirm whether the feature values are usable.

```{r}
#| label: experiment

exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = top_return_strategy,
  features = features,
  opening = ledgr_opening(cash = 10000),
  cost_model = ledgr_cost_zero()
)

top_return_run <- ledgr_run(
  exp,
  params = list(lookback = 5, n = 1, equity_fraction = 0.1),
  run_id = "top_return"
)

ledgr_results(top_return_run, what = "fills") |>
  select(ts_utc, instrument_id, side, qty) |>
  slice_head(n = 4)
```

## Feature Maps For Readable Feature Access

The examples above keep the exact feature ID contract visible:
`ctx$feature(id, feature_id)` reads one registered feature for one instrument
at one pulse. That contract remains the foundation.

For cross-sectional strategies, `ctx$vec$feature(feature_id)` returns the same
feature at the same pulse for every instrument in `ctx$universe`. Warmup for a
known feature remains `NA`; an unknown feature ID fails loudly. The scalar
helper stays the clearest teaching surface, while the vector helper is the
lower-overhead surface for decision-axis inspection.

When a strategy reads several features per instrument, repeating feature ID
strings can obscure the trading idea. A feature map bundles indicator objects
with strategy-facing aliases. The same object can be registered with the
experiment and used by the strategy for pulse-time lookup.

```{r}
#| label: feature-map-setup

mapped_features <- ledgr_feature_map(
  ret_5 = ledgr_ind_returns(5),
  sma_10 = ledgr_ind_sma(10)
)

ledgr_feature_id(mapped_features)
```

The strategy closes over `mapped_features`. Inside the universe loop,
`ctx$features(id, mapped_features)` returns a named numeric vector keyed by the
aliases. `ledgr_passed_warmup()` is a guard for that vector. For values returned
by `ctx$features()`, it means every requested indicator is usable at this pulse.
It is not a signal pipeline transformation, and it is not a data-quality
diagnostic for arbitrary vectors.

```{r}
#| label: feature-map-strategy

mapped_return_strategy <- function(ctx, params) {
  targets <- ctx$flat()

  for (id in ctx$universe) {
    x <- ctx$features(id, mapped_features)

    if (
      ledgr_passed_warmup(x) &&
        x[["ret_5"]] > params$min_return &&
        ctx$close(id) > x[["sma_10"]]
    ) {
      targets[id] <- params$qty
    }
  }

  targets
}
```

Read that as one pulse-time decision:

1. `ctx$features()` reads the mapped feature values for one instrument.
2. `ledgr_passed_warmup()` keeps the rule inactive until the mapped indicators
   are usable.
3. The condition states the trading idea.
4. The strategy still returns an ordinary target vector.

Plain `features = list(...)` remains valid. Use it when exact IDs are clearest.
Use a feature map when aliases make a feature-heavy strategy easier to read.
`ledgr_experiment(features = ...)` accepts indicators, lists, named lists, and
feature maps. The strategy context then uses either the exact-ID scalar
accessor `ctx$feature()`, the exact-ID vector accessor `ctx$vec$feature()`, or
the mapped accessor `ctx$features()`. When in doubt, prefer the
experiment-first workflow.

```{r}
#| label: feature-map-experiment

mapped_exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = mapped_return_strategy,
  features = mapped_features,
  opening = ledgr_opening(cash = 10000),
  cost_model = ledgr_cost_zero()
)
```

Run it the same way as any other experiment. The strategy still returns target
quantities; the feature map only changes how the strategy reads features.

```{r}
#| label: run-feature-map

bt_mapped <- mapped_exp |>
  ledgr_run(
    params = list(min_return = 0, qty = 5),
    run_id = "mapped_return"
  )

summary(bt_mapped)
```

Feature-map strategies commonly close over the feature map object. Keep that
construction code with the research record. The experiment store records the
registered feature definitions, but recovered strategy source may still
reference the original alias-map object by name.

## Keep Feature Declaration Outside Strategy

Do not declare or rebuild features inside a strategy:

```r
strategy <- function(ctx, params) {
  features <- ledgr_feature_map(
    fast = ledgr_ind_sma(params$fast_n),
    slow = ledgr_ind_sma(params$slow_n)
  )
  x <- ctx$features("AAA", features)
  ctx$flat()
}
```

That code puts feature declaration inside the execution loop. Strategy code
should read pulse-known values from the context; experiment code owns feature
declaration. Duplicating a parameterized feature map inside the strategy
creates a drift risk between the experiment's feature declaration and the
strategy lookup map.

For exploratory sweeps over indicator parameters, use active aliases and
feature grids. The canonical walkthrough is
`vignette("sweeps", package = "ledgr")`.

## Check Target Names Before A Full Run

When a hand-written target fails validation, compare its names with the
decision axis before rerunning the experiment:

```{r}
#| label: helper-debug-checklist

setdiff(pulse$universe, names(equal_target))
setdiff(names(equal_target), pulse$universe)
```

An empty first result means no instrument is missing. An empty second result
means no unknown instrument was added. For invalid return shapes, strategy
preflight tiers, stored source, hash verification, and trust boundaries, use
the canonical guide: read
`vignette("reproducibility", package = "ledgr")`.


## Cleanup

```{r cleanup}
close(bt_mapped)
close(top_return_run)
close(pulse)
ledgr_snapshot_close(snapshot)
```

## Where Next

- `vignette("strategy-development", package = "ledgr")` is the shorter
  first-pass strategy tutorial.
- `vignette("ttr-and-adapter-indicators", package = "ledgr")` covers
  adapter-backed indicator declarations.
- `vignette("walk-forward", package = "ledgr")` shows how strategies and
  sweeps feed the walk-forward workflow.
- `?ledgr_feature_map` and `?ledgr_passed_warmup` are the function-level
  references for mapped feature access and warmup filtering.
