---
title: "Strategy Development And Comparison"
output:
  rmarkdown::html_vignette:
    toc: false
  github_document:
    toc: false
    html_preview: false
vignette: >
  %\VignetteIndexEntry{Strategy Development And Comparison}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 90)
options(cli.unicode = FALSE)
default_output_hook <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(
  output = function(x, options) {
    default_output_hook(gsub("[ \t]+(?=\n)", "", x, perl = TRUE), options)
  }
)
demo_data_path <- if (file.exists("data/ledgr_demo_bars.rda")) {
  "data/ledgr_demo_bars.rda"
} else {
  file.path("..", "data", "ledgr_demo_bars.rda")
}
if (!exists("ledgr_demo_bars") && file.exists(demo_data_path)) {
  load(demo_data_path)
}
```

The examples use `dplyr` for demo-data preparation. Strategy functions use
ledgr's pulse context rather than data-frame operations.

```{r library, message=FALSE}
library(ledgr)
library(dplyr)
data("ledgr_demo_bars", package = "ledgr")
```

A ledgr strategy is an ordinary R function, but the important idea is not the
function syntax. The important idea is the pulse.

A backtest in ledgr is a sequence of decision moments. At each pulse, ledgr
shows the strategy only what could have been known at that time. The strategy
answers with desired holdings. ledgr records the decision, applies the fill
model, and moves to the next pulse.

This matters because leakage is easy. If a strategy accidentally sees tomorrow's
price while deciding today's trade, the backtest can look profitable for the
wrong reason. ledgr's strategy interface is built to make that mistake harder:
your strategy receives one pulse context, not the whole future.

## Wrong And Right: Leakage

The tempting vectorized pattern is to compute a future-looking column first and
then trade from it. In the example below, `lead(close)` shifts tomorrow's close
onto today's row. The resulting `buy_signal` looks like an ordinary column, but
it answers a question the strategy could not have answered at today's decision
time: "will tomorrow's close be higher than today's close?" Trading from that
column lets the backtest use future market data as if it were already known.

```{r leakage-wrong, eval=FALSE}
leaky_signals <- ledgr_demo_bars |>
  group_by(instrument_id) |>
  arrange(ts_utc, .by_group = TRUE) |>
  mutate(
    tomorrow_close = lead(close),
    buy_signal = tomorrow_close > close
  )
```

The ledgr version expresses the rule at one pulse. The strategy can read the
current bar for the current instrument. Later sections add registered features
to the same pulse model. The strategy cannot reach into the next row of the
market-data table. That is the same information shape a live trading strategy
gets as time passes: each pulse is a new slice of the knowable universe.

```{r leakage-right}
no_leak_bar_strategy <- function(ctx, params) {
  targets <- ctx$flat()

  for (id in ctx$universe) {
    if (ctx$close(id) > ctx$open(id)) {
      targets[id] <- 1
    }
  }

  targets
}
```

The ledgr strategy has no object from which it can accidentally read tomorrow's
close.

## A Strategy That Does Nothing

The simplest economic policy is: hold cash and own no instruments.

```{r flat-strategy}
flat_strategy <- function(ctx, params) {
  ctx$flat()
}
```

`ctx$flat()` creates a full target vector with one entry for every instrument in
the run and every value set to zero. Economically, this means: after the next
fill opportunity, hold no positions.

This is a complete ledgr strategy. It is not useful for making money, but it is
useful for understanding the contract: at every pulse, return target holdings.

The return value is a named numeric vector. Names are instrument IDs from
`ctx$universe`, values are desired quantities. `ctx$flat()` produces the
full-universe shape with every entry at zero.

The deeper mental model is that a strategy is a **policy**, not a sequence of
orders. At each pulse, the strategy declares a desired state: "I want to hold
this many shares of each instrument." The engine compares that against current
holdings, computes the gap, and fills accordingly. The strategy never says
"buy 3 shares"; it says "I want to hold 3 shares." This distinction is what
keeps strategies free from execution-state bookkeeping, and it is what makes
a ledgr strategy composable, testable, and directly readable as financial
reasoning.

## What Is `ctx`?

`ctx` is the pulse context. It is the information packet ledgr gives your
strategy at one decision time.

It contains the current timestamp, current bars, current features, current
positions, cash, equity, and small helper functions for accessing those values.
It is deliberately not the full dataset.

| Expression | Meaning at one pulse |
|---|---|
| `ctx$ts_utc` | current decision timestamp |
| `ctx$universe` | instruments in the run |
| `ctx$open(id)`, `ctx$close(id)` | current bar values for one instrument |
| `ctx$feature(id, name)` | current indicator value for one instrument |
| `ctx$position(id)` | current simulated position |
| `ctx$cash`, `ctx$equity` | current simulated portfolio state |
| `ctx$flat()` | target zero positions unless changed |
| `ctx$hold()` | target current positions unless changed |

Use `ctx$flat()` when the strategy should be flat unless it sees a reason to
act. Use `ctx$hold()` when the strategy should keep existing positions unless
it sees a reason to change them.

For example, this policy starts from current holdings and only changes the book
when it sees an exit reason. Economically, it means: "keep what I already own,
unless today's bar gives me a reason to leave."

```{r hold-example}
hold_unless_down <- function(ctx, params) {
  targets <- ctx$hold()

  for (id in ctx$universe) {
    if (ctx$close(id) < ctx$open(id)) {
      targets[id] <- 0
    }
  }

  targets
}
```

This loop style is fine while the mechanics are still visible. Once you have
helpers like `signal_*()` and `select_*()`, most strategy logic is easier to
express at the whole-universe level instead of one instrument at a time.

## A First Trading Rule

Now add one small economic idea:

> If an instrument closes above its open, own one share. Otherwise own nothing.

This is still a teaching strategy, not investment advice. It shows how
observable data becomes a target.

```{r buy-if-up}
buy_if_up <- function(ctx, params) {
  targets <- ctx$flat()

  for (id in ctx$universe) {
    if (ctx$close(id) > ctx$open(id)) {
      targets[id] <- 1
    }
  }

  targets
}
```

Targets are desired quantities, not orders, signals, or portfolio weights. A
target of `1` means "after the next fill opportunity, hold one share." A target
of `0` means "hold no shares."

In these examples, decisions fill at the next open: a decision made at pulse
`t` fills at the next available bar. That keeps the strategy from deciding and
filling on the same close.

The loop is intentionally plain because this is the first example. Once the
economic idea is clear, ledgr strategies are usually easier to read when they
use helper functions that operate on the whole universe at once. The later
sections make that transition.

## Why `params` Exists

Hard-coded constants make experiments awkward. Parameters let one economic idea
run under different assumptions.

```{r buy-if-up-param}
buy_if_up_qty <- function(ctx, params) {
  targets <- ctx$flat()

  for (id in ctx$universe) {
    if (ctx$close(id) > ctx$open(id)) {
      targets[id] <- params$qty
    }
  }

  targets
}
```

Strategies use `function(ctx, params)`. `ctx` is the pulse. `params` is the
experimenter's chosen configuration for this run. Keeping them separate makes
the strategy easier to test, compare, and store.

## Prepare A Small Experiment

Use two instruments from the offline demo data so the examples run anywhere.

```{r data}
bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% c("DEMO_01", "DEMO_02"),
    between(
      ts_utc,
      ledgr::ledgr_utc("2019-01-01"),
      ledgr::ledgr_utc("2019-06-30")
    )
  )

snapshot <- ledgr_snapshot_from_df(
  bars,
  snapshot_id = "strategy_chapter_snapshot"
)
```

The snapshot seals the market data. That is the evidence base for the
experiment. Strategies and indicators can derive from it, but the underlying
bars do not change mid-research.

## Indicators And Feature IDs

Indicators are feature definitions. Before a strategy uses a feature, ask ledgr
for the exact ID.

```{r feature-ids}
features <- list(ledgr_ind_returns(5))

ledgr_feature_id(features)
```

Those strings are the names used inside `ctx$feature()`. They are exact. A typo
such as `"returns_5"` is not treated as a warmup value; it is an unknown feature
and ledgr fails loudly.

Warmup is different. A known feature can be `NA` early in the sample because
there are not enough prior bars yet. Strategy code should treat that as "no
signal yet."

## Debug One Pulse Before Running

Before running a full backtest, inspect one pulse. This is the fastest way to
understand what your strategy will see.

```{r pulse}
pulse <- ledgr_pulse_snapshot(
  snapshot,
  universe = c("DEMO_01", "DEMO_02"),
  ts_utc = ledgr::ledgr_utc("2019-03-01"),
  features = features
)

pulse$ts_utc
pulse$universe
pulse$close("DEMO_01")
pulse$feature("DEMO_01", "return_5")
pulse$hold()
```

The same pulse can also be viewed as one wide row. This is useful when you want
to see prices, portfolio state, and computed features together.

```{r pulse-wide}
ledgr_pulse_wide(pulse) |>
  glimpse()
```

The wide row and the scalar accessors are two ways of looking at the same
pulse-known data. The wide row is good for inspection and model-like thinking.
The rest of this vignette uses the non-wide accessors because they keep the
step-by-step strategy logic easier to read.

Now build the strategy logic one transformation at a time.

The economic idea:

> Rank instruments by recent return, keep the top names, split capital equally,
> and convert those weights into share quantities.

`signal_return()` is a thin helper around the same feature you inspected above:
it reads `return_N` for every instrument in the pulse and returns one
universe-wide signal object.

The helper pipeline has four stages:

| Stage | Input | Output | Question answered |
|---|---|---|---|
| signal | pulse context | numeric scores with origin metadata | What looks attractive? |
| selection | signal | logical inclusion with the same origin | What should be considered? |
| weights | selection | allocation weights with the same origin | How should capital be split? |
| target | weights and context | full-universe share quantities | What should the portfolio hold? |

Execution semantics begin only at the target stage. `signal`, `selection`, and
`weights` are research objects that help author the strategy; `target` is the
ordinary full named target vector shape the runner validates and executes.

```{r pulse-pipeline}
signal <- signal_return(pulse, lookback = 5)
signal

selection <- select_top_n(signal, n = 1)
selection

weights <- weight_equal(selection)
weights

target <- target_rebalance(weights, pulse, equity_fraction = 0.1)
target
```

`target_rebalance()` sizes with current pulse equity and current close prices,
then floors to whole shares. For the selected `DEMO_01` pulse above, 10% of
equity is allocated to the one selected instrument:

```{r floor-sizing}
raw_qty <- weights[["DEMO_01"]] * 0.1 * pulse$equity / pulse$close("DEMO_01")
c(pre_floor = raw_qty, target_qty = unclass(target)[["DEMO_01"]])
```

The helper objects are not a second execution path. They are authoring aids.
The pipeline still ends in a `ledgr_target`, which unwraps to the same target
quantity vector the runner has always consumed.

Interactive pulse snapshots and backtest handles can be closed when you are
done inspecting them. This releases DuckDB resources in long sessions; completed
run artifacts are already durable when `ledgr_run()` returns.

```{r close-pulse}
close(pulse)
```

## Turn The Idea Into A Strategy

The same transformations become an ordinary strategy function.

The full backtest replays every bar, including the earliest warmup pulses.
During those first pulses, `return_5` is `NA` for every instrument because five
prior bars do not exist yet. `select_top_n()` warns when a signal has no usable
values; that warning is useful during pulse debugging, but expected during a
full run. The strategy suppresses that specific warmup noise while still
returning an empty selection, empty weights, and a flat target.

The warning exists because an empty selection can mean either "normal warmup"
or "my signal is broken." The helper cannot know which one is true, so it warns
and lets the strategy decide how to handle that case. The warning includes the
signal origin and non-missing count; if a run finishes with zero trades, inspect
a late pulse without suppression before assuming the empty selection was only
early warmup.

```{r helper-strategy}
top_return_strategy <- function(ctx, params) {
  signal <- signal_return(ctx, lookback = params$lookback)
  selection <- suppressWarnings(
    select_top_n(signal, n = params$n),
    classes = "ledgr_empty_selection"
  )

  weights <- weight_equal(selection)
  target_rebalance(weights, ctx, equity_fraction = params$equity_fraction)
}
```

Read it economically:

1. `signal_return()` scores each instrument by recent return.
2. `select_top_n()` keeps the highest scores and ignores warmup `NA`.
3. `weight_equal()` splits the chosen allocation equally.
4. `target_rebalance()` converts weights into floored full target quantities.

No helper registers indicators automatically. The experiment must say which
features exist.

Class-specific suppression is deliberately narrow: it keeps expected warmup
noise out of a full run while still allowing unrelated warnings to surface.
There is no `warn_empty = FALSE` argument in v0.1.7.4; use a diagnostic pulse
and call `select_top_n()` without suppression when a strategy produces no
fills or no closed trades.

```{r experiment}
exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = top_return_strategy,
  features = features,
  opening = ledgr_opening(cash = 10000)
)
```

## Feature Maps For Readable Feature Access

The examples above keep the exact feature ID contract visible:
`ctx$feature(id, feature_id)` reads one registered feature for one instrument at
one pulse. That contract remains the foundation.

When a strategy reads several features per instrument, repeating feature ID
strings can obscure the trading idea. A feature map bundles indicator objects
with strategy-facing aliases. The same object can be registered with the
experiment and used by the strategy for pulse-time lookup.

```{r feature-map-setup}
mapped_features <- ledgr_feature_map(
  ret_5 = ledgr_ind_returns(5),
  sma_10 = ledgr_ind_sma(10)
)

ledgr_feature_id(mapped_features)
```

The strategy closes over `mapped_features`. Inside the universe loop,
`ctx$features(id, mapped_features)` returns a named numeric vector keyed by the
aliases. `passed_warmup()` is a guard for that vector: for values returned by
`ctx$features()`, it means every requested indicator is usable at this pulse.
It is not a signal pipeline transformation, and it is not a data-quality
diagnostic for arbitrary vectors. A zero-length input is a classed error because
an empty feature bundle cannot prove warmup has passed. The specific error
classes are `ledgr_empty_warmup_input` and `ledgr_invalid_warmup_input`.

```{r feature-map-strategy}
mapped_return_strategy <- function(ctx, params) {
  targets <- ctx$flat()

  for (id in ctx$universe) {
    x <- ctx$features(id, mapped_features)

    if (
      passed_warmup(x) &&
        x[["ret_5"]] > params$min_return &&
        ctx$close(id) > x[["sma_10"]]
    ) {
      targets[id] <- params$qty
    }
  }

  targets
}
```

Read that as one pulse-time decision:

1. `ctx$features()` reads the mapped feature values for one instrument.
2. `passed_warmup()` keeps the rule inactive until the mapped indicators are
   usable.
3. The condition states the trading idea.
4. The strategy still returns an ordinary target vector.

Plain `features = list(...)` remains valid. Use it when exact IDs are clearest.
Use a feature map when aliases make a feature-heavy strategy easier to read.

```{r feature-map-experiment}
mapped_exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = mapped_return_strategy,
  features = mapped_features,
  opening = ledgr_opening(cash = 10000)
)
```

Run it the same way as any other experiment. The strategy still returns target
quantities; the feature map only changes how the strategy reads features.

```{r run-feature-map}
bt_mapped <- mapped_exp |>
  ledgr_run(
    params = list(min_return = 0, qty = 5),
    run_id = "mapped_return"
  )

summary(bt_mapped)
```

Feature-map strategies commonly close over the feature map object. That is the
single-definition pattern: one object defines aliases, registers indicators,
and drives pulse-time lookup.

This has a provenance consequence. Tier 2 is common for strategy functions that
call package helpers or depend on external symbols; the helper-pipeline strategy
below is also tier 2. Feature maps add a more specific concern: the recovered
strategy source may reference `mapped_features`, but the alias map object itself
is not recovered as a standalone object from strategy provenance. The experiment
store records the registered feature definitions, while the human-readable alias
construction remains part of your research code. Keep the feature-map
construction code with the research record when you need to rerun the strategy
later.

## Run One Backtest

```{r run-one}
bt_top_1 <- exp |>
  ledgr_run(
    params = list(lookback = 5, n = 1, equity_fraction = 0.1),
    run_id = "top_return_1"
  )

summary(bt_top_1)
```

The summary is portfolio-level: total return, max drawdown, and trade count are
computed from the completed run. In ledgr, trades are closed round trips; the
fills table can contain more rows because opening fills and closing fills are
both recorded.

The annualized volatility is high because this toy strategy switches positions
often on a tiny two-instrument demo universe. Treat it as a warning about the
example, not as a property you should expect from the same idea on real data.
The drawdown is disproportionate to the final loss for the same reason: a small,
concentrated portfolio can swing hard during the run even if it ends roughly
flat.

Do not expect a teaching strategy to be good. A weak or unattractive result is
still useful evidence: ledgr records failed ideas with the same care as
successful ones, which is part of not fooling yourself.

Inspecting trades shows the actions produced by the target decisions.

```{r trades}
ledgr_results(bt_top_1, what = "trades")
```

The trade table only includes closed round trips. Small one-share rows appear
when integer sizing and price movement leave a tiny adjustment after a previous
target. Larger rows are the ordinary position exits. `realized_pnl` is the
profit or loss booked when that position closes.

## Compare Parameter Variants

Now keep the economic idea fixed and change one assumption: hold the top two
instruments instead of the top one.

```{r run-two}
bt_top_2 <- exp |>
  ledgr_run(
    params = list(lookback = 5, n = 2, equity_fraction = 0.1),
    run_id = "top_return_2"
  )

ledgr_compare_runs(snapshot, run_ids = c("top_return_1", "top_return_2"))
```

Comparison is most useful when the compared runs differ for a reason you can
explain. Here the reason is simple: one run concentrates the allocation in the
single strongest recent-return instrument, and the other splits it across two.

The comparison table is not asking "which number is biggest?" in isolation. It
is asking whether the change in assumption improved the run in ways you can
defend: return, drawdown, number of closed trades, and win rate all matter.

These runs share the same sealed data, initial cash, feature set, and cost
assumptions. That keeps the teaching example narrow. A real comparison would
also ask whether the conclusion survives different samples, execution costs,
starting capital, and parameter choices.

## Compare Against A Baseline

The flat strategy is a sanity baseline, not a market benchmark. It tells you
what the result table looks like when the strategy deliberately does nothing,
and it keeps the comparison honest: if an active strategy cannot beat doing
nothing on the same sealed data, that is valuable information.

```{r baseline}
flat_exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = flat_strategy,
  opening = ledgr_opening(cash = 10000)
)

bt_flat <- flat_exp |>
  ledgr_run(params = list(), run_id = "flat_baseline")

ledgr_compare_runs(snapshot, run_ids = c("top_return_1", "top_return_2", "flat_baseline"))
```

The flat baseline has no win rate because it has no closed trades. That is not
missing data; there are no wins or losses to count.

## When ledgr Complains

ledgr tries to fail loudly when an error would make a backtest misleading. If a
strategy returns a vector with the wrong names or length, ledgr rejects it
instead of silently treating missing instruments as zero. If a helper reads an
unregistered feature, `ctx$feature()` reports the unknown feature ID and lists
the available IDs. If `target_rebalance()` receives negative or over-allocated
weights, it fails before turning them into target quantities.

Those errors are part of the design. They are meant to catch research mistakes
while the mistake is still small enough to understand.

## Inspect Stored Source

The experiment store lets you come back later and know exactly what code and
parameters produced a result. When you return to a six-month-old run, this is
the artifact that tells you what was actually tested, not what you remember
testing.

ledgr stores strategy provenance with the run. Inspection is read-only by
default: `trust = FALSE` returns stored text and metadata without parsing or
evaluating it.

Hash verification proves stored-text identity, not safety. That is why the
read-only path is the default.

You usually do not read the hashes yourself. They are the durable fingerprints
ledgr uses to verify that recovered source and parameters are identical to the
artifacts that produced the run.

```{r extract}
ledgr_extract_strategy(snapshot, "top_return_1", trust = FALSE)
```

Use `trust = TRUE` only when you explicitly trust the experiment store and want
to recover a function object.

## Cleanup

These calls release DuckDB connections. Forgetting them in a short interactive
session is not a data-safety issue; completed run artifacts are already durable
when `ledgr_run()` returns. In long sessions and scripts, releasing resources
explicitly avoids lock and resource warnings.

```{r cleanup}
close(bt_top_1)
close(bt_top_2)
close(bt_flat)
close(bt_mapped)
ledgr_snapshot_close(snapshot)
```

## What's Next?

If you want the formal contract, read the strategy and context sections in
`inst/design/contracts.md`. If you want the indicator story, read
`vignette("indicators", package = "ledgr")`. For formal feature-map help, see
`?ledgr_feature_map` and `?passed_warmup`. If you want the deployment story,
continue with `vignette("research-to-production", package = "ledgr")`. If you
want to inspect or compare durable runs, read `vignette("experiment-store",
package = "ledgr")`.
