---
title: "Strategy Authoring Tools"
format:
  html:
    toc: false
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{Strategy Authoring Tools}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 90)
options(cli.unicode = FALSE)
# Local renders load ledgr from source; installed builds use the library call
# in the prerequisites chunk.
pkg_root <- if (file.exists("DESCRIPTION")) {
  "."
} else if (file.exists("../DESCRIPTION")) {
  ".."
} else {
  NA_character_
}
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
default_output_hook <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(
  output = function(x, options) {
    default_output_hook(gsub("[ \t]+(?=\n)", "", x, perl = TRUE), options)
  }
)
demo_data_path <- if (file.exists("data/ledgr_demo_bars.rda")) {
  "data/ledgr_demo_bars.rda"
} else {
  file.path("..", "data", "ledgr_demo_bars.rda")
}
if (!exists("ledgr_demo_bars") && file.exists(demo_data_path)) {
  load(demo_data_path)
}
```

`vignette("strategy-development", package = "ledgr")` showed the contract: a
strategy is a function of `ctx` and `params` that returns the holdings you want
after the next fill. That is enough to run a first idea. It is not yet enough to
trust one. Before you spend a backtest on a strategy, you want to see what it
decides on a single day, know exactly how a ranking became a number of shares,
choose what happens when an input is missing, and sometimes let today's
decision depend on yesterday's.

This article builds one momentum strategy three times. Each version fixes a
problem the previous one showed:

1. rank the instruments and hold the top two;
2. hold only the instruments that are actually rising;
3. rebalance once a week, but exit a position early when it breaks its trend.

By the end you will have tested each version on one pulse, checked its share
sizing by hand, run all three, and compared how often they trade.

## Set Up

```{r}
#| label: library
#| message: false

library(ledgr)
library(dplyr)
library(tibble)
data("ledgr_demo_bars", package = "ledgr")
```

The examples use four instruments from the offline demo data for the first half
of 2019.

```{r}
#| label: data

instruments <- c("DEMO_01", "DEMO_02", "DEMO_03", "DEMO_04")

bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% instruments,
    between(ts_utc, ledgr_utc("2019-01-01"), ledgr_utc("2019-06-30"))
  )

snapshot <- ledgr_snapshot_from_df(
  bars,
  snapshot_id = "strategy_authoring_snapshot"
)

features <- list(ledgr_ind_returns(5), ledgr_ind_sma(10))
ledgr_feature_id(features)
```

The strategies read two registered features: the 5-day return, used as
momentum, and the 10-day simple moving average, used as a trend line. Features
are declared on the experiment, never inside the strategy.
`vignette("indicators", package = "ledgr")` covers feature IDs, warmup, and
readable aliases.

## Test A Strategy On One Pulse

The first version is the plainest momentum rule: every day, hold the two
instruments with the highest 5-day return, using half of the account.

```{r}
#| label: top-momentum

top_momentum <- function(ctx, params) {
  weights <- ctx |>
    ledgr_signal_return(lookback = 5) |>
    ledgr_select_top_n(params$n) |>
    ledgr_weight_equal()

  weights |>
    ledgr_target_rebalance(ctx, equity_fraction = params$invested)
}

params <- list(n = 2, invested = 0.5)
```

The next section takes that pipeline apart. First, check what it decides.
`ledgr_pulse_snapshot()` builds an inspection context for one timestamp. Its
cash, positions, and equity follow the same rules as the `ctx` a dense backtest
passes in; equity is cash plus the current value of the supplied positions.
You can therefore call the strategy yourself. Start by looking at the inputs:

```{r}
#| label: study-pulse

pulse <- ledgr_pulse_snapshot(
  snapshot,
  universe = instruments,
  ts_utc = ledgr_utc("2019-03-12"),
  features = features
)

tibble(
  instrument_id = pulse$universe,
  close = pulse$vec$close,
  momentum = pulse$vec$feature("return_5"),
  trend = pulse$vec$feature("sma_10")
)
```

Each `ctx$vec` value is a plain vector in the order of `ctx$universe`, which is
why the four columns line up. Now ask the strategy for its decision:

```{r}
#| label: test-top-momentum

top_momentum(pulse, params)
```

On 12 March the strategy wants `DEMO_02` and `DEMO_03`, the two highest 5-day
returns, and nothing else. The pulse computes every declared feature through
the same feature path as a dense run, so the backtest sees these same values at
this timestamp. The share counts depend on the account, which here is the
pulse's default of 100,000 in cash. Calling a strategy on a pulse is the fastest
way to catch a wrong feature ID, a reversed comparison, or an unexpected
position size before you run months of bars.

::: {.ledgr-callout .ledgr-callout-tip}
**Read whole vectors, not one instrument at a time**

`ctx$vec$close` and `ctx$vec$feature()` return one value per instrument in a
single read. The scalar forms `ctx$close(id)`, `ctx$feature(id, feature_id)`,
and `ctx$position(id)` are for an instrument you have already singled out.
Calling them in a loop over `ctx$universe` repeats per-instrument work at every
pulse and is the most common way to make a strategy slow. For universes of at
least 100 instruments, a loop over the bar, price, position, or exact-ID
feature accessor triggers one `ledgr_scalar_accessor_loop` warning per run that
names the vector to use; the warning changes no result. `ctx$features(id)`
stays unrecorded because its active-alias bundle has no whole-universe
equivalent yet.
:::

## From Scores To Shares

`top_momentum()` is a chain of small objects. Build it on the pulse one step at
a time:

```{r}
#| label: pipeline-steps

signal <- ledgr_signal_return(pulse, lookback = 5)
signal

selection <- ledgr_select_top_n(signal, n = 2)
selection

weights <- ledgr_weight_equal(selection)
weights
```

`ledgr_signal_return()` reads the registered `return_5` feature; it never
registers a feature for you. Use `ledgr_signal_feature()` for another
registered feature. Use `ledgr_signal(ctx, values = ...)` only when you have
already transformed the values into a genuinely custom score.

| Object | What it holds | Made by |
|---|---|---|
| signal | one score per instrument | `ledgr_signal_return()`, `ledgr_signal_feature()` |
| selection | `TRUE` or `FALSE`: which instruments are chosen | `ledgr_select_top_n()`, `ledgr_selection()` |
| weights | each chosen instrument's share of the budget | `ledgr_weight_equal()`, `ledgr_weights()` |
| target | whole-share quantities for every instrument | `ledgr_target_rebalance()` |

Only the target is something ledgr executes. The other three are working
objects: returning a signal, selection, or weights from a strategy is an error.

A ranking can come up short. When fewer than `n` scores are usable,
`ledgr_select_top_n()` selects what it can and warns, so an unexpectedly short
ranking is visible. Opt out only when selecting fewer names is an intended
policy:

```{r}
#| label: short-selection-policy

short_signal <- ledgr_signal(c(
  DEMO_01 = 0.02,
  DEMO_02 = NA_real_,
  DEMO_03 = NA_real_,
  DEMO_04 = NA_real_
))

tryCatch(
  ledgr_select_top_n(short_signal, n = 2),
  warning = conditionMessage
)
ledgr_select_top_n(short_signal, n = 2, partial = "allow")
```

Both calls select `DEMO_01`, the only usable score. The first returns the
warning, which names the opt-out; the second makes the same selection silently.

Weights are relative. `0.5` means "half of the budget", not a number of shares.
The weights say how to divide a budget. They do not contain the budget, current
prices, or current holdings. Turning them into shares therefore needs the
current `ctx` again:

```{r}
#| label: pipeline-target

target <- weights |>
  ledgr_target_rebalance(pulse, equity_fraction = 0.5)
target
```

You can reproduce that sizing yourself. Each chosen instrument receives
`weight * equity_fraction * equity` in cash, divided by its current close and
rounded down to whole shares:

```{r}
#| label: sizing-check

chosen <- names(weights)
chosen_close <- pulse$vec$close[match(chosen, pulse$universe)]

tibble(
  instrument_id = chosen,
  budget = as.numeric(weights) * 0.5 * pulse$equity,
  close = chosen_close,
  shares = floor(budget / chosen_close),
  target = as.numeric(target[chosen])
)
```

The remainder left by rounding stays in cash. Every instrument that was not
chosen gets a target of zero, so a rebalance sells anything it no longer wants.

### Keep A Position While Rebalancing The Rest

Sometimes you want to rebalance around a position you already own. Name it in
`keep`: ledgr keeps its current quantity, reserves its value once, and sizes
the new weights from the capital that remains. This pulse holds four shares of
`DEMO_04` and has 600 in cash. Keeping `DEMO_04` leaves that 600 for the new
position:

```{r}
#| label: keep-position

keep_pulse <- ledgr_pulse_snapshot(
  snapshot,
  universe = instruments,
  ts_utc = ledgr_utc("2019-03-12"),
  features = features,
  cash = 600,
  positions = c(DEMO_04 = 4)
)
keep_weights <- ledgr_weights(c(DEMO_02 = 1), universe = instruments)
keep_weights |>
  ledgr_target_rebalance(keep_pulse, keep = "DEMO_04")
```

`DEMO_04` keeps its four shares. `DEMO_02` gets eight shares at about 67.7,
which uses about 542 of the 600; the rest stays in cash.

`equity_fraction` applies to that remaining capital. Do not shrink it by hand
to make room for the kept position; doing both would shrink the remaining
budget a second time and underinvest the account.

## Run The First Version

```{r}
#| label: run-top-momentum

top_momentum_run <- ledgr_experiment(
  snapshot = snapshot,
  strategy = top_momentum,
  features = features,
  opening = ledgr_opening(cash = 10000),
  cost_model = ledgr_cost_zero()
) |>
  ledgr_run(params = params, run_id = "v1_ranking")

top_momentum_fills <- ledgr_results(top_momentum_run, what = "fills")
nrow(top_momentum_fills)

top_momentum_fills |>
  select(ts_utc, instrument_id, side, qty) |>
  slice_head(n = 6)
```

Nothing trades during the first week. Until five prior bars exist, every 5-day
return is `NA`; ranking skips missing scores, finds nothing to choose, and
returns an empty selection without a warning, so the target is all zeros. After
warmup, the strategy trades almost every day because the two highest returns
keep changing. Each decision fills at the next open, so a fill's date is the
pulse after the decision.

If a whole run finishes with no fills, test the strategy on a late pulse. A
feature that never warms up looks exactly like this first week, only it never
ends.

When you run this yourself, R also prints a `LEDGR_LAST_BAR_NO_FILL` warning.
The strategy changed its target on the final bar, and there is no later open at
which to fill it. Expect it at the end of most samples; `?LEDGR_LAST_BAR_NO_FILL`
explains when it matters.

## Choose Members With A Rule Instead Of A Rank

The pulse test already showed a weakness. On 12 March the first version bought
`DEMO_03` even though its 5-day return was negative and its price was below its
10-day average. A ranking always picks two names, even when both are falling.

A better rule: hold every instrument that is rising, meaning positive momentum
and a close above its trend, and split the budget equally among however many
qualify. That is a yes-or-no decision per instrument, so it goes into
`ledgr_selection(ctx, where = ...)` instead of a ranking. Written directly, the
rule fails at the start of the sample:

```{r}
#| label: missing-decision

first_pulse <- ledgr_pulse_snapshot(
  snapshot,
  universe = instruments,
  ts_utc = min(bars$ts_utc),
  features = features
)

rising <- first_pulse$vec$feature("return_5") > 0 &
  first_pulse$vec$close > first_pulse$vec$feature("sma_10")
rising

tryCatch(
  ledgr_selection(first_pulse, where = rising),
  error = conditionMessage
)
```

At the first pulse neither feature has enough history, so every comparison is
`NA`. ledgr refuses a missing decision instead of quietly treating it as
`FALSE`, because "not chosen" and "could not tell" are different statements.

Ranking treated missing values differently. In `top_momentum()` a missing
score was simply left out of the ranking. A missing number means "no score";
a missing `TRUE`/`FALSE` means "no decision", and the strategy has to resolve
it. Here, "not rising" is the right answer when any input is missing, including
the close:

```{r}
#| label: trend-momentum

trend_momentum <- function(ctx, params) {
  momentum <- ctx$vec$feature("return_5")
  trend <- ctx$vec$feature("sma_10")
  close <- ctx$vec$close
  rising <- momentum > 0 & close > trend

  weights <- ctx |>
    ledgr_selection(where = rising, missing = "exclude") |>
    ledgr_weight_equal()

  weights |>
    ledgr_target_rebalance(ctx, equity_fraction = params$invested)
}

trend_momentum(first_pulse, params)
trend_momentum(pulse, params)
```

The first pulse now selects nothing. `missing = "exclude"` is the deliberate
policy: an unknown decision counts as not selected. On 12 March the rule keeps
only `DEMO_02` and gives it the whole invested half of the account.

```{r}
#| label: run-trend-momentum

trend_momentum_run <- ledgr_experiment(
  snapshot = snapshot,
  strategy = trend_momentum,
  features = features,
  opening = ledgr_opening(cash = 10000),
  cost_model = ledgr_cost_zero()
) |>
  ledgr_run(params = params, run_id = "v2_rule")

nrow(ledgr_results(trend_momentum_run, what = "fills"))
```

The rule trades even more than the ranking did. It stopped buying falling
instruments, but every time an instrument starts or stops rising, the equal
split changes, and the rebalance resizes every position to its new share, often
by a share or two. The third version deals with that churn.

## Decide What A Missing Input Should Do

Both versions go flat while their inputs warm up. That was harmless above
because the account started in cash. It is not harmless when you start with
existing holdings:

```{r}
#| label: warmup-with-holdings

holding_pulse <- ledgr_pulse_snapshot(
  snapshot,
  universe = instruments,
  ts_utc = min(bars$ts_utc),
  features = features,
  positions = c(DEMO_01 = 0, DEMO_02 = 40, DEMO_03 = 0, DEMO_04 = 0)
)

trend_momentum(holding_pulse, params)
holding_pulse$hold()
```

`trend_momentum()` returns zero for `DEMO_02`, which would sell all 40 shares on
the first day only because the features have not warmed up. `ctx$hold()`
returns the current quantities, which means "change nothing". What a strategy
returns for an instrument is always a quantity to hold after the next fill:

| The strategy returns | Meaning |
|---|---|
| the current quantity, as `ctx$hold()` does | keep the position |
| zero, including an instrument left out of the selection | sell down to zero |
| a positive number | hold that many shares |

These are requests. If the experiment has a `risk_chain`, it still applies, so
returning `ctx$hold()` does not guarantee that no fill occurs.

If missing input should pause the strategy rather than empty it, say so
explicitly. `ledgr_passed_warmup()` is `TRUE` only when every value it is given
is present, so the guard is one line at the top of the strategy:

```r
if (!ledgr_passed_warmup(ctx$vec$feature("sma_10"))) return(ctx$hold())
```

The third version uses it.

## Give The Strategy A Memory

`trend_momentum()` still decides from scratch every day, so it trades whenever
an instrument crosses its trend line. Suppose you want to rebalance only every
fifth pulse, roughly once a week on daily bars, but still exit a position early
if it drops below its trend. The strategy then needs to remember something
between pulses: how many pulses have passed.

Strategy state is that memory:

- Instead of a bare target, return `list(targets = ..., state_update = ...)`.
  `state_update` is a list of plain data you choose: numbers, strings, logical
  values, and lists of them.
- At the next pulse, that list arrives as `ctx$state_prev`. On the first pulse
  there is nothing to remember yet, so `ctx$state_prev$pulses_seen` is `NULL`.
- Returning a bare target keeps the previous state unchanged.
- ledgr stores the state with the run but does not interpret it. The names
  inside it are yours.
- State records decisions, not fills. A target returned at one pulse fills at
  the next open.

```{r}
#| label: weekly-trend-momentum

weekly_trend_momentum <- function(ctx, params) {
  pulses_seen <- ctx$state_prev$pulses_seen
  if (is.null(pulses_seen)) pulses_seen <- 0

  momentum <- ctx$vec$feature("return_5")
  trend <- ctx$vec$feature("sma_10")
  close <- ctx$vec$close

  if (!ledgr_passed_warmup(trend)) {
    targets <- ctx$hold()
  } else if (pulses_seen %% params$every == 0) {
    rising <- momentum > 0 & close > trend
    weights <- ctx |>
      ledgr_selection(where = rising, missing = "exclude") |>
      ledgr_weight_equal()
    targets <- weights |>
      ledgr_target_rebalance(ctx, equity_fraction = params$invested)
  } else {
    targets <- ctx$hold()
    targets[which(close < trend)] <- 0
  }

  list(targets = targets, state_update = list(pulses_seen = pulses_seen + 1))
}

weekly_params <- list(invested = 0.5, every = 5)
```

`pulses_seen` counts pulses. While the trend line is still warming up, the
strategy holds. On every fifth pulse it rebalances with the rule from the second
version. In between, it starts from `ctx$hold()`, keeps every position, and
changes one thing: an instrument that closed below its trend is set to zero.
An instrument without a close is neither bought nor sold. That is the
hold-then-edit pattern: start from what you own and change only what today's
evidence justifies.

You might be tempted to call `trend_momentum()` from inside
`weekly_trend_momentum()` instead of repeating the rule. Don't: ledgr only runs
self-contained strategies, and a call to your own function defined outside the
strategy, such as `trend_momentum()`, is rejected before the run starts.
`vignette("reproducibility", package = "ledgr")` explains why.

Calling the strategy on the study pulse shows the shape it returns. There is no
previous state, so `pulses_seen` starts at zero and this is a rebalance pulse:

```{r}
#| label: test-weekly

weekly_trend_momentum(pulse, weekly_params)
```

Supply previous state and holdings to test the other branch without running a
backtest. Here `pulses_seen = 1` makes this a holding pulse; the strategy keeps
the book except for a position whose close broke below its trend:

```{r}
#| label: test-weekly-exit

exit_pulse <- ledgr_pulse_snapshot(
  snapshot,
  universe = instruments,
  ts_utc = ledgr_utc("2019-03-12"),
  features = features,
  positions = c(DEMO_02 = 10, DEMO_03 = 10),
  state_prev = list(pulses_seen = 1)
)

weekly_trend_momentum(exit_pulse, weekly_params)
```

Now run it and compare the three versions:

```{r}
#| label: run-weekly

weekly_run <- ledgr_experiment(
  snapshot = snapshot,
  strategy = weekly_trend_momentum,
  features = features,
  opening = ledgr_opening(cash = 10000),
  cost_model = ledgr_cost_zero()
) |>
  ledgr_run(params = weekly_params, run_id = "v3_weekly")

ledgr_run_compare(snapshot)
```

`ledgr_run_compare()` summarizes every run stored with the snapshot.
`n_trades` counts sales that close earlier purchases, fully or in part;
purchases alone are not counted, so it is smaller than the number of fills.

Each change did what it was designed to do: the rule stopped buying falling
instruments, and the schedule cut the number of trades by more than half. That
is not evidence that the weekly version is a better strategy; in this sample the
plain ranking even finished with the most equity. Three variants of one idea on
four instruments over six months say nothing reliable about performance.
`vignette("research-workflow", package = "ledgr")` covers how to compare
candidates without fooling yourself.

::: {.ledgr-callout .ledgr-callout-tip}
**Try it**

Run `weekly_trend_momentum()` with `every = 1`. It should reproduce
`trend_momentum()` fill for fill. Why? Checks like this are a cheap way to
confirm that a rewrite changed only what you meant to change.
:::

The fills show both kinds of pulse:

```{r}
#| label: weekly-fills

ledgr_results(weekly_run, what = "fills") |>
  select(ts_utc, instrument_id, side, qty) |>
  slice_head(n = 8)
```

Purchases happen only after rebalance pulses, because between them the strategy
can only keep or exit. On 30 January two more instruments qualified, so the
budget was split four ways instead of two: `DEMO_01` and `DEMO_04` were cut
roughly in half to make room for `DEMO_02` and `DEMO_03`. The sales on 1 and 4
February are early exits: those instruments closed below their trend between
rebalances.

## In Availability-Aware Runs

Everything above uses a dense panel, where every instrument can be selected at
every pulse. In an availability-aware run, the set of instruments you may invest
in changes over time; `vignette("survivorship-bias", package = "ledgr")` shows
one end to end. Four things then work differently:

- `ctx$universe` can include an instrument you still hold but may no longer
  select, a *held nonmember*. `ctx$members` lists the instruments that are
  currently eligible.
- `ledgr_selection()`, `ledgr_signal()`, and `ledgr_select_top_n()` choose only
  among members. Values you supply for held nonmembers are ignored, not ranked.
  `ledgr_signal_return()` and `ledgr_signal_feature()` also mark scores of
  target-inadmissible members as missing. A score you build yourself with
  `ledgr_signal(ctx, values = ...)` deliberately bypasses that convenience
  mask.
- `ledgr_target_rebalance()` keeps each held nonmember's quantity and reserves
  its value before sizing, so weights apply to what remains. With equity 100
  and a held nonmember worth 40, weight 1 allocates 60, and weight 0.6
  allocates 36. If your weights are fractions of total equity, for example
  from an optimizer, convert them first: 60% of total equity is weight 1 here.
- `ctx$state_prev` also carries `asset_state`, one entry per instrument in
  `ctx$universe` for state that belongs to that instrument. An instrument that
  leaves the universe loses its entry; `?ledgr_strategy_context` states the
  rules.

Neither membership nor `ctx$tradable()` guarantees that an order fills at the
next open.

## Cleanup

```{r}
#| label: cleanup

close(top_momentum_run)
close(trend_momentum_run)
close(weekly_run)
close(pulse)
close(first_pulse)
close(holding_pulse)
close(keep_pulse)
close(exit_pulse)
ledgr_snapshot_close(snapshot)
```

## Where Next

- `vignette("strategy-development", package = "ledgr")` covers the strategy
  contract, `ctx$flat()` and `ctx$hold()`, and a first run.
- `vignette("indicators", package = "ledgr")` covers feature IDs, warmup, and
  feature maps with readable aliases.
- `vignette("sweeps", package = "ledgr")` compares parameter values such as
  `invested` and `every` systematically.
- `vignette("reproducibility", package = "ledgr")` explains preflight tiers and
  why strategies must be self-contained.
- `vignette("survivorship-bias", package = "ledgr")` runs an
  availability-aware strategy end to end.
