---
title: "Strategy Basics"
format:
  html:
    toc: false
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{Strategy Basics}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 90)
options(cli.unicode = FALSE)
# Local renders load ledgr from source; installed builds use the library call
# in the prerequisites chunk.
pkg_root <- if (file.exists("DESCRIPTION")) "." else if (file.exists("../DESCRIPTION")) ".." else NA_character_
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
default_output_hook <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(
  output = function(x, options) {
    default_output_hook(gsub("[ \t]+(?=\n)", "", x, perl = TRUE), options)
  }
)
demo_data_path <- if (file.exists("data/ledgr_demo_bars.rda")) {
  "data/ledgr_demo_bars.rda"
} else {
  file.path("..", "data", "ledgr_demo_bars.rda")
}
if (!exists("ledgr_demo_bars") && file.exists(demo_data_path)) {
  load(demo_data_path)
}
```

```{=html}
<style>
.ledgr-diagram {
  margin: 1.25rem auto 1.5rem auto;
  text-align: center;
}
.ledgr-diagram .mermaid {
  display: inline-block;
  max-width: 760px;
  width: 100%;
}
.ledgr-diagram .node text,
.ledgr-diagram .edgeLabel {
  font-size: 18px !important;
}
</style>
```

This article teaches the strategy contract from the user side: what a
strategy can see, what it must return, and how to run a first backtest
without hidden lookahead. For testing a strategy on one pulse, helper
pipelines, share sizing, and strategy state, read
`vignette("strategy-authoring-tools", package = "ledgr")`.

## Prerequisites

The examples use `dplyr` for demo-data preparation. Strategy functions use
ledgr's pulse context rather than data-frame operations. The article assumes
basic familiarity with sealed snapshots (`vignette("data-input-and-snapshots",
package = "ledgr")`); it introduces feature IDs where the first strategy needs
one.

```{r}
#| label: library
#| message: false

library(ledgr)
library(dplyr)
data("ledgr_demo_bars", package = "ledgr")
```

A backtest in ledgr is a sequence of decision moments. At each pulse, ledgr
shows the strategy only what could have been known at that time. The strategy
answers with desired holdings. ledgr records the decision, applies the fill
model, and moves to the next pulse.

This matters because leakage is easy. If future information enters a historical
decision, the backtest can look profitable for the wrong reason. ledgr's
strategy interface is built to make one common mistake harder: your strategy
receives one pulse context, not the whole future. For the broader leakage model,
including feature-construction leakage and remaining user responsibilities, see
`vignette("leakage", package = "ledgr")`.

## The Causal Boundary

Strategy code receives one pulse context rather than a table containing future
bars. That removes one easy route to accidental look-ahead, but it does not
certify snapshot construction, feature logic, event timestamps, universe
history, or parameter selection. The Leakage article owns the worked
`lead(close)` example and the broader causal review; this article now stays on
the strategy contract itself.

With that boundary in mind, start with the simplest possible strategy.

## What Is `ctx`?

`ctx` is the **pulse context**: the information packet ledgr gives your
strategy at one decision time. It contains the current timestamp, current bars,
current features, positions, cash, equity, and small helper functions for
accessing those values. It is deliberately not the full future dataset.

| Expression | Meaning at one pulse |
|---|---|
| `ctx$ts_utc` | current decision timestamp |
| `ctx$universe` | instruments in the run |
| `ctx$idx(id)` | 1-based universe position for one instrument |
| `ctx$open(id)`, `ctx$close(id)` | current bar values for one instrument |
| `ctx$vec$close` | current close values for the full universe |
| `ctx$feature(id, feature_id)` | current indicator value for one instrument by engine feature ID |
| `ctx$vec$feature(feature_id)` | current indicator values for the full universe by engine feature ID |
| `ctx$features(id, feature_map)` | mapped indicator values for one instrument by alias, using the supplied feature map |
| `ctx$position(id)` | current simulated position |
| `ctx$vec$position` | current simulated positions aligned to `ctx$universe` |
| `ctx$cash`, `ctx$equity` | current simulated portfolio state |
| `ctx$state_prev` | what the strategy returned as `state_update` at the previous pulse |
| `ctx$flat()` | target zero positions unless changed |
| `ctx$hold()` | target current positions unless changed |

For the installed accessor reference, see `?ledgr_strategy_context`.

::: {.ledgr-callout .ledgr-callout-tip}
**One instrument or everyone**

The `ctx$vec$` forms return a value for the whole universe in one read. The
scalar forms answer for one instrument you have already singled out. Both are
correct; only the first is cheap when you want every name, because the scalar
form costs per call and a universe-wide loop pays it once per instrument per
pulse. `vignette("strategy-authoring-tools", package = "ledgr")` shows the
plane-based pattern and documents the `ledgr_scalar_accessor_loop` warning
emitted for that shape on large universes.
:::

The pulse loop is the contract in motion:

::: {.ledgr-diagram .ledgr-pulse-loop}

```{mermaid}
%%| echo: false

flowchart TB
  state_t["pulse t state<br/>bars through t<br/>positions, cash, equity"]
  ctx_node["ctx<br/>pulse-known projection"]
  strategy_node["strategy(ctx, params)"]
  target_node["target vector<br/>desired holdings"]
  fill_node["next-open fill<br/>at t + 1"]
  state_next["pulse t + 1 state<br/>ledger updated"]

  state_t --> ctx_node --> strategy_node --> target_node --> fill_node --> state_next
  state_next -. next pulse .-> state_t
```

:::

`ctx` is the no-lookahead handoff. It gives the strategy the current pulse
projection, the strategy returns a target vector, and ledgr handles validation,
fill timing, ledger events, and the next pulse.

The two target starters have different economic meanings:

| Helper | Starts from | Economic meaning |
|---|---|---|
| `ctx$flat()` | zero positions | only hold what this pulse explicitly selects |
| `ctx$hold()` | current positions | keep existing positions unless changed |

For example, this policy starts from current holdings and only changes the book
when it sees an exit reason. Economically, it means: "keep what I already own,
unless today's bar gives me a reason to leave."

```{r}
#| label: hold-example

hold_unless_down <- function(ctx, params) {
  down <- ctx$vec$close < ctx$vec$open
  targets <- ctx$hold()
  targets[down] <- 0
  targets
}
```

`ctx$vec$close < ctx$vec$open` compares every instrument at once, in the order
of `ctx$universe`, and gives the condition a name: `down` holds one `TRUE` or
`FALSE` per instrument. Assigning through it changes exactly the instruments
where it is `TRUE`; every other instrument keeps its current quantity.

## A Strategy That Does Nothing

The simplest economic policy is: hold cash and own no instruments.

```{r}
#| label: flat-strategy

flat_strategy <- function(ctx, params) {
  ctx$flat()
}
```

`ctx$flat()` creates a full target vector with one entry for every instrument in
the run and every value set to zero. Economically, this means: after the next
fill opportunity, hold no positions.

This is a complete ledgr strategy: useful for understanding the contract, not
for making money.

The return value is a named numeric vector. Names are instrument IDs from
`ctx$universe`, values are desired quantities. `ctx$flat()` produces the
full-universe shape with every entry at zero.

A **target vector** is the strategy's requested holdings for the full universe
at one pulse. It is named by instrument ID, numeric, and complete. It is not an
order list, a signal table, or a partial update. The deeper mental model is
that a strategy is a **policy**, not a sequence of orders. At each pulse, it
declares a desired state: "I want to hold this many shares of each instrument."
The engine compares that against current holdings and fills the gap.

That distinction keeps strategies free from execution-state bookkeeping. A
target is intent: the layers after the strategy decide what is allowed, when
it executes and what it costs, as
`vignette("risk-and-cost", package = "ledgr")` explains.

::: {.ledgr-callout .ledgr-callout-warning}
**Affordability is not automatic**

A dense run fills a target whether or not the account can pay for it, so cash
can go negative. `vignette("execution-semantics", package = "ledgr")` states
the rule, including what availability-aware runs do instead.
:::

## A First Trading Rule

Now add one small economic idea:

> If an instrument closes above its open, own one share. Otherwise own nothing.

This is still a teaching strategy, not investment advice. It shows how
observable data becomes a target.

```{r}
#| label: buy-if-up

buy_if_up <- function(ctx, params) {
  up <- ctx$vec$close > ctx$vec$open
  targets <- ctx$flat()
  targets[up] <- 1
  targets
}
```

Read it as one decision for the whole universe: start from zero for every
instrument, then set the instruments whose close is above their open to one
share.

Targets are desired quantities, not orders, signals, or portfolio weights. A
target of `1` means "after the next fill opportunity, hold one share." A target
of `0` means "hold no shares."

In these examples, decisions fill at the next open: a decision made at pulse
`t` fills at the next available bar. That keeps the strategy from deciding and
filling on the same close.

A rule like this reads current values directly. When a rule needs to rank
instruments or size positions from the account, the helper pipeline in
`vignette("strategy-authoring-tools", package = "ledgr")` does that work.

::: {.ledgr-callout .ledgr-callout-tip}
**Try it**

Change `buy_if_up()` so it starts from `ctx$hold()` instead of `ctx$flat()`.
Which positions would persist after a down bar, and why does that change the
economic meaning of the strategy?
:::

## Why `params` Exists

`params` is the run's **strategy configuration**. Put research choices you want
to compare, store, or sweep into `params`; do not hide them in globals or
inside feature declarations. Hard-coded constants make experiments awkward.
Parameters let one economic idea run under different assumptions.

```{r}
#| label: buy-if-up-param

buy_if_up_qty <- function(ctx, params) {
  up <- ctx$vec$close > ctx$vec$open
  targets <- ctx$flat()
  targets[up] <- params$qty
  targets
}
```

Strategies use `function(ctx, params)`. `ctx` is the pulse. `params` is the
experimenter's chosen configuration for this run. Keeping them separate makes
the strategy easier to test, compare, and store. `vignette("sweeps", package =
"ledgr")` runs one strategy across many `params` values.

## Remembering Between Pulses

Most strategies decide from the current pulse alone. A strategy that has to
remember something between pulses, such as how many pulses have passed or the
price at which it bought, returns `list(targets = ..., state_update = ...)`
instead of a bare target. ledgr stores the `state_update` with the run and
hands it back as `ctx$state_prev` at the next pulse.

The state must be plain data: numbers, strings, logical values, and lists of
them. A new `state_update` replaces the previous one; returning a bare target
keeps it. On the first pulse there is nothing to remember yet, so check each
field for `NULL` before using it.

`vignette("strategy-authoring-tools", package = "ledgr")` builds a strategy
that uses state step by step. `?ledgr_strategy_context` states the full rules,
including the per-instrument `asset_state` of availability-aware runs.

## Prepare A Small Experiment

Use two instruments from the offline demo data so the first full backtest can
run anywhere.

```{r}
#| label: data

bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% c("DEMO_01", "DEMO_02"),
    between(
      ts_utc,
      ledgr_utc("2019-01-01"),
      ledgr_utc("2019-06-30")
    )
  )

snapshot <- ledgr_snapshot_from_df(
  bars,
  snapshot_id = "strategy_chapter_snapshot"
)

features <- list(ledgr_ind_returns(5))

ledgr_feature_id(features)
```

The snapshot seals the market data. `features` declares the pulse-known return
input the strategy will read. Feature IDs are exact: a typo is an unknown
feature, not warmup.

```{r}
#| label: return-strategy

positive_return <- function(ctx, params) {
  ret <- ctx$vec$feature("return_5")
  if (!ledgr_passed_warmup(ret)) return(ctx$flat())
  rising <- ret > params$min_return
  targets <- ctx$flat()
  targets[rising] <- params$qty
  targets
}
```

Economically, this holds `qty` shares of every instrument whose 5-bar return is
above `min_return`, and nothing otherwise. A 5-bar return needs five earlier
closes, so for the first bars it is still `NA`. Both instruments share one
calendar, so the return is missing for all of them at once, and
`ledgr_passed_warmup()` holds nothing until every value exists. Handle warmup explicitly like this rather than letting a missing value slip
through the assignment unseen. `vignette("indicators", package = "ledgr")`
covers warmup in full.

## The Same Rule With Helpers

Every strategy, however it is written, returns the same thing: shares per
instrument. ledgr's strategy helpers build that named vector for you in named
steps. Here is the same rule as a helper pipeline:

```{r}
#| label: return-strategy-helpers

positive_return_helpers <- function(ctx, params) {
  rising <- ledgr_signal_return(ctx, lookback = 5) > params$min_return
  ctx |>
    ledgr_selection(where = rising, missing = "exclude") |>
    ledgr_target_quantity(ctx, params$qty)
}
```

`ledgr_signal_return()` reads the registered 5-bar return, `ledgr_selection()`
keeps the instruments where the condition holds, and `missing = "exclude"`
says what a warmup `NA` means: not selected. `ledgr_target_quantity()` turns the
selection into `qty` shares for the selected instruments and zero for the rest.

To see that both forms agree, ask each for its target on one decision date.
`ledgr_pulse_snapshot()` builds the `ctx` a strategy would receive at one
timestamp; `vignette("strategy-authoring-tools", package = "ledgr")` covers it
in detail.

```{r}
#| label: helpers-agree

pulse <- ledgr_pulse_snapshot(
  snapshot,
  universe = c("DEMO_01", "DEMO_02"),
  ts_utc = ledgr_utc("2019-01-29"),
  features = features
)
params <- list(min_return = 0, qty = 10)

hand_built <- positive_return(pulse, params)
piped <- positive_return_helpers(pulse, params)

hand_built
identical(unclass(piped), hand_built)

close(pulse)
```

The helpers return the same named vector, with a class that prints a short
summary. Use whichever form makes the rule easier to read. The later articles
use the helpers for rules like this one and write the vector by hand when
manipulating it is the lesson.

The pipeline has four kinds of step. A rule that ranks instruments or sizes
positions from the account swaps in the ranking and weighting steps:

| Step | Use |
|---|---|
| `ledgr_signal_return()` or `ledgr_signal_feature()` | read one registered feature as comparable scores |
| `ledgr_select_top_n(..., partial = )` or `ledgr_selection(..., missing = )` | choose instruments and state how incomplete inputs behave |
| `ledgr_weight_equal()` or `ledgr_weights()` | divide the allocation budget |
| `ledgr_target_rebalance(..., keep = )` | convert weights into shares using the current account context |

The weights carry only relative allocations. Rebalancing needs `ctx` again
because current equity, prices, and holdings determine the share quantities.
`vignette("strategy-authoring-tools", package = "ledgr")` builds a strategy
with each step and its missing-input policy.

```{r}
#| label: experiment

exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = positive_return,
  features = features,
  opening = ledgr_opening(cash = 10000),
  cost_model = ledgr_cost_zero()
)
```

## Run One Backtest

```{r}
#| label: run-one

bt_first <- exp |>
  ledgr_run(
    params = list(min_return = 0, qty = 10),
    run_id = "positive_return_1"
  )

summary(bt_first)
```

The run prints a `LEDGR_LAST_BAR_NO_FILL` warning. The strategy changed its
target on the last bar of the sample, and there is no later open at which to
fill that change, so ledgr reports it rather than dropping it silently. Expect
it at the end of most samples; `vignette("execution-semantics", package = "ledgr")` explains when it matters.

The summary is portfolio-level: total return, max drawdown, and trade count are
computed from the completed run. In ledgr, trades are closed round trips; the
fills table can contain more rows because opening fills and closing fills are
both recorded.

The rendered result is a short, tiny-universe teaching run. Its annualized
volatility, drawdown and return describe only that fixture; none is evidence
that the rule is attractive or that the annualization is stable. Read the
numbers together: a positive final return can coexist with an intra-run
drawdown, and 20 closed trades from a tiny synthetic fixture are still far too
little evidence for a strategy judgment.

Do not expect a teaching strategy to be good. A weak or unattractive result is
still useful evidence: ledgr records failed ideas with the same care as
successful ones, which is part of not fooling yourself.

Inspecting trades shows the actions produced by the target decisions.

```{r}
#| label: trades

ledgr_results(bt_first, what = "trades")
```

The trade table only includes closed round trips. Small one-share rows appear
when integer sizing and price movement leave a tiny adjustment after a previous
target. Larger rows are the ordinary position exits. `realized_pnl` is the
profit or loss booked when that position closes.

If a run has zero trades, inspect fills before assuming nothing happened:

```{r}
#| label: fills-diagnostic

ledgr_results(bt_first, what = "fills")
```

Zero fills means no execution occurred. Non-empty fills with zero trades means
positions opened but did not close. `n_trades` counts closed round trips, while
the fills table shows both opening and closing execution rows. For the full
checklist when a run makes no trades, see
`vignette("indicators", package = "ledgr")`.

If you want to compare variants, keep the strategy authoring question separate
from the research-comparison question. Use
`vignette("experiment-store", package = "ledgr")` for stored-run comparison and
`vignette("sweeps", package = "ledgr")` for promotion and review.

## When ledgr Complains

ledgr tries to fail loudly when an error would make a backtest misleading. A
strategy that forgets an instrument is the most common case. ledgr rejects the
target instead of silently treating the missing instrument as zero:

```{r}
#| label: incomplete-target

only_one_name <- function(ctx, params) {
  c(DEMO_01 = 1)
}

tryCatch(
  ledgr_experiment(
    snapshot = snapshot,
    strategy = only_one_name,
    opening = ledgr_opening(cash = 10000),
    cost_model = ledgr_cost_zero()
  ) |>
    ledgr_run(run_id = "incomplete_target"),
  error = conditionMessage
)
```

The message names the instrument that is missing. Starting from `ctx$flat()` or
`ctx$hold()` avoids this mistake, because both already contain every
instrument. The other common failures are just as specific: a mistyped feature
ID is reported together with the IDs that are registered, and
`ledgr_target_rebalance()` refuses negative or over-allocated weights before
turning them into share quantities.

Those errors are part of the design. They are meant to catch research mistakes
while the mistake is still small enough to understand.

## Cleanup

```{r}
#| label: cleanup

close(bt_first)
ledgr_snapshot_close(snapshot)
```

## Where Next

- `vignette("indicators", package = "ledgr")` declares the features a strategy
  reads and explains their IDs and warmup.
- `vignette("strategy-authoring-tools", package = "ledgr")` goes deeper on
  testing on one pulse, helper pipelines, share sizing, and strategy state.
- `vignette("execution-semantics", package = "ledgr")` explains when and at
  what price a target fills.
