---
title: "Survivorship Bias And Point-In-Time Universes"
format:
  html:
    toc: true
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{Survivorship Bias And Point-In-Time Universes}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 92)
options(cli.unicode = FALSE)
options(pillar.sigfig = 5)
pkg_root <- if (file.exists("DESCRIPTION")) {
  "."
} else if (file.exists("../DESCRIPTION")) {
  ".."
} else {
  NA_character_
}
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
default_output_hook <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(
  output = function(x, options) {
    x <- gsub(
      "Database:(\\s+)[^\\r\\n]+[.]duckdb",
      "Database:\\1<temporary DuckDB path>",
      x,
      perl = TRUE
    )
    x <- gsub("[ \t]+(?=\n)", "", x, perl = TRUE)
    default_output_hook(x, options)
  }
)
```

```{r}
#| label: prerequisites
#| message: false

library(ledgr)
library(dplyr)
```

You pull today's index constituents, fetch history for each one, and backtest a
simple rule. The equity curve looks good. You check the arithmetic, reseal the
snapshot, confirm the hashes. Everything reproduces.

The number is still wrong, and nothing in the output says so.

The file you downloaded lists the companies that are in the index *today*. The
ones that were dropped along the way never appear in it. Your strategy was
never offered the chance to lose money on them. You didn't choose to exclude
them. The data did, using information from the end of your sample.

::: {.ledgr-callout .ledgr-callout-warning}
**Survivorship bias**

Selection on what remained observable. When a sample is truncated by survival,
the surviving history can show returns the strategy could not have earned.
Brown, Goetzmann, Ibbotson, and Ross (1992) measured the effect on performance
studies. Shumway (1997) documented the related delisting bias that appears when
large negative delisting returns are missing from equity data.
:::

By the end of this article you will be able to declare a universe-aware
experiment, explain the assumptions it rests on, and say why its positions,
fills, and reported horizon differ from what you asked for.

## Two Companies, Ten Sessions

The example is synthetic and small enough to read in full. `DEMO` is an
illustrative venue, not a real exchange, and none of these rows are
redistributed vendor data.

Two instruments are listed on January 6. `AAA` loses more than half its value
over five sessions and stops trading. `BBB` grinds upward and is still listed
at the end of the window.

```{r}
#| label: prices

session_dates <- seq(as.Date("2020-01-06"), as.Date("2020-01-17"), by = "day")
open_day <- as.integer(format(session_dates, "%u")) <= 5L
open_dates <- session_dates[open_day]

prices <- tibble(
  date = open_dates,
  AAA = c(100, 92, 78, 61, 44, NA, NA, NA, NA, NA),
  BBB = c(50, 51, 52, NA, 54, 55, 56, 57, 58, 59)
)

prices
```

```{r}
#| label: fig-prices
#| echo: false
#| fig-height: 3.6
#| fig-width: 8
#| fig-alt: >
#|   Observed closes. AAA falls from 100 to 44 and its line ends there. BBB
#|   rises from 50 to 59 with a visible break at the missing 9 January session.

price_long <- bind_rows(
  tibble(date = prices$date, instrument_id = "AAA", close = prices$AAA),
  tibble(date = prices$date, instrument_id = "BBB", close = prices$BBB)
)

ggplot2::ggplot(price_long, ggplot2::aes(date, close, colour = instrument_id)) +
  ggplot2::geom_vline(
    xintercept = as.Date("2020-01-13"), linetype = "dashed",
    colour = "grey55", linewidth = 0.5
  ) +
  ggplot2::geom_line(linewidth = 1, na.rm = TRUE) +
  ggplot2::geom_point(size = 2, na.rm = TRUE) +
  ggplot2::annotate(
    "text", x = as.Date("2020-01-13"), y = 96,
    label = "AAA leaves the index", hjust = -0.04, size = 3.4, colour = "grey35"
  ) +
  ggplot2::annotate(
    "text", x = as.Date("2020-01-10"), y = 37,
    label = "last AAA observation", hjust = 0.9, size = 3.4, colour = "#B4472A"
  ) +
  ggplot2::scale_colour_manual(values = c(AAA = "#B4472A", BBB = "#4F9B63")) +
  ggplot2::labs(
    title = "What was actually observed",
    subtitle = "Lines stop where observations stop. Gaps are not interpolated.",
    x = NULL, y = "Close", colour = NULL
  ) +
  ggplot2::theme_minimal(base_size = 13) +
  ggplot2::theme(
    legend.position = "top",
    panel.grid.minor = ggplot2::element_blank(),
    plot.title.position = "plot",
    plot.title = ggplot2::element_text(face = "bold")
  )
```

Two holes matter later, and they are not the same kind of hole. `AAA` stops
printing prices after January 10. `BBB` is missing a single session on
January 9 and then resumes.

::: {.ledgr-callout .ledgr-callout-note}
**A missing price is not a reason**

Nothing in this table says *why* an observation is absent. A delisting, a
trading halt, and a vendor feed gap all look identical here. This example
declares no lifetime or trading-status facts, so ledgr is never told which one
occurred. The prose below supplies that story; the data does not.
:::

## The Selection Problem

Before any ledgr call, do the arithmetic by hand.

A researcher who starts from today's index list sees only `BBB`, because `AAA`
is no longer a constituent. A researcher who asks what the index actually held
on January 6 sees both.

```{r}
#| label: arithmetic

first_price <- function(x) x[is.finite(x)][1]
last_price <- function(x) x[is.finite(x)][sum(is.finite(x))]
simple_return <- function(x) last_price(x) / first_price(x) - 1

returns <- c(AAA = simple_return(prices$AAA), BBB = simple_return(prices$BBB))
round(returns, 3)

tibble(
  view = c("Survivor list (BBB only)", "Point-in-time (AAA and BBB)"),
  equal_weight_return = c(returns[["BBB"]], mean(returns))
)
```

One view reports a gain, the other a loss, from the same file. That is the
shape of the problem.

It is only a sketch, and one assumption is doing quiet work: it treats `AAA`'s
last observed price as its ending value. No settlement was observed. Nobody
told us what a holder actually received. Hold onto that, because it is exactly
the point at which ledgr later refuses to produce a number.

## The Objects You Are About To Build

ledgr splits this problem across a few objects. Each has one job, and they fall
into three groups: what the world supplied, what you chose, and what the run
recorded.

| Object | Its job in this example |
| --- | --- |
| **Bars** | The prices actually observed for `AAA` and `BBB`. |
| **Facts** | Declared expected sessions and dated membership, including when each assertion became knowable. |
| **Snapshot** | A sealed version of those inputs, so a run can be reproduced and reopened. |
| **Strategy** | The positions you want at each decision. |
| **Experiment** | The snapshot and strategy combined with a universe rule, a valuation policy, and cost rules. |
| **Run** | The execution of that configuration, plus the evidence and explanations it retained. |

Bars and facts are **source evidence**. The universe rule, valuation policy,
and strategy are **your choices**. Everything a run reports afterwards is a
**recorded outcome** of those two. Most confusion about point-in-time
backtesting comes from mixing the three, which a single wide price table
encourages you to do.

## Declaring The Facts

Everything below is something *you* assert about the world. The mechanical
reshaping is folded away; the claims are not.

```{r}
#| label: bars-from
#| code-fold: true
#| code-summary: "Turning the wide price table into bars (mechanical)"

bars_from <- function(prices) {
  bind_rows(
    tibble(instrument_id = "AAA", date = prices$date, close = prices$AAA),
    tibble(instrument_id = "BBB", date = prices$date, close = prices$BBB)
  ) |>
    filter(is.finite(close)) |>
    transmute(
      instrument_id,
      ts_utc = date,
      open = close,
      high = close + 1,
      low = close - 1,
      close,
      volume = 1000
    )
}
```

### Which sessions were expected?

A **session** is one trading day at one venue: the day the venue was open, and
the hours it was open for. It is a property of the venue, not of your data. A
venue has sessions on days when nothing you own happened to trade.

So the calendar is declared, not inferred from the rows you received. If you
infer sessions from observations, a day with no rows never existed, and your
history silently shortens. Build it as an ordinary table first:

```{r}
#| label: calendar-table

calendar <- tibble(
  session_date = session_dates,
  status = if_else(open_day, "open", "closed"),
  session_open = if_else(open_day, "14:30:00", NA_character_),
  session_close = if_else(open_day, "21:00:00", NA_character_)
)

calendar
```

Read the columns as four separate assertions:

| Column | What you are asserting |
| --- | --- |
| `session_date` | This calendar day existed at the venue. |
| `status` | Whether it was `open` or `closed`. Closed days are supplied too — that is how a weekend is distinguished from a missing file. |
| `session_open` | The time execution prices are taken from. Fills are priced at the next session's open. |
| `session_close` | The time decisions are made, and the pulse each fill is booked on. |

Supplying the closed days matters. Had you listed only the ten open days, a
holiday and a failed download would look identical. Here they do not.

In production this table comes from the venue or an exchange-calendar package,
covering the whole period you intend to run — not from the instrument rows you
happen to hold.

```{r}
#| label: sessions

session_facts <- ledgr_facts_sessions(
  calendar |> mutate(knowledge_time = ledgr_utc("2020-01-01")),
  venue_id = "DEMO",
  timezone = "UTC"
)
```

`venue_id` names the venue these sessions belong to. `knowledge_time` says the
calendar was published in advance, on January 1 — a calendar is known before
the period it describes, which is what makes it usable for decisions inside it.

January 9 is now a real session on which `BBB` had no observation, rather than
a date that never happened.

### Which companies belonged, and when was that knowable?

Start from what a provider actually hands you: a constituent list, as of a
date. Here there are two. On January 6 the universe was both companies; from
January 13, after `AAA` was removed, it was `BBB` alone.

```{r}
#| label: membership-lists

members_from_06 <- c("AAA", "BBB")
members_from_13 <- c("BBB")
```

Each list also needs a second date: when that list became *knowable*. The
January 6 list was published with the calendar on January 1. The removal was
announced on January 13, the day it took effect.

Two lists become three rows — one row per member per list:

```{r}
#| label: membership-rows

membership_rows <- bind_rows(
  tibble(
    effective_from = ledgr_utc("2020-01-06"),
    knowledge_time = ledgr_utc("2020-01-01"),
    instrument_id = members_from_06
  ),
  tibble(
    effective_from = ledgr_utc("2020-01-13"),
    knowledge_time = ledgr_utc("2020-01-13"),
    instrument_id = members_from_13
  )
) |>
  mutate(source = "synthetic_membership")

membership_rows
```

The two dates answer two different questions:

- **`effective_from`** — when does this membership apply?
- **`knowledge_time`** — when could a strategy first have known it?

Those clocks are usually different in real data. An index change is decided,
announced, and takes effect on three separate days. A downloaded constituent
list carries neither timestamp, which is why it leaks.

::: {.ledgr-callout .ledgr-callout-tip}
**Which timestamp goes in `knowledge_time`?**

Use the earliest moment you can defend as "we could have read this":

- The provider timestamps the announcement — use that.
- The file has a publication or delivery date — use that.
- You only know when you downloaded it — use the download time. It is later
  than the truth, which makes the run more conservative, not less.

The one answer that is always wrong is copying `effective_from`. That asserts
every change was knowable the instant it took effect, which is exactly the
assumption survivorship bias hides behind.
:::

```{r}
#| label: membership-facts

membership_facts <- ledgr_facts_membership_snapshots(
  membership_rows,
  universe_id = "demo_members",
  complete = TRUE
)
```

`complete = TRUE` is the load-bearing argument. It says each `effective_from`
group is the *entire* membership set on that date, not a partial report of some
members. That is why the January 13 row removes `AAA`: `AAA` is absent from a
set that claims to be complete. Without `complete = TRUE` these rows would only
add members and `AAA` would never leave.

```{r}
#| label: facts

instruments <- tibble(instrument_id = c("AAA", "BBB"))
facts <- ledgr_facts(session_facts, membership_facts)
ledgr_facts_validate(facts, bars_from(prices), instruments)
```

`Can seal: yes` means this bundle and these observations pass the structural
checks. It does not certify that your vendor supplied every relevant fact.

### What the knowledge clock actually changes

This is worth seeing rather than asserting. Build the same membership with the
removal knowable one day later, and ask what the strategy can see at the
January 13 decision.

```{r}
#| label: knowledge-counterfactual

late_membership <- ledgr_facts_membership_snapshots(
  membership_rows |>
    mutate(
      knowledge_time = if_else(
        effective_from == ledgr_utc("2020-01-13"),
        ledgr_utc("2020-01-14"),
        knowledge_time
      )
    ),
  universe_id = "demo_members",
  complete = TRUE
)

visible_membership <- function(member_facts, label) {
  snap <- ledgr_snapshot_from_df(
    bars_from(prices),
    instruments_df = instruments,
    facts = ledgr_facts(session_facts, member_facts),
    db_path = ledgr_temp_store(),
    snapshot_id = paste0("knowledge-", label)
  )
  on.exit(ledgr_snapshot_close(snap), add = TRUE)
  run <- ledgr_run(
    ledgr_experiment(
      snap,
      function(ctx, params) ctx$hold(),
      universe = ledgr_universe_members("demo_members"),
      valuation_policy = ledgr_valuation_stale(max_sessions = 2),
      cost_model = ledgr_cost_zero(),
      # Hold both companies so each one stays on the decision axis even after
      # it leaves the universe. Otherwise there is no row to compare.
      opening = ledgr_opening(
        cash = 10000,
        positions = c(AAA = 1, BBB = 1),
        cost_basis = c(AAA = 100, BBB = 50)
      )
    ),
    run_id = paste0("knowledge-", label)
  )
  on.exit(close(run), add = TRUE)
  ledgr_results(run, "availability") |>
    filter(instrument_id == "AAA", as.Date(ts_utc) == as.Date("2020-01-13")) |>
    transmute(decision = ts_utc, removal_knowable = label, member)
}

bind_rows(
  visible_membership(membership_facts, "13 January"),
  visible_membership(late_membership, "14 January")
)
```

Same effective date, same prices, different answer. When the removal is
knowable on the 13th, `AAA` is already outside the universe at that decision.
When the same change only becomes knowable on the 14th, `AAA` is still a member
on the 13th, because that is all the strategy could have known. Moving one
timestamp changes the opportunity set.

Note how much machinery that answer cost: a snapshot, an experiment, and a
full run, to read back which instruments were members on one date. There is
currently no lighter way to ask. `ledgr_experiment_plan()` reports which fact
families *participate*, not what they *resolve to* on a given day. Until an
inspection helper exists, running a holding strategy and reading the
`availability` evidence is the supported way to check your facts say what you
think they say.

## What A Strategy Returns

Before running anything, be precise about the contract, because the execution
behaviour later depends on it.

- `ctx$hold()` starts from the positions you currently hold.
- The numbers you return are **desired position quantities**, not orders.
- A target is evaluated against execution rules. It does not guarantee a fill.
- An unfilled target does **not** persist. It is not a standing order waiting
  for liquidity.

The strategy below is deliberately plain: on the first session, put equal
weight into every company the declared universe offers, then hold.

```{r}
#| label: strategy

equal_weight_once <- function(ctx, params) {
  if (substr(ctx$ts_utc, 1, 10) != "2020-01-06") {
    return(ctx$hold())
  }
  ids <- ctx$universe
  weights <- ledgr_weights(
    stats::setNames(rep(1 / length(ids), length(ids)), ids),
    universe = ids
  )
  ledgr_target_rebalance(weights, ctx, equity_fraction = params$invested)
}
```

`equity_fraction` is deliberately below 1. Sizing happens at the decision
price, but the fill happens at the next session's open, so the cash a target
actually requires is not yet known when you ask for it. Leaving headroom is a
research choice, and the exercise at the end of this article shows what happens
without it.

## Two Universes, One Everything Else

Now the comparison the opening promised, run properly. Same snapshot, same
strategy, same valuation policy, same starting cash. The only difference is
which universe is declared.

Seal the evidence first. After this point the inputs are frozen and
identified by a hash.

```{r}
#| label: runs

store_path <- ledgr_temp_store(
  file.path(tempdir(), "ledgr_survivorship_bias.duckdb")
)

snapshot <- ledgr_snapshot_from_df(
  bars_from(prices),
  instruments_df = instruments,
  facts = facts,
  db_path = store_path,
  snapshot_id = "survivorship-demo"
)
```

An experiment is the sealed snapshot plus every choice you are making about it.
Here is the whole thing, written out once:

```{r}
#| label: experiment

point_in_time_experiment <- ledgr_experiment(
  snapshot,
  equal_weight_once,
  universe = ledgr_universe_members("demo_members"),
  valuation_policy = ledgr_valuation_stale(max_sessions = 2),
  cost_model = ledgr_cost_zero(),
  opening = ledgr_opening(cash = 10000)
)
```

Read those arguments as four separate decisions. `universe` selects the
membership history you declared, rather than a fixed list. `valuation_policy`
says how long a stale mark may be carried before the run must stop.
`cost_model` and `opening` fix the trading frictions and the starting account.
Only the first of them differs between the two runs below, so the comparison
isolates the universe.

```{r}
#| label: plan

ledgr_experiment_plan(point_in_time_experiment)
```

The plan states which checks actually participate. Membership and sessions are
declared. Trading-status and lifetime are **omitted**, and omitted does not
mean the source said every instrument was active and listed. It means no such
check ran, which is why ledgr never learns that `AAA`'s silence after
January 10 was a delisting rather than a halt or a feed gap.

```{r}
#| label: runs-execute

# The survivor experiment repeats every argument above and changes one, so
# wrap the shared choices rather than retyping them.
declare <- function(universe) {
  ledgr_experiment(
    snapshot,
    equal_weight_once,
    universe = universe,
    valuation_policy = ledgr_valuation_stale(max_sessions = 2),
    cost_model = ledgr_cost_zero(),
    opening = ledgr_opening(cash = 10000)
  )
}

point_in_time <- ledgr_run(
  point_in_time_experiment, params = list(invested = 0.9), run_id = "pit-universe"
)
survivor <- ledgr_run(
  declare(c("BBB")), params = list(invested = 0.9), run_id = "survivor-universe"
)
```

The point-in-time run bought both companies. The survivor run only ever saw
one.

```{r}
#| label: fills

ledgr_results(point_in_time, "fills") |>
  select(ts_utc, instrument_id, side, qty, price)
```

The decision was made on January 6, but the fills are stamped January 7 at
21:00 — the session *close*. That is not the moment of execution, and the two
are worth separating.

::: {.ledgr-callout .ledgr-callout-note}
**Two clocks in one fill row**

- `price` is the **open** of the next session. That is where execution happened.
- `ts_utc` is that session's **close** — the pulse at which the fill is
  recorded and first visible to the strategy.

A target decided at session *t* is priced at session *t+1*'s open and booked on
session *t+1*'s pulse. `ts_utc` answers "which pulse does this belong to",
not "what time did the trade print".
:::

### Compare only what covers the same period

```{r}
#| label: comparison

pit_equity <- ledgr_results(point_in_time, "equity")
survivor_equity <- ledgr_results(survivor, "equity")

last_common <- max(pit_equity$ts_utc)
return_at <- function(equity, ts) equity$equity[equity$ts_utc == ts] / 10000 - 1

tibble(
  universe = c("Survivor list (BBB only)", "Point-in-time (AAA and BBB)"),
  return_to_common_session = c(
    return_at(survivor_equity, last_common),
    return_at(pit_equity, last_common)
  )
)

ledgr_run_info(snapshot, survivor$run_id)$status
ledgr_run_info(snapshot, point_in_time$run_id)$status
```

At the last session both runs cover, January 14, the survivor universe reports
**+9.0%** and the point-in-time universe **−17.1%**: a gap of **26.1 percentage
points**, produced by nothing but the universe declaration.

The two runs do not cover the same horizon. The survivor run finishes and
reports `DONE`, reaching January 17 at +14.4%. The point-in-time run ends early
and reports `INCOMPLETE`. Those full-horizon figures are not comparable to each
other, which is why the number above is taken at the last session they share.

```{r}
#| label: fig-comparison
#| echo: false
#| fig-height: 4
#| fig-width: 8
#| fig-alt: >
#|   Two equity lines from 10,000. The survivor line rises and continues to
#|   17 January. The point-in-time line falls and ends at 14 January, where a
#|   marker shows the run stopped. The remaining intended horizon is shaded.

paths <- bind_rows(
  survivor_equity |> transmute(ts_utc, equity, universe = "Survivor list (BBB only)"),
  pit_equity |> transmute(ts_utc, equity, universe = "Point-in-time (AAA and BBB)")
)
stop_point <- pit_equity[nrow(pit_equity), ]

ggplot2::ggplot(paths, ggplot2::aes(ts_utc, equity, colour = universe)) +
  ggplot2::annotate(
    "rect",
    xmin = last_common, xmax = ledgr_utc("2020-01-17 21:00:00"),
    ymin = -Inf, ymax = Inf, fill = "grey88", alpha = 0.55
  ) +
  ggplot2::annotate(
    "text", x = ledgr_utc("2020-01-16 12:00:00"), y = 10250,
    label = "no point-in-time\nevidence here", size = 3.2, colour = "grey35"
  ) +
  ggplot2::geom_hline(yintercept = 10000, linewidth = 0.4, colour = "grey70") +
  ggplot2::geom_line(linewidth = 1.1) +
  ggplot2::geom_point(
    data = stop_point, ggplot2::aes(ts_utc, equity),
    shape = 4, size = 3.6, stroke = 1.2, colour = "#B4472A", inherit.aes = FALSE
  ) +
  ggplot2::annotate(
    "text", x = last_common, y = 8000, hjust = 1.05, size = 3.4,
    colour = "#B4472A", label = "run stops here"
  ) +
  ggplot2::scale_colour_manual(
    values = c("Survivor list (BBB only)" = "#4F9B63",
               "Point-in-time (AAA and BBB)" = "#B4472A")
  ) +
  ggplot2::labs(
    title = "What the two experiments reported",
    subtitle = "Same strategy and policies. Only the declared universe differs.",
    x = NULL, y = "Equity", colour = NULL,
    caption = "Compare the lines only where both exist. The shaded region has no point-in-time evidence."
  ) +
  ggplot2::theme_minimal(base_size = 13) +
  ggplot2::theme(
    legend.position = "top",
    panel.grid.minor = ggplot2::element_blank(),
    plot.title.position = "plot",
    plot.title = ggplot2::element_text(face = "bold")
  )
```

The universe declaration explains the difference between *these two runs*. It
does not follow that point-in-time backtests cannot finish. This one stops for
a specific, inspectable reason, which the next two sections work through.

## What The Strategy Could See

The availability view reports, for every instrument at every decision, what the
configured logic concluded.

```{r}
#| label: availability

availability <- ledgr_results(point_in_time, "availability")

availability |>
  filter(instrument_id == "AAA") |>
  select(ts_utc, member, held, admissible, priced, mark_source, mark_age)
```

These columns are not stages of one process. They answer four different
questions:

| Field | The question it answers |
| --- | --- |
| `member` | Does this asset belong to the declared universe at this decision? |
| `held` | Do I already own a nonzero quantity of it? |
| `admissible` | Does the configured availability logic permit *new* exposure? |
| `priced` | Can the valuation policy assign it a permissible mark? |

January 13 is the row worth studying. `AAA` is outside the universe, still
owned, and still valued — using a mark that is one session old. Membership,
ownership, and the ability to value a position are three separate things, and
a survivor-filtered file cannot represent any of them, because the instrument
is simply absent.

::: {.ledgr-callout .ledgr-callout-note}
**`priced = TRUE` is not a tradable price**

A permissible mark values a holding on the equity curve. It says nothing about
whether an executable price exists. Fills come only from observed bars.
:::

```{r}
#| label: fig-timeline
#| echo: false
#| fig-height: 3.4
#| fig-width: 8
#| fig-alt: >
#|   Three tracks for AAA against the same dates. Membership goes from member
#|   to non-member on 13 January. Position goes from unheld to held and stays
#|   held. Valuation goes current, then stale, then expired, where the run stops.

aaa <- availability |> filter(instrument_id == "AAA")
tracks <- bind_rows(
  tibble(ts_utc = aaa$ts_utc, track = "Membership",
         state = if_else(aaa$member, "member", "non-member")),
  tibble(ts_utc = aaa$ts_utc, track = "Position",
         state = if_else(aaa$held, "held", "unheld")),
  tibble(ts_utc = aaa$ts_utc, track = "Valuation",
         state = sub("_close$", "", aaa$mark_source))
)
tracks$track <- factor(tracks$track, levels = c("Valuation", "Position", "Membership"))
palette <- c(member = "#4F9B63", `non-member` = "#C8B04F", held = "#4F7FC8",
             unheld = "grey80", current = "#4F9B63", stale = "#C8B04F",
             expired = "#B4472A")

ggplot2::ggplot(tracks, ggplot2::aes(ts_utc, track, fill = state)) +
  ggplot2::geom_tile(width = 60000, height = 0.62, colour = "white", linewidth = 0.7) +
  ggplot2::geom_text(
    ggplot2::aes(label = state), size = 2.9, colour = "grey15", angle = 90
  ) +
  ggplot2::scale_fill_manual(values = palette) +
  ggplot2::labs(
    title = "AAA: three independent tracks",
    subtitle = "Leaving the index, still owning it, and being able to value it are separate.",
    x = NULL, y = NULL
  ) +
  ggplot2::theme_minimal(base_size = 13) +
  ggplot2::theme(
    legend.position = "none",
    panel.grid = ggplot2::element_blank(),
    plot.title.position = "plot",
    plot.title = ggplot2::element_text(face = "bold")
  )
```

## Why The Point-In-Time Run Is Incomplete

`AAA` is delisted, still held, and by January 15 its last observation is three
sessions old. The valuation policy permits two. ledgr will not carry the number
forward and will not quietly drop the position.

```{r}
#| label: incomplete

ledgr_results(point_in_time, "diagnostics") |>
  filter(outcome == "stopped") |>
  select(ts_utc, stage, outcome, reason_code, instrument_id, mark_age)

tibble(
  intended_end = ledgr_utc("2020-01-17 21:00:00"),
  achieved_end = max(pit_equity$ts_utc)
)
```

This is the hand calculation's quiet assumption coming due. Earlier, marking
`AAA` at its last observed price produced a tidy ending number. ledgr declines
to make that assumption for you: no settlement was observed, and no lifetime or
terminal-event fact was declared, so the run keeps the prefix it could value
and labels the rest missing. Inspect the prefix, correct evidence or policy,
then rerun under a new identity.

Corporate actions, delisting cash flows, borrow and short financing,
order-management behaviour, and general imputation are outside this first
availability surface.

::: {.ledgr-callout .ledgr-callout-tip}
**Try it: move the valuation boundary**

Change `max_sessions = 2` to `max_sessions = 1` in `declare()`, use a new
`run_id`, and rerun the point-in-time universe. Predict first: does the run
stop earlier, later, or on the same session? Then compare `achieved_end` and
the `mark_age` on the stop row. One fewer permitted stale session moves the
boundary. It does not change when `AAA` was delisted.
:::

## When The Bar You Need Is Not There

`BBB`'s missing January 9 session teaches a different lesson, so give it its
own small run: ask to halve the `BBB` position on January 8, and again on
January 10.

```{r}
#| label: trim-run

trim_twice <- function(ctx, params) {
  target <- ctx$hold()
  day <- substr(ctx$ts_utc, 1, 10)
  if (day == "2020-01-06") target[["BBB"]] <- 100
  if (day %in% c("2020-01-08", "2020-01-10")) target[["BBB"]] <- 50
  target
}

trim_run <- ledgr_run(
  ledgr_experiment(
    snapshot,
    trim_twice,
    universe = c("BBB"),
    valuation_policy = ledgr_valuation_stale(max_sessions = 2),
    cost_model = ledgr_cost_zero(),
    opening = ledgr_opening(cash = 10000)
  ),
  run_id = "trim-story"
)

bind_rows(lapply(
  ledgr_utc(c("2020-01-08 21:00:00", "2020-01-09 21:00:00", "2020-01-10 21:00:00")),
  function(ts) ledgr_run_explain(trim_run, "BBB", ts)
)) |>
  select(ts_utc, target_before_risk, execution_outcome, execution_reason,
         resulting_position)
```

Follow the dates against the contract stated earlier:

- **January 8 decision** asks for 50 units. Its execution opportunity is the
  next session's open, January 9. `BBB` has no January 9 bar, so the
  explanation records `execution_bar_missing` and the position stays at 100.
- **January 9 decision** holds. The outcome is `no_action` with reason
  `no_target_change`. The January 8 request did not survive as a standing
  order. A target states the position you want *now*.
- **January 10 decision** asks again. Its execution opportunity is the
  January 13 open, where a `BBB` bar exists, so the trim fills.

That is why a missing fill cannot silently become a later liquidation.

## Evidence Survives Reopening

Closing a run handle releases resources. It does not erase what was recorded.

```{r}
#| label: reopen

pit_run_id <- point_in_time$run_id
expected <- ledgr_run_explain(point_in_time, "AAA", ledgr_utc("2020-01-13 21:00:00"))

close(point_in_time)
close(survivor)
close(trim_run)
ledgr_snapshot_close(snapshot)

reopened_snapshot <- ledgr_snapshot_open(
  store_path, "survivorship-demo", verify = TRUE
)
reopened_run <- ledgr_run_open(reopened_snapshot, pit_run_id)

identical(
  ledgr_run_explain(reopened_run, "AAA", ledgr_utc("2020-01-13 21:00:00")),
  expected
)
```

The same recorded decision comes back from a fresh handle, without running the
strategy again.

## Invalid Observations Are Refused By Default

Survivorship is selection you did not make. Malformed data is a different
problem, and ledgr's default is to refuse it rather than guess: a bar whose
high sits below its open prevents sealing, with
`ledgr_availability_validation_failed`. Explicit quarantine
(`invalid_observations = "quarantine"`) is the deliberate alternative, which
hashes the rejected row as evidence and excludes it from runtime bars while the
expected session stays on the calendar. For that workflow, read
`vignette("data-input-and-snapshots", package = "ledgr")`.

::: {.ledgr-callout .ledgr-callout-tip}
**Try it: spend the whole account**

Rerun `point_in_time` with `params = list(invested = 1)` and a new `run_id`.
Predict first: which fills still happen? Sizing uses the January 6 close, but
execution happens at the January 7 open, where `BBB` costs 51 rather than 50.
Check `ledgr_results(run, "diagnostics")` for the execution rows and find the
reason code.
:::

## What This Does And Does Not Establish

The example shows five distinctions worth preserving:

- membership is a dated assertion about opportunity, not a sell order;
- a held former member stays visible until an explicit target changes it;
- valuation marks and execution prices answer different questions;
- expected sessions remain visible when observations are absent;
- incomplete terminal evidence stays incomplete after reopening.

It does not show that ledgr detects survivorship bias. A company omitted
entirely from the evidence you supply stays invisible, and no amount of
downstream machinery recovers it. The narrower claim is the one demonstrated
here: a holding that *is* represented cannot silently disappear because
membership ended or observations stopped.

Nor does sealing certify your source. If membership knowledge time is assumed
from effective time, `ledgr_experiment_plan()` labels the family
`assumption_backed`. If trading-status or lifetime facts are omitted, their
checks remain omitted. The 26.1-point gap belongs to one synthetic fixture
chosen to make the mechanism legible; it is not an estimate of survivorship
bias in any real index.

For sealing details, read
`vignette("data-input-and-snapshots", package = "ledgr")`. For execution timing
and target semantics, read
`vignette("execution-semantics", package = "ledgr")`. For durable run handling,
read `vignette("experiment-store", package = "ledgr")`.

## References

- Brown, S. J., Goetzmann, W. N., Ibbotson, R. G., and Ross, S. A. (1992).
  [Survivorship Bias in Performance Studies](https://doi.org/10.1093/rfs/5.4.553).
  *The Review of Financial Studies*, 5(4), 553-580.
- Shumway, T. (1997).
  [The Delisting Bias in CRSP Data](https://doi.org/10.1111/j.1540-6261.1997.tb03818.x).
  *The Journal of Finance*, 52(1), 327-340.
- Center for Research in Security Prices.
  [Research Data Products](https://www.crsp.org/research/). The permanent-ID
  discussion is useful context for keeping stable instrument identity separate
  from mutable symbols and membership.
- Nasdaq Trader.
  [Trading Halt Codes](https://www.nasdaqtrader.com/Trader.aspx?id=TradeHaltCodes).
  The status vocabulary illustrates why a valid price row and positive evidence
  of tradability are different source claims.

```{r}
#| label: cleanup
#| include: false

close(reopened_run)
ledgr_snapshot_close(reopened_snapshot)
```
