---
title: "Exploratory Sweeps And Candidate Promotion"
output:
  rmarkdown::html_vignette:
    toc: false
  github_document:
    toc: false
    html_preview: false
vignette: >
  %\VignetteIndexEntry{Exploratory Sweeps And Candidate Promotion}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE, purl = FALSE)
```

The examples assume the usual vignette setup:

```{r attach, message=FALSE}
library(ledgr)
library(dplyr)
```

# Sweep Is Exploration

`ledgr_sweep()` is the lightweight exploration surface. It evaluates a
`ledgr_param_grid()` against a `ledgr_experiment()` and returns a
`ledgr_sweep_results` table. It does not write candidate runs to the experiment
store and it does not choose a winner.

`ledgr_run()` is the committed artifact surface. A promoted candidate becomes a
durable run only when you explicitly select it with `ledgr_candidate()` and call
`ledgr_promote()`, which in turn calls `ledgr_run()`.

```text
ledgr_sweep()                 explore
ledgr_candidate()             select one row deliberately
ledgr_promote() / ledgr_run() commit an auditable run
```

Sweep results carry an `evaluation_scope` attribute that defaults to
`"exploratory"`. That label is intentional. A sweep table records what was run;
it does not prove that the selected parameters were evaluated on held-out data.

# Normal Train/Test Discipline

The recommended workflow is:

```text
source bars
  -> train snapshot
  -> test snapshot
  -> sweep on train snapshot
  -> select candidate and lock params
  -> evaluate the locked params on the test snapshot
```

You can create the train and test snapshots manually by filtering source bars
before calling `ledgr_snapshot_from_df()`.

```r
split_at <- ledgr_utc("2022-01-01")

train_bars <- bars |>
  filter(ts_utc < split_at)

test_bars <- bars |>
  filter(ts_utc >= split_at)

train_snapshot <- ledgr_snapshot_from_df(train_bars)
test_snapshot  <- ledgr_snapshot_from_df(test_bars)

train_exp <- ledgr_experiment(
  snapshot = train_snapshot,
  strategy = strategy,
  features = features,
  opening = ledgr_opening(cash = 100000)
)

test_exp <- ledgr_experiment(
  snapshot = test_snapshot,
  strategy = strategy,
  features = features,
  opening = ledgr_opening(cash = 100000)
)

grid <- ledgr_param_grid(
  conservative = list(qty = 10, threshold = 0.010, sma_n = 20),
  moderate     = list(qty = 10, threshold = 0.005, sma_n = 20),
  fast         = list(qty = 20, threshold = 0.005, sma_n = 10)
)

precomputed <- ledgr_precompute_features(train_exp, grid)
results <- ledgr_sweep(train_exp, grid, precomputed_features = precomputed, seed = 2026)
```

Rank explicitly with ordinary R tools. ledgr does not own objective functions or
automatic candidate ranking.

```r
ranked <- results |>
  filter(status == "DONE") |>
  arrange(desc(sharpe_ratio))

candidate <- ledgr_candidate(ranked, 1)
```

The sweep table has one metric context for all candidate metrics. Inspect it
before ranking if the annualization or risk-free-rate assumption matters:

```r
ledgr_metric_context(results)
```

To evaluate out of sample, promote against the held-out experiment and make the
cross-snapshot choice explicit:

```r
test_run <- ledgr_promote(
  test_exp,
  candidate,
  run_id = "momentum_locked_test",
  note = "Locked params selected from train sweep.",
  require_same_snapshot = FALSE
)
```

`require_same_snapshot = FALSE` is required because the candidate came from the
train snapshot and the committed run is intentionally executed on the test
snapshot.

# Same-Snapshot Replay Is Secondary

Sometimes you want to commit the selected candidate on the same snapshot that
produced the sweep result. That is useful for audit, diagnostics, or comparing a
single candidate against the sweep summary. It remains in-sample.

```r
train_run <- ledgr_promote(
  train_exp,
  candidate,
  run_id = "momentum_locked_train",
  note = "Same-snapshot replay of selected exploratory candidate."
)
```

Same-snapshot promotion is protected by default. `ledgr_promote()` requires the
candidate snapshot hash to match the target experiment unless you deliberately
set `require_same_snapshot = FALSE`.

Same-snapshot replay is the direct way to verify that a selected row is
commit-ready. Compare the sweep summary with the committed run's result tables
and metrics:

```r
selected_row <- candidate$row
train_metrics <- ledgr_compute_metrics(train_run)
train_equity <- ledgr_results(train_run, what = "equity")

stopifnot(isTRUE(all.equal(
  selected_row$final_equity[[1]],
  tail(train_equity$equity, 1)
)))
stopifnot(isTRUE(all.equal(
  selected_row$sharpe_ratio[[1]],
  train_metrics$sharpe_ratio
)))
```

# Parameter Grids And Feature Factories

`ledgr_param_grid()` stores candidate labels and parameter lists. Named entries
become sweep row labels. Unnamed entries receive stable `grid_<hash>` labels
derived from canonical params JSON. These labels identify sweep candidates; they
are not committed run IDs.

Indicator parameters are ordinary sweep parameters when the experiment uses a
feature factory:

```r
features <- function(params) {
  list(
    ledgr_ind_sma(params$sma_n),
    ledgr_ind_rsi(params$rsi_n)
  )
}

grid <- ledgr_param_grid(
  sma20_rsi14 = list(sma_n = 20, rsi_n = 14, threshold = 0.010, qty = 10),
  sma50_rsi14 = list(sma_n = 50, rsi_n = 14, threshold = 0.010, qty = 10),
  sma50_rsi21 = list(sma_n = 50, rsi_n = 21, threshold = 0.005, qty = 10)
)
```

Each row stores both `params` and `feature_fingerprints`: the requested
parameters and the resolved feature identities actually used by that candidate.
The row-level `provenance$feature_set_hash` is the compact identity for that
candidate's resolved feature set. Candidates with different indicator
parameters should have different feature-set hashes even when the strategy code
is the same.

Strategy lookup is still explicit. For parameterized built-in indicators, the
current bridge pattern is exact-ID lookup from `params` inside the strategy.
Do not call the feature factory from inside the strategy; user factory helpers
are Tier 3 under preflight. Static feature maps remain useful when aliases do
not vary by candidate. See `vignette("strategy-development", package =
"ledgr")` for the bridge patterns and the active-alias API gap.

# Precompute Larger Grids

`precomputed_features` is optional. For small grids, `ledgr_sweep()` can compute
features internally. When the grid has more than 20 combinations and no
precomputed feature payload is supplied, ledgr warns and suggests
`ledgr_precompute_features()`.

Precompute resolves candidate-specific feature factories, deduplicates shared
indicator definitions by fingerprint, and validates the payload against the
snapshot, universe, scoring range, feature engine version, and grid labels.

The precompute payload is intentionally tied to its inputs. `ledgr_sweep()`
rejects a payload built for a different snapshot hash, universe, scoring range,
grid labels, parameter hashes, feature engine version, or resolved feature
fingerprints. Recompute the payload after changing any of those inputs.

Feature-factory failures are candidate-level when they occur while resolving a
specific candidate's features and `stop_on_error = FALSE`. The failed row keeps
the candidate label, params, error class, and error message so the bad
combination can be inspected without losing the rest of the sweep.

# Failure Rows

By default `stop_on_error = FALSE`. Candidate-level failures become rows with:

- `status = "FAILED"`
- `error_class`
- `error_msg`
- metric columns set to `NA`

Use `stop_on_error = TRUE` when debugging a single failing candidate and you
want ledgr to rethrow the first failure.
When asserting on a rethrown strategy failure in tests, prefer
`inherits(e, "ledgr_strategy_error")` rather than exact class-vector equality;
the original strategy error class is preserved alongside ledgr's generic class.

`ledgr_candidate()` rejects failed rows by default. Use
`ledgr_candidate(results, which, allow_failed = TRUE)` only for diagnostic
extraction of the failed candidate's params, error, warnings, and provenance.
`ledgr_promote()` still rejects failed candidates; promotion is only for
committing a runnable candidate through `ledgr_run()`.

Contract errors still abort before or during the sweep. Invalid grids, invalid
precomputed feature payloads, including snapshot-hash mismatches, and Tier 3
strategy preflight failures are not candidate results.

# Seeds, Provenance, And Promotion Context

When `seed` is supplied to `ledgr_sweep()`, each candidate receives a stable
`execution_seed` derived from the master seed, the candidate label, and the
candidate params. `ledgr_promote()` forwards that exact seed into `ledgr_run()`
so a promoted same-snapshot candidate can reproduce the selected sweep result.
If `seed = NULL`, row-level `execution_seed` is `NA` and promotion does not
inject a replay seed.

Each row also carries compact `provenance`, including snapshot hash, strategy
hash, feature-set hash, master seed, seed contract, and evaluation scope.
Provenance records what ran. It does not prove that your selection process was
out-of-sample.

The candidate table itself is not a full artifact store. Treat sweep columns as
candidate summaries: `run_id`, `status`, `params`, `execution_seed`, metric
columns, `feature_fingerprints`, `provenance`, and warning/error fields are the
programmatic inspection surface. Full equity, fills, trades, and ledger rows are
created only by committed runs.

Warnings are part of candidate inspection. A candidate can finish with warnings,
including `LEDGR_LAST_BAR_NO_FILL`, when a strategy changes target on the final
available bar and there is no later bar for a next-open fill. Inspect the row's
warning fields before promotion; after promotion, inspect the committed run with
`summary(test_run)`, `ledgr_results(test_run, what = "fills")`, and
`ledgr_promotion_context(test_run)`. If the final-bar warning is expected,
extend the snapshot by one executable bar to verify that the intended target
would fill.

Promoted runs write durable promotion context. You can inspect it later:

```r
context <- ledgr_promotion_context(test_run)
context$selected_candidate
context$source_sweep
ledgr_metric_context(context)
ledgr_metric_context(test_run)

ledgr_run_promotion_context(test_exp, "momentum_locked_test")
ledgr_run_info(test_exp$snapshot, "momentum_locked_test")$promotion_context
```

The context stores a compact selected-candidate record, source-sweep metadata,
and the filtered/sorted candidate-summary view that was passed to
`ledgr_candidate()`. It does not store full ledger rows, full equity curves, or a
complete sweep artifact.

The source sweep metric context and the committed run metric context remain
separate. Use `ledgr_metric_context(ledgr_promotion_context(test_run))` for the
context that ranked the candidate, and `ledgr_metric_context(test_run)` for the
committed run's default analysis context.

For result inspection after promotion, use the normal run tools such as
`summary(test_run)`, `ledgr_compute_metrics(test_run)`, and
`ledgr_compare_runs()`. See `vignette("metrics-and-accounting", package =
"ledgr")` for the event, fills, equity, and metric views.

# Explicit Non-Goals

Current sweep mode does not ship:

- automatic ranking, objectives, or `ledgr_tune()`
- parallel sweep execution
- walk-forward, PBO, or CSCV helpers
- risk-layer insertion
- public cost-model factories
- paper/live trading adapters
- intraday-specific support
- `ledgr_save_sweep()` or full sweep artifact persistence

The full sweep-artifact idea is deferred. Current sweep mode stores selection
context on promoted runs instead.
