---
title: "Exploratory Sweeps And Candidate Promotion"
format:
  html:
    toc: false
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{Exploratory Sweeps And Candidate Promotion}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

options(width = 90)
options(cli.unicode = FALSE)
pkg_root <- if (file.exists("DESCRIPTION")) "." else if (file.exists("../DESCRIPTION")) ".." else NA_character_
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
default_output_hook <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(
  output = function(x, options) {
    x <- gsub("Database:(\\s+)[^\\r\\n]+[.]duckdb", "Database:\\1<temporary DuckDB path>", x, perl = TRUE)
    x <- gsub("[ \t]+(?=\n)", "", x, perl = TRUE)
    default_output_hook(x, options)
  }
)
```

```{=html}
<style>
.ledgr-diagram {
  margin: 1.25rem auto 1.5rem auto;
  text-align: center;
}

.ledgr-diagram .mermaid {
  display: inline-block;
  max-width: 100%;
}

.ledgr-diagram .mermaid svg {
  display: block;
  height: auto !important;
  margin-left: auto;
  margin-right: auto;
}

.ledgr-grid-diagram .mermaid svg {
  max-width: 760px !important;
}

.ledgr-alias-diagram .mermaid svg {
  max-width: 820px !important;
}

.ledgr-validation-diagram .mermaid svg {
  max-width: 520px !important;
}
</style>
```

A sweep answers a parameter question: which values of a declared set of knobs
did what, on the same sealed evidence? This article turns such a question into
candidate rows, inspects them, and promotes one candidate into a committed run
you can reopen later. It assumes you have already written and run the strategy
once; `vignette("research-workflow", package = "ledgr")` covers that loop and
when a parameter question is the right next step.

Sweeps are for exploration. Promotion records one selected candidate as a
committed run. Neither step proves that the selected candidate will generalize.

## Prerequisites

```{r}
#| label: prerequisites

library(ledgr)
library(dplyr)
data("ledgr_demo_bars", package = "ledgr")
```

This article uses `dplyr` for tabular inspection. The sweep itself is ledgr's
job.

::: {.ledgr-callout .ledgr-callout-note}
**Running this yourself**

This article is evaluated when it is rendered, so it writes to a temporary
store. In a real project, replace `store_path` with a durable path such as
`artifacts/ledgr_store.duckdb`.
:::

```{r}
#| label: store

store_path <- ledgr_temp_store()
```

## Sweep Is Exploration

`ledgr_sweep()` evaluates a declared grid against a `ledgr_experiment()`. The
result is an exploratory candidate table: it records what each candidate did,
does not choose a winner, and does not write candidate runs to the experiment
store.

Each **candidate** is one row of that table: resolved feature parameters,
strategy parameters, execution seed, status, metrics, warnings or errors, and
provenance.

That separation is the workflow boundary:

```text
ledgr_sweep()                 explore declared candidates
ledgr_candidate()             select one row deliberately
ledgr_promote() / ledgr_run() commit an auditable run
```

::: {.ledgr-callout .ledgr-callout-warning}
**Selection is not validation**

A sweep table records what was run. It does not prove that the selected
parameters were evaluated on held-out data. Promotion records a choice; it does
not make that choice out-of-sample.
:::

## Declare Parameterized Features

Seal the evidence first. Here you use two demo instruments over the first half
of 2019. Once the data is sealed, the snapshot becomes the identity of the
evidence: every run below refers back to the same immutable input.

```{r}
#| label: experiment-setup

bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% c("DEMO_01", "DEMO_02"),
    between(ts_utc, ledgr_utc("2019-01-01"), ledgr_utc("2019-06-30"))
  )

snapshot <- ledgr_snapshot_from_df(
  bars,
  db_path = store_path,
  snapshot_id = "demo_2019_h1"
)

features <- ledgr_feature_map(
  fast = ledgr_ind_sma(ledgr_param("fast_n")),
  slow = ledgr_ind_sma(ledgr_param("slow_n"))
)

sma_crossover_strategy <- function(ctx, params) {
  targets <- ctx$flat()

  for (id in ctx$universe) {
    values <- ctx$features(id)
    if (
      ledgr_passed_warmup(values) &&
        ((values[["fast"]] / values[["slow"]]) - 1) > params$threshold
    ) {
      targets[id] <- params$qty
    }
  }

  targets
}

exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = sma_crossover_strategy,
  features = features,
  opening = ledgr_opening(cash = 10000),
  cost_model = ledgr_cost_zero()
)

exp
```

The strategy function itself does not change across candidates. At each pulse,
ledgr calls the same `function(ctx, params)`: `ctx` contains the current
pulse-known market data, positions, cash, equity, and resolved alias values;
`params` contains the strategy-parameter values for this candidate.

The function above is the strategy the experiment executes; it is not a
separate illustrative approximation.

During a sweep, `ctx$features(id)` returns values for the concrete indicators
resolved for that candidate. `params$threshold` and `params$qty` come from the
strategy grid. For the full strategy contract, read
`vignette("strategy-development", package = "ledgr")`.

The per-instrument read is necessary for active aliases today: ledgr does not
yet expose an alias-aware whole-universe vector accessor. Fixed feature IDs
should use `ctx$vec$feature(feature_id)`; this loop is a disclosed boundary,
not the preferred pattern for ordinary maps.

An **active alias** is a stable strategy-facing feature name whose concrete
indicator can vary by candidate. The strategy can keep reading
`values[["fast"]]` and `values[["slow"]]` even when one candidate uses SMA(5)
and SMA(20) and another candidate uses SMA(10) and SMA(40). The alias is the
strategy-facing contract. The concrete feature IDs and fingerprints are
provenance.

Feature parameters vary the knobs exposed by a feature constructor. For
`ledgr_ind_sma()`, the knob is `n`, the moving-average window. For TTR-backed
features, the knobs are the supported arguments of the wrapped TTR indicator.
You still need to understand the feature you are tuning; ledgr separates the
parameter namespaces, but it does not decide which indicator arguments are
economically meaningful.

Only knobs declared with `ledgr_param("name")` need values in the feature grid.
Concrete arguments stay fixed. For example, `ledgr_ind_sma(20)` needs no
`feature_grid` entry, while `ledgr_ind_sma(ledgr_param("fast_n"))` requires a
scalar `fast_n` value for each candidate.

::: {.ledgr-diagram .ledgr-alias-diagram}

```{mermaid}
%%| echo: false

flowchart TB
  params["feature params<br/>fast_n = 5<br/>slow_n = 20"]
  indicators["concrete indicators<br/>SMA 5 and SMA 20"]
  aliases["stable aliases<br/>fast and slow"]
  strategy["strategy reads<br/>fast and slow"]

  params --> indicators --> aliases --> strategy
```

:::

For a second candidate with `fast_n = 10` and `slow_n = 40`, the same arrows
resolve to SMA(10) and SMA(40), but the strategy still reads `values[["fast"]]`
and `values[["slow"]]`. The aliases stay stable across candidates.

## Build The Candidate Grid

**Feature parameters** materialize indicators before execution. **Strategy
parameters** are passed to `strategy(ctx, params)` during execution. Keeping
those namespaces separate is what lets a strategy read stable aliases while the
sweep varies indicator windows. Use `ledgr_feature_grid()` for feature knobs,
`ledgr_strategy_grid()` for strategy-code knobs, and `ledgr_grid_cross()` to
combine them.

```{r}
#| label: grids

feature_grid <- ledgr_feature_grid(
  fast_n = c(5L, 10L),
  slow_n = c(20L, 40L),
  .filter = fast_n < slow_n
)

strategy_grid <- ledgr_strategy_grid(
  threshold = c(0, 0.01),
  qty = c(5, 10)
)

grid <- ledgr_grid_cross(features = feature_grid, strategy = strategy_grid)
grid
```

The `.filter` expression is a structural grid constraint. Here it says the
fast moving average must be shorter than the slow moving average. Filter
expressions are evaluated against grid columns; they do not read run state,
feature data, or caller globals.

::: {.ledgr-callout .ledgr-callout-warning}
**Mind the combinatorial explosion**

`ledgr_grid_cross()` multiplies grid dimensions. Four parameters with five
values each produce 625 candidates before you add another axis. Keep early
sweeps deliberately small, then expand only after the candidate table is
readable and the feature payload cost is understood.
:::

The cross product is explicit:

::: {.ledgr-diagram .ledgr-grid-diagram}

```{mermaid}
%%| echo: false

flowchart TB
  fg["feature_grid<br/>fast_n, slow_n"]
  sg["strategy_grid<br/>threshold, qty"]
  cross["grid_cross"]
  rows["candidate rows<br/>feature params + strategy params"]

  fg --> cross
  sg --> cross
  cross --> rows
```

:::

::: {.ledgr-callout .ledgr-callout-tip}
**Try it**

Run this larger grid. The strategy code and its `fast` and `slow` aliases do
not change; only the concrete feature parameters do.

```{r}
expanded_grid <- ledgr_grid_cross(
  features = ledgr_feature_grid(
    fast_n = c(5L, 10L),
    slow_n = c(20L, 40L, 60L),
    .filter = fast_n < slow_n
  ),
  strategy = strategy_grid
)

tibble(
  grid_name = c("original", "expanded"),
  candidates = c(length(grid$params), length(expanded_grid$params)),
  feature_combinations = c(
    nrow(distinct(bind_rows(lapply(grid$params, `[[`, "feature_params")))),
    nrow(distinct(bind_rows(lapply(
      expanded_grid$params,
      `[[`,
      "feature_params"
    ))))
  )
)
```
:::

## Precompute Shared Features

Precomputing is an execution optimization, not a separate research decision.
The same declared feature grid is resolved once, deduplicated by fingerprint,
and reused across sweep candidates.

```{r}
#| label: precompute

precomputed <- ledgr_precompute_features(exp, grid)
precomputed
```

`ledgr_sweep()` can compute features internally for small grids. For larger
grids, precompute first so feature resolution and payload validation are
explicit. When a grid has more than 20 combinations and no precomputed payload,
ledgr warns because feature computation may be repeated per candidate.

## Run The Sweep

Give the sweep a master seed when reproducible stochastic strategy behavior
matters. Each row receives its own derived `execution_seed`. The default sweep
is scalar-only: it keeps candidate summary rows but not per-pulse return
series.

```{r}
#| label: run-scalar-sweep

sweep <- ledgr_sweep(
  exp,
  grid,
  precomputed_features = precomputed,
  seed = 2026L
)

sweep
candidate_table <- bind_cols(
  sweep |> select(candidate_id, status),
  bind_rows(lapply(sweep$params, as.data.frame)),
  bind_rows(lapply(sweep$feature_params, as.data.frame))
)
candidate_table
```

The second view connects each status row to the strategy and feature parameters
that produced it. The table is not a full artifact store and it
does not write durable candidate ledgers, equity curves, feature panels, or
telemetry rows. Each row keeps the compact reproduction key needed for later
materialization: snapshot identity, selector, strategy identity, feature
fingerprints, seed metadata, and candidate params. Use
`ledgr_candidate_reproduction_key()` when you want to inspect that key directly.
Full equity, fills, trades, and ledger rows are created only by committed runs.

`ledgr_sweep()` is the memory-backed exploration mode. It evaluates candidate
rows through the same fold semantics as committed runs, but it keeps compact
candidate evidence instead of writing a durable ledger and equity curve for
every row. Promotion is the point where one selected candidate pays the
durable-materialization cost.

If the experiment declares a `risk_chain`, sweep candidates also carry
`risk_chain_hash` and row-level provenance carries `risk_plan_json`. These fields
are execution identity: they say which target-risk plan transformed the
validated strategy targets before fill timing and cost resolution. They do not
rank candidates, select winners, estimate liquidity, enforce broker policy, or
turn risk settings into a separate grid-composition surface. Parameterized risk
arguments use ordinary candidate params through `ledgr_param()`.

## Retain Candidate Return Series

When you need per-pulse net portfolio equity or adjacent-period returns for
completed candidates, opt in explicitly with `ledgr_sweep_retention()`.

```{r}
#| label: retain-candidate-returns

retained_sweep <- ledgr_sweep(
  exp,
  grid,
  precomputed_features = precomputed,
  seed = 2026L,
  retain = ledgr_sweep_retention("completed")
)

retained_long <- ledgr_sweep_returns(retained_sweep)

retained_long |>
  select(sweep_id, candidate_id, ts_utc, equity, period_return) |>
  slice_head(n = 8)
```

`period_return` is `NA_real_` on the first retained row for each candidate
because there is no prior equity value to compare against. Drop that first
return before handing the series to an external metric package:

```{r}
#| label: retained-returns-for-external-metrics

external_metric_input <- retained_long |>
  filter(!is.na(period_return)) |>
  select(ts_utc, candidate_id, period_return)

external_metric_input |>
  slice_head(n = 8)
```

The retained series are net strategy returns. They include the execution costs
resolved by the experiment's cost model, but they are not benchmark-relative
returns and they do not contain gross-vs-net attribution. Failed candidates
remain in the sweep summary table but have no retained return rows.

Final-bar no-fill warnings are row-level warnings on the candidate summary. They
do not remove the final equity row from retained return series.

Use the wide accessor when a downstream tool expects one return or equity
column per candidate:

```{r}
#| label: retained-returns-wide

ledgr_sweep_returns_wide(
  retained_sweep,
  candidates = retained_sweep$candidate_id[1:3],
  value = "returns"
) |>
  slice_head(n = 5)
```

## Save And Reopen Sweep Artifacts

A saved sweep is a compact artifact. It stores candidate summary evidence and,
when requested, retained net equity/return series. It is not a batch of
committed runs: full ledgers, fills, trades, and per-instrument artifacts remain
available only after explicit promotion.

```{r}
#| label: save-and-reopen-sweep

saved_id <- ledgr_sweep_save(
  retained_sweep,
  snapshot,
  sweep_id = "sma_retained_sweep",
  note = "Exploratory SMA sweep with retained return series."
)

ledgr_sweep_list(snapshot)

reopened_sweep <- ledgr_sweep_open(snapshot, saved_id)

ledgr_sweep_info(reopened_sweep)
```

Reopened sweeps behave like sweep result objects for candidate extraction,
ordinary dplyr inspection, retained return access, and promotion. Promotion from
a reopened saved sweep re-executes the selected candidate from its reproduction
key against the sealed snapshot; it does not replay precomputed retained return
rows as if they were a committed ledger.

```{r}
#| label: reopened-retained-returns

ledgr_sweep_returns(reopened_sweep) |>
  filter(candidate_id == reopened_sweep$candidate_id[[1]]) |>
  slice_head(n = 5)
```

## Three Retention Levels

Sweeps give you three retention levels. Use the cheapest level
that answers the question in front of you, then promote only the candidate that
needs committed-run artifacts.

| Level | What it keeps | Typical use |
| --- | --- | --- |
| Scalar row | Candidate params, status, scalar metrics, warnings, and reproduction key | Screen and debug a declared grid |
| Retained series | Scalar row plus net portfolio equity and period returns for completed candidates | Inspect return shape, rolling behavior, or external metric-package inputs |
| Promoted run | Full committed run artifacts: ledger, fills, trades, equity, metrics, and promotion context | Audit and reopen the selected candidate as a durable run |

Retention does not change execution identity. It changes which derived evidence
is kept after the same candidate execution.

Saved sweeps persist the same risk identity fields on the parent sweep row and
candidate rows. Opening an older store may perform an additive saved-sweep
schema migration for `risk_chain_hash` and `risk_plan_json`; it does not rewrite
candidate results or create committed run artifacts.

## What Retained Returns Can And Cannot Validate

Retained returns make triage better because they let you inspect the path of a
candidate, not just its final scalar score. They can help you notice unstable
return profiles, drawdown concentration, missing warmup behavior, and candidates
whose scalar score hides a bad path.

They do not make the sweep statistically valid. If you selected the candidate
from the same sample, the retained path is still in-sample evidence. Generalized
validation belongs to a held-out evaluation, walk-forward analysis, or the
selection-integrity diagnostics taught in
`vignette("selection-integrity", package = "ledgr")`.

::: {.ledgr-callout .ledgr-callout-warning}
**Return paths are not validation by themselves**

Retained series preserve more evidence from a sweep. They do not prove that the
candidate-selection process was sound.
:::

## Why ledgr And PerformanceAnalytics Metrics May Differ

Retained return series can be shaped for packages such as PerformanceAnalytics,
but ledgr metrics and external package metrics may differ. Annualization,
calendar assumptions, treatment of the leading missing return, and return-shape
conventions are package contracts, not universal truths.

Keep ledgr's scalar metric rows as the canonical ledgr evidence. Use external
metric packages as additional analysis over an explicit return series, and label
any overlapping headline metric when it comes from a different convention.

## Compiled Accounting And Workers

The default sweep path is memory-backed and uses the canonical R accounting
fold. When your workload is spot-asset FIFO and you want the scoped compiled
accelerator, opt in explicitly:

```{r}
#| label: sweep-spot-fifo-shape
#| eval: false

sweep <- ledgr_sweep(
  exp,
  grid,
  precomputed_features = precomputed,
  seed = 2026L,
  compiled_accounting_model = "spot_fifo"
)
```

`compiled_accounting_model = NULL` remains the default. `"spot_fifo"` is a
memory-backed sweep accelerator only: it is not a general compiled execution engine,
not the durable `ledgr_run()` path, not a non-spot accounting model, and not
enabled by default. A committed `ledgr_run()` that requests it fails closed.
Unsupported model names raise `ledgr_unsupported_accounting_model`, and missing
compiled support raises `ledgr_compiled_spot_fifo_unavailable`.

For independent candidates, `workers` can dispatch sweep rows in parallel when
the required backend and worker package dependencies are available. Parallelism
changes candidate dispatch, not strategy semantics; interrupted parallel sweeps
discard the partial table instead of returning partially promotable rows.
That means parallel sweep execution is a dispatch choice over independent
candidate rows, not a second execution engine.

## Failure Rows And Contract Errors

By default, candidate-level failures become rows with `status = "FAILED"`.
Inspect those rows before selecting anything.

This debug example uses no features so the failure is only about strategy
parameters. In ordinary research, the experiment usually declares features just
like the main sweep above.

```{r}
#| label: failed-rows

debug_strategy <- function(ctx, params) {
  if (params$qty < 0) {
    stop("qty must be non-negative")
  }
  ctx$flat()
}

debug_exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = debug_strategy,
  features = list(),
  opening = ledgr_opening(cash = 10000),
  cost_model = ledgr_cost_zero()
)

debug_grid <- ledgr_strategy_grid(qty = c(5, -1))

failed_sweep <- ledgr_sweep(debug_exp, debug_grid)

failed_sweep |>
  select(candidate_id, candidate_row, status, error_class, error_msg, params)
```

Use the failed row as an interactive debugging handle: inspect `error_class`,
`error_msg`, and `params`, then reproduce the single candidate in a smaller
session before rerunning the sweep. For pulse-level strategy debugging patterns,
read `vignette("strategy-development", package = "ledgr")`.

Contract errors still abort before a candidate table exists. Invalid experiment
or grid shape is not a failed strategy idea; it is a setup problem that ledgr
stops immediately so the sweep does not mix incomparable rows.

`ledgr_candidate()` rejects failed rows by default. Use
`ledgr_candidate(results, which, allow_failed = TRUE)` only when you are
extracting a failed row for diagnostics; `ledgr_promote()` still rejects failed
candidates.

## Inspect Before Promotion

Name the ranking rule before selecting. Here the rule is deliberately simple:
among completed candidates, sort by Sharpe ratio descending. ledgr ships
business-objective helpers for explicit selection rules, as taught in
`vignette("selection-integrity", package = "ledgr")`; it does not silently
choose a winner. This example keeps the rule visible as ordinary R code, and it
reads the reopened saved sweep to show that a saved sweep reopens to the same
dplyr-friendly surface as the in-session result.

```{r}
#| label: inspect-candidates

review <- ledgr_sweep_review(reopened_sweep, rank_by = desc(sharpe_ratio), n = 5)
review$top |>
  select(candidate_id, sharpe_ratio, params, feature_params)
review$issues
ranked <- review$ranked
```

`ledgr_sweep_review()` packages the inspection shape while keeping the ranking
rule explicit in the call. `review$top` is for the candidates you want to read
closely, `review$issues` is for warnings and failures, and `review$ranked`
retains the full completed-candidate table for deliberate selection.

::: {.ledgr-callout .ledgr-callout-tip}
**Try it**

Sort by `total_return` instead of `sharpe_ratio`. Does the first candidate
change? If it does, your "best" candidate depends on the metric, not only on
the strategy.
:::

## Promote One Candidate With A Note

Promotion replays one selected candidate as a committed run and attaches the
human selection note. This is the slow path that explicitly pays to
materialize durable ledger and equity artifacts.

```{r}
#| label: promote-selected

candidate <- ledgr_candidate(ranked, 1)

candidate_table |>
  filter(candidate_id == candidate$candidate_id)

promoted_run <- ledgr_promote(
  exp,
  candidate,
  run_id = "workflow_promoted_candidate",
  note = paste(
    "Promoted from an exploratory same-snapshot sweep for workflow review.",
    "This note records the selection rationale; it is not statistical validation."
  )
)

summary(promoted_run)
```

The promoted run is now a committed run with its own run ID. The selection note
travels with the run. Same-snapshot promotion is useful for audit, debugging,
and durable storage of the chosen candidate. It remains in-sample.

## Reopen The Promoted Run

The payoff for the project store is that you can come back later without
re-running the strategy. Release the live handles, reopen the snapshot with
hash verification, and open the promoted run by its ID.

```{r}
#| label: reopen

close(promoted_run)
ledgr_snapshot_close(snapshot)

snapshot <- ledgr_snapshot_open(
  store_path,
  snapshot_id = "demo_2019_h1",
  verify = TRUE
)

ledgr_run_list(snapshot)

reopened <- ledgr_run_open(snapshot, "workflow_promoted_candidate")
reopened
```

Three public readers recover what ledgr recorded about the promoted run:

- `ledgr_run_info()` for the run's identity and completion evidence;
- `ledgr_promotion_context()` for the selected sweep row, its parameters and
  the promotion note;
- `ledgr_run_strategy()` for strategy source and parameter provenance.

```{r}
#| label: recover-provenance

promotion <- ledgr_promotion_context(reopened)
list(
  source = promotion$source,
  selected_candidate = promotion$selected_candidate$candidate_id,
  selected_candidate_row = promotion$selected_candidate$candidate_row,
  strategy_params_json = promotion$selected_candidate$params_json,
  feature_params_json = promotion$selected_candidate$feature_params_json,
  note = promotion$note
)

extracted <- ledgr_run_strategy(
  snapshot,
  "workflow_promoted_candidate",
  trust = FALSE
)

extracted$strategy_params
extracted$reproducibility_level
extracted$hash_verified
```

The result is compact but load-bearing: ledgr stores which sweep candidate was
selected, which strategy parameters and feature parameters produced it, what
note justified the promotion, and which strategy source metadata was captured.

::: {.ledgr-callout .ledgr-callout-note}
**What recovery means**

Recovery is provenance, not magic. Tier 1 strategy source can usually be
inspected, hash-checked, and optionally evaluated with `trust = TRUE`. Tier 2
strategies may depend on external functions or package state, so ledgr records
the source text, hashes, parameters, dependency metadata, and warnings that
explain what was captured and what remains outside the run artifact.
:::

## Plot The Promoted Evidence

A report should show the promoted run's path, not only its scalar metrics.
The equity table already contains the information needed for a compact equity
and drawdown view.

```{r}
#| label: promoted-equity-plot
#| fig.width: 7
#| fig.height: 4

promoted_equity <- ledgr_results(reopened, what = "equity") |>
  mutate(drawdown_axis = drawdown * equity[[1]] + equity[[1]])

ggplot2::ggplot(promoted_equity, ggplot2::aes(x = ts_utc)) +
  ggplot2::geom_line(
    ggplot2::aes(y = equity),
    linewidth = 0.8,
    color = "#1f77b4"
  ) +
  ggplot2::geom_area(
    ggplot2::aes(y = drawdown_axis),
    fill = "#d55e00",
    alpha = 0.20
  ) +
  ggplot2::labs(
    title = "Promoted candidate equity path",
    x = NULL,
    y = "Equity",
    caption = "Orange area scales drawdown onto the equity axis for compact review."
  ) +
  ggplot2::theme_minimal(base_size = 12)
```

## What Promotion Does Not Prove

You now have a promoted candidate. It means you can reopen the exact evidence
trail: the sealed data, feature declarations, parameter values, run identity,
candidate row, and promotion note. It does not mean the strategy will
generalize.

A single-window sweep is exploratory evidence with an audit trail. Naive
sweep-and-pick selection is a selection-bias risk because every candidate was
compared on the same evidence window. Picking the highest metric from that
window can overfit the sample even when every individual run was deterministic
and leakage-safe at the pulse boundary. The more candidates you try, the more
opportunity you create for sample-specific luck to look like skill.

::: {.ledgr-diagram .ledgr-validation-diagram}

```{mermaid}
%%| echo: false
%%{init: {"theme": "base", "flowchart": {"nodeSpacing": 18, "rankSpacing": 22, "curve": "linear"}, "themeVariables": {"fontFamily": "system-ui, -apple-system, Segoe UI, sans-serif", "fontSize": "15px", "primaryColor": "#f8fafc", "primaryTextColor": "#1f2937", "primaryBorderColor": "#64748b", "lineColor": "#64748b", "secondaryColor": "#eef2ff", "secondaryTextColor": "#1f2937", "secondaryBorderColor": "#64748b", "tertiaryColor": "#fff7ed", "tertiaryTextColor": "#1f2937", "tertiaryBorderColor": "#fb923c"}}}%%

flowchart TB
  window["Same evidence window<br/>Candidate 1<br/>Candidate 2<br/>Candidate 3<br/>Candidate ..."]
  pick["Pick highest metric"]
  caveat["Selection recorded<br/>not validated"]

  window --> pick --> caveat
```

:::

If the question is generalization rather than artifact reproducibility, use
`vignette("selection-integrity", package = "ledgr")` and
`vignette("walk-forward", package = "ledgr")`.

This is the same selection-bias boundary that separates reproducible sweep
evidence from walk-forward or later validation-toolkit evidence.

::: {.ledgr-callout .ledgr-callout-tip}
**Try it**

Write one sentence explaining why you promoted the candidate and one sentence
explaining why that promotion is not validation. If the second sentence feels
hard to write, the selection rule probably needs more work.
:::

## Cost Models Are Fixed Inputs

Cost models are part of the experiment identity in this release. A sweep varies
feature parameters and strategy parameters across the declared grid; it does not
compose cost models as another grid dimension. If you want to compare different
cost assumptions, run separate experiments or separate sweeps with explicit
`cost_model` values and compare the resulting evidence.

A future `ledgr_cost_grid()` may make cost assumptions participate in candidate
identity deliberately. That API has not shipped, so do not expect
`ledgr_grid_cross()` to accept cost-model dimensions.

## Explicit Non-Goals

Sweep mode intentionally leaves some decisions outside the API. It does not
ship:

- automatic winner selection or `ledgr_tune()`;
- cost-grid composition such as `ledgr_cost_grid()`;
- paper/live trading adapters;
- intraday-specific support;
- full per-candidate committed-run artifacts.

## Where Next

- For selection-bias diagnostics over a sweep, read
  `vignette("selection-integrity", package = "ledgr")`.
- For held-out evaluation of a selection rule, read
  `vignette("walk-forward", package = "ledgr")`.
- For feature maps, indicator identity, and active aliases, read
  `vignette("indicators", package = "ledgr")`.
- For durable run comparison after promotion, read
  `vignette("experiment-store", package = "ledgr")`.

```{r}
#| label: cleanup
#| include: false

close(reopened)
ledgr_snapshot_close(snapshot)
```
