---
title: "Exploratory Sweeps And Candidate Promotion"
format:
  html:
    toc: false
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{Exploratory Sweeps And Candidate Promotion}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

options(width = 90)
options(cli.unicode = FALSE)
pkg_root <- if (file.exists("DESCRIPTION")) "." else if (file.exists("../DESCRIPTION")) ".." else NA_character_
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
default_output_hook <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(
  output = function(x, options) {
    x <- gsub("Database:(\\s+)[^\\r\\n]+[.]duckdb", "Database:\\1<temporary DuckDB path>", x, perl = TRUE)
    x <- gsub("[ \t]+(?=\n)", "", x, perl = TRUE)
    default_output_hook(x, options)
  }
)
```

```{=html}
<style>
.ledgr-diagram {
  margin: 1.25rem auto 1.5rem auto;
  text-align: center;
}

.ledgr-diagram .mermaid {
  display: inline-block;
  max-width: 100%;
}

.ledgr-diagram .mermaid svg {
  display: block;
  height: auto !important;
  margin-left: auto;
  margin-right: auto;
}

.ledgr-grid-diagram .mermaid svg {
  max-width: 760px !important;
}

.ledgr-alias-diagram .mermaid svg {
  max-width: 820px !important;
}
</style>
```

This article explains how declared parameter variation becomes candidate rows.
The central idea is active aliases: one strategy can read stable feature names
such as `fast` and `slow` while the sweep varies the concrete indicators behind
those names.

Sweeps are for exploration. Promotion records one selected candidate as a
committed run. Neither step proves that the selected candidate will generalize.

## Prerequisites

```{r}
#| label: prerequisites

library(ledgr)
library(dplyr)
data("ledgr_demo_bars", package = "ledgr")
```

This article uses `dplyr` for tabular inspection. The sweep itself is ledgr's
job.

## Sweep Is Exploration

::: {.callout-note}
## Definition

A sweep is an evaluated candidate table over a declared grid. It is
exploratory: it returns candidate summaries, does not choose a winner, and does
not write candidate runs to the experiment store.
:::

`ledgr_sweep()` evaluates a grid against a `ledgr_experiment()`. It tells you
what each declared candidate did. It does not decide which candidate matters.

::: {.callout-note}
## Definition

A sweep usually contains many candidates. Each candidate is one row of the
sweep: resolved feature parameters, strategy parameters, execution seed,
status, metrics, warnings or errors, and provenance.
:::

That separation is the workflow boundary:

```text
ledgr_sweep()                 explore declared candidates
ledgr_candidate()             select one row deliberately
ledgr_promote() / ledgr_run() commit an auditable run
```

::: {.callout-warning}
## Selection is not validation

A sweep table records what was run. It does not prove that the selected
parameters were evaluated on held-out data. Promotion records a choice; it does
not make that choice out-of-sample.
:::

## Declare Parameterized Features

Start with one sealed snapshot and one strategy. This article stays on sweep
mechanics rather than train/test or walk-forward evaluation.

```{r}
#| label: experiment-setup

bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% c("DEMO_01", "DEMO_02"),
    between(ts_utc, ledgr_utc("2019-01-01"), ledgr_utc("2019-06-30"))
  )

snapshot <- ledgr_snapshot_from_df(
  bars,
  snapshot_id = "sweep_alias_demo",
  db_path = tempfile(fileext = ".duckdb")
)

features <- ledgr_feature_map(
  fast = ledgr_ind_sma(ledgr_param("fast_n")),
  slow = ledgr_ind_sma(ledgr_param("slow_n"))
)

strategy <- ledgr_demo_sma_crossover_strategy()

exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = strategy,
  features = features,
  opening = ledgr_opening(cash = 100000)
)

exp
```

The strategy function itself does not change across candidates. At each pulse,
ledgr calls the same `function(ctx, params)`: `ctx` contains the current
pulse-known market data, positions, cash, equity, and resolved alias values;
`params` contains the strategy-parameter values for this candidate.

The demo strategy you assigned above follows this shape internally:

```r
sma_crossover_body <- function(ctx, params) {
  targets <- ctx$flat()

  for (id in ctx$universe) {
    values <- ctx$features(id)
    if (
      passed_warmup(values) &&
        ((values[["fast"]] / values[["slow"]]) - 1) > params$threshold
    ) {
      targets[id] <- params$qty
    }
  }

  targets
}
```

During a sweep, `ctx$features(id)` returns values for the concrete indicators
resolved for that candidate. `params$threshold` and `params$qty` come from the
strategy grid. For the full strategy contract, read
`vignette("strategy-development", package = "ledgr")`.

::: {.callout-note}
## Definition

An active alias is a stable strategy-facing feature name whose concrete
indicator can vary by candidate. The strategy reads aliases such as `fast` and
`slow`; ledgr resolves the concrete SMA windows for each candidate before
execution.
:::

The strategy can keep reading `values[["fast"]]` and `values[["slow"]]` even
when one candidate uses SMA(5) and SMA(20) and another candidate uses SMA(10)
and SMA(40). The alias is the strategy-facing contract. The concrete feature
IDs and fingerprints are provenance.

Feature parameters vary the knobs exposed by a feature constructor. For
`ledgr_ind_sma()`, the knob is `n`, the moving-average window. For TTR-backed
features, the knobs are the supported arguments of the wrapped TTR indicator.
You still need to understand the feature you are tuning; ledgr separates the
parameter namespaces, but it does not decide which indicator arguments are
economically meaningful.

Only knobs declared with `ledgr_param("name")` need values in the feature grid.
Concrete arguments stay fixed. For example, `ledgr_ind_sma(20)` needs no
`feature_grid` entry, while `ledgr_ind_sma(ledgr_param("fast_n"))` requires a
scalar `fast_n` value for each candidate.

::: {.ledgr-diagram .ledgr-alias-diagram}

```{mermaid}
%%| echo: false

flowchart TB
  params["feature params<br/>fast_n = 5<br/>slow_n = 20"]
  indicators["concrete indicators<br/>SMA 5 and SMA 20"]
  aliases["stable aliases<br/>fast and slow"]
  strategy["strategy reads<br/>fast and slow"]

  params --> indicators --> aliases --> strategy
```

:::

For a second candidate with `fast_n = 10` and `slow_n = 40`, the same arrows
resolve to SMA(10) and SMA(40), but the strategy still reads `values[["fast"]]`
and `values[["slow"]]`. The aliases stay stable across candidates.

## Build The Candidate Grid

::: {.callout-note}
## Definition

Feature parameters materialize indicators before execution. Strategy parameters
are passed to `strategy(ctx, params)` during execution. Keeping those namespaces
separate is what lets a strategy read stable aliases while the sweep varies
indicator windows.
:::

Use `ledgr_feature_grid()` for the feature knobs you decided to vary and
`ledgr_strategy_grid()` for the knobs in your own strategy code. Then cross
them with `ledgr_grid_cross()`.

```{r}
#| label: grids

feature_grid <- ledgr_feature_grid(
  fast_n = c(5L, 10L),
  slow_n = c(20L, 40L),
  .filter = fast_n < slow_n
)

strategy_grid <- ledgr_strategy_grid(
  threshold = c(0, 0.01),
  qty = c(5, 10)
)

grid <- ledgr_grid_cross(features = feature_grid, strategy = strategy_grid)
grid
```

The `.filter` expression is a structural grid constraint. Here it says the
fast moving average must be shorter than the slow moving average. Filter
expressions are evaluated against grid columns; they do not read run state,
feature data, or caller globals.

::: {.callout-warning}
## Mind the combinatorial explosion

`ledgr_grid_cross()` multiplies grid dimensions. Four parameters with five
values each produce 625 candidates before you add another axis. Keep early
sweeps deliberately small, then expand only after the candidate table is
readable and the feature payload cost is understood.
:::

The cross product is explicit:

::: {.ledgr-diagram .ledgr-grid-diagram}

```{mermaid}
%%| echo: false

flowchart TB
  fg["feature_grid<br/>fast_n, slow_n"]
  sg["strategy_grid<br/>threshold, qty"]
  cross["grid_cross"]
  rows["candidate rows<br/>feature params + strategy params"]

  fg --> cross
  sg --> cross
  cross --> rows
```

:::

::: {.callout-tip}
## Try it

Change `slow_n` to `c(20L, 40L, 60L)` and rerun the sweep. The strategy code
does not change. How many candidates appear after `.filter`, and which
candidate rows still use the same `fast` and `slow` aliases?
:::

## Precompute Shared Features

Precomputing is an execution optimization, not a separate research decision.
The same declared feature grid is resolved once, deduplicated by fingerprint,
and reused across sweep candidates.

```{r}
#| label: precompute

precomputed <- ledgr_precompute_features(exp, grid)
precomputed
```

`ledgr_sweep()` can compute features internally for small grids. For larger
grids, precompute first so feature resolution and payload validation are
explicit. When a grid has more than 20 combinations and no precomputed payload,
ledgr warns because feature computation may be repeated per candidate.

## Run The Sweep

Give the sweep a master seed when reproducible stochastic strategy behavior
matters. Each row receives its own derived `execution_seed`.

```{r}
#| label: run-sweep

sweep <- ledgr_sweep(
  exp,
  grid,
  precomputed_features = precomputed,
  seed = 2026L
)

sweep
```

The table contains candidate summaries. It is not a full artifact store. Full
equity, fills, trades, and ledger rows are created by committed runs.

## Inspect Before Promotion

Name the ranking rule before selecting. Here the rule is deliberately simple:
among completed candidates, sort by Sharpe ratio descending. ledgr does not own
objective functions or automatic candidate ranking; ordinary R code should make
the selection rule visible.

```{r}
#| label: inspect-candidates

ranked <- sweep |>
  filter(status == "DONE") |>
  arrange(desc(sharpe_ratio))

top_n <- ranked |>
  slice_head(n = 5) |>
  select(
    run_id, status, final_equity, total_return, sharpe_ratio,
    params, feature_params, execution_seed
  )

glimpse(top_n)

issues <- sweep |>
  filter(status != "DONE") |>
  select(any_of(c("run_id", "status", "error_class", "error_msg", "warnings"))) |>
  as_tibble()

issues
```

Some columns are list columns. `glimpse()` keeps the table readable while still
showing that params, feature params, warnings, and provenance remain attached to
the rows.

::: {.callout-note}
## Design note

This explicit table code keeps the selection rule visible. The v0.1.8.6 cycle
plans a sweep-review helper that ranks completed candidates, returns a compact
review table, separates issue rows, and preserves the same explicit selection
rule.
:::

## Promote One Candidate

Promotion replays one selected candidate as a committed run. For the full
research loop around promotion notes, reopen, and human review, read
`vignette("research-workflow", package = "ledgr")`.

```{r}
#| label: promote-selected

candidate <- ledgr_candidate(ranked, 1)

promoted_run <- ledgr_promote(
  exp,
  candidate,
  run_id = "sweep_selected_candidate",
  note = "Selected highest-Sharpe completed candidate from the exploratory sweep."
)

summary(promoted_run)
```

## What A Sweep Does Not Prove

A sweep is exploratory evidence with an audit trail. It does not answer whether
the selected rule will generalize.

The more candidates you try, the more opportunity you create for sample-specific
luck to look like skill. If the question is generalization rather than artifact
reproducibility, use walk-forward evaluation when that layer lands in
v0.1.9.x.

## Failure Rows And Contract Errors

By default, candidate-level failures become rows with `status = "FAILED"`.
Inspect those rows before selecting anything.

This debug example uses no features so the failure is only about strategy
parameters. In ordinary research, the experiment usually declares features just
like the main sweep above.

```{r}
#| label: failed-rows

debug_strategy <- function(ctx, params) {
  if (params$qty < 0) {
    stop("qty must be non-negative")
  }
  ctx$flat()
}

debug_exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = debug_strategy,
  features = list(),
  opening = ledgr_opening(cash = 100000)
)

debug_grid <- ledgr_strategy_grid(qty = c(5, -1))

failed_sweep <- ledgr_sweep(debug_exp, debug_grid)

failed_sweep |>
  select(run_id, status, error_class, error_msg, params)
```

Use the failed row as an interactive debugging handle: inspect `error_class`,
`error_msg`, and `params`, then reproduce the single candidate in a smaller
session before rerunning the sweep. For pulse-level strategy debugging patterns,
read `vignette("strategy-development", package = "ledgr")`.

Contract errors still abort before a candidate table exists. Invalid experiment
or grid shape is not a failed strategy idea; it is a setup problem that ledgr
stops immediately so the sweep does not mix incomparable rows.

`ledgr_candidate()` rejects failed rows by default. Use
`ledgr_candidate(results, which, allow_failed = TRUE)` only when you are
extracting a failed row for diagnostics; `ledgr_promote()` still rejects failed
candidates.

## Explicit Non-Goals

Current sweep mode intentionally stays small. It does not currently ship:

- automatic ranking, objective functions, or `ledgr_tune()`;
- parallel sweep execution;
- walk-forward, PBO, or CSCV helpers;
- risk-layer insertion;
- public cost-model factories;
- paper/live trading adapters;
- intraday-specific support;
- `ledgr_save_sweep()` or full sweep artifact persistence.

## Where Next

- For the full research loop around sweeps, read
  `vignette("research-workflow", package = "ledgr")`.
- For feature maps, indicator identity, and active aliases, read
  `vignette("indicators", package = "ledgr")`.
- For no-lookahead pulse timing, next-open fills, and final-bar warnings, read
  `vignette("execution-semantics", package = "ledgr")`.
- For durable run comparison after promotion, read
  `vignette("experiment-store", package = "ledgr")`.

```{r}
#| label: cleanup
#| include: false

close(promoted_run)
ledgr_snapshot_close(snapshot)
```
