---
title: "TTR Indicators In ledgr"
output:
  rmarkdown::html_vignette:
    toc: false
  github_document:
    toc: false
    html_preview: false
vignette: >
  %\VignetteIndexEntry{TTR Indicators In ledgr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(width = 90)
demo_data_path <- if (file.exists("data/ledgr_demo_bars.rda")) {
  "data/ledgr_demo_bars.rda"
} else {
  file.path("..", "data", "ledgr_demo_bars.rda")
}
if (!exists("ledgr_demo_bars") && file.exists(demo_data_path)) {
  load(demo_data_path)
}
```

`ledgr_ind_ttr()` is the low-code path for using TTR indicators in ledgr.

TTR stays outside the core engine. The adapter translates a TTR call into a
normal `ledgr_indicator`:

```text
TTR -> ledgr_ind_ttr() -> ledgr_indicator -> deterministic pulse engine
```

The engine does not need to know which external package produced the feature.
It only sees the ledgr indicator contract: `fn`, `series_fn`, `requires_bars`,
`stable_after`, and deterministic `params`.

## Inspect Supported Warmup Rules

This vignette uses `dplyr` for compact display. TTR examples require the
suggested `TTR` package; chunks that call TTR are skipped when it is not
installed.

```{r rules, message=FALSE}
library(ledgr)
library(dplyr)
data("ledgr_demo_bars", package = "ledgr")

as.data.frame(ledgr_ttr_warmup_rules()[, c("ttr_fn", "input", "formula")])
```

`ledgr_ttr_warmup_rules()` is the inspectable contract for automatic warmup
inference. A TTR function is listed only when ledgr can infer the first stable
row from explicit arguments alone. Unsupported or ambiguous functions can still
be used by supplying `requires_bars` manually.

## Simple Close-Based Indicators

```{r simple, eval=requireNamespace("TTR", quietly = TRUE)}
rsi_14 <- ledgr_ind_ttr("RSI", input = "close", n = 14)
wma_10 <- ledgr_ind_ttr("WMA", input = "close", n = 10)
mom_10 <- ledgr_ind_ttr("momentum", input = "close", n = 10)

ledgr_feature_id(rsi_14)
ledgr_feature_id(wma_10)
ledgr_feature_id(mom_10)
```

The generated ID is derived from the TTR function and explicit arguments. ledgr
does not rely on TTR defaults for warmup inference or ID construction.

Built-in indicators follow the same convention:

```{r builtin-ids}
builtins <- list(
  ledgr_ind_sma(20),
  ledgr_ind_ema(20),
  ledgr_ind_rsi(14),
  ledgr_ind_returns(5)
)

ledgr_feature_id(builtins)
```

Those strings are the names you use inside a strategy:

```r
ctx$feature("AAA", "sma_20")
ctx$feature("AAA", "return_5")
```

## Multi-Input And Multi-Output Indicators

```{r multi, eval=requireNamespace("TTR", quietly = TRUE)}
atr_20 <- ledgr_ind_ttr("ATR", input = "hlc", output = "atr", n = 20)
bb_up <- ledgr_ind_ttr("BBands", input = "close", output = "up", n = 20)
macd_line <- ledgr_ind_ttr(
  "MACD",
  input = "close",
  output = "macd",
  nFast = 12,
  nSlow = 26,
  nSig = 9,
  percent = FALSE
)
macd_signal <- ledgr_ind_ttr(
  "MACD",
  input = "close",
  output = "signal",
  nFast = 12,
  nSlow = 26,
  nSig = 9
)
aroon_osc <- ledgr_ind_ttr("aroon", input = "hl", output = "oscillator", n = 20)

ledgr_feature_id(list(rsi_14, atr_20, bb_up, macd_line, macd_signal, aroon_osc))
```

Some TTR indicators return several columns. For those indicators, choose the
column with `output`. The available outputs are checked at construction time so
errors happen before a backtest starts.

```{r id-reference-table, eval=requireNamespace("TTR", quietly = TRUE)}
data.frame(
  constructor = c(
    "ledgr_ind_sma(20)",
    "ledgr_ind_returns(5)",
    'ledgr_ind_ttr("RSI", input = "close", n = 14)',
    'ledgr_ind_ttr("BBands", input = "close", output = "up", n = 20)',
    'ledgr_ind_ttr("MACD", output = "macd", nFast = 12, nSlow = 26, nSig = 9, percent = FALSE)',
    'ledgr_ind_ttr("MACD", output = "signal", nFast = 12, nSlow = 26, nSig = 9)'
  ),
  feature_id = c(
    ledgr_feature_id(ledgr_ind_sma(20)),
    ledgr_feature_id(ledgr_ind_returns(5)),
    ledgr_feature_id(rsi_14),
    ledgr_feature_id(bb_up),
    ledgr_feature_id(macd_line),
    ledgr_feature_id(macd_signal)
  ),
  stringsAsFactors = FALSE
)
```

Use this pattern as a reference table in your own project: construct the
indicator objects first, call `ledgr_feature_id()`, and copy only those returned
strings into strategy code.

The ID format is deterministic:

```text
ttr_<function>_<explicit args>_<output>
```

For example, the IDs above include `ttr_rsi_14`, `ttr_atr_20_atr`,
`ttr_bbands_20_up`, `ttr_macd_12_26_9_false_macd`, and
`ttr_macd_12_26_9_signal`. Use
`ledgr_feature_id(ind)` or print the indicator object instead of guessing the
string.

The two MACD examples intentionally use different explicit arguments:
`macd_line` sets `percent = FALSE`, while `macd_signal` leaves TTR's default
`percent` argument implicit. Explicit extra arguments become part of the
feature ID, so combine MACD outputs in one strategy only when their argument
sets match the computation you intend.

## Warmup And Short Histories

Every ledgr indicator declares `requires_bars` and `stable_after`.
`requires_bars` is the first row where the indicator may produce a non-`NA`
value. `stable_after` is the first row ledgr treats as stable; built-in and
supported TTR indicators use the same value unless an adapter has a reason to
be more conservative.

```{r warmup-values, eval=requireNamespace("TTR", quietly = TRUE)}
c(
  rsi = rsi_14$requires_bars,
  momentum = mom_10$requires_bars,
  bbands = bb_up$requires_bars,
  macd = macd_line$requires_bars,
  macd_signal = macd_signal$requires_bars
)
```

Warmup `NA` is expected and should be handled in the strategy with `is.na()`.
Unknown feature IDs are different: `ctx$feature()` fails loudly when the ID was
not configured for the run.

For MACD, ledgr verifies the supported warmup rules against direct TTR output.
With current TTR behavior, the `macd` output is first valid at `nSlow`, while
`signal` and the derived ledgr `histogram` output are first valid at
`nSlow + nSig - 1`. The same rule is verified for both `percent = TRUE` and
`percent = FALSE`.

## Feature Computation In A Backtest

```{r backtest, eval=requireNamespace("TTR", quietly = TRUE)}
bars <- ledgr_demo_bars |>
  filter(
    instrument_id == "DEMO_01",
    between(ts_utc, ledgr_utc("2019-01-01"), ledgr_utc("2019-06-30"))
  )

features <- list(
  ledgr_ind_ttr("RSI", input = "close", n = 14),
  ledgr_ind_ttr("momentum", input = "close", n = 10),
  ledgr_ind_ttr("BBands", input = "close", output = "up", n = 20),
  ledgr_ind_ttr(
    "MACD",
    input = "close",
    output = "macd",
    nFast = 12,
    nSlow = 26,
    nSig = 9,
    percent = FALSE
  )
)
ledgr_feature_id(features)

strategy <- function(ctx, params) {
  targets <- ctx$hold()
  rsi <- ctx$feature("DEMO_01", "ttr_rsi_14")
  mom <- ctx$feature("DEMO_01", "ttr_momentum_10")
  bb_up <- ctx$feature("DEMO_01", "ttr_bbands_20_up")
  macd <- ctx$feature("DEMO_01", "ttr_macd_12_26_9_false_macd")

  # This article uses one demo instrument, so only DEMO_01 is targeted.
  if (
    !is.na(rsi) &&
      !is.na(mom) &&
      !is.na(bb_up) &&
      !is.na(macd) &&
      rsi > 50 &&
      mom > 0 &&
      macd > 0 &&
      ctx$close("DEMO_01") > bb_up
  ) {
    targets["DEMO_01"] <- params$qty
  }
  targets
}

snapshot <- ledgr_snapshot_from_df(bars)
exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = strategy,
  features = features,
  opening = ledgr_opening(cash = 10000)
)

bt <- exp |>
  ledgr_run(params = list(qty = 10), run_id = paste0("ttr-article-demo-", Sys.getpid()))

nrow(tibble::as_tibble(bt, what = "trades"))
close(bt)
ledgr_snapshot_close(snapshot)
```

TTR-backed indicators use `series_fn`, so ledgr computes the full feature
series once per instrument before the pulse loop and then performs lookups
during execution. The indicator fingerprint includes TTR metadata, including
the installed TTR version.

## Unsupported TTR Functions

```{r unsupported, eval=requireNamespace("TTR", quietly = TRUE)}
ledgr_ind_ttr(
  "DEMA",
  input = "close",
  n = 10,
  requires_bars = 20
)$id
```

When a TTR function is not in the warmup rules table, provide `requires_bars`
explicitly. To measure it, run the TTR function on a sufficiently long series
and count the leading `NA` rows in the selected output, then add one.

For non-TTR sources or more specialized logic, use `ledgr_indicator()` directly
with a `series_fn`. That is the adapter escape hatch: external logic remains at
the boundary, while the engine keeps the same deterministic indicator contract.
