---
title: "TTR Indicators In ledgr"
output:
  github_document:
    toc: false
    html_preview: false
vignette: >
  %\VignetteIndexEntry{TTR Indicators In ledgr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(width = 90)
```

`ledgr_ind_ttr()` is the low-code path for using TTR indicators in ledgr.

TTR stays outside the core engine. The adapter translates a TTR call into a
normal `ledgr_indicator`:

```text
TTR -> ledgr_ind_ttr() -> ledgr_indicator -> deterministic pulse engine
```

The engine does not need to know which external package produced the feature.
It only sees the ledgr indicator contract: `fn`, `series_fn`, `requires_bars`,
`stable_after`, and deterministic `params`.

## Inspect Supported Warmup Rules

```{r rules}
library(ledgr)

as.data.frame(ledgr_ttr_warmup_rules()[, c("ttr_fn", "input", "formula")])
```

`ledgr_ttr_warmup_rules()` is the inspectable contract for automatic warmup
inference. A TTR function is listed only when ledgr can infer the first stable
row from explicit arguments alone. Unsupported or ambiguous functions can still
be used by supplying `requires_bars` manually.

## Simple Close-Based Indicators

```{r simple, eval=requireNamespace("TTR", quietly = TRUE)}
rsi_14 <- ledgr_ind_ttr("RSI", input = "close", n = 14)
wma_10 <- ledgr_ind_ttr("WMA", input = "close", n = 10)

rsi_14$id
wma_10$id
```

The generated ID is derived from the TTR function and explicit arguments. ledgr
does not rely on TTR defaults for warmup inference or ID construction.

## Multi-Input And Multi-Output Indicators

```{r multi, eval=requireNamespace("TTR", quietly = TRUE)}
atr_20 <- ledgr_ind_ttr("ATR", input = "hlc", output = "atr", n = 20)
bb_up <- ledgr_ind_ttr("BBands", input = "close", output = "up", n = 20)
aroon_osc <- ledgr_ind_ttr("aroon", input = "hl", output = "oscillator", n = 20)

atr_20$id
bb_up$id
aroon_osc$id
```

Some TTR indicators return several columns. For those indicators, choose the
column with `output`. The available outputs are checked at construction time so
errors happen before a backtest starts.

## Feature Computation In A Backtest

```{r backtest, eval=requireNamespace("TTR", quietly = TRUE)}
bars <- data.frame(
  ts_utc = as.POSIXct("2020-01-01", tz = "UTC") + 86400 * 0:39,
  instrument_id = "AAA",
  open = 100 + seq_len(40),
  high = 101 + seq_len(40),
  low = 99 + seq_len(40),
  close = 100 + seq_len(40),
  volume = 1000 + seq_len(40)
)

features <- list(
  ledgr_ind_ttr("RSI", input = "close", n = 14),
  ledgr_ind_ttr("BBands", input = "close", output = "up", n = 20)
)

strategy <- function(ctx) {
  targets <- ctx$current_targets()
  rsi <- ctx$feature("AAA", "ttr_rsi_14")

  # This article uses one synthetic instrument, so only AAA is targeted.
  if (!is.na(rsi) && rsi > 50) {
    targets["AAA"] <- 10
  }
  targets
}

bt <- ledgr_backtest(
  data = bars,
  strategy = strategy,
  features = features,
  initial_cash = 10000,
  run_id = paste0("ttr-article-demo-", Sys.getpid())
)

nrow(tibble::as_tibble(bt, what = "trades"))
close(bt)
```

TTR-backed indicators use `series_fn`, so ledgr computes the full feature
series once per instrument before the pulse loop and then performs lookups
during execution. The indicator fingerprint includes TTR metadata, including
the installed TTR version.

## Unsupported TTR Functions

```{r unsupported, eval=requireNamespace("TTR", quietly = TRUE)}
ledgr_ind_ttr(
  "DEMA",
  input = "close",
  n = 10,
  requires_bars = 20
)$id
```

When a TTR function is not in the warmup rules table, provide `requires_bars`
explicitly. To measure it, run the TTR function on a sufficiently long series
and count the leading `NA` rows in the selected output, then add one.

For non-TTR sources or more specialized logic, use `ledgr_indicator()` directly
with a `series_fn`. That is the adapter escape hatch: external logic remains at
the boundary, while the engine keeps the same deterministic indicator contract.
