---
title: "TTR Indicators And Bundles"
format:
  html:
    toc: false
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{TTR Indicators And Bundles}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(width = 90)
options(cli.unicode = FALSE)
# Local renders load ledgr from source; installed builds use the library call
# in the prerequisites chunk.
pkg_root <- if (file.exists("DESCRIPTION")) "." else if (file.exists("../DESCRIPTION")) ".." else NA_character_
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
demo_data_path <- if (file.exists("data/ledgr_demo_bars.rda")) {
  "data/ledgr_demo_bars.rda"
} else {
  file.path("..", "data", "ledgr_demo_bars.rda")
}
if (!exists("ledgr_demo_bars") && file.exists(demo_data_path)) {
  load(demo_data_path)
}
```

```{=html}
<style>
.ledgr-diagram {
  margin: 1.25rem auto 1.5rem auto;
  text-align: center;
}
.ledgr-diagram .mermaid {
  display: inline-block;
  max-width: 760px;
  width: 100%;
}
.ledgr-diagram .node text,
.ledgr-diagram .edgeLabel {
  font-size: 18px !important;
}
</style>
```

You want a TTR indicator, such as RSI, Bollinger bands, or MACD, as a
feature. By the end of this article you can declare TTR indicators and
multi-output bundles, predict their feature IDs and warmup, and know which TTR
shapes availability-aware runs accept. You need it only if you use TTR; the
built-in indicators need no extra package. R functions and CSV-backed external
features live in `vignette("custom-indicators", package = "ledgr")`; the
feature lifecycle lives in `vignette("indicators", package = "ledgr")`.

## Prerequisites

```{r}
#| label: packages
#| message: false

library(ledgr)
library(dplyr)
data("ledgr_demo_bars", package = "ledgr")
```


```{r}
#| label: demo-bars

bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% c("DEMO_01", "DEMO_02"),
    between(ts_utc, ledgr_utc("2019-01-01"), ledgr_utc("2019-06-30"))
  )
```

## TTR-Backed Indicators

`ledgr_ind_ttr()` is the adapter for supported indicators from the suggested
`TTR` package. TTR stays outside the core engine:

::: {.ledgr-diagram .ledgr-ttr-adapter}

```{mermaid}
%%| echo: false

flowchart LR
  ttr["TTR"]
  adapter["ledgr_ind_ttr()"]
  indicator["ledgr_indicator"]
  engine["deterministic<br/>pulse engine"]

  ttr --> adapter --> indicator --> engine
```

:::

The engine sees a normal `ledgr_indicator`. That means TTR-backed indicators
follow the same feature-ID, warmup, and pulse-view rules as built-in
indicators. The TTR-backed examples below are skipped when `TTR` is not
installed. In your own project, install TTR with `install.packages("TTR")`
before creating TTR-backed indicators.

### Single-Output TTR Indicators

```{r}
#| label: ttr-backed-indicators
#| eval: !expr requireNamespace("TTR", quietly = TRUE)

ttr_features <- ledgr_feature_map(
  ret_5 = ledgr_ind_returns(5),
  ttr_rsi = ledgr_ind_ttr("RSI", input = "close", n = 14),
  bb_up = ledgr_ind_ttr("BBands", input = "close", output = "up", n = 20),
  macd = ledgr_ind_ttr(
    "MACD",
    input = "close",
    output = "macd",
    nFast = 12,
    nSlow = 26,
    nSig = 9,
    percent = FALSE
  ),
  macd_signal = ledgr_ind_ttr(
    "MACD",
    input = "close",
    output = "signal",
    nFast = 12,
    nSlow = 26,
    nSig = 9,
    percent = FALSE
  )
)

ledgr_feature_contracts(ttr_features)
ledgr_feature_id(ttr_features)
```

This mixed feature map combines a built-in return feature with TTR-backed RSI,
BBands, and MACD features. The MACD ID embeds the explicit arguments because
they are part of the calculation identity.

### Native RSI vs TTR RSI

ledgr also includes a native RSI helper. It does not require TTR and follows the
same ID and warmup contract as other built-in indicators:

```{r}
#| label: native-rsi-contract

native_rsi_features <- ledgr_feature_map(
  rsi_14 = ledgr_ind_rsi(14)
)

ledgr_feature_contracts(native_rsi_features)
ledgr_feature_id(native_rsi_features)
```

The native RSI feature ID is `rsi_14`. The TTR-backed RSI feature ID above is
`ttr_rsi_14`. Those are different feature definitions and should not be treated
as interchangeable without checking that their calculation and warmup behavior
match your research intent.

Some TTR functions return several columns. For those functions, choose one
column with `output` when you need exactly one output, or use
`ledgr_ind_ttr_outputs()` to declare several outputs from one shared TTR
configuration. `BBands` exposes `dn`, `mavg`, `up`, and `pctB`. `MACD` exposes
`macd` and `signal`; ledgr also supports a derived `histogram`.

### Multi-Output TTR Indicators

```{r}
#| label: ttr-output-examples
#| eval: !expr requireNamespace("TTR", quietly = TRUE)

ledgr_feature_contracts(ledgr_feature_map(
  bb_dn = ledgr_ind_ttr("BBands", input = "close", output = "dn", n = 20),
  bb_mavg = ledgr_ind_ttr("BBands", input = "close", output = "mavg", n = 20),
  bb_up = ledgr_ind_ttr("BBands", input = "close", output = "up", n = 20),
  bb_pctB = ledgr_ind_ttr("BBands", input = "close", output = "pctB", n = 20)
))
```

### Bundle Naming Rules

For multi-output authoring, prefer the bundle helper. By default, bundle feature
IDs use a normalized prefix derived from the TTR function name. For `BBands`,
that produces `bbands_dn`, `bbands_mavg`, `bbands_up`, and `bbands_pctb`.
The helper returns a `ledgr_indicator_bundle`, but the experiment sees ordinary
single-output indicators after feature declaration is materialized.

Those bundle IDs are shorter than the hand-written single-output TTR IDs such
as `ttr_bbands_20_up`. That asymmetry is intentional: bundle defaults optimize
for readable output names. Use `naming = c(up = "ttr_bbands_20_up")` or
hand-written `ledgr_ind_ttr(output = ...)` calls when you need the exact IDs of the single-output form.

```{r}
#| label: ttr-output-bundle
#| eval: !expr requireNamespace("TTR", quietly = TRUE)

bbands_bundle <- ledgr_ind_ttr_outputs("BBands", input = "close", n = 20)
ledgr_feature_id(bbands_bundle)
ledgr_feature_contracts(bbands_bundle)
```

When an unnamed bundle is placed inside `ledgr_feature_map()`, its entries
expand using their feature IDs as aliases. A named outer entry such as
`bands = bbands_bundle` is refused: one alias cannot name several outputs.
Control names with the bundle's `prefix` or `naming` argument instead.

Use `outputs` as a filter. The derived or explicit prefix still applies to
selected outputs, so a subset remains collision-resistant:

```{r}
#| label: ttr-output-bundle-filter
#| eval: !expr requireNamespace("TTR", quietly = TRUE)

bbands_subset <- ledgr_ind_ttr_outputs(
  "BBands",
  input = "close",
  outputs = c("dn", "up"),
  prefix = "bb",
  n = 20
)
ledgr_feature_id(bbands_subset)
```

`naming` renames selected outputs; it is not itself an output filter. When
renaming only part of a bundle, make the filter explicit:

```{r}
#| label: ttr-output-bundle-naming
#| eval: !expr requireNamespace("TTR", quietly = TRUE)

bbands_named_subset <- ledgr_ind_ttr_outputs(
  "BBands",
  input = "close",
  outputs = c("dn", "up"),
  naming = c(dn = "lower_band", up = "upper_band"),
  n = 20
)
ledgr_feature_id(bbands_named_subset)
```

Set `prefix = NULL` only when you explicitly want raw normalized output names
such as `dn`, `up`, or `pctb`. Raw names are short and can collide when one
experiment combines several bundles or parameterizations.

### MACD Argument Consistency

The two MACD entries in `ttr_features` both set `percent = FALSE`. Explicit
arguments become part of the feature ID, so combine MACD outputs in one strategy
only when their argument sets match the computation you intend. If one MACD
output uses `percent = FALSE`, the paired `signal` output should usually set
`percent = FALSE` too.

### TTR Warmup Rules

TTR warmup inference is inspectable:

```{r}
#| label: ttr-warmup-rules
#| message: false
#| eval: !expr requireNamespace("TTR", quietly = TRUE)

ledgr_ind_ttr_warmup_rules() |>
  select(ttr_fn, input, formula)
```

For MACD, ledgr verifies the supported warmup rules against direct TTR output.
TTR computes the signal EMA internally even when you select only the `macd`
column. In a pulse-by-pulse backtest, all supported MACD outputs are therefore
first callable at `nSlow + nSig - 1`. The same rule is verified for `macd`,
`signal`, the derived ledgr `histogram`, and both `percent = TRUE` and
`percent = FALSE`.

When a function is absent from the table, the constructor's error explains the
measurement route: run the TTR function over at least 50 representative bars,
inspect the first row where the selected output is finite, then provide that
deliberately as `requires_bars`. Do not guess from the parameter name alone.
An explicit warmup makes dense execution possible; it does not certify a
recursive or otherwise unknown TTR shape across availability gaps.

To debug a TTR-backed feature at one decision time, use an active snapshot
handle, choose a timestamp late enough for the indicator warmup, and pass the
same TTR feature map to `ledgr_pulse_snapshot()`. Pulse inspection uses the
dense run's feature computation over snapshot history through that timestamp,
including the TTR adapter's parameters, recursive series calculation and
warmup boundary. A completed backtest proves the run succeeded, but it does not
replace the snapshot handle needed for interactive pulse inspection.

```{r}
#| label: ttr-pulse-snapshot
#| eval: !expr requireNamespace("TTR", quietly = TRUE)

ttr_snapshot <- ledgr_snapshot_from_df(
  bars,
  snapshot_id = "ttr-pulse-inspection"
)

ttr_pulse <- ledgr_pulse_snapshot(
  ttr_snapshot,
  universe = c("DEMO_01", "DEMO_02"),
  ts_utc = ledgr_utc("2019-06-03"),
  features = ttr_features
)

ledgr_pulse_features(ttr_pulse, ttr_features)
close(ttr_pulse)
ledgr_snapshot_close(ttr_snapshot)
```

## Unsupported Or Custom Indicators

The explicit measurement route above can admit an unknown TTR function for a
dense study. This example records the measured boundary rather than pretending
that `n = 10` establishes it:

```{r}
#| label: unsupported-ttr
#| eval: !expr requireNamespace("TTR", quietly = TRUE)

ledgr_ind_ttr(
  "DEMA",
  input = "close",
  n = 10,
  requires_bars = 20
)$id
```

For non-TTR sources or more specialized logic, use `ledgr_indicator()` directly
with a `series_fn`, as `vignette("custom-indicators", package = "ledgr")`
shows. That is the adapter escape hatch: external logic remains at
the boundary, while the engine keeps the same deterministic indicator contract.

Availability-aware runs accept only the exact single-output
`SMA(close, n)` TTR shape; every other TTR shape is refused with
`ledgr_indicator_gap_unsupported`, whatever `requires_bars` says. The support
matrix in `vignette("indicators", package = "ledgr")` lists every accepted
form. The general warmup and zero-trade checklist lives in
`vignette("indicators", package = "ledgr")`.


## Where Next

- `vignette("custom-indicators", package = "ledgr")` covers writing your own
  indicators and R or CSV adapters.
- `vignette("indicators", package = "ledgr")` covers the feature lifecycle
  and strategy-time access patterns.
- `vignette("strategy-authoring-tools", package = "ledgr")` shows how
  registered features feed a strategy's helper pipeline.
