---
title: "Walk-Forward Evaluation"
format:
  html:
    toc: false
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{Walk-Forward Evaluation}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 90)
options(cli.unicode = FALSE)
pkg_root <- if (file.exists("DESCRIPTION")) "." else if (file.exists("../DESCRIPTION")) ".." else NA_character_
if (!is.na(pkg_root) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(pkg_root, quiet = TRUE)
}
```

You ran a sweep and picked the best candidate. Was that a real edge, or the
luckiest parameter on this slice of history? A sweep selects and scores on the
same data, so its winner is in-sample evidence. It cannot tell you whether the
rule generalizes.

Walk-forward evaluation answers the generalization question. It splits time into
folds, and in each one it selects a candidate on a train window, then tests only
that candidate on the next, untouched window. The test window is held-out
evidence: the selection never saw it.

```text
for each fold:
  train snapshot window -> sweep candidates -> scalar selection
  selected candidate -> test window run -> recorded evidence
```

Walk-forward uses the same `ledgr_sweep()` and `ledgr_run()` machinery as the
rest of ledgr; it does not add a separate execution engine. By the end of this
article you will run a multi-fold evaluation, read its train-versus-test
degradation table, and extract a candidate for review.

## Setup

```{r}
#| label: prerequisites
#| message: false

library(ledgr)
library(dplyr)

data("ledgr_demo_bars", package = "ledgr")
```

::: {.ledgr-callout .ledgr-callout-note}
**Running this yourself**

The code writes to a temporary DuckDB store, so the article leaves no project
files behind. The demo-data window and fold sizes are kept small so the vignette
runs quickly; they are not a statistical recommendation.
:::

```{r}
#| label: data

bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% c("DEMO_01", "DEMO_02"),
    between(ts_utc, ledgr_utc("2019-01-01"), ledgr_utc("2019-12-31"))
  )

snapshot <- ledgr_snapshot_from_df(
  bars,
  db_path = tempfile(fileext = ".duckdb"),
  snapshot_id = "walk_forward_demo"
)
```

The experiment is the same kind of object used by `ledgr_run()` and
`ledgr_sweep()`.

```{r}
#| label: experiment

features <- ledgr_feature_map(
  fast = ledgr_ind_sma(ledgr_param("fast_n")),
  slow = ledgr_ind_sma(ledgr_param("slow_n"))
)

exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = ledgr_demo_sma_crossover_strategy(),
  features = features,
  opening = ledgr_opening(cash = 10000),
  cost_model = ledgr_cost_zero()
)
```

## Folds And The Train/Test Boundary

Folds are calendar-time value objects. They describe train and test scoring
windows; they do not execute anything by themselves.

```{r}
#| label: folds

folds <- ledgr_folds_rolling(
  start = "2019-01-01",
  end = "2019-07-31",
  train_window = "3 months",
  test_window = "1 month",
  step = "2 months"
)

folds
```

Plotting the folds makes the core contract easier to see. The blue train
window sweeps and scores the grid. At the boundary, ledgr locks one selected
candidate. The green test window then runs only that candidate.

```{r}
#| label: fold-timeline
#| echo: false
#| fig-width: 9.5
#| fig-height: 4.8
#| fig-dpi: 180

fold_windows <- do.call(
  rbind,
  lapply(folds, function(fold) {
    data.frame(
      fold = paste("Fold", fold$fold_seq),
      y = as.integer(fold$fold_seq),
      train_start = as.Date(fold$train_start_utc),
      boundary = as.Date(fold$test_start_utc),
      test_end = as.Date(fold$test_end_utc),
      stringsAsFactors = FALSE
    )
  })
)
fold_windows$y <- max(fold_windows$y) - fold_windows$y + 1L

timeline_rects <- rbind(
  data.frame(
    fold = fold_windows$fold,
    y = fold_windows$y,
    phase = "train",
    xmin = fold_windows$train_start,
    xmax = fold_windows$boundary,
    label = "train sweep\nscore grid"
  ),
  data.frame(
    fold = fold_windows$fold,
    y = fold_windows$y,
    phase = "test",
    xmin = fold_windows$boundary,
    xmax = fold_windows$test_end,
    label = "test run\nselected only"
  )
)
timeline_rects$xmid <- timeline_rects$xmin +
  (timeline_rects$xmax - timeline_rects$xmin) / 2

selection_boundaries <- data.frame(
  y = fold_windows$y,
  x = fold_windows$boundary,
  label = c("selection\nlocked", rep("", nrow(fold_windows) - 1L))
)

ggplot2::ggplot() +
  ggplot2::geom_rect(
    data = timeline_rects,
    ggplot2::aes(
      xmin = xmin,
      xmax = xmax,
      ymin = y - 0.28,
      ymax = y + 0.28,
      fill = phase
    ),
    color = "white",
    linewidth = 0.55
  ) +
  ggplot2::geom_segment(
    data = selection_boundaries,
    ggplot2::aes(x = x, xend = x, y = y - 0.36, yend = y + 0.36),
    linewidth = 0.65,
    color = "#1F4E79"
  ) +
  ggplot2::geom_text(
    data = timeline_rects,
    ggplot2::aes(x = xmid, y = y, label = label),
    color = "white",
    fontface = "bold",
    size = 3.8,
    lineheight = 0.9
  ) +
  ggplot2::geom_text(
    data = selection_boundaries[selection_boundaries$label != "", ],
    ggplot2::aes(x = x, y = y - 0.50, label = label),
    size = 3.2,
    lineheight = 0.9,
    color = "#1F4E79"
  ) +
  ggplot2::scale_fill_manual(
    values = c(train = "#4F7FC8", test = "#4F9B63")
  ) +
  ggplot2::scale_y_continuous(
    breaks = fold_windows$y,
    labels = fold_windows$fold,
    limits = c(0.35, max(fold_windows$y) + 0.65),
    expand = c(0, 0)
  ) +
  ggplot2::scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b",
    expand = ggplot2::expansion(mult = c(0.015, 0.035))
  ) +
  ggplot2::labs(
    title = "Walk-forward evaluation",
    subtitle = paste(
      "Each fold selects on the train window, then tests only",
      "the selected candidate on the next window."
    ),
    x = NULL,
    y = NULL,
    caption = paste(
      "The test window is not another sweep. It runs the candidate",
      "selected from the immediately preceding train window."
    )
  ) +
  ggplot2::theme_minimal(base_size = 13) +
  ggplot2::theme(
    legend.position = "none",
    panel.grid.major.y = ggplot2::element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    plot.title.position = "plot",
    plot.title = ggplot2::element_text(face = "bold"),
    axis.text.y = ggplot2::element_text(face = "bold"),
    plot.caption = ggplot2::element_text(hjust = 0, color = "#555555"),
    plot.margin = ggplot2::margin(14, 18, 14, 14)
  )
```

## Build The Grid And Run It

Build a small grid. Feature parameters vary the concrete SMA windows. Strategy
parameters vary the target quantity and threshold.

```{r}
#| label: grid

feature_grid <- ledgr_feature_grid(
  fast_n = c(5L, 10L),
  slow_n = 20L,
  .filter = fast_n < slow_n
)

strategy_grid <- ledgr_strategy_grid(
  qty = 5,
  threshold = 0
)

grid <- ledgr_grid_cross(features = feature_grid, strategy = strategy_grid)
```

The selection rule is explicit and scalar: ledgr selects one candidate per fold
from the train-window score rows.

```{r}
#| label: walk-forward

wf <- ledgr_walk_forward(
  exp,
  grid = grid,
  folds = folds,
  selection_rule = ledgr_select_argmax("sharpe_ratio"),
  seed = 2026L
)

wf
```

## Read The Degradation Table First

The degradation table is the first surface to inspect. Printing it shows the
core train-versus-test columns; the full table, including the window ranges and
percentage change, is one `as_tibble()` away.

```{r}
#| label: degradation

wf$degradation
```

`metric_diff_abs` is `test_metric_value - train_metric_value`. Whether a
positive value means improvement or degradation depends on the selection rule
direction.

Short test windows may carry warning flags. Those warnings are not failures;
they are metadata telling you how to interpret the evidence.

## Extract A Candidate With A Rationale

Use `ledgr_candidate()` to extract the selected candidate for a specific fold.
Using `"latest"` is allowed only with a rationale.

```{r}
#| label: candidate

candidate <- ledgr_candidate(
  wf,
  fold_seq = "latest",
  selection_rationale = "Documentation example: inspect the latest completed fold."
)

candidate$candidate_id
```

Promotion uses the same candidate object as sweep promotion. In a real research
project, promote only after you have written down why this fold and candidate
are being reviewed.

```{r}
#| label: promotion-sketch
#| eval: false

promoted <- ledgr_promote(
  exp,
  candidate,
  run_id = "wf_latest_review",
  note = "Manual review accepted the latest completed fold."
)
```

## Reading The Evidence Honestly

A degradation table is easy to over-read. Three things bound what it can tell
you.

Walk-forward evidence is only as survivorship-safe as the sealed snapshot and
universe semantics it evaluates.

::: {.ledgr-callout .ledgr-callout-warning}
**Selection integrity is a separate question**

Reproducibility and selection integrity are orthogonal. Walk-forward records
what was selected and tested; it does not prove that the search space, metric,
or selection procedure was statistically sound. Scalar scores are not PBO,
CSCV, CPCV, DSR, benchmark diagnostics, or a multiple-testing correction. For
what provenance does and does not prove, see
`vignette("reproducibility", package = "ledgr")`.
:::

::: {.ledgr-callout .ledgr-callout-warning}
**Test folds are not independent**

The default `opening_state_policy = "carry_test_state"` is path-dependent. Each
completed selected test run may seed the next test opening state, so per-fold
test metrics are not independent observations. The explicit
`opening_state_policy = "flat_test_state"` opt-in starts every test window from
the original experiment opening state and marks the result as
`cold_start_distorted`.
:::

Anchored folds grow their train window over time. That is useful for some
research workflows, but later folds can cost more to compute than early folds,
especially with large candidate grids and wide feature sets.

## Try It

::: {.ledgr-callout .ledgr-callout-tip}
**Try it**

Re-run `ledgr_walk_forward()` with `opening_state_policy = "flat_test_state"`.
How does the degradation table change, and why does `cold_start_distorted`
appear? Then widen the `test_window` in `ledgr_folds_rolling()` to `"2 months"`
and re-run. Do the short-test-window warning flags go away?
:::

## What This Does Not Do

::: {.ledgr-callout .ledgr-callout-important}
**Outside this surface**

ledgr does not yet implement paper/live walk-forward, OMS behavior,
cross-snapshot walk-forward, benchmark-relative metrics, gross-vs-net
attribution, signal decay, implementation/cost decay, or selection-integrity
diagnostics. Those require later surfaces over the stable session and candidate
identity introduced here.
:::

## Where Next

- For the sweep and selection mechanics this builds on, see
  `vignette("sweeps", package = "ledgr")`.
- For the full research arc from idea to reviewed candidate, see
  `vignette("research-workflow", package = "ledgr")`.
- For hashes, source capture, and the limits of provenance, see
  `vignette("reproducibility", package = "ledgr")`.
- For the Sharpe ratio and annualization conventions behind these scores, see
  `vignette("metric-contexts-and-conventions", package = "ledgr")`.

```{r}
#| label: cleanup
#| include: false

invisible(lapply(wf$test_runs, close))
ledgr_snapshot_close(snapshot)
```
