---
title: "Walk-Forward Evaluation"
format:
  html:
    toc: false
    theme: cosmo
    code-copy: true
execute:
  echo: true
vignette: >
  %\VignetteIndexEntry{Walk-Forward Evaluation}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: setup
#| include: false

options(width = 90)
```

Walk-forward evaluation asks a narrower question than a sweep:

```text
for each fold:
  train snapshot window -> sweep candidates -> scalar selection
  selected candidate -> test window run -> recorded evidence
```

The v1 surface is deliberately procedural. It uses the same `ledgr_sweep()`
and `ledgr_run()` machinery as the rest of ledgr; it does not add a separate
execution engine.

## Design-only Workflow Sketch

The shape below is design-only. It uses short synthetic windows so the object
names are visible, not because those windows are statistically meaningful.
Building a runnable walk-forward example needs a sealed snapshot and multi-fold
experiment fixture that is too large for this MVP vignette; a later teaching
release should add an executable compact example.

```{r}
#| eval: false
folds <- ledgr_folds_rolling(
  start = "2020-01-01",
  end = "2022-12-31",
  train_window = "1 year",
  test_window = "3 months",
  step = "3 months"
)

wf <- ledgr_walk_forward(
  exp,
  grid = ledgr_param_grid(fast = c(10, 20), slow = c(50, 100)),
  folds = folds,
  selection_rule = ledgr_select_argmax("sharpe_ratio"),
  seed = 42L
)

wf$degradation
scores <- ledgr_walk_forward_scores(snapshot, wf$session_id)
candidate <- ledgr_walk_forward_extract_candidate(
  snapshot,
  wf$session_id,
  fold_seq = "latest",
  selection_rationale = "Manual review accepted the latest completed fold."
)
promoted <- ledgr_promote(exp, candidate, run_id = "wf-latest-review")
```

The default print starts with the train-vs-test degradation table. That table is
the first surface to inspect. It shows the selected train metric beside the
matching test metric, the absolute and percentage difference, and warning flags
such as `short_test_window` and `cold_start_distorted`.
`metric_diff_abs` is `test_metric_value - train_metric_value`, so whether a
positive value means improvement or degradation depends on the selection rule
direction.

## Caveats

Walk-forward evidence is only as survivorship-safe as the sealed snapshot and universe semantics it evaluates.

Reproducibility and selection integrity are orthogonal. Walk-forward records
what was selected and tested; it does not prove that the search space, metric,
or selection procedure was statistically sound. v1 scalar scores are not PBO,
CSCV, CPCV, DSR, benchmark diagnostics, or a multiple-testing correction.

The default `opening_state_policy = "carry_test_state"` is path-dependent.
Each completed selected test run may seed the next test opening state, so
per-fold test metrics are not independent observations. The explicit
`opening_state_policy = "flat_test_state"` opt-in starts every test window from
the original experiment opening state and marks the result as
`cold_start_distorted`.

Anchored folds grow their train window over time. That is useful for some
research workflows, but later folds can cost more to compute than early folds,
especially with large candidate grids and wide feature sets.

## Non-Scope

This MVP does not implement paper/live walk-forward, OMS behavior,
cross-snapshot walk-forward, benchmark-relative metrics, gross-vs-net
attribution, signal decay, implementation/cost decay, or selection-integrity
diagnostics. Those require later surfaces over the stable session and candidate
identity introduced here.
