---
title: "Getting Started with ledgr"
output:
  rmarkdown::html_vignette:
    toc: false
  github_document:
    toc: false
    html_preview: false
vignette: >
  %\VignetteIndexEntry{Getting Started with ledgr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "figures/getting-started-",
  out.width = "100%"
)
options(width = 90)
options(cli.unicode = FALSE)
default_output_hook <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(
  output = function(x, options) {
    default_output_hook(gsub("[ \t]+(?=\n)", "", x, perl = TRUE), options)
  }
)
demo_data_path <- if (file.exists("data/ledgr_demo_bars.rda")) {
  "data/ledgr_demo_bars.rda"
} else {
  file.path("..", "data", "ledgr_demo_bars.rda")
}
if (!exists("ledgr_demo_bars") && file.exists(demo_data_path)) {
  load(demo_data_path)
}
```

This vignette walks through the v0.1.7 research loop:

1. start from deterministic demo bars;
2. seal them into a snapshot;
3. define an experiment;
4. run one or more parameter sets;
5. inspect the event-derived results;
6. reopen and compare stored runs.

## Data And Snapshot

The examples use `dplyr` and `tibble` for data preparation and compact display.
They are suggested packages used by the vignettes, not part of the strategy
contract.

```{r attach, message=FALSE}
library(ledgr)
library(dplyr)
library(tibble)
data("ledgr_demo_bars", package = "ledgr")
```

`ledgr_demo_bars` is bundled with the package so examples do not need local
files or network access.

```{r demo-data}
bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% c("DEMO_01", "DEMO_02"),
    between(ts_utc, ledgr_utc("2019-01-01"), ledgr_utc("2019-06-30"))
  )

bars |>
  slice_head(n = 6)
```

Create a sealed snapshot. This is the immutable input artifact for all runs in
the experiment.

```{r snapshot}
snapshot <- ledgr_snapshot_from_df(bars)
snapshot
```

For durable work, pass a stable `db_path` at snapshot creation. After that,
normal run and store operations use the snapshot handle.

## Strategy Contract

A ledgr strategy is `function(ctx, params)`. `ctx` is the pulse context; it
contains only state observable at the current decision point. `params` is the
JSON-safe parameter list passed to `ledgr_run()`.

```{r strategy}
strategy <- function(ctx, params) {
  targets <- ctx$flat()

  for (id in ctx$universe) {
    sma <- ctx$feature(id, "sma_20")
    if (is.finite(sma) && ctx$close(id) > sma) {
      targets[id] <- params$qty
    }
  }

  targets
}
```

Targets are desired holdings, not signals. A return value like
`c(DEMO_01 = 10, DEMO_02 = 0)` means "hold 10 units of `DEMO_01` and 0 units
of `DEMO_02`." The names must match `ctx$universe`.

Use `ctx$flat()` when each pulse should restate the whole portfolio from flat.
Use `ctx$hold()` when a rule should keep current positions unless it emits a
new target.

## Experiment And Run

Indicators are feature definitions. Ask ledgr for the feature IDs before using
them in `ctx$feature()`.

```{r features}
features <- list(ledgr_ind_sma(20))
ledgr_feature_id(features)
```

Bundle the reusable parts into an experiment. Construction validates the
snapshot, strategy, features, opening state, universe, and execution options;
it does not execute the strategy.

```{r experiment}
exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = strategy,
  features = features,
  opening = ledgr_opening(cash = 10000)
)

exp
```

Run one parameter set.

```{r run}
bt <- exp |>
  ledgr_run(params = list(qty = 10), run_id = "getting_started_qty_10")

bt
```

A backtest handle points to stored run artifacts. The artifacts are already
durable when `ledgr_run()` returns, and ordinary result inspection opens and
closes read connections per operation. Use `close(bt)` as explicit resource
cleanup in long sessions, tests, explicit-open workflows, and lazy result
cursors.

```r
on.exit(close(bt), add = TRUE)
```

## Inspect Results

```{r summary}
summary(bt)
```

Results are derived views over recorded events.

```{r result-tables}
ledgr_results(bt, what = "trades")
tail(ledgr_results(bt, what = "equity"), 4)
```

The ledger is the source of truth.

```{r ledger}
head(ledgr_results(bt, what = "ledger"), 6)
```

## Debug One Pulse

`ledgr_pulse_snapshot()` builds the same pulse context a strategy sees at a
single timestamp. It is useful for explaining one decision without rerunning an
entire experiment.

```{r pulse}
pulse <- ledgr_pulse_snapshot(
  snapshot = snapshot,
  universe = c("DEMO_01", "DEMO_02"),
  ts_utc = "2019-03-01T00:00:00Z",
  features = features
)

pulse$close("DEMO_01")
pulse$feature("DEMO_01", "sma_20")
strategy(pulse, list(qty = 10))
close(pulse)
```

## Compare Variants

Run a second parameter set into the same snapshot-backed store.

```{r second-run}
bt_qty_20 <- exp |>
  ledgr_run(params = list(qty = 20), run_id = "getting_started_qty_20")

ledgr_compare_runs(snapshot, run_ids = c("getting_started_qty_10", "getting_started_qty_20"))
```

## Durable Store Workflow

Use a stable DuckDB path for research you want to keep. The vignette uses
`tempfile()` so it does not leave files in your project; in real work, replace
that with a project path such as `"research.duckdb"`.

```{r durable}
artifact_db <- tempfile("ledgr_getting_started_", fileext = ".duckdb")
durable_snapshot <- ledgr_snapshot_from_df(
  bars,
  db_path = artifact_db,
  snapshot_id = "getting_started_snapshot"
)

durable_exp <- ledgr_experiment(
  snapshot = durable_snapshot,
  strategy = strategy,
  features = features,
  opening = ledgr_opening(cash = 10000)
)

durable_bt <- durable_exp |>
  ledgr_run(params = list(qty = 10), run_id = "durable_qty_10")

close(durable_bt)
ledgr_snapshot_close(durable_snapshot)

reloaded <- ledgr_snapshot_load(artifact_db, "getting_started_snapshot", verify = TRUE)
ledgr_run_list(reloaded)
ledgr_run_info(reloaded, "durable_qty_10")
ledgr_snapshot_close(reloaded)
```

`ledgr_snapshot_load()` is the new-session resumption path. Store APIs such as
`ledgr_run_list()`, `ledgr_run_info()`, `ledgr_compare_runs()`,
`ledgr_run_label()`, and `ledgr_run_archive()` operate on the snapshot handle.

## Scope

v0.1.7 is a research release. It does not provide live trading, broker
integrations, short-selling semantics, or parameter sweep execution. Sweep and
tune APIs are reserved for later versions.

```{r cleanup}
close(bt)
close(bt_qty_20)
ledgr_snapshot_close(snapshot)
```
