---
title: "Strategy Development And Comparison"
output:
  rmarkdown::html_vignette:
    toc: false
  github_document:
    toc: false
    html_preview: false
vignette: >
  %\VignetteIndexEntry{Strategy Development And Comparison}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(width = 90)
options(cli.unicode = FALSE)
default_output_hook <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(
  output = function(x, options) {
    default_output_hook(gsub("[ \t]+(?=\n)", "", x, perl = TRUE), options)
  }
)
demo_data_path <- if (file.exists("data/ledgr_demo_bars.rda")) {
  "data/ledgr_demo_bars.rda"
} else {
  file.path("..", "data", "ledgr_demo_bars.rda")
}
if (!exists("ledgr_demo_bars") && file.exists(demo_data_path)) {
  load(demo_data_path)
}
```

A ledgr strategy is a function that receives a pulse context and returns target
holdings.

```{r library}
library(ledgr)
library(dplyr)
data("ledgr_demo_bars", package = "ledgr")
```

At each pulse, ledgr gives the strategy the current observable state. The
strategy answers with a full named numeric vector: the desired share quantity
for every instrument in `ctx$universe`.

## The Strategy Function

The simplest form is `function(ctx, params)`.

```{r ctx-strategy}
flat_strategy <- function(ctx, params) {
  ctx$flat()
}
```

Parameterized strategies use `function(ctx, params)`. Parameters arrive as the
second argument; they are not stored on the context object.

```{r params-strategy}
threshold_strategy <- function(ctx, params) {
  targets <- ctx$flat()
  for (id in ctx$universe) {
    if (ctx$close(id) > params$threshold[[id]]) {
      targets[id] <- params$qty
    }
  }
  targets
}
```

## What Is `ctx`?

`ctx` is the pulse context. It represents one decision point. The main fields
and helpers are:

- `ctx$ts_utc`: current pulse timestamp;
- `ctx$universe`: character vector of instrument IDs;
- `ctx$bars`: current OHLCV rows;
- `ctx$features`: long feature table at the current pulse;
- `ctx$features_wide`: wide feature table for scanning;
- `ctx$positions`: current position vector;
- `ctx$cash`: current simulated cash;
- `ctx$equity`: current simulated equity;
- `ctx$open(id)`, `ctx$high(id)`, `ctx$low(id)`, `ctx$close(id)`,
  `ctx$volume(id)`: scalar OHLCV accessors;
- `ctx$position(id)`: current quantity for one instrument;
- `ctx$feature(id, name)`: feature lookup;
- `ctx$flat(default = 0)`: full target vector initialized to `default`;
- `ctx$hold()`: full target vector initialized from current
  positions.

Use `ctx$flat()` when the strategy should go flat unless it explicitly emits
a target. Use `ctx$hold()` when the strategy should hold unless it
explicitly changes a target.

## Targets

Targets must be full named numeric vectors. Names must exactly match
`ctx$universe`, and values are desired quantities, not portfolio weights or
orders. A target of `10` means hold 10 shares after the next fill; a target
of `0` means hold no shares.

```{r target-example}
buy_one_if_up <- function(ctx, params) {
  targets <- ctx$flat()
  if (ctx$close("AAA") > ctx$open("AAA")) {
    targets["AAA"] <- 1
  }
  targets
}
```

v0.1.x is long-only. Negative targets are outside the supported public workflow
until explicit shorting semantics are specified.

The default fill model is next-open: a target decided at pulse `t` fills at the
next available bar. A target change on the final pulse cannot fill.

For multi-asset strategies, build the full vector first and then assign the
quantities you want to hold. This example uses the built-in one-bar return
indicator, whose feature ID is `return_1`; the next section shows how to ask
ledgr for feature IDs before using them in a strategy.

```{r sizing-example}
top_two_equal_quantity <- function(ctx, params) {
  targets <- ctx$flat()
  scores <- numeric()

  for (id in ctx$universe) {
    score <- ctx$feature(id, "return_1")
    if (!is.na(score)) {
      scores[id] <- score
    }
  }

  if (length(scores) == 0) return(targets)

  selected <- names(sort(scores, decreasing = TRUE))[seq_len(min(2, length(scores)))]
  targets[selected] <- params$qty_per_instrument
  targets
}
```

## Indicators And Feature IDs

Built-in indicators are ordinary `ledgr_indicator` objects.

```{r builtins}
sma_3 <- ledgr_ind_sma(3)
ret_1 <- ledgr_ind_returns(1)
ledgr_feature_id(list(sma_3, ret_1))
```

TTR indicators use `ledgr_ind_ttr()`.

```{r ttr}
rsi_3 <- ledgr_ind_ttr("RSI", input = "close", n = 3)
bb_up <- ledgr_ind_ttr("BBands", input = "close", output = "up", n = 3)
ledgr_feature_id(list(rsi_3, bb_up))
```

Feature IDs are exact strings. Ask ledgr for them before writing
`ctx$feature()` calls.

Feature values can be `NA` during warmup. Unknown feature IDs fail loudly;
warmup `NA` for known features is normal. This distinction matters: a typo such
as `"returns_1"` fails because no configured feature has that ID, while
`"return_1"` may return `NA` on early bars if the indicator is not ready.

```{r indicator-strategy}
rsi_strategy <- function(ctx, params) {
  targets <- ctx$hold()
  rsi <- ctx$feature("AAA", "ttr_rsi_3")
  if (is.na(rsi)) return(targets)

  if (rsi < params$buy_below) {
    targets["AAA"] <- params$qty
  } else if (rsi > params$sell_above) {
    targets["AAA"] <- 0
  }
  targets
}
```

`requires_bars` is the first row where an indicator may emit a value.
`stable_after` is the first row ledgr treats as stable. Built-in and supported
TTR indicators set these for you. When a strategy sees `NA`, the ordinary
pattern is to return `ctx$hold()` for hold-unless-signal strategies or
`ctx$flat()` for flat-unless-signal strategies.

```{r warmup-flat}
sma_breakout <- function(ctx, params) {
  targets <- ctx$flat()
  sma <- ctx$feature("AAA", "sma_3")
  if (is.na(sma)) return(targets)

  if (ctx$close("AAA") > sma) {
    targets["AAA"] <- params$qty
  }
  targets
}
```

On very short histories, this is not an error: the run can complete with no
trades because the feature never becomes ready. That is different from an
insufficient-history custom indicator that returns `NA` after `stable_after`;
post-warmup `NA` values are invalid and ledgr aborts rather than silently using
bad features.

## Debug One Pulse

`ledgr_pulse_snapshot()` freezes the context at one timestamp so you can inspect
what a strategy saw.

```{r data}
bars <- ledgr_demo_bars |>
  filter(
    instrument_id %in% c("DEMO_01", "DEMO_02"),
    between(ts_utc, ledgr_utc("2019-01-01"), ledgr_utc("2019-04-30"))
  )

snapshot <- ledgr_snapshot_from_df(
  bars,
  snapshot_id = "strategy_demo_snapshot"
)
```

```{r pulse}
pulse <- ledgr_pulse_snapshot(
  snapshot,
  universe = c("DEMO_01", "DEMO_02"),
  ts_utc = "2019-03-01T00:00:00Z",
  features = list(sma_3, rsi_3)
)

pulse$close("DEMO_01")
pulse$feature("DEMO_01", "sma_3")
pulse$hold()
threshold_strategy(
  pulse,
  list(threshold = c(DEMO_01 = 55, DEMO_02 = 75), qty = 1)
)
close(pulse)
```

## Compare Parameter Variants

Run related strategy variants through one experiment.

```{r compare-params}
exp <- ledgr_experiment(
  snapshot = snapshot,
  strategy = threshold_strategy,
  opening = ledgr_opening(cash = 10000)
)

bt_qty_1 <- exp |>
  ledgr_run(
    params = list(threshold = c(DEMO_01 = 55, DEMO_02 = 75), qty = 1),
    run_id = "threshold_qty_1"
  )

bt_qty_3 <- exp |>
  ledgr_run(
    params = list(threshold = c(DEMO_01 = 55, DEMO_02 = 75), qty = 3),
    run_id = "threshold_qty_3"
  )

ledgr_compare_runs(snapshot, run_ids = c("threshold_qty_1", "threshold_qty_3"))
```

## Compare Different Strategies

The same comparison table also works across different strategy functions.

```{r compare-strategies}
flat_exp <- ledgr_experiment(
  snapshot,
  strategy = flat_strategy,
  opening = ledgr_opening(cash = 10000)
)

bt_flat <- flat_exp |>
  ledgr_run(params = list(), run_id = "flat")

ledgr_compare_runs(snapshot, run_ids = c("threshold_qty_1", "flat"))
```

## Inspect Stored Source

`ledgr_extract_strategy(..., trust = FALSE)` returns stored source and metadata
without parsing or evaluating it.

```{r extract}
extracted <- ledgr_extract_strategy(snapshot, "threshold_qty_1", trust = FALSE)
extracted
```

Use `trust = TRUE` only when you explicitly trust the experiment store and want
to recover a function object. Hash verification proves stored-text identity, not
safety.

```{r cleanup}
close(bt_qty_1)
close(bt_qty_3)
close(bt_flat)
close(snapshot)
```
