---
title: "ledgr Peer Parity And Performance Benchmark Report"
format: gfm
output-file: peer_benchmark.md
execute:
  echo: false
  warning: false
  message: true
params:
  run_harness: false
  preset: "smoke"
  results_preset: "record"
  results_prefix: "peer_benchmark_record_20260926T080725Z"
---

This tracked report records the v0.2.1.0 internal same-host peer workload.
Parity is interpreted before timing. The report preserves phase boundaries,
unavailable engines, and classified divergences without turning one host and
one fixture into a public engine ranking.

The exact sampled record command was:

```powershell
& dev/bench/peer_benchmark/run_record.ps1 `
  -RepoRoot "C:\tmp\ledgr-ws12" `
  -Release "v0.2.1.0" `
  -LedgrOrder "ttr-first"
```

The exact ignored local prefix is
`dev/bench/results/peer_benchmark_record_20260926T080725Z`. It was measured
from commit `c7d936ddccbdf7affbdcd1c9be22f7d8fdaa7eb0`; the exact benchmark
harness has SHA-256
`9ac0c8f2b5f276000f10028104de9824f106719705307329853f6a2de283edaf`.
It ran under R 4.6.1 ucrt on Windows build 26200. Quantstrat 0.25 completed
from the isolated library with the ticket-pinned blotter 0.17.0,
FinancialInstrument 1.3.0, xts 0.14.2, and TTR 0.24.4. Backtrader and
zipline-reloaded completed through pinned `uv` environments. Local LEAN was
unavailable because its CLI root uses an obsolete organization layout. No
hosted LEAN service was used.

The report checks every ledgr row before interpreting its timing. The canonical
public sweep completed in 39.68 seconds cold and 32.31 seconds warm; the public
compiled sweep completed in 27.19 seconds cold and 17.60 seconds warm. The two
public sweep rows are exact across all 1,260 equity rows, 68,201 fills, and the
realized-trade row after normalizing only the engine label. Against durable
ledgr, fills and trades are exact; compensated production-inline equity passes
the existing relative `all.equal()` tolerance of `1e-8`, with maximum absolute
residual `1.9744e-07`, maximum relative residual `1.971e-14`, and affected
column `equity` across 1,222 rows. Earlier diagnostic
peer CSVs used reconstructed equity; this final record uses compensated
production-inline equity as the memory reference. Richer surfaces come from an
explicitly untimed private oracle and are used only for parity. The report
retains peer residuals and partial surfaces rather than turning a numerical
tolerance into a claim of full parity. Zipline now pairs all 1,260 source
sessions positionally and records daily-return correlation `0.987974`; the
v0.2.0.1 report's `0.146064` was produced by the disclosed one-day timestamp
misalignment and remains historical evidence rather than being rewritten.

Workstream 17 also corrects the indicator boundary. All published TTR rows now
use exported `ledgr_ind_ttr("SMA")`; the retired private wrapper is absent.
Durable public TTR and built-in SMA ran in separate fresh R processes with
identical R, library-path and package-version records. Their feature axes and
NA masks are exact; 1,260,000 numeric cells pass the existing `1e-8` tolerance
with maximum absolute residual `7.567e-10`. Equity, fills and realized trades
are exact. Timing does not identify a source effect: the promoted TTR-first
record reports warm clocks of 53.02 and 68.33 seconds. The six-order
attribution spike independently places every source in every process position
twice; each contrast changes direction. That spike classifies the old
9.57-second chart gap as order-confounded rather than a package-speed
difference. A later built-in-first diagnostic is excluded because its command
was not recorded and its first-position clock did not reproduce the spike.

> **Scope of this report**
>
> Internal repo-local benchmark. Same host, same seed, single fixture.
> A B2 row, when explicitly requested, remains an opt-in
> (`compiled_accounting_model = "spot_fifo"`) on ledgr's public
> one-candidate sweep path. Default ledgr execution remains canonical R.
> Numbers are for maintainer use and release closeout, not a public ranking or
> a universal performance claim.

```{r}
#| echo: false
candidates <- c(".", "..", file.path("..", ".."), file.path("..", "..", ".."))
repo_root <- candidates[file.exists(file.path(candidates, "DESCRIPTION"))][[1L]]
repo_root <- normalizePath(repo_root, mustWork = TRUE)
peer_harness <- new.env(parent = globalenv())
sys.source(
  file.path(repo_root, "dev", "bench", "peer_benchmark", "peer_benchmark.R"),
  envir = peer_harness
)
result_dir <- file.path(repo_root, "dev", "bench", "results")
if (nzchar(params$results_prefix)) {
  status_files <- file.path(
    result_dir,
    paste0(params$results_prefix, "_status.csv")
  )
} else {
  status_files <- Sys.glob(file.path(
    result_dir,
    sprintf("peer_benchmark_%s_*_status.csv", params$results_preset)
  ))
  status_files <- status_files[!grepl("_surface_status[.]csv$", status_files)]
}
if (!length(status_files) || !all(file.exists(status_files))) {
  stop(
    "No peer benchmark status file found for prefix/preset: ",
    params$results_prefix,
    if (nzchar(params$results_prefix)) "" else params$results_preset
  )
}
status_file <- tail(sort(status_files), 1L)
stem <- sub("_status[.]csv$", "", status_file)
read_result <- function(suffix) {
  path <- paste0(stem, suffix)
  if (!file.exists(path)) stop("Missing expected result file: ", path)
  utils::read.csv(path, stringsAsFactors = FALSE)
}
status <- read_result("_status.csv")
parity <- read_result("_parity.csv")
feature_parity <- read_result("_feature_parity.csv")
performance <- read_result("_performance.csv")
report_performance <- peer_harness$peer_prepare_report_performance(performance)
if (!requireNamespace("ggplot2", quietly = TRUE)) {
  stop("Rendering the peer benchmark report requires ggplot2.")
}
compiled_engine <- "ledgr_ttr_compiled_spot_fifo_sweep"
has_compiled <- compiled_engine %in% performance$engine
if (!"cost_model" %in% names(performance)) {
  performance$cost_model <- NA_character_
}
if (!"risk_chain" %in% names(performance)) {
  performance$risk_chain <- NA_character_
}
if (!"compiled_accounting_model" %in% names(performance)) {
  performance$compiled_accounting_model <- NA_character_
}
divergence_summary <- read_result("_divergence_summary.csv")
surface <- read_result("_surface_status.csv")
env <- jsonlite::read_json(paste0(stem, "_environment.json"), simplifyVector = TRUE)
ttr_metadata <- jsonlite::read_json(
  paste0(stem, "_ledgr_ttr_canonical_metadata.json"),
  simplifyVector = FALSE
)
builtin_metadata <- jsonlite::read_json(
  paste0(stem, "_ledgr_builtin_sma_metadata.json"),
  simplifyVector = FALSE
)
if (!identical(ttr_metadata$runtime, builtin_metadata$runtime)) {
  stop("Fresh public TTR and built-in rows resolved different runtimes.")
}
stopifnot(
  isTRUE(ttr_metadata$fresh_process),
  isTRUE(builtin_metadata$fresh_process),
  isTRUE(feature_parity$axis_identical[[1L]]),
  isTRUE(feature_parity$na_mask_identical[[1L]]),
  isTRUE(feature_parity$within_tolerance[[1L]]),
  feature_parity$max_relative[[1L]] <= feature_parity$tolerance[[1L]]
)
peer_peak_path <- paste0(stem, "_working_set_peak.csv")
if (!file.exists(peer_peak_path)) {
  stop("The final peer record requires a process-tree peak-working-set sidecar.")
}
peer_peak <- utils::read.csv(peer_peak_path, stringsAsFactors = FALSE)
quantstrat_env_path <- paste0(stem, "_quantstrat_environment.json")
quantstrat_env <- if (file.exists(quantstrat_env_path)) {
  jsonlite::read_json(quantstrat_env_path, simplifyVector = FALSE)
} else {
  NULL
}
if (any(status$engine == "quantstrat" & status$status == "DONE") &&
    is.null(quantstrat_env)) {
  stop("A DONE quantstrat row requires its isolated-environment sidecar.")
}
equity_files <- Sys.glob(paste0(stem, "_*_equity.csv"))
equity <- do.call(rbind, lapply(equity_files, utils::read.csv, stringsAsFactors = FALSE))

fmt_num <- function(x, digits = 6) {
  ifelse(is.na(x), NA, formatC(as.numeric(x), digits = digits, format = "fg", flag = "#"))
}
fmt_sec <- function(x) ifelse(is.na(x), "NULL", sprintf("%.3f", as.numeric(x)))
fmt_rate <- function(x) {
  ifelse(is.na(x), NA, formatC(as.numeric(x), digits = 6, format = "fg", big.mark = ","))
}
fmt_pct <- function(x, digits = 4) {
  pct <- 100 * as.numeric(x)
  ifelse(
    is.na(pct),
    NA,
    ifelse(abs(pct) < 0.0001 & pct != 0, "<0.0001%", paste0(formatC(pct, digits = digits, format = "fg", flag = "#"), "%"))
  )
}
fmt_corr <- function(x) ifelse(is.na(x), NA, sprintf("%.6f", as.numeric(x)))
fmt_signed <- function(x, digits = 6) {
  ifelse(is.na(x), NA, formatC(as.numeric(x), digits = digits, format = "fg", flag = "#"))
}
fmt_pp <- function(x, digits = 5) {
  pp <- 100 * as.numeric(x)
  ifelse(
    is.na(pp),
    NA,
    ifelse(abs(pp) < 0.00001 & pp != 0, "<0.00001 pp", paste0(formatC(pp, digits = digits, format = "fg", flag = "#"), " pp"))
  )
}
parity_surface <- function(peer) {
  trade_surface <- surface$trades[match(peer, surface$engine)]
  fill_surface <- surface$fills[match(peer, surface$engine)]
  if (identical(fill_surface, "available") &&
      identical(trade_surface, "trade_count_available_only")) {
    return("partial: equity + fills + trade count; no realized trade P&L")
  }
  if (identical(fill_surface, "available") && identical(trade_surface, "available_realized_pnl")) {
    return("equity + fills + realized trades")
  }
  if (identical(trade_surface, "available_realized_pnl")) {
    return("equity + realized trades; fills unavailable")
  }
  trade_surface %||% "unavailable"
}
```

## Optional B2 timing context

When requested, the B2 row uses ledgr's explicit
`compiled_accounting_model = "spot_fifo"` hot frame on the same fixture, seed,
shared bars, and SMA crossover strategy as the other rows. If it is absent,
this section reports that state rather than synthesizing a comparison.

```{r}
#| echo: false
engine_labels <- c(
  ledgr_ttr_compiled_spot_fifo_sweep = "ledgr B2\npublic sweep",
  backtrader = "Backtrader",
  ledgr_ttr_canonical_sweep = "ledgr canonical\npublic sweep",
  `zipline-reloaded-full` = "Zipline",
  quantstrat = "Quantstrat"
)
compare_engines <- names(engine_labels)
compare <- performance[match(compare_engines, performance$engine), , drop = FALSE]
compare <- compare[!is.na(compare$engine), , drop = FALSE]
if (!compiled_engine %in% compare$engine) {
  knitr::kable(data.frame(
    Note = "No B2 spot-FIFO peer-shaped row found beside this record bundle. Run the record command with --compiled-accounting-model spot_fifo to populate this section."
  ))
} else {
  b2 <- compare[compare$engine == compiled_engine, , drop = FALSE]
  bt <- compare[compare$engine == "backtrader", , drop = FALSE]
  speed_table <- data.frame(
    Surface = c("Cold end to end", "Warm research iteration", "Engine phase"),
    `B2 seconds` = fmt_sec(c(
      b2$cold_end_to_end, b2$warm_research_iteration, b2$engine_sec
    )),
    `Backtrader seconds` = fmt_sec(c(
      bt$cold_end_to_end, bt$warm_research_iteration, bt$engine_sec
    )),
    `B2 / Backtrader` = sprintf("%.2fx", c(
      b2$cold_end_to_end / bt$cold_end_to_end,
      b2$warm_research_iteration / bt$warm_research_iteration,
      b2$engine_sec / bt$engine_sec
    )),
    `B2 reduction vs Backtrader` = paste0(sprintf("%.1f", 100 * c(
      1 - b2$cold_end_to_end / bt$cold_end_to_end,
      1 - b2$warm_research_iteration / bt$warm_research_iteration,
      1 - b2$engine_sec / bt$engine_sec
    )), "%"),
    check.names = FALSE
  )
  knitr::kable(speed_table, row.names = FALSE)
}
```

The three clocks answer different questions. Cold includes the reusable sealed
snapshot. Warm measures the public research iteration over that unchanged
snapshot. Engine isolates the fold slice replaced by the compiled hot frame.
The timed sweep results phase uses retained inline facts and public accessors;
the private full-surface parity oracle runs outside every reported clock.

```{r}
#| echo: false
#| fig.width: 10
#| fig.height: 5.4
if (compiled_engine %in% compare$engine) {
  timing_plot <- rbind(
    data.frame(
      engine = compare$engine,
      surface = "Cold end to end",
      seconds = compare$cold_end_to_end
    ),
    data.frame(
      engine = compare$engine,
      surface = "Warm research iteration",
      seconds = compare$warm_research_iteration
    ),
    data.frame(
      engine = compare$engine,
      surface = "Engine phase",
      seconds = compare$engine_sec
    )
  )
  timing_plot <- timing_plot[is.finite(timing_plot$seconds), , drop = FALSE]
  timing_plot$label <- unname(engine_labels[timing_plot$engine])
  timing_plot$label <- factor(
    timing_plot$label,
    levels = rev(unname(engine_labels[compare$engine]))
  )
  timing_plot$surface <- factor(
    timing_plot$surface,
    levels = c("Cold end to end", "Warm research iteration", "Engine phase")
  )
  timing_plot$kind <- ifelse(
    timing_plot$engine == compiled_engine,
    "Compiled sweep",
    ifelse(timing_plot$engine == "backtrader", "Backtrader", "Other")
  )
  backtrader_lines <- timing_plot[
    timing_plot$engine == "backtrader",
    c("surface", "seconds"),
    drop = FALSE
  ]
  print(
    ggplot2::ggplot(
      timing_plot,
      ggplot2::aes(x = seconds, y = label, fill = kind)
    ) +
      ggplot2::geom_col(width = 0.72) +
      ggplot2::geom_text(
        ggplot2::aes(label = sprintf("%.2fs", seconds)),
        hjust = -0.08,
        size = 3
      ) +
      ggplot2::geom_vline(
        data = backtrader_lines,
        ggplot2::aes(xintercept = seconds),
        linetype = 2,
        colour = "#555555",
        inherit.aes = FALSE
      ) +
      ggplot2::facet_wrap(~surface, scales = "free_x", nrow = 1) +
      ggplot2::scale_fill_manual(values = c(
        "Compiled sweep" = "#1F77B4",
        "Backtrader" = "#7F7F7F",
        "Other" = "#C7C7C7"
      )) +
      ggplot2::scale_x_continuous(expand = ggplot2::expansion(mult = c(0, 0.22))) +
      ggplot2::labs(
        title = "Optional B2 timing on the same fixture",
        subtitle = "Interpret only after parity; dashed line marks Backtrader",
        x = "Seconds (lower is better)",
        y = NULL,
        fill = NULL
      ) +
      ggplot2::theme_minimal(base_size = 10) +
      ggplot2::theme(legend.position = "bottom")
  )
}
```

## Indicator source boundary

The old report always ran a private TTR wrapper before built-in SMA in one R
process. That row order could not attribute its 9.57-second difference. This
record replaces the wrapper with exported `ledgr_ind_ttr("SMA")` everywhere
and runs the two durable rows in independent child processes. LTB-0075 also
reverses their launch order on a small executable fixture; that detector is
not another peer record.

| Surface | Standard | Result | Max absolute residual | Max relative residual |
|---|---|---:|---:|---:|
| feature axis | exact | PASS | 0 | 0 |
| feature NA mask | exact | PASS | 0 | 0 |
| feature values | relative `all.equal()` tolerance `1e-8` | PASS | 7.567e-10 | 1.182e-13 |

| Order | Source | Cold s | Warm s | Engine s |
|---|---|---:|---:|---:|
| TTR first | public TTR SMA | 60.82 | 53.02 | 47.44 |
| TTR first | built-in SMA | 76.16 | 68.33 | 61.67 |

The six-permutation attribution spike carries the source conclusion. Its
first-position means were 50.635 seconds for built-in SMA, 51.050 for the
private wrapper and 51.295 for public TTR, and every pair changed direction by
process position. The old 9.57-second gap therefore did not reproduce as an
indicator-source cost.

A later built-in-first diagnostic exists beside the promoted record, but its
command was not recorded and its first-launched built-in row took 65.50
seconds against the spike's matching first-position observations of 50.530
and 50.740 seconds. It is not loaded by this report and supports no attribution
claim. The public-adapter
correction is a measurement-boundary fix, not a package speedup.

## Why parity comes before timing

A speed number only matters if the engine produces the same answer. When the
compiled row is present, the following check requires exact identity for
fills, trades, and every non-floating equity field. Floating equity uses only
the already registered relative `all.equal()` tolerance of `1e-8` and reports
its residuals.

```{r}
#| echo: false
canonical_ephemeral_engine <- "ledgr_ttr_canonical_sweep"
read_surface <- function(engine, surface_name) {
  path <- paste0(stem, "_", engine, "_", surface_name, ".csv")
  if (!file.exists(path)) stop("Missing ledgr parity surface: ", path)
  out <- utils::read.csv(path, stringsAsFactors = FALSE, check.names = FALSE)
  if ("engine" %in% names(out)) out$engine <- "ledgr"
  out
}
if (!has_compiled) {
  knitr::kable(data.frame(
    Note = "Compiled spot-FIFO row absent; no compiled timing is interpreted."
  ))
} else {
  canonical_equity <- read_surface(canonical_ephemeral_engine, "equity")
  compiled_equity <- read_surface(compiled_engine, "equity")
  equity_residual <- peer_harness$peer_equity_parity(
    canonical_equity,
    compiled_equity,
    floating_columns = c(
      "equity", "cash", "positions_value", "position_proxy",
      "realized_pnl", "unrealized_pnl"
    ),
    context = "Rendered compiled-versus-canonical equity"
  )
  canonical_fills <- read_surface(canonical_ephemeral_engine, "fills")
  compiled_fills <- read_surface(compiled_engine, "fills")
  canonical_trades <- read_surface(canonical_ephemeral_engine, "trades")
  compiled_trades <- read_surface(compiled_engine, "trades")
  peer_harness$peer_exact_surface(
    canonical_fills,
    compiled_fills,
    "Rendered compiled-versus-canonical fills"
  )
  peer_harness$peer_exact_surface(
    canonical_trades,
    compiled_trades,
    "Rendered compiled-versus-canonical trades"
  )
  b2_parity <- data.frame(
    Surface = c("equity", "fills", "trades"),
    Standard = c("relative all.equal tolerance 1e-8", "exact", "exact"),
    Result = c("PASS", "PASS", "PASS"),
    `Max absolute residual` = c(equity_residual$max_abs, 0, 0),
    `Max relative residual` = c(equity_residual$max_relative, 0, 0),
    `Affected columns` = c(
      paste(equity_residual$affected_columns, collapse = ", "),
      "none",
      "none"
    ),
    `Affected rows` = c(length(equity_residual$affected_rows), 0, 0),
    check.names = FALSE
  )
  knitr::kable(b2_parity, row.names = FALSE)
}
```

The report renders compiled timing only after the registered comparisons pass.
Peer-by-peer tolerance results against durable ledgr follow below; timing does
not waive a failed semantic comparison.

## Full peer comparison

A compiled row, when requested, uses the same fixture as the other engines.
Backtrader, Zipline, quantstrat, and LEAN retain their declared lifecycle
boundaries and availability states; no row is treated as universally
representative of its engine or language.

Canonical durable and public-sweep rows track the audited run workflow and the
interactive research workflow ledgr ships. An optional B2 sweep row tracks the
opt-in compiled hot frame. Keeping those labels distinct prevents reading a
compiled opt-in result as a default ledgr claim.

```{r}
#| echo: false
display_performance <- report_performance$display
display_full_row_sec <- display_performance$full_row_sec
perf_table <- data.frame(
  `Engine row` = display_performance$engine,
  Status = display_performance$status,
  `Full row` = fmt_sec(display_full_row_sec),
  `Snapshot prepare` = fmt_sec(display_performance$snapshot_prepare_sec),
  `Setup / orchestration` = fmt_sec(display_performance$experiment_setup_sec),
  `Engine phase` = fmt_sec(display_performance$engine_sec),
  Results = fmt_sec(display_performance$results_sec),
  Cold = fmt_sec(display_performance$cold_end_to_end),
  Warm = fmt_sec(display_performance$warm_research_iteration),
  `Bars/sec` = fmt_rate(ifelse(
    display_performance$status == "DONE",
    display_performance$core_bars_per_sec,
    NA_real_
  )),
  check.names = FALSE
)
has_cost_risk_columns <- any(!is.na(performance$cost_model)) ||
  any(!is.na(performance$risk_chain)) ||
  any(!is.na(performance$compiled_accounting_model))
if (has_cost_risk_columns) {
  perf_table <- cbind(
    perf_table[, c("Engine row", "Status"), drop = FALSE],
    Cost = performance$cost_model,
    Risk = performance$risk_chain,
    Compiled = performance$compiled_accounting_model,
    perf_table[, setdiff(names(perf_table), c("Engine row", "Status")), drop = FALSE]
  )
}
knitr::kable(perf_table, row.names = FALSE)
```

Read `Cold` as the sum of snapshot preparation, setup/orchestration, engine,
and results. Read `Warm` as the same row without reusable snapshot preparation.
`Full row` is the wrapper clock around the engine row; phase definitions are in
the Methodology section.

For an unavailable engine, every displayed performance clock is `NULL`. A
failed wrapper attempt may have elapsed time in its raw metadata, but that is
diagnostic latency rather than a completed benchmark result.

The public-sweep setup field is broader than experiment-object construction.
It is the complete public `ledgr_sweep()` wall not covered by its candidate
fold and inline-results clocks, plus experiment and grid construction. It
therefore includes warm snapshot reads, bars matrices and views, validation,
feature and task preparation, and sweep retention/orchestration. The durable
row exposes no matching subclock: analogous work inside `ledgr_run()` remains
in that row's engine phase. Compare cold and warm totals across those rows;
use the colored segments only as each row's declared decomposition.

For peer comparison, this complete boundary is the cold end-to-end view
when the engines perform meaningfully comparable preparation and produce
the required outputs. ledgr's repeated research workflow has a separate
warm clock over an existing unchanged, verified snapshot. Snapshot preparation
is paid once per unchanged data-and-facts identity and again after that
identity changes.

```{r}
#| echo: false
#| fig.width: 9
#| fig.height: 4.8
phase_plot <- report_performance$phase_plot
phase_totals <- report_performance$phase_totals
unavailable_labels <- report_performance$unavailable_labels
print(
  ggplot2::ggplot(
    phase_plot,
    ggplot2::aes(x = seconds, y = engine, fill = phase)
  ) +
    ggplot2::geom_col(
      width = 0.72,
      position = ggplot2::position_stack(reverse = TRUE)
    ) +
    ggplot2::geom_text(
      data = phase_totals,
      ggplot2::aes(
        x = seconds,
        y = engine,
        label = paste0(sprintf("%.3f", seconds), "s")
      ),
      hjust = -0.08,
      size = 3,
      inherit.aes = FALSE
    ) +
    ggplot2::geom_text(
      data = unavailable_labels,
      ggplot2::aes(x = seconds, y = engine, label = label),
      hjust = -0.08,
      colour = "#666666",
      size = 3,
      inherit.aes = FALSE
    ) +
    ggplot2::scale_fill_manual(values = c(
      "Snapshot prepare" = "#7A9E9F",
      "Setup / orchestration" = "#4C78A8",
      "Engine" = "#D19C1D",
      "Results" = "#B85C5C"
    )) +
    ggplot2::scale_x_continuous(expand = ggplot2::expansion(mult = c(0, 0.16))) +
    ggplot2::scale_y_discrete(drop = FALSE) +
    ggplot2::labs(
      title = "Record performance timing",
      subtitle = "Reusable preparation / warm orchestration / engine / results",
      x = "Seconds",
      y = NULL,
      fill = NULL
    ) +
    ggplot2::theme_minimal(base_size = 10) +
    ggplot2::theme(legend.position = "bottom")
)
```

## Equity curves

Each engine's equity is normalized to its first observed value, so curve
shape is comparable without implying an absolute return ranking. A
flat-overlapping bundle of curves is the expected pattern when parity
holds; visible separation is a divergence-source signal worth
investigating in the parity tables below.

```{r}
#| echo: false
#| fig.width: 9
#| fig.height: 5
equity$ts_utc <- as.POSIXct(equity$ts_utc, tz = "UTC")
equity <- equity[order(equity$engine, equity$ts_utc), ]
equity$normalized_equity <- ave(
  equity$equity,
  equity$engine,
  FUN = function(x) x / x[[1L]]
)
print(
  ggplot2::ggplot(
    equity,
    ggplot2::aes(x = ts_utc, y = normalized_equity, colour = engine)
  ) +
    ggplot2::geom_line(linewidth = 0.75, alpha = 0.9) +
    ggplot2::scale_x_datetime(date_breaks = "6 months", date_labels = "%Y-%m") +
    ggplot2::labs(
      title = "Record equity curves",
      subtitle = "Normalized within each engine",
      x = "Month",
      y = "Equity / first observed equity",
      colour = NULL
    ) +
    ggplot2::theme_minimal(base_size = 10) +
    ggplot2::theme(
      legend.position = "bottom",
      axis.text.x = ggplot2::element_text(angle = 35, hjust = 1)
    )
)
```

## Peer parity

ledgr's canonical TTR row is the parity reference. Each peer is compared
against it across three tiers: per-bar equity behavior, derived top-line
metrics, and trade-level surface. A peer that disagrees with canonical
TTR gets a `review` status and a documented divergence attribution
rather than being treated as wrong by default.

The mental move when a parity check fails is to consider three
candidates in order: ledgr is wrong, the peer is wrong, the harness is
wrong. The residual difference is then attributed to a named source:
indicator initialization window, fill-timing edge, cost or margin
default, position-sizing rounding, timestamp alignment, or
float-ordering rounding.

### Tier 1: per-bar equity behavior

```{r}
#| echo: false
surface_idx <- match(parity$peer, surface$engine)
status_idx <- match(parity$peer, status$engine)
full_surface <-
  surface$equity[surface_idx] == "available" &
  surface$fills[surface_idx] == "available" &
  surface$trades[surface_idx] == "available_realized_pnl"
partial_surface <-
  status$status[status_idx] == "DONE" &
  !is.na(full_surface) &
  !full_surface
weak_return_correlation <-
  is.finite(parity$tier1_daily_return_cor) &
  parity$tier1_daily_return_cor < 0.99
retained_share <- if ("tier1_retained_share" %in% names(parity)) {
  parity$tier1_retained_share
} else {
  rep(NA_real_, nrow(parity))
}
tier1 <- data.frame(
  Peer = parity$peer,
  `Tier 1` = ifelse(
    partial_surface,
    "review (partial surface)",
    ifelse(
      weak_return_correlation,
      "review (weak return correlation)",
      ifelse(parity$attribution == "passes Tier 1 tolerance", "pass", "review")
    )
  ),
  `Parity surface` = vapply(parity$peer, parity_surface, character(1)),
  `Retained rows` = ifelse(
    is.finite(retained_share),
    sprintf("%.1f%%", 100 * retained_share),
    "not recorded"
  ),
  `Equity corr` = fmt_corr(parity$tier1_equity_cor),
  `Max div` = fmt_pct(parity$tier1_max_single_bar_divergence_pct, digits = 4),
  `Return corr` = fmt_corr(parity$tier1_daily_return_cor),
  check.names = FALSE
)
knitr::kable(tier1, row.names = FALSE)
```

Rows in `review` carry an explicit attribution:

```{r}
#| echo: false
review_mask <- parity$attribution != "passes Tier 1 tolerance" |
  partial_surface |
  weak_return_correlation
review <- parity[review_mask, c("peer", "attribution"), drop = FALSE]
review_idx <- match(review$peer, parity$peer)
review_partial <- partial_surface[review_idx]
review_weak <- weak_return_correlation[review_idx] & !review_partial
review$attribution[review_partial] <-
  "partial surface: no comparable realized-trade P&L"
review$attribution[review_weak] <- sprintf(
  "equity-level tolerance passes, but daily-return correlation is %.6f",
  parity$tier1_daily_return_cor[review_idx[review_weak]]
)
names(review) <- c("Peer", "Review attribution")
if (nrow(review) == 0L) {
  review <- data.frame(Peer = "(none)", `Review attribution` = "All peer rows passed Tier 1 tolerance.", check.names = FALSE)
}
knitr::kable(review, row.names = FALSE)
```

```{r}
#| echo: false
#| fig.width: 8
#| fig.height: 4.5
div <- 100 * as.numeric(parity$tier1_max_single_bar_divergence_pct)
plot_div <- pmax(div, 1e-12)
ord <- order(plot_div)
divergence_plot <- data.frame(
  peer = parity$peer[ord],
  display_value = plot_div[ord],
  label = ifelse(div[ord] == 0, "0%", fmt_pct(div[ord] / 100, digits = 4)),
  status = ifelse(
    parity$attribution[ord] == "passes Tier 1 tolerance",
    "Pass",
    "Review"
  ),
  stringsAsFactors = FALSE
)
divergence_plot$peer <- factor(
  divergence_plot$peer,
  levels = divergence_plot$peer
)
print(
  ggplot2::ggplot(
    divergence_plot,
    ggplot2::aes(x = display_value, y = peer, colour = status)
  ) +
    ggplot2::geom_point(size = 2.4) +
    ggplot2::geom_text(
      ggplot2::aes(label = label),
      hjust = -0.12,
      size = 3,
      show.legend = FALSE
    ) +
    ggplot2::scale_x_log10(
      expand = ggplot2::expansion(mult = c(0.03, 0.28))
    ) +
    ggplot2::scale_colour_manual(values = c("Pass" = "#5B8C5A", "Review" = "#B85C5C")) +
    ggplot2::labs(
      title = "Tier 1 max divergence vs ledgr canonical",
      subtitle = "Log scale keeps tiny and zero results visible",
      x = "Max divergence percent; zeros shown at display floor",
      y = NULL,
      colour = NULL
    ) +
    ggplot2::theme_minimal(base_size = 10) +
    ggplot2::theme(legend.position = "bottom")
)
```

### Tier 2: derived top-line metrics

```{r}
#| echo: false
tier2 <- data.frame(
  Peer = parity$peer,
  `Total return delta` = fmt_pp(parity$tier2_total_return_diff, digits = 5),
  `Sharpe diff` = fmt_signed(parity$tier2_sharpe_diff, digits = 6),
  `Max DD delta` = fmt_pp(parity$tier2_max_drawdown_diff, digits = 5),
  check.names = FALSE
)
knitr::kable(tier2, row.names = FALSE)
```

### Tier 3: trade-level surface

```{r}
#| echo: false
tier3 <- data.frame(
  Peer = parity$peer,
  `Trade count diff` = parity$tier3_trade_count_diff,
  `Trade surface` = surface$trades[match(parity$peer, surface$engine)],
  `Parity note` = vapply(parity$peer, parity_surface, character(1)),
  check.names = FALSE
)
knitr::kable(tier3, row.names = FALSE)
```

### Per-bar divergence attribution

Every DONE peer row writes a per-bar `_divergence.csv` plus a
`_divergence_summary.csv`. The attribution columns below decompose total
absolute equity divergence against `ledgr_ttr_canonical` into named
sources.

```{r}
#| echo: false
div_table <- divergence_summary
pct_cols <- grep("^pct_attributable_", names(div_table), value = TRUE)
zero_divergence <- div_table$total_abs_divergence == 0
for (col in pct_cols) {
  div_table[[col]] <- fmt_pct(div_table[[col]], digits = 4)
  div_table[[col]][zero_divergence] <- "n/a (zero divergence)"
}
names(div_table) <- c(
  "Peer",
  "Total abs divergence",
  "Diverging bars",
  "First divergence",
  "Indicator warmup",
  "Fill timing",
  "Calendar",
  "Position size",
  "Float rounding",
  "Other"
)
div_table[["Total abs divergence"]] <- fmt_num(div_table[["Total abs divergence"]], digits = 7)
knitr::kable(div_table, row.names = FALSE)
```

## Methodology

### Harness

The harness lives at `dev/bench/peer_benchmark/peer_benchmark.R`. It
runs two checks per peer: numerical parity against ledgr canonical TTR,
and same-host phase timing. The parity check asks whether the engines
agree on this strategy and data. The timing check asks how long this
local same-host harness took under the declared boundary.

```{r}
if (isTRUE(params$run_harness)) {
  candidates <- c(".", "..", file.path("..", ".."), file.path("..", "..", ".."))
  repo_root <- candidates[file.exists(file.path(candidates, "DESCRIPTION"))][[1L]]
  repo_root <- normalizePath(repo_root, mustWork = TRUE)
  source(file.path(repo_root, "dev", "bench", "peer_benchmark", "peer_benchmark.R"))
  peer_main(peer_parse_args(c("--preset", params$preset)))
} else {
  message("Harness execution disabled. Render with -P run_harness:true to run.")
}
```

Smoke command:

```powershell
& "C:\Program Files\R\R-4.6.1\bin\x64\Rscript.exe" dev/bench/peer_benchmark/peer_benchmark.R --preset smoke
```

Record command:

```powershell
& "C:\Program Files\R\R-4.6.1\bin\x64\Rscript.exe" dev/bench/peer_benchmark/peer_benchmark.R --preset record
```

Record command with the opt-in B2 row:

```powershell
& "C:\Program Files\R\R-4.6.1\bin\x64\Rscript.exe" dev/bench/peer_benchmark/peer_benchmark.R --preset record --compiled-accounting-model spot_fifo
```

Current-surface cost/risk-chain command:

```powershell
& "C:\Program Files\R\R-4.6.1\bin\x64\Rscript.exe" dev/bench/peer_benchmark/peer_benchmark.R --preset record --engine-set ledgr-cost --release v0.1.9.6
```

The harness writes ignored local artifacts under `dev/bench/results/`:
one shared bars CSV with input hash, per-engine canonical equity curves
where available, fills and trade summary tables where available, engine
surface-status rows for unavailable outputs, Tier 1/2/3 parity CSVs,
performance timing CSV, environment metadata JSON, and append-only
parity history JSON under `dev/bench/results/parity_history/`.

### Required rows

| Row | Status policy |
| --- | --- |
| ledgr canonical TTR | Required; errors if `TTR` is missing. |
| ledgr canonical TTR public sweep | Required; calls public `ledgr_sweep()` with one candidate and retained returns/trades. Full equity/fill parity is checked afterward by an untimed internal oracle. |
| ledgr current-surface cost/risk sweep | Optional; enabled with `--engine-set ledgr-cost`; measures public cost/risk-chain sweeps on the same fixture and seed. The legacy resolver row remains an internal diagnostic. |
| ledgr B2 spot-FIFO public sweep | Optional; enabled with `--compiled-accounting-model spot_fifo`; calls the same public one-candidate sweep with the closed compiled selector and must match the canonical sweep before timing is interpreted. |
| ledgr built-in SMA diagnostic | Required; runs through ledgr built-ins. |
| quantstrat | Runs when local R packages are installed; otherwise explicit unavailable row. |
| Backtrader | Managed through `dev/bench/peer_benchmark/python/backtrader/` and run through `python -m uv`. |
| zipline-reloaded full engine | Managed through `dev/bench/peer_benchmark/python/zipline/`; temporary csvdir bundle ingestion plus `zipline.run_algorithm()`. |
| LEAN CLI | Managed through `dev/bench/peer_benchmark/python/lean/`; invokes the local LEAN CLI or reports unavailable with the CLI failure reason. |

Optional peers must emit a status row. Missing fills/trade surfaces are
labeled with unavailable metadata rather than silently dropped.

### Phase definitions

Each DONE row in the performance table is split into four measured
phases:

- **Snapshot preparation**: reusable native data or bundle preparation.
- **Setup / orchestration** (`experiment_setup_sec`): construction and other
  non-engine work repeated for each experiment or public workflow call.
- **Engine**: from ready-to-iterate until strategy execution completes
  and engine state is final.
- **Results**: from final engine state through the result surface declared by
  the row. Public sweep rows retain and read returns/trades; their richer
  parity-oracle surfaces are generated outside the reported clock.

The four phase columns must reconcile to the cold clock within 0.5 seconds or
the harness aborts. The warm clock is setup/orchestration plus engine plus
results.

The internal phase boundary is not identical across rows. Durable ledgr times
the complete public `ledgr_run()` call as engine work. Public sweep rows expose
candidate fold and inline-results clocks; their remaining public-call wall is
recorded as setup/orchestration. The latter includes warm snapshot reads,
matrix/view preparation and retention assembly, not merely creation of the
experiment object. Cold and warm totals are the cross-row comparison surfaces.
The components explain each row and must not be compared as identical internal
operations.
LEAN is the one explicit exception: when the local CLI runs, the whole CLI
subprocess is bucketed as engine time because preparation, execution, and
extraction are not separable from outside the CLI.

| Engine | Snapshot preparation | Setup / orchestration | Engine | Results |
| --- | --- | --- | --- | --- |
| `ledgr_ttr_canonical` | CSV read, timestamp normalization, and sealed DuckDB snapshot | experiment and run identity | `ledgr_run()` | `ledgr_results()` plus canonical materialization |
| `ledgr_ttr_canonical_sweep` | CSV read, timestamp normalization, and sealed DuckDB snapshot | experiment, one-candidate grid, and public sweep orchestration outside its candidate clocks | public `ledgr_sweep()` candidate fold | inline summary plus public retained returns/trades; full parity oracle excluded |
| `ledgr_ttr_canonical_sweep_with_cost_risk` | Same public sweep boundary | experiment, one-candidate grid, cost model, risk chain, and sweep orchestration | public sweep with cost and risk chains | inline summary plus public retained returns/trades; full parity oracle excluded |
| `ledgr_ttr_compiled_spot_fifo_sweep` | Same public sweep boundary | experiment, one-candidate grid, compiled-model selection, and sweep orchestration | public sweep with `compiled_accounting_model = "spot_fifo"` | inline summary plus public retained returns/trades; full parity oracle excluded |
| `ledgr_builtin_sma` | Same durable ledgr boundary | experiment with built-in SMA features | `ledgr_run()` | `ledgr_results()` plus canonical materialization |
| `quantstrat` | CSV read, xts construction, and global data assignment | portfolio, account, orders, and strategy | `applyStrategy()` plus account updates | equity and transaction extraction plus canonical writes |
| `backtrader` | CSV read and grouped pandas bars | feeds, broker, and strategy construction | `cerebro.run()` | canonical equity, fill, and trade writes |
| `zipline-reloaded-full` | CSV read, csvdir creation, bundle registration, and ingest | algorithm callbacks and calendar | `run_algorithm()` | performance-frame extraction plus canonical writes |
| `LEAN` | Not separable from outside the CLI | Not separable from outside the CLI | whole CLI subprocess if available | Not separable from outside the CLI |

Boundary ambiguity decisions: quantstrat data/xts preparation is reusable while
portfolio, account, order, and strategy construction is experiment-specific;
zipline bundle registration/ingest is reusable native-data preparation; LEAN
CLI phases are unavailable from outside the CLI.

```{r}
#| echo: false
boundary_table <- data.frame(
  Engine = performance$engine,
  Boundary = performance$boundary,
  check.names = FALSE
)
knitr::kable(boundary_table, row.names = FALSE)
```

### Parity tiers

Tier 1 per-bar parity: equity-curve correlation, max single-bar equity
divergence as a fraction of ledgr equity, daily-return correlation, cash
trajectory match where both engines expose cash, and position proxy
match where both engines expose comparable positions.

Tier 2 derived top-line parity: total return, annualized return,
annualized volatility, Sharpe ratio where defined, and max drawdown.

Tier 3 trade-level parity: trade count, win rate and average trade
where the peer exposes comparable realized trade data, and explicit
unavailable metadata where the peer does not.

The ledgr crossover strategy used by this report is vectorized over
`ctx$features_wide`. It keeps previous crossover state in
`ctx$state_prev`, but the current feature comparison and target updates
are vector operations rather than a per-instrument R loop.

### Parity glossary

| Label | Meaning |
| --- | --- |
| `Tier 1` | Per-bar equity/return parity status. `pass` means the row passed the current tolerance; `review` means the row has a documented divergence attribution. |
| `Parity surface` | Which comparable artifacts exist for that peer. Quantstrat is partial because this harness gets account equity, transactions, and a trade count, but not comparable realized trade P&L. |
| `Equity corr` | Correlation between peer equity and ledgr canonical equity. Closer to 1 is better. |
| `Max div` | Largest single-bar equity difference versus ledgr canonical, shown as a percent of ledgr equity. Lower is better. |
| `Return corr` | Correlation between one-bar equity changes. Closer to 1 is better. |
| `Total return delta` | Peer total return minus ledgr canonical total return, in percentage points. |
| `Sharpe diff` | Peer Sharpe minus ledgr canonical Sharpe. |
| `Max DD delta` | Peer max drawdown minus ledgr canonical max drawdown, in percentage points. |
| `Trade count diff` | Peer closed-trade count minus ledgr canonical closed-trade count. |

### Engine status

```{r}
#| echo: false
status_table <- status
status_table <- data.frame(
  Engine = status_table$engine,
  Status = status_table$status,
  Reason = ifelse(is.na(status_table$reason), "", status_table$reason),
  check.names = FALSE
)
status_table[["Reason"]] <- ifelse(
  status_table[["Engine"]] == "LEAN" & status_table[["Status"]] == "UNAVAILABLE",
  "LEAN CLI unavailable: lean backtest rejected the temporary project root; local lean init organization setup is required.",
  status_table[["Reason"]]
)
knitr::kable(status_table[, c("Engine", "Status", "Reason"), drop = FALSE], row.names = FALSE)
```

### Surface availability

```{r}
#| echo: false
knitr::kable(surface, row.names = FALSE)
```

### Latest record bundle

```{r}
#| echo: false
data.frame(
  field = c(
    "created_at", "release", "preset", "input_hash", "git_sha",
    "process_tree_peak_mib", "peak_sampling_interval_sec"
  ),
  value = c(
    env$created_at, env$release, env$preset, env$input_hash, env$git_sha,
    sprintf("%.1f", peer_peak$peak_working_set_mib[[1L]]),
    peer_peak$sampling_interval_sec[[1L]]
  ),
  stringsAsFactors = FALSE
) |>
  knitr::kable()
```

### Isolated quantstrat environment

```{r}
#| echo: false
quantstrat_packages <- c(
  "quantstrat", "blotter", "FinancialInstrument", "xts", "TTR"
)
metadata_text <- function(x) {
  if (is.null(x) || !length(x)) "" else as.character(x[[1L]])
}
quantstrat_table <- do.call(rbind, lapply(quantstrat_packages, function(pkg) {
  metadata <- quantstrat_env$packages[[pkg]]
  data.frame(
    Package = pkg,
    Version = metadata_text(metadata$version),
    `GitHub SHA` = metadata_text(metadata$remote_sha),
    Library = metadata_text(metadata$library),
    check.names = FALSE
  )
}))
knitr::kable(quantstrat_table, row.names = FALSE)
```

## Caveats

### Same-host scope

Hardware, R version, OS, and toolchain versions are pinned to the
maintainer's environment. The numbers are comparable within this record
bundle. They are not a portable speed ranking across hosts or a public
benchmark claim.

### Interpreting the B2 row

The B2 row is interpreted only when measured through the public one-candidate
sweep on the same peer fixture
as the external engines: same shared bars CSV, same fast/slow SMA
settings, same seed. It is also an opt-in compiled accelerator
measurement, not a default ledgr claim. ledgr's default execution path
remains canonical R. Do not combine B2 numbers arithmetically with
workload-grid B2 rows, which use a different fixture.

### Engines deliberately excluded

VectorBT is excluded from the event-driven peer table because it is a
vectorized engine and a different paradigm. Published external LEAN and
Zipline reference rows remain context-only. The same-host LEAN row is
driven only through the real LEAN CLI; if that CLI is not usable
locally, the row is marked unavailable.

### Historical-shape context

Do not compare bars/sec here directly to the v0.1.8.7 `peer_sma_crossover`
row as a pure version regression. That row used the same 500 x 1260
scale but a continuous `fast > slow` target strategy and far fewer
fills. This row uses high-turnover crossover-state semantics plus
per-engine parity artifacts and surface-status outputs. The four-phase
table is the useful comparison artifact, not a single bars/sec headline.
