---
title: "ledgr Peer Parity And Performance Benchmark Report"
format: gfm
output-file: peer_benchmark.md
execute:
  echo: false
  warning: false
  message: true
params:
  run_harness: false
  preset: "smoke"
  results_preset: "record"
---

## Scope

This report is a repo-local maintainer artifact for v0.1.8.8 / LDG-2476. It is
not package documentation, not pkgdown content, and not a public release-note
performance claim.

The report has two separate jobs:

- compute internal parity checks against the ledgr canonical TTR-backed row;
- record local performance timings with the measured boundary stated for each
  engine row.

The parity benchmark asks whether the engines agree on this strategy and data.
The performance benchmark asks how long this local same-host harness took under
the declared boundary. Those are related but not the same question.

Published external LEAN and Ziplime reference rows remain context-only. The
same-host LEAN row is driven only through the real LEAN CLI; if that CLI is not
usable locally, the row is marked unavailable. VectorBT remains excluded from
the event-driven peer table because it is a vectorized engine and a different
paradigm.

## Harness

The current harness is `dev/bench/peer_benchmark/peer_benchmark.R`.

```{r}
if (isTRUE(params$run_harness)) {
  candidates <- c(".", "..", file.path("..", ".."), file.path("..", "..", ".."))
  repo_root <- candidates[file.exists(file.path(candidates, "DESCRIPTION"))][[1L]]
  repo_root <- normalizePath(repo_root, mustWork = TRUE)
  source(file.path(repo_root, "dev", "bench", "peer_benchmark", "peer_benchmark.R"))
  peer_main(peer_parse_args(c("--preset", params$preset)))
} else {
  message("Harness execution disabled. Render with -P run_harness:true to run.")
}
```

Smoke command:

```powershell
& "C:\Program Files\R\R-4.5.2\bin\x64\Rscript.exe" dev/bench/peer_benchmark/peer_benchmark.R --preset smoke
```

The harness writes ignored local artifacts under `dev/bench/results/`:

- one shared bars CSV with input hash;
- per-engine canonical equity curves where available;
- fills and trade summary tables where available;
- engine surface-status rows for unavailable outputs;
- Tier 1, Tier 2, and Tier 3 parity CSV;
- performance timing CSV;
- environment metadata JSON;
- append-only parity history JSON under `dev/bench/results/parity_history/`.

## Required Rows

| Row | Status Policy |
| --- | --- |
| ledgr canonical TTR | Required; errors if `TTR` is missing. |
| ledgr canonical TTR ephemeral | Required; runs the same fold core through an in-memory output handler and must match durable ledgr equity/fills before the run is accepted. |
| ledgr built-in SMA diagnostic | Required; runs through ledgr built-ins. |
| quantstrat | Runs when local R packages are installed; otherwise explicit unavailable row. |
| Backtrader | Managed through `dev/bench/peer_benchmark/python/backtrader/` and run through `python -m uv`. |
| zipline-reloaded full engine | Managed through `dev/bench/peer_benchmark/python/zipline/`; temporary csvdir bundle ingestion plus `zipline.run_algorithm()`. |
| LEAN CLI | Managed through `dev/bench/peer_benchmark/python/lean/`; invokes the local LEAN CLI or reports unavailable with the CLI failure reason. |

Optional peers must emit a status row. Missing fills/trade surfaces are labeled
with unavailable metadata rather than silently dropped.

## Parity Tiers

Tier 1 per-bar parity:

- equity-curve correlation;
- max single-bar equity divergence as a fraction of ledgr equity;
- daily-return correlation;
- cash trajectory match where both engines expose cash;
- position proxy match where both engines expose comparable positions.

Tier 2 derived top-line parity:

- total return;
- annualized return;
- annualized volatility;
- Sharpe ratio where defined;
- max drawdown.

Tier 3 trade-level parity:

- trade count;
- win rate and average trade where the peer exposes comparable realized trade
  data;
- explicit unavailable metadata where the peer does not expose a comparable
  trade surface.

The ledgr crossover strategy used by this report is vectorized over
`ctx$features_wide`. It keeps previous crossover state in `ctx$state_prev`, but
the current feature comparison and target updates are vector operations rather
than a per-instrument R loop.

## Divergence Attribution

When a parity check fails, the candidate explanations are:

1. ledgr is wrong;
2. the peer is wrong;
3. the harness is wrong.

The default mental move is to consider all three, not to assume ledgr first.
Residual divergences are attributed to one of:

- indicator initialization window;
- fill-timing edges;
- cost/margin defaults;
- position-sizing rounding;
- timestamp alignment;
- float-ordering rounding.

## Current Record Results

The record preset is the primary LDG-2476 artifact. Tables and plots below read
the newest matching artifact bundle from `dev/bench/results/`. They are still
internal same-host evidence, not public ranking material.

Record command:

```powershell
& "C:\Program Files\R\R-4.5.2\bin\x64\Rscript.exe" dev/bench/peer_benchmark/peer_benchmark.R --preset record
```

```{r}
#| echo: false
candidates <- c(".", "..", file.path("..", ".."), file.path("..", "..", ".."))
repo_root <- candidates[file.exists(file.path(candidates, "DESCRIPTION"))][[1L]]
repo_root <- normalizePath(repo_root, mustWork = TRUE)
result_dir <- file.path(repo_root, "dev", "bench", "results")
status_files <- Sys.glob(file.path(
  result_dir,
  sprintf("peer_benchmark_%s_*_status.csv", params$results_preset)
))
status_files <- status_files[!grepl("_surface_status[.]csv$", status_files)]
if (!length(status_files)) {
  stop("No peer benchmark status files found for preset: ", params$results_preset)
}
status_file <- tail(sort(status_files), 1L)
stem <- sub("_status[.]csv$", "", status_file)
read_result <- function(suffix) {
  path <- paste0(stem, suffix)
  if (!file.exists(path)) stop("Missing expected result file: ", path)
  utils::read.csv(path, stringsAsFactors = FALSE)
}
status <- read_result("_status.csv")
parity <- read_result("_parity.csv")
performance <- read_result("_performance.csv")
divergence_summary <- read_result("_divergence_summary.csv")
surface <- read_result("_surface_status.csv")
env <- jsonlite::read_json(paste0(stem, "_environment.json"), simplifyVector = TRUE)
equity_files <- Sys.glob(paste0(stem, "_*_equity.csv"))
equity <- do.call(rbind, lapply(equity_files, utils::read.csv, stringsAsFactors = FALSE))

fmt_num <- function(x, digits = 6) {
  ifelse(is.na(x), NA, formatC(as.numeric(x), digits = digits, format = "fg", flag = "#"))
}
fmt_sec <- function(x) ifelse(is.na(x), NA, sprintf("%.3f", as.numeric(x)))
fmt_rate <- function(x) {
  ifelse(is.na(x), NA, formatC(as.numeric(x), digits = 6, format = "fg", big.mark = ","))
}
fmt_pct <- function(x, digits = 4) {
  pct <- 100 * as.numeric(x)
  ifelse(
    is.na(pct),
    NA,
    ifelse(abs(pct) < 0.0001 & pct != 0, "<0.0001%", paste0(formatC(pct, digits = digits, format = "fg", flag = "#"), "%"))
  )
}
fmt_corr <- function(x) ifelse(is.na(x), NA, sprintf("%.6f", as.numeric(x)))
fmt_signed <- function(x, digits = 6) {
  ifelse(is.na(x), NA, formatC(as.numeric(x), digits = digits, format = "fg", flag = "#"))
}
fmt_pp <- function(x, digits = 5) {
  pp <- 100 * as.numeric(x)
  ifelse(
    is.na(pp),
    NA,
    ifelse(abs(pp) < 0.00001 & pp != 0, "<0.00001 pp", paste0(formatC(pp, digits = digits, format = "fg", flag = "#"), " pp"))
  )
}
tier1_pass <- function(attribution) {
  ifelse(identical(attribution, "passes Tier 1 tolerance"), "pass", "review")
}
parity_surface <- function(peer) {
  trade_surface <- surface$trades[match(peer, surface$engine)]
  fill_surface <- surface$fills[match(peer, surface$engine)]
  if (identical(peer, "quantstrat")) {
    return("partial: equity + trade count only")
  }
  if (identical(fill_surface, "available") && identical(trade_surface, "available_realized_pnl")) {
    return("equity + fills + realized trades")
  }
  if (identical(trade_surface, "available_realized_pnl")) {
    return("equity + realized trades; fills unavailable")
  }
  trade_surface %||% "unavailable"
}
```

Latest record bundle:

```{r}
#| echo: false
data.frame(
  field = c("created_at", "release", "preset", "input_hash", "git_sha"),
  value = c(env$created_at, env$release, env$preset, env$input_hash, env$git_sha),
  stringsAsFactors = FALSE
) |>
  knitr::kable()
```

### Engine Status

This table only shows whether each row completed. Timing interpretation is in
the separate performance section below.

```{r}
#| echo: false
status_table <- status
names(status_table) <- c("Engine", "Status", "Wall seconds", "Ingestion seconds", "Engine seconds", "Results seconds", "Reason")
status_table[["Wall seconds"]] <- fmt_sec(status_table[["Wall seconds"]])
status_table[["Reason"]] <- ifelse(is.na(status_table[["Reason"]]), "", status_table[["Reason"]])
status_table[["Reason"]] <- ifelse(
  status_table[["Engine"]] == "LEAN" & status_table[["Status"]] == "UNAVAILABLE",
  "LEAN CLI unavailable: lean backtest rejected the temporary project root; local lean init organization setup is required.",
  status_table[["Reason"]]
)
knitr::kable(status_table[, c("Engine", "Status", "Reason"), drop = FALSE], row.names = FALSE)
```

### Performance Benchmark

The performance benchmark uses the same record input as the parity benchmark:
`r env$preset`, `r unique(performance$n_instruments)` instruments, `r unique(performance$n_days)`
daily bars, and `r format(unique(performance$n_bars), big.mark = ",")` total bars.

The report now splits each DONE row into three measured phases:

- `Ingestion`: from timed-window start until the engine has native data
  structures ready to iterate.
- `Engine`: from ready-to-iterate until strategy execution completes and engine
  state is final.
- `Results`: from final engine state until canonical equity/fills/trades are
  materialized for this harness.

The three phase columns must reconcile to `Total` within 0.5 seconds or the
harness aborts. LEAN is the one explicit exception: the local CLI row is
unavailable here; if it becomes available, the CLI subprocess boundary is
bucketed as engine time because ingestion/run/extraction are not separable from
outside the CLI.

This is a real local performance benchmark under stated boundaries. It is still
not a public speed ranking because the rows intentionally expose different
engine surfaces. Ledgr now appears in two rows: durable and ephemeral. The
ephemeral row removes DuckDB snapshot/event-log persistence while keeping the
same fold core, so the durability cost is visible as the delta between those two
ledgr rows.

Use `Total` for the engine-row boundary and `Full row` for the current wrapper
cost around that row. Large full-row overhead, especially in Python rows,
reflects per-invocation process/environment setup outside the phase total.

Do not compare the bars/sec here directly to the v0.1.8.7 `peer_sma_crossover`
row as a pure version regression. The old row used the same 500 x 1260 scale but
a continuous `fast > slow` target strategy and far fewer fills. This row uses
high-turnover crossover-state semantics plus per-engine parity artifacts and
surface-status outputs.

```{r}
#| echo: false
perf_table <- data.frame(
  Engine = performance$engine,
  Status = performance$status,
  `Full row` = fmt_sec(performance$full_row_sec),
  Ingestion = fmt_sec(performance$ingestion_sec),
  Engine = fmt_sec(performance$engine_sec),
  Results = fmt_sec(performance$results_sec),
  Total = fmt_sec(performance$reported_core_sec),
  `Bars/sec` = fmt_rate(performance$core_bars_per_sec),
  check.names = FALSE
)
knitr::kable(perf_table, row.names = FALSE)
```

### Phase Definitions

| Engine | Ingestion | Engine | Results |
| --- | --- | --- | --- |
| `ledgr_ttr_canonical` | `read.csv`, timestamp normalization, DuckDB snapshot creation, experiment construction | `ledgr_run()` | `ledgr_results()` for equity/fills plus canonical materialization |
| `ledgr_ttr_canonical_ephemeral` | `read.csv`, timestamp normalization, in-memory bars/features/projection construction | `ledgr_execute_fold()` with `ledgr_memory_output_handler()` | event-stream equity/fills reconstruction plus canonical materialization |
| `ledgr_builtin_sma` | Same durable ledgr boundary with built-in SMA features | `ledgr_run()` | `ledgr_results()` for equity/fills plus canonical materialization |
| `quantstrat` | `read.csv`, xts/globalenv construction, portfolio/account/orders/strategy setup | `applyStrategy()` plus account updates | account/transaction extraction plus canonical materialization |
| `backtrader` | `read.csv`, `PandasData` construction, `cerebro.adddata` loop | `cerebro.run()` | CSV writes from in-memory observer/fill rows |
| `zipline-reloaded-full` | `read.csv`, temporary csvdir creation, bundle registration and ingest | `run_algorithm()` | performance-frame transaction/equity extraction plus CSV writes |
| `LEAN` | Not separable from outside the CLI | whole CLI subprocess if available | Not separable from outside the CLI |

Boundary ambiguity decisions: quantstrat initialization is ingestion because it
builds native portfolio/order state before strategy iteration; zipline bundle
registration/ingest is ingestion because it creates the native bundle consumed
by `run_algorithm()`; LEAN CLI phases are unavailable from outside the CLI and
are disclosed as such.

Timing boundary notes:

```{r}
#| echo: false
boundary_table <- data.frame(
  Engine = performance$engine,
  Boundary = performance$boundary,
  check.names = FALSE
)
knitr::kable(boundary_table, row.names = FALSE)
```

```{r}
#| echo: false
#| fig.width: 9
#| fig.height: 4.8
old_par <- par(no.readonly = TRUE)
par(mar = c(5, 10, 3, 1))
perf_done <- performance[is.finite(performance$full_row_sec), , drop = FALSE]
ord <- order(perf_done$full_row_sec)
perf_done <- perf_done[ord, , drop = FALSE]
mat <- rbind(
  Ingestion = perf_done$ingestion_sec,
  Engine = perf_done$engine_sec,
  Results = perf_done$results_sec
)
mat[!is.finite(mat)] <- 0
bp <- barplot(
  mat,
  beside = FALSE,
  horiz = TRUE,
  names.arg = perf_done$engine,
  las = 1,
  col = c("#7A9E9F", "#D19C1D", "#B85C5C"),
  border = NA,
  xlab = "Seconds",
  main = "Record Performance Timing\ningestion / engine / results phases"
)
legend("bottomright", legend = rownames(mat), fill = c("#7A9E9F", "#D19C1D", "#B85C5C"), bty = "n")
grid(nx = NA, ny = NULL, col = "gray85")
totals <- colSums(mat)
text(totals, bp, labels = paste0(sprintf("%.3f", totals), "s"), pos = 4, cex = 0.72)
par(old_par)
```

The performance comparison is phase-explicit for DONE rows. The LEAN row is
unavailable in this local run because the CLI failed before the engine could
start.

### Historical Shape Context

The v0.1.8.7 closeout row was 500 x 1260 with continuous target semantics and
about 13k fills. This record is 500 x 1260 with crossover-event semantics and
about 68k ledgr fills. The three-phase table is therefore the useful artifact:
it shows ingestion, engine, and result materialization separately instead of
collapsing a high-turnover workload into one headline number.

### Parity Glossary

The raw CSV keeps machine-readable field names for downstream scripts. The
tables below use shorter labels:

| Label | Meaning |
| --- | --- |
| `Tier 1` | Per-bar equity/return parity status. `pass` means the row passed the current tolerance; `review` means the row has a documented divergence attribution. |
| `Parity surface` | Which comparable artifacts exist for that peer. Quantstrat is partial because this harness currently gets account equity and transaction count, not comparable fills or realized trade P&L. |
| `Equity corr` | Correlation between peer equity and ledgr canonical equity. Closer to 1 is better. |
| `Max div` | Largest single-bar equity difference versus ledgr canonical, shown as a percent of ledgr equity. Lower is better. |
| `Return corr` | Correlation between one-bar equity changes. Closer to 1 is better. |
| `Total return delta` | Peer total return minus ledgr canonical total return, in percentage points. |
| `Sharpe diff` | Peer Sharpe minus ledgr canonical Sharpe. |
| `Max DD delta` | Peer max drawdown minus ledgr canonical max drawdown, in percentage points. |
| `Trade count diff` | Peer closed-trade count minus ledgr canonical closed-trade count. |

### Tier 1 Per-Bar Parity

```{r}
#| echo: false
tier1 <- data.frame(
  Peer = parity$peer,
  `Tier 1` = ifelse(
    parity$peer == "quantstrat",
    "review (partial surface)",
    ifelse(parity$attribution == "passes Tier 1 tolerance", "pass", "review")
  ),
  `Parity surface` = vapply(parity$peer, parity_surface, character(1)),
  `Equity corr` = fmt_corr(parity$tier1_equity_cor),
  `Max div` = fmt_pct(parity$tier1_max_single_bar_divergence_pct, digits = 4),
  `Return corr` = fmt_corr(parity$tier1_daily_return_cor),
  check.names = FALSE
)
knitr::kable(tier1, row.names = FALSE)
```

Rows marked `review` keep the standard divergence-source attribution:

```{r}
#| echo: false
review <- parity[parity$attribution != "passes Tier 1 tolerance", c("peer", "attribution"), drop = FALSE]
names(review) <- c("Peer", "Review attribution")
if (nrow(review) == 0L) {
  review <- data.frame(Peer = "(none)", `Review attribution` = "All peer rows passed Tier 1 tolerance.", check.names = FALSE)
}
knitr::kable(review, row.names = FALSE)
```

### Per-Bar Divergence Attribution

Every DONE peer row writes a per-bar `_divergence.csv` plus a
`_divergence_summary.csv`. Attribution is computed from equity divergence and
same-timestamp fill/transaction comparisons against `ledgr_ttr_canonical`.

```{r}
#| echo: false
div_table <- divergence_summary
pct_cols <- grep("^pct_attributable_", names(div_table), value = TRUE)
for (col in pct_cols) div_table[[col]] <- fmt_pct(div_table[[col]], digits = 4)
names(div_table) <- c(
  "Peer",
  "Total abs divergence",
  "Diverging bars",
  "First divergence",
  "Indicator warmup",
  "Fill timing",
  "Calendar",
  "Position size",
  "Float rounding",
  "Other"
)
div_table[["Total abs divergence"]] <- fmt_num(div_table[["Total abs divergence"]], digits = 7)
knitr::kable(div_table, row.names = FALSE)
```

```{r}
#| echo: false
#| fig.width: 8
#| fig.height: 4.5
old_par <- par(no.readonly = TRUE)
par(mar = c(5, 12, 3, 2))
div <- 100 * as.numeric(parity$tier1_max_single_bar_divergence_pct)
names(div) <- parity$peer
plot_div <- pmax(div, 1e-12)
ord <- order(plot_div)
plot_div <- plot_div[ord]
labels <- ifelse(div[ord] == 0, "0%", fmt_pct(div[ord] / 100, digits = 4))
plot(
  log10(plot_div),
  seq_along(plot_div),
  yaxt = "n",
  ylab = "",
  xlab = "log10(max divergence percent); zeros shown at display floor",
  pch = 19,
  col = ifelse(parity$attribution[ord] == "passes Tier 1 tolerance", "#5B8C5A", "#B85C5C"),
  main = "Tier 1 Max Divergence vs ledgr Canonical\nlog scale so tiny LEAN/zipline float differences remain visible"
)
axis(2, at = seq_along(plot_div), labels = names(plot_div), las = 1)
grid(col = "gray85")
text(log10(plot_div), seq_along(plot_div), labels = paste0(" ", labels), pos = 4, cex = 0.8)
par(old_par)
```

### Tier 2 Top-Line Metrics

```{r}
#| echo: false
tier2 <- data.frame(
  Peer = parity$peer,
  `Total return delta` = fmt_pp(parity$tier2_total_return_diff, digits = 5),
  `Sharpe diff` = fmt_signed(parity$tier2_sharpe_diff, digits = 6),
  `Max DD delta` = fmt_pp(parity$tier2_max_drawdown_diff, digits = 5),
  check.names = FALSE
)
knitr::kable(tier2, row.names = FALSE)
```

### Tier 3 Trade Surface

```{r}
#| echo: false
tier3 <- data.frame(
  Peer = parity$peer,
  `Trade count diff` = parity$tier3_trade_count_diff,
  `Trade surface` = surface$trades[match(parity$peer, surface$engine)],
  `Parity note` = vapply(parity$peer, parity_surface, character(1)),
  check.names = FALSE
)
knitr::kable(tier3, row.names = FALSE)
```

### Equity Curves

The plot normalizes equity for each engine to its first observed value, so shape
differences are visible without implying an absolute performance ranking between
engines.

```{r}
#| echo: false
#| fig.width: 9
#| fig.height: 5
equity$ts_utc <- as.POSIXct(equity$ts_utc, tz = "UTC")
equity <- equity[order(equity$engine, equity$ts_utc), ]
engines <- unique(equity$engine)
cols <- c("#2F4858", "#5B8C5A", "#B85C5C", "#7A9E9F", "#8D6A9F", "#D19C1D", "#4C6A92")
plot(NULL, xlim = range(equity$ts_utc), ylim = range(unlist(lapply(engines, function(engine) {
  y <- equity$equity[equity$engine == engine]
  y / y[[1L]]
})), na.rm = TRUE), xaxt = "n", xlab = "Month", ylab = "Equity / first observed equity", main = "Record Equity Curves\nnormalized within each engine")
axis.POSIXct(1, at = pretty(equity$ts_utc, n = 8), format = "%Y-%m")
grid(col = "gray88")
for (i in seq_along(engines)) {
  d <- equity[equity$engine == engines[[i]], ]
  lines(d$ts_utc, d$equity / d$equity[[1L]], col = cols[((i - 1L) %% length(cols)) + 1L], lwd = 2)
}
legend_cols <- cols[((seq_along(engines) - 1L) %% length(cols)) + 1L]
legend("topleft", legend = engines, col = legend_cols, lwd = 2, bty = "n", cex = 0.75)
```

### Surface Availability

```{r}
#| echo: false
knitr::kable(surface, row.names = FALSE)
```
